---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  dpi = 300
)
```

# sdbuildR: Easily Build and Simulate Stock-and-Flow Models in R

<!-- badges: start -->
[![R-CMD-check](https://github.com/KCEvers/sdbuildR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/KCEvers/sdbuildR/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/KCEvers/sdbuildR/graph/badge.svg)](https://app.codecov.io/gh/KCEvers/sdbuildR)
<!-- badges: end -->

Model systems as stock-and-flow models in R. Stock-and-flow models are foundational to system dynamics, and help to understand systems complicated by nonlinearities, delays, and feedback loops. sdbuildR aims to make stock-and-flow modeling accessible and effortless, enabling you to dedicate your expertise to what matters most: building insightful, high-quality models. Get started at [https://kcevers.github.io/sdbuildR/](https://kcevers.github.io/sdbuildR/)!

## Features

* **Accessibility**: Get started with stock-and-flow models with limited knowledge.
* **Flexibility**: Modify your models easily and robustly.
* **Insight Maker integration**: Import models from [Insight Maker](https://insightmaker.com/).
* **Julia backend**: Use Julia for high-performance simulations, without any knowledge of Julia.
* **Unit support**: Use standard or custom units to verify and interpret your model. 

All package capabilities are described in the vignettes:

* `vignette("build")` Learn how to build, modify, and debug stock-and-flow models.
* `vignette("ensemble")` Learn how to assess a model's sensitivity, uncertainty and robustness with ensemble simulations.
* `vignette("units")` Learn why and how to use (custom) units.
<!-- 3. **Delays**: Not all dynamics are instantaneous. Learn about delays (fixed, smooth, material and information) and how to implement them with sdbuildR. -->
<!-- 4. **Sensitivity**:  -->


## Installation

You can install the development version of sdbuildR like so:

```{r eval=FALSE, include=TRUE}
# install.packages("remotes")
remotes::install_github("KCEvers/sdbuildR")
```

sdbuildR offers two simulation engines: R and Julia (supported by [JuliaConnectoR](https://github.com/stefan-m-lenz/JuliaConnectoR/)). If you would like to run ensemble simulations, use units and implement delay functions, you will need to set up the Julia environment. Run `use_julia()` to install Julia and required packages (initial setup may take 5–15 minutes):

```{r installing Julia, eval=FALSE, include=TRUE}
sdbuildR::use_julia()
```


## Limitations

Unlike in other System Dynamics software, sdbuildR minimally supports non-negative stocks and flows. Setting stocks to non-negative will not adjust flows. In any case, enforcing either stocks or flows to be non-negative is not recommended, as it may mask model misspecification. Stocks and flows that logically cannot be negative (e.g., animals or deaths) should ideally remain non-negative as a result of the model's equations and parameters, rather than by forcing them to be non-negative. In addition, sdbuildR does not support vectorized operations. 

## Other System Dynamics software

sdbuildR is heavily based on common System Dynamics software such as [Vensim](https://vensim.com/), [Powersim](https://powersim.com/), [Stella](https://www.iseesystems.com/), and [Insight Maker](https://insightmaker.com/). To translate xmile models to R, see the R package [readsdr](https://github.com/jandraor/readsdr). To build stock-and-flow models with the R package [deSolve](https://cran.r-project.org/web/packages/deSolve/index.html), the book [System Dynamics Modeling with R](https://link.springer.com/book/10.1007/978-3-319-34043-2) by Jim Duggan will prove useful. In Python, stock-and-flow models are supported by [PySD](https://pysd.readthedocs.io/en/master/).


## Troubleshooting

sdbuildR is under active development. While thoroughly tested, the package may have bugs, particularly in complex model translations. We encourage users to report issues on GitHub (https://github.com/KCEvers/sdbuildR/issues) - your input helps the package improve! Use `debugger()` to diagnose model errors, and use the vignettes for guidance.
