---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  dpi = 300
)
```

# sdbuildR: Stock-and-flow modelling with R

<!-- badges: start -->
<!-- badges: end -->

sdbuildR makes it easy to create and simulate stock-and-flow models in R. With Julia as a backend, it combines fast performance with flexibility in a user-friendly R environment.

## Installation

You can install the development version of sdbuildR like so:

```{r eval=FALSE, include=TRUE}
# install.packages("remotes")
remotes::install_github("KCEvers/sdbuildR")
```

sdbuildR uses Julia as its primary backend, with limited R-based simulation support. To enable full functionality, install Julia and required packages via JuliaCall. Run `sdbuildR_setup()` to configure Julia (initial setup may take 5–15 minutes):

```{r installing Julia, eval=FALSE, include=TRUE}
sdbuildR::sdbuildR_setup()
```
