---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  dpi = 300
)
```

# sdbuildR: Stock-and-flow modelling in R

<!-- badges: start -->
[![R-CMD-check](https://github.com/KCEvers/sdbuildR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/KCEvers/sdbuildR/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

Easily build and simulate stock-and-flow models with sdbuildR in R. With Julia as a backend, sdbuildR combines fast performance with flexibility in a user-friendly R environment.

## Installation

You can install the development version of sdbuildR like so:

```{r eval=FALSE, include=TRUE}
# install.packages("remotes")
remotes::install_github("KCEvers/sdbuildR")
```

sdbuildR offers two simulation engines: R and Julia. If you would like to use units and delay functions, you will need to set up the Julia environment. Run `use_julia()` to install Julia and required packages (initial setup may take 5–15 minutes):

```{r installing Julia, eval=FALSE, include=TRUE}
sdbuildR::use_julia()
```
