---
title: "Building stock-and-flow models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Building stock-and-flow models}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: inline
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6, # Set default figure width to 6 inches
  fig.height = 4, # Set default figure height to 4 inches
  fig.align = "center", # Center all figures
  out.width = "80%" # Display figures at 80% of container width
  # fig.retina = 2,         # Optimize for retina displays
  # dpi = 96               # Standard DPI for web
)
```

```{r, echo=TRUE, message=FALSE}
library(sdbuildR)

# Disable WebGL: many plotly widgets per HTML page can exceed the browser WebGL
# context limit and render blank. SVG always renders.
options(sdbuildR.webgl = FALSE)
```

```{r libraries, include=FALSE}
library(knitr)
library(kableExtra)
```


```{r include=FALSE}
# Set this to TRUE to recreate the figures in the paper
recreate_figs <- interactive()
```

```{r include=FALSE}
# Helper function
sprintf_arg <- function(template, values) {
  for (nm in names(values)) {
    template <- gsub(paste0("%\\(", nm, "\\)s"), values[[nm]], template)
  }
  return(template)
}
font_family <- "Times New Roman"

# Plotting parameters
filepath_base <- this.path::this.dir()
filepath_figs <- file.path(filepath_base, "figs_tutorial2026")

if (recreate_figs) {
  dir.create(filepath_figs, showWarnings = FALSE)
}

VARS <- c("engagement", "performance", "energy", "demands", "resources", "motivation_rate")
PALETTE <- c("Geyser", "Temps", "Spectral", "Zissou 1", "Roma")[5]
COLORS <- grDevices::hcl.colors(n = 50, palette = PALETTE)
# scales::show_col(COLORS)

inflow_col <- adjustcolor(COLORS[10], alpha.f = 1)
outflow_col <- adjustcolor(COLORS[3], alpha.f = 1)

inflow_fill <- adjustcolor(inflow_col, alpha.f = .5)
outflow_fill <- adjustcolor(outflow_col, alpha.f = .5)

# scales::show_col( grDevices::hcl.colors(n = 50, palette = "Greens"))
increase_col <- adjustcolor(
  grDevices::hcl.colors(n = 50, palette = "Greens")[20],
  alpha.f = 1)
decrease_col <- adjustcolor(COLORS[2], alpha.f = 1)
# scales::show_col(c(increase_col, decrease_col))
increase_fill <- adjustcolor(increase_col, alpha.f = .8)
decrease_fill <- adjustcolor(decrease_col, alpha.f = .4)

COLORS_alpha <- adjustcolor(COLORS, alpha.f = .25)

COLORS <- COLORS[c(37, 9, 32, 1, 47, 17)]
stock_col <- adjustcolor(COLORS[1], alpha.f = .75)
flow_col <- COLORS[6]

health_path <- adjustcolor(COLORS[4], alpha.f = .75)
motivation_path <- adjustcolor(COLORS[1], alpha.f = .85)
health_path_fill <- adjustcolor(COLORS[4], alpha.f = .25)
motivation_path_fill <- adjustcolor(COLORS[1], alpha.f = .25)
performance_path <- adjustcolor(COLORS[2], alpha.f = .75)
performance_path_fill <- adjustcolor(COLORS[2], alpha.f = .25)

line_width <- 1.5
line_width_stock <- line_width + 2

font_size <- 16
font_size_xlab <- round(font_size * 1.4)
font_size_legend <- round(font_size * 1.1)
font_size_subplot <- font_size # round(font_size * 1.5)
xlab <- "Time (days)"

# Finer-grained simulation
save_at_fine <- .1 # .01
save_at_coarse <- 1

# Annotation template
annot <- list(xanchor = "center", yanchor = "middle", showarrow = FALSE, xref = "paper", yref = "paper", font = list(size = font_size_subplot, family = font_family))
annot_font <- list(family = font_family, size = font_size)
annot_legend <- list(
  orientation = "h", # orientation
  x = 0.5, # Center horizontally
  y = -0.15, # Below the plot
  xanchor = "center",
  yanchor = "top",
  # traceorder = "normal",
  # tracegroupgap = 0,
  # itemwidth = 15,
  font = list(size = font_size_legend)
)
annot_xlab <- list(
  text = xlab,
  font = list(
    family = font_family,
    size = font_size_xlab
  ),
  xref = "paper",
  yref = "paper",
  xanchor = "center",
  yanchor = "top",
  x = 0.5, y = -.1,
  showarrow = FALSE
)

rm_grid <- function(pl) {
  plotly::layout(pl,
    xaxis = list(showgrid = FALSE, showline = TRUE),
    yaxis = list(showgrid = FALSE, showline = TRUE)
  )
}
```


Stock-and-flow models represent systems as states (stocks) that accumulate over time with processes (flows) that change these variables. In this vignette, we will demonstrate how to create stock-and-flow models from scratch using sdbuildR. It covers the basics of stock-and-flow modelling in the context of psychology with an example of burnout. Note that this vignette serves as online supplemental material A accompanying the paper *Formalizing Psychological Theory with sdbuildR: A Stock-and-Flow Modelling Tutorial in R* by Evers et al. (under review). To reproduce the figures in the paper, please see the corresponding .Rmd file.

## Stock-and-flow models

Stock-and-flow models conceptualize systems in terms of quantities that accumulate (i.e., stocks) and the processes (i.e., flows) that change them over time. Stocks are like the amount of water in a bathtub: they store the effects of past and present flows. Stocks must be able to increase and decrease, and should be measurable at a single moment in time. Inflows -- water from the tap -- raise the stock, while outflows -- water through the drain -- lower it. As such, flows represent the rates at which stocks change, measured in units per time (e.g., litre per minute). The net rate of change in the water level is determined by the difference between the inflows and outflows. In this way, a stock functions as a memory of past activity: it increases when inflows exceed outflows and decreases when outflows outpace inflows. Without an outflow, the water remains in the bathtub; without an inflow, the bathtub stays empty. This structure is the foundation of stock-and-flow models, where stocks represent the state of a system, and flows represent the processes that alter that state over time.


```{r, echo = FALSE}
colors <- c(stock_col, inflow_col, outflow_col)
sfm_list <- list()

sfm <- stockflow() |>
  sim_settings(only_stocks = FALSE, stop = 1, save_at = save_at_fine, dt = .01) |>
  stock(Stock)

sfm1 <- sfm |>
  flow(Inflow, eqn = 1, to = Stock)
sfm_list[["A"]] <- sfm1

sfm2 <- sfm |>
  flow(Outflow, eqn = 1, from = Stock)
sfm_list[["B"]] <- sfm2

sfm3 <- sfm |>
  flow(Inflow, eqn = 1, to = Stock) |>
  flow(Outflow, eqn = 5 * Stock, from = Stock)
sfm_list[["C"]] <- sfm3

sfm_big <- stockflow() |>
  stock(stockA, label = "Stock") |>
  flow(inflowA, eqn = 1, label = "Inflow", to = stockA) |>
  stock(stockB, label = "Stock") |>
  flow(outflowB, eqn = 1, label = "Outflow", from = stockB) |>
  stock(Stock, label = "Stock") |>
  flow(inflowC, eqn = 1, label = "Inflow", to = Stock) |>
  flow(outflowC, eqn = 5 * Stock, label = "Outflow", from = Stock)

# copy cat(plot(sfm_big, minlen = 1, nodesep= .8, pad=.1, order = c("stockA", "stockB", "stockC"), align = c("stockA", "stockB", "stockC"))$x$diagram) and paste into viz_str below


# {
#   viz_str <- "
#   digraph sfm {

#     graph [layout = dot, rankdir = LR, center=true, outputorder='edgesfirst', pad=.1, nodesep= .8]

#     # Rank groupings
#     {rank=same; stockA; stockB; stockC}

#     # Invisible edge to fix order
#     stockA -> stockB -> stockC [style=invis]

#     # Define stock nodes
#     'stockA' [id='stockA',label='Stock',tooltip = 'eqn = 0',shape=box,style=filled,fillcolor='#09B0ACBF',fontsize=20,fontname='Times New Roman']
#     'stockB' [id='stockB',label='Stock',tooltip = 'eqn = 0',shape=box,style=filled,fillcolor='#09B0ACBF',fontsize=20,fontname='Times New Roman']
#     'stockC' [id='stockC',label='Stock',tooltip = 'eqn = 0',shape=box,style=filled,fillcolor='#09B0ACBF',fontsize=20,fontname='Times New Roman']

#     # Define flow nodes (intermediate nodes for flows)
#     'inflowA' [id='inflowA',label='Inflow', tooltip = 'eqn = 0', shape = plaintext, fontsize=18, fontname='Times New Roman', width=0.6, height=0.3]
#     'outflowB' [id='outflowB',label='Outflow', tooltip = 'eqn = 0', shape = plaintext, fontsize=18, fontname='Times New Roman', width=0.6, height=0.3]
#     'inflowC' [id='inflowC',label='Inflow', tooltip = 'eqn = 0', shape = plaintext, fontsize=18, fontname='Times New Roman', width=0.6, height=0.3]
#     'outflowC' [id='outflowC',label='Outflow', tooltip = 'eqn = 0', shape = plaintext, fontsize=18, fontname='Times New Roman', width=0.6, height=0.3]

#     # Define external cloud nodes
#     'Cloud1' [label='', tooltip = 'Unspecified source',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
#     'Cloud2' [label='', tooltip = 'Unspecified source',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
#     'Cloud3' [label='', tooltip = 'Unspecified sink',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
#     'Cloud4' [label='', tooltip = 'Unspecified sink',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]

#     # Define flow edges (stock -> flow_node -> stock)
#     'Cloud3' -> 'inflowA' [arrowhead='none', color='black:%(inflow_col)s:black', penwidth=1.1, minlen=1.5, splines=false]
#     'inflowA' -> 'stockA' [arrowhead='normal', color='black:%(inflow_col)s:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]
#     'stockB' -> 'outflowB' [arrowhead='none', color='black:%(outflow_col)s:black', penwidth=1.1, minlen=1.5, splines=false]
#     'outflowB' -> 'Cloud1' [arrowhead='normal', color='black:%(outflow_col)s:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]
#     'Cloud4' -> 'inflowC' [arrowhead='none', color='black:%(inflow_col)s:black', penwidth=1.1, minlen=1.5, splines=false]
#     'inflowC' -> 'stockC' [arrowhead='normal', color='black:%(inflow_col)s:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]
#     'stockC' -> 'outflowC' [arrowhead='none', color='black:%(outflow_col)s:black', penwidth=1.1, minlen=1.5, splines=false]
#     'outflowC' -> 'Cloud2' [arrowhead='normal', color='black:%(outflow_col)s:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]

#   }

# "
# viz_str <- sprintf_arg(
#     viz_str,
#     list(
#       inflow_col = inflow_col,
#       outflow_col = outflow_col
#     )
#   )
#
# pl <- DiagrammeR::grViz(viz_str)
# }

pl <- plot(sfm_big, minlen = 1, nodesep = .8, pad = .1, order = c("stockA", "stockB", "Stock"), align = c("stockA", "stockB", "Stock"), show_eqn = FALSE, show_dependencies = FALSE)


if (recreate_figs) {
  export_plot(pl, file.path(filepath_figs, "sfm.pdf"))
}

print(pl)
```



```{r, echo = FALSE}
vars <- c("Stock", "Inflow", "Outflow")
colors <- c(stock_col, inflow_col, outflow_col)
pl_list <- list()

for (i in names(sfm_list)) {
  if (i == "A") {
    idx <- c(1, 2)
  } else if (i == "B") {
    idx <- c(1, 3)
  } else {
    idx <- 1:3
  }
  # idx <- rev(idx)

  sim <- simulate(sfm_list[[i]], save_at = save_at_fine)
  pl_list[[i]] <- plot(sim,
    vars = vars[idx],
    xlab = "", main = "", # line_width = 3,
    colors = colors[idx],
    font_family = font_family,
    font_size = font_size_subplot,
    showlegend = (i == "C")
  ) |> rm_grid()
}

# Simulation
pl_sim <- plotly::subplot(pl_list,
  nrows = 3,
  shareX = FALSE, shareY = FALSE,
  margin = 0.05, heights = c(.25, .3, .25)
) |>
  plotly::layout(
    # margin = 0.05,
    font = annot_font,
    legend = utils::modifyList(annot_legend, list(traceorder = "normal", y = -0.02))
  ) |>
  plotly::layout(
    yaxis = list(range = c(0, 1.1)),
    yaxis3 = list(range = c(0, 1.1)),
    annotations = list(
      utils::modifyList(annot_xlab, list(x = 0.5, y = 0.05))
    )
  )
# print(pl_sim)
```

```{r, echo = FALSE, eval = has_internet(), error = TRUE}
if (recreate_figs) {
  export_plot(pl_sim, file.path(filepath_figs, "sim.pdf"),
    width = 3.25, height = 5.5
  )


  # Simulation
  pl_sim <- plotly::subplot(pl_list,
    nrows = 1,
    shareX = FALSE, shareY = TRUE,
    margin = 0.025, widths = c(.25, .3, .25)
  ) |>
    plotly::layout(
      # margin = 0.05,
      font = annot_font,
      legend = utils::modifyList(annot_legend, list(traceorder = "normal", y = -0.45))
    ) |>
    plotly::layout(
      yaxis = list(range = c(-1.1, 1.1)),
      # yaxis2 = list(range = c(-1.1, 1.1)),
      # yaxis3 = list(range = c(0, 1.1)),
      annotations = list(
        utils::modifyList(annot_xlab, list(x = 0.5, y = -0.22))
      )
    )

  export_plot(pl_sim, file.path(filepath_figs, "sim_row.pdf"),
    width = 6, height = 2
  )
}
```




```{r, echo = FALSE}
colors <- c(stock_col, inflow_col, outflow_col)

sfm <- stockflow() |>
  sim_settings(only_stocks = FALSE, stop = 1, save_at = save_at_fine, dt = .01) |>
  stock(Stock) |>
  flow(Inflow, eqn = 1, to = Stock) |>
  flow(Outflow, eqn = 5 * Stock, from = Stock)

pl <- plot(sfm, minlen = 1, nodesep = .8, pad = .1, show_eqn = FALSE, show_dependencies = FALSE)
# cat(pl$x$diagram)

# {
#   viz_str <- '
#      digraph sfm {

#       graph [layout = dot, rankdir = LR, center=true, outputorder="edgesfirst", pad=0.1, nodesep=0.8]

#       # Shared across all nodes (persists until overridden)
#       node [fontsize=18,fontname="Times New Roman",fontcolor="black"]

#       # Define stock nodes
#       node [shape=box,style=filled,fillcolor="#83d3d4"]
#       "Stock" [id="Stock",label="Stock", tooltip = "Stock: Stock\nInitial value: 0\nInflows: Inflow\nOutflows: Outflow"]

#       # Define flow nodes (intermediate nodes for flows)
#       node [style = "",shape=plaintext, fontsize=16, width=0.6, height=0.3]
#       "Inflow" [id="Inflow",label="Inflow", tooltip = "Flow: Inflow\nEquation: 1\nFrom: outside model boundary\nTo: Stock"]
#         "Outflow" [id="Outflow",label="Outflow", tooltip = "Flow: Outflow\nEquation: 5 * Stock\nFrom: Stock\nTo: outside model boundary"]

#       # Define external cloud nodes
#       node [shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
#       "Cloud1" [label="", tooltip = "Outside model boundary\nSink of: Outflow"]
#         "Cloud2" [label="", tooltip = "Outside model boundary\nSource of: Inflow"]

#       # Define auxiliary nodes
      
      

#       # Define constant nodes
      
      

#       # Define flow edges (stock -> flow_node)
#       edge [style = "", arrowhead="none", color="black:%(inflow_col)s:black", penwidth=1.1, minlen=1, splines=false, tailport="e", headport="w"]
#       "Cloud2" -> "Inflow"
#       edge [color="black:%(outflow_col)s:black"]
#       "Stock" -> "Outflow"

#       # Define flow edges (flow_node -> stock)
#       edge [style = "", arrowhead="normal", color="black:%(inflow_col)s:black", arrowsize=1.5, penwidth=1.1, minlen=1, splines=ortho, tailport="e", headport="w"]
#       "Inflow" -> "Stock"
#       edge [color="black:%(outflow_col)s:black"]
#       "Outflow" -> "Cloud1"

#       # Define dependency edges
      
      

      


#       # Rank groupings
      

#     }'
# }

# viz_str <- sprintf_arg(
#   viz_str,
#   list(
#     inflow_col = inflow_col,
#     outflow_col = outflow_col
#   )
# )

# pl <- DiagrammeR::grViz(viz_str)


if (recreate_figs) {
  export_plot(pl, file.path(filepath_figs, "build_sfm_inflow_outflow.pdf"))
}

print(pl)
```



```{r}

# # --- Adjustable settings -----------------------------------------------------
# t0 <- 0   # start (inflow == outflow, stock flat)
# t1 <- 2   # equal -> rise: inflow > outflow
# t2 <- 4   
# t3 <- 6   # rise  -> fall: outflow > inflow
# t4 <- 8   


# sfm <- stockflow() |>
# stock(Stock, eqn = 1) |>
# custom_func(func1, eqn = pulse(times, start = !!t1, width = !!t2 - !!t1, height = .5)) |>
# custom_func(func2, eqn = pulse(times, start = !!t3, width = !!t4 - !!t3, height = .5)) |>
# # custom_func(func3, eqn = pulse(times, start = 4, 
# # # Make width large just in case
# # width = 4, height = .5)) |>
# flow(Inflow, eqn = 1 + func1(t), to = Stock) |>
# flow(Outflow, eqn = 1 + func2(t), from = Stock) |>
# sim_settings(only_stocks=FALSE)

# sim <- simulate(sfm, stop = 10)
# df_stock <- as.data.frame(sim, vars = "Stock")
# df_in    <- as.data.frame(sim, vars = "Inflow")
# df_out   <- as.data.frame(sim, vars = "Outflow")



# increase_col   <- inflow_col # 'rgba(46, 134, 171, 1)'   # inflow > outflow: arrow + shade
# rise_fill  <- inflow_fill # 'rgba(46, 134, 171, 0.3)'
# decrease_col   <- outflow_col # 'rgba(231, 76, 60, 1)'    # outflow > inflow: arrow + shade
# fall_fill  <- outflow_fill # 'rgba(231, 76, 60, 0.3)'
# size_marker <- 8
# line_width  <- 5

# # Stock value at each time point (robust to non-exact time matches)
# stock_at <- function(tt) df_stock$value[which.min(abs(df_stock$time - tt))]
# y0 <- stock_at(t0); y1 <- stock_at(t1); y2 <- stock_at(t2); y3 <- stock_at(t3); y4 <- stock_at(t4)

# # Closed polygon spanning the gap between inflow and outflow over [a, b]
# band <- function(a, b){
#   ti <- df_in$time  >= a & df_in$time  <= b
#   to <- df_out$time >= a & df_out$time <= b
#   list(x = c(df_in$time[ti], rev(df_out$time[to])),
#        y = c(df_in$value[ti], rev(df_out$value[to])))
# }
# band_rise <- band(t1, t2)   # inflow > outflow
# band_fall <- band(t3, t4)   # outflow > inflow

# # --- Top panel: stock with a change indicator per phase ----------------------
# fig1 <- plot_ly(df_stock, x = ~time, y = ~value, type = 'scatter',
# showlegend = FALSE,
#                 mode = 'lines', name = 'Stock', line = list(color = stock_col, width = line_width)) 
                
                
# # markers at the phase boundaries
# fig1 <- fig1 |>
#   add_trace(x = c(t1), y = c(y1), mode = 'markers', name = 't1', showlegend = FALSE,
#             marker = list(color = increase_col, size = size_marker)) |>
#   add_trace(x = c(t2), y = c(y2), mode = 'markers', name = 't2', showlegend = FALSE,
#             marker = list(color = increase_col, size = size_marker)) |>
#   add_trace(x = c(t3), y = c(y3), mode = 'markers', name = 't3', showlegend = FALSE,
#             marker = list(color = decrease_col, size = size_marker)) |>
#   add_trace(x = c(t4), y = c(y4), mode = 'markers', name = 't4', showlegend = FALSE,
#             marker = list(color = decrease_col, size = size_marker))
            

# # # Horizontal reference lines at the end-of-phase levels (rise and fall phases)            
# # fig1 <- fig1 |>
# #   # horizontal reference line at the end-of-phase level (rise phase)
# #   add_trace(x = c(t1, t2), y = c(y1, y1), mode = 'lines', name = 'rise',
# #             line = list(width = line_width,
# #               # dash = 'dot', 
# #             color = increase_col), showlegend = FALSE) |>
# #   # horizontal reference line at the end-of-phase level (fall phase)
# #   add_trace(x = c(t3, t4), y = c(y3, y3), mode = 'lines', name = 'fall',
# #             line = list(width = line_width,
# #               # dash = 'dot', 
# #             color = decrease_col), showlegend = FALSE)




# # --- Bottom panel: inflow & outflow with the difference shaded ----------------
# fig2 <- plot_ly() |>
#   add_trace(x = band_rise$x, y = band_rise$y, type = 'scatter', mode = 'lines',
#             fill = 'toself', fillcolor = rise_fill, line = list(width = 0),
#             showlegend = FALSE,
#             name = 'Inflow > Outflow', hoverinfo = 'skip') |>
#   add_trace(x = band_fall$x, y = band_fall$y, type = 'scatter', mode = 'lines',
#             fill = 'toself', fillcolor = fall_fill, line = list(width = 0),
#             showlegend = FALSE,
#             name = 'Outflow > Inflow', hoverinfo = 'skip') |>
#   add_trace(data = df_in, x = ~time, y = ~value, type = 'scatter', mode = 'lines',
#             name = 'Inflow', line = list(color = increase_col, width = line_width)) |>
#   add_trace(data = df_out, x = ~time, y = ~value, type = 'scatter', mode = 'lines',
#             name = 'Outflow', line = list(color = decrease_col, width = line_width, dash = 'dash'))


# # Vertical reference lines at the end-of-phase levels (rise and fall phases)            
# add_vert_lines <- function(fig, y_range = c(0, 2.5)) {
#   fig |> 
#     add_trace(x = c(t1, t1), y = y_range, mode = 'lines', name = 't1',
#              line = list(width = 2, dash = 'dot', color = increase_col), showlegend = FALSE) |>
#     add_trace(x = c(t2, t2), y = y_range, mode = 'lines', name = 't2',
#               line = list(width = 2, dash = 'dot', color = increase_col), showlegend = FALSE) |>
#     add_trace(x = c(t3, t3), y = y_range, mode = 'lines', name = 't3',
#               line = list(width = 2, dash = 'dot', color = decrease_col), showlegend = FALSE) |>
#     add_trace(x = c(t4, t4), y = y_range, mode = 'lines', name = 't4',
#               line = list(width = 2, dash = 'dot', color = decrease_col), showlegend = FALSE)
# }

# fig1 <- add_vert_lines(fig1)
# fig2 <- add_vert_lines(fig2)

# # --- Combine panels ----------------------------------------------------------
# sp_margin <- 0.12                       # vertical gap between the two panels (paper)
# top_margin <- 100
# pad_axis <- .2
# fig <- plotly::subplot(fig1, fig2, nrows = 2, shareX = TRUE, titleY = TRUE,
#                        margin = sp_margin)

# # --- Phase labels above each panel -------------------------------------------
# annot_ <- utils::modifyList(annot, list(
#   font = list(size = font_size_subplot - 2, family = font_family)
# ))
# top_y    <- 1.0                         # top labels sit just above the top panel
# bot_y    <- (0.5 - sp_margin) + 0.01    # bottom labels sit just above bottom panel
# # bot_y    <- (0.5 - sp_margin) - 0.1

# # phase mid-points on the (shared) data x-axis, used to centre each label
# centers  <- c((t0 + t1)/2, (t1 + t2)/2, (t2 + t3)/2, (t3 + t4)/2, (t4 + 10)/2)
# top_text <- paste0("<i>", c("Static", "Increases", "Static", "Decreases", "Static"), "</i>")
# bot_text <- paste0("<i>", c("Inflow ≡ Outflow", "Inflow > Outflow", "Inflow ≡ Outflow",
#               "Inflow < Outflow", "Inflow == Outflow"), "</i>")

# phase_label <- function(text, x, y) list(
#   x = x, y = y, text = text, xref = 'x', yref = 'paper',
#   xanchor = 'center', yanchor = 'bottom', showarrow = FALSE, font = annot_
# )

# # arrows live in the combined layout: xref 'x' = shared axis, yref 'y' = top panel
# arrow_anns <- list(
#   # vertical arrow: change while inflow > outflow (stock rises)
#   list(x = t2, y = y2, ax = t2, ay = y1,
#        xref = 'x', yref = 'y', axref = 'x', ayref = 'y',
#        showarrow = TRUE, arrowhead = 3, arrowsize = 1.5,
#        arrowwidth = 2, arrowcolor = increase_col, text = ''),
#   # vertical arrow: change while outflow > inflow (stock falls)
#   list(x = t4, y = y4, ax = t4, ay = y3,
#        xref = 'x', yref = 'y', axref = 'x', ayref = 'y',
#        showarrow = TRUE, arrowhead = 3, arrowsize = 1.5,
#        arrowwidth = 2, arrowcolor = decrease_col, text = '')
# )

# label_anns <- c(
#   Map(phase_label, top_text, centers, MoreArgs = list(y = bot_y + .22)),
#   Map(phase_label, bot_text, centers, MoreArgs = list(y = bot_y - .07))
# )

# fig <- fig |> layout(
#   # unname(): Map() returns a named list, which plotly mis-parses
#   annotations = unname(c(
#     # arrow_anns, 
#   label_anns)),
#   margin = list(t = top_margin)                 # headroom for the top labels
# ) |>
#   plotly::layout(
#     xaxis = list(title = ""),
#     yaxis = list(title = "Units", range = c(1 - pad_axis, 2 + pad_axis),
#      tickvals = c(1, 2), ticktext = c("1", "2")),
#     yaxis2 = list(title = "Units / Time", range = c(1 - pad_axis, 1.5 + pad_axis), 
#                   tickvals = c(1, 1.5), ticktext = c("1", "1.5")),
#     font = annot_font,
#     annotations = list(
#       utils::modifyList(annot_xlab, list(text = "<i>Timeseries of Stock</i>", x = 0.5, y = top_y + .15)),
#       utils::modifyList(annot_xlab, list(text = "<i>Timeseries of Flows</i>", x = 0.5, y = bot_y + .15)),
#       phase_label("<i>Area = Increase\nin stock</i>", x = centers[2], y = bot_y - .25),
#       phase_label("<i>Area = Decrease\nin stock</i>", x = centers[4], y = bot_y - .25),
#       utils::modifyList(annot_xlab, list(text = "Time", x = 0.5, y = -0.05))
#     ),
#     legend = utils::modifyList(annot_legend, list(
#       itemwidth = 50,
#       traceorder = "normal", 
#       # y = -0.15
#       y = bot_y+.09
#       ))
#   )



# # # --- Vertical lines spanning BOTH panels -------------------------------------
# # # x positions and colors for the vertical lines
# # x_lines  <- c(t1, t2, t3, t4)
# # col_lines <- c(increase_col, increase_col, decrease_col, decrease_col)

# # # build a list of shape objects: xref 'x' = shared data axis,
# # # yref 'paper' with y0=0/y1=1 spans the whole figure height (across the gap)
# # shapes_list <- lapply(seq_along(x_lines), function(i) {
# #   list(
# #     type = "line",
# #     x0 = x_lines[i], x1 = x_lines[i],
# #     y0 = 0, y1 = 1,       # span entire figure vertically
# #     xref = "x",
# #     yref = "paper",
# #     line = list(
# #       color = col_lines[i],
# #       width = 2,
# #       dash  = "dot"
# #     )
# #   )
# # })

# # # NOTE: `fig |> layout(shapes = shapes_list)` silently DROPS a list of shapes
# # # when `fig` is a subplot, and attaching them before subplot() rescales
# # # yref="paper" to a single panel. Assigning x$layout$shapes directly is the only
# # # route that keeps every shape AND spans the full figure height.
# # fig$x$layout$shapes <- shapes_list


# if (recreate_figs) {
#   export_plot(fig, file.path(filepath_figs, "build_stock_flow_phases2.pdf"),
#     width = 6.2, height = 4.75
#   )
# }


# fig
```




```{r}

# Time points for the four phases of stock change (rise and fall)
t0 <- 0   # start (inflow == outflow, stock flat)
t1 <- 2   # equal -> rise: inflow > outflow
t2 <- 4   
t3 <- 6   # rise  -> fall: outflow > inflow
t4 <- 8   


sfm <- stockflow() |>
stock(Stock, eqn = 1) |>
custom_func(func1, eqn = pulse(times, start = !!t1, width = !!t2 - !!t1, height = .5)) |>
custom_func(func2, eqn = pulse(times, start = !!t3, width = !!t4 - !!t3, height = .5)) |>
# custom_func(func3, eqn = pulse(times, start = 4, 
# # Make width large just in case
# width = 4, height = .5)) |>
flow(Inflow, eqn = 1 + func1(t), to = Stock) |>
flow(Outflow, eqn = 1 + func2(t), from = Stock) |>
sim_settings(only_stocks=FALSE)

sim <- simulate(sfm, stop = 10)
df_stock <- as.data.frame(sim, vars = "Stock")
df_in    <- as.data.frame(sim, vars = "Inflow")
df_out   <- as.data.frame(sim, vars = "Outflow")
df_change <- df_in
df_change$value <- df_change$value - df_out$value


size_marker <- 8
line_width_thick  <- 5

# Stock value at each time point (robust to non-exact time matches)
stock_at <- function(tt) df_stock$value[which.min(abs(df_stock$time - tt))]
y0 <- stock_at(t0); y1 <- stock_at(t1); y2 <- stock_at(t2); y3 <- stock_at(t3); y4 <- stock_at(t4)

# --- Top panel: stock with a change indicator per phase ----------------------
fig1 <- plot_ly(df_stock, x = ~time, y = ~value, type = 'scatter',
showlegend = FALSE,
                mode = 'lines', name = 'Stock', line = list(color = stock_col, width = line_width_thick)) 
                
                
# markers at the phase boundaries
fig1 <- fig1 |>
  add_trace(x = c(t1), y = c(y1), mode = 'markers', name = 't1', showlegend = FALSE,
            marker = list(color = increase_col, size = size_marker)) |>
  add_trace(x = c(t2), y = c(y2), mode = 'markers', name = 't2', showlegend = FALSE,
            marker = list(color = increase_col, size = size_marker)) |>
  add_trace(x = c(t3), y = c(y3), mode = 'markers', name = 't3', showlegend = FALSE,
            marker = list(color = decrease_col, size = size_marker)) |>
  add_trace(x = c(t4), y = c(y4), mode = 'markers', name = 't4', showlegend = FALSE,
            marker = list(color = decrease_col, size = size_marker))


# --- Bottom panel: inflow & outflow with the difference shaded ----------------
fig2 <- plot_ly() 

# Vertical reference lines at the end-of-phase levels (rise and fall phases)            
add_vert_lines <- function(fig, y_range = c(-2.5, 2.5)) {
  fig |> 
    add_trace(x = c(t1, t1), y = y_range, mode = 'lines', name = 't1',
             line = list(width = 2, dash = 'dot', color = increase_col), showlegend = FALSE) |>
    add_trace(x = c(t2, t2), y = y_range, mode = 'lines', name = 't2',
              line = list(width = 2, dash = 'dot', color = increase_col), showlegend = FALSE) |>
    add_trace(x = c(t3, t3), y = y_range, mode = 'lines', name = 't3',
              line = list(width = 2, dash = 'dot', color = decrease_col), showlegend = FALSE) |>
    add_trace(x = c(t4, t4), y = y_range, mode = 'lines', name = 't4',
              line = list(width = 2, dash = 'dot', color = decrease_col), showlegend = FALSE)
}

fig1 <- add_vert_lines(fig1)
fig2 <- add_vert_lines(fig2)

# Net flow
fig2 <- fig2 |>
  add_trace(data = df_change, x = ~time, y = ~value, type = 'scatter', mode = 'lines',
  showlegend = FALSE, fill = 'tozeroy',
  fillcolor = ifelse(df_change$value >= 0, increase_fill, decrease_fill),
            name = 'Net Flow', line = list(width = 0)) |>

  add_trace(data = df_change, x = ~time, y = ~value, type = 'scatter', mode = 'lines',
  showlegend = FALSE, 
            name = 'Net Flow', line = list(color = flow_col, width = line_width_thick)) 


# --- Combine panels ----------------------------------------------------------
sp_margin <- 0.12                       # vertical gap between the two panels (paper)
top_margin <- 80 #40
pad_axis <- .1
pad_axis2 <- .25
fig <- plotly::subplot(fig1, fig2, nrows = 2, shareX = FALSE, titleY = TRUE,
heights = c(.5, .5),
                       margin = sp_margin)

# --- Phase labels above each panel -------------------------------------------
annot_ <- utils::modifyList(annot, list(
  font = list(size = font_size_subplot - 2, family = font_family)
))
top_y    <- 1.0                         # top labels sit just above the top panel
bot_y    <- (0.5 - sp_margin) + 0.01    # bottom labels sit just above bottom panel
top_text_y <- c(bot_y + c(.265, .35, .58, .35, .265))
bot_text_y <- c(bot_y - c(.195, .065, .195, .39, .195))
top_angles <- c(0, -43, 0, 43, 0)
bot_angles <- c(0, 0, 0, 0, 0)
# bot_y    <- (0.5 - sp_margin) - 0.1

# phase mid-points on the (shared) data x-axis, used to centre each label
centers  <- c((t0 + t1)/2, (t1 + t2)/2, (t2 + t3)/2, (t3 + t4)/2, (t4 + 10)/2)
centers_top <- centers
centers_top[c(2, 4)] <- centers_top[c(2, 4)] + c(-.17, .17)   # shift the bottom labels for the rise/fall phases
top_text <- paste0("<i>", c("Static", "Increases", "Static", "Decreases", "Static"), "</i>")
bot_text <- paste0("<i>", c("Inflow ≡ Outflow", "Inflow > Outflow", "Inflow ≡ Outflow",
              "Inflow < Outflow", "Inflow ≡ Outflow"), "</i>")

phase_label <- function(text, x, y, angle = 0) list(
  x = x, y = y, text = text, xref = 'x', yref = 'paper',
  textangle = angle,
  xanchor = 'center', yanchor = 'bottom', showarrow = FALSE, font = annot_
)


label_anns <- c(
  lapply(seq_along(top_text), function(i) phase_label(top_text[i], centers_top[i], y = top_text_y[i], angle = top_angles[i])),
  lapply(seq_along(bot_text), function(i) phase_label(bot_text[i], centers[i], y = bot_text_y[i], angle = bot_angles[i]))
)

fig <- fig |> layout(
  # unname(): Map() returns a named list, which plotly mis-parses
  annotations = unname(c(
  label_anns)),
  margin = list(t = top_margin)                 # headroom for the top labels
) |>
  plotly::layout(
    xaxis = list(title = "", showline = TRUE, zerolinewidth = 0, showgrid = FALSE),
    xaxis2 = list(title = "", showline = TRUE, showgrid = FALSE,
                              zerolinecolor = '#ffff',
 zerolinewidth = 0),
    yaxis = list(title = "Units", showgrid = FALSE,
    # showline = TRUE,
     range = c(1 - pad_axis, 2 + pad_axis),
     tickvals = c(1, 2), ticktext = c("1", "2")),
    yaxis2 = list(title = "Units / Time", 
    zerolinewidth = 0, showgrid = FALSE,
    showline = TRUE, 
    range = c(-.5 - pad_axis2, .5 + pad_axis2), 
                  tickvals = c(-.5, .5), ticktext = c("-0.5", "0.5")),
    font = annot_font,
    annotations = list(
      utils::modifyList(annot_xlab, list(text = "<i>A. Stock-and-Flow Diagram</i>", x = 0.5, y = top_y + .22)),
      utils::modifyList(annot_xlab, list(text = "<i>B. Timeseries of Stock</i>", x = 0.5, y = top_y + .13)),
      utils::modifyList(annot_xlab, list(text = "<i>C. Timeseries of Change in Stock</i>", x = 0.5, 
      y = bot_y + .11)),
      phase_label("<i>Area = Increase\nin stock</i>", x = centers[2], 
      
      y = bot_y - .2
      ),
      phase_label("<i>Area = Decrease\nin stock</i>", x = centers[4], y = bot_y - .325),
      utils::modifyList(annot_xlab, list(text = "Time", x = 0.5, y = -0.06))
    )
  )


if (recreate_figs) {
  export_plot(fig, file.path(filepath_figs, "build_stock_flow_phases.pdf"),
    width = 6, height = 4.2
  )
}


fig
```


Stock-and-flow models provide an intuitive way to formalize psychological theories as many are fundamentally concerned with change over time. Despite the physical connotation of the term, stocks need not be tangible: emotions, knowledge, beliefs, perceptions, stress, motivation, and trust are all examples of psychological constructs that accumulate over time, often in response to experience or behaviour. The processes that drive these changes -- such as emotion regulation, coping, and learning -- are the flows, specifying what causes psychological states to increase or decrease.

To demonstrate stock-and-flow modelling in a psychological context, we will create a simplified model of burnout. 
The model represents a stylized pattern of how burnout develops over time on a within-person level.
It contains a single stock representing the current level of energy, an inflow for recovery, and an outflow for depletion from work. We first initialize an empty stock-and-flow model:

```{r , fig.width=4, fig.height=2}
sfm <- stockflow()
print(sfm)
```

Though the model contains no elements, it includes default simulation settings, such as the duration of the simulation and a solver (`euler`) specifying the numerical technique used to generate output from the model. We change the simulation settings to model energy over the course of 16 weeks (note that the time unit merely changes the labels on the axes of the resulting plots, and does not affect the model's behaviour). In addition, we set `only_stocks = FALSE` to return all model variables in the simulation output, not just the stocks:

```{r}
sfm <- sim_settings(sfm, stop = 16, time_units = "weeks",
 only_stocks = FALSE
 )
```

A model name can be supplied with `meta()`.

```{r}
sfm <- meta(sfm, name = "Burnout")
```

Next, we introduce a stock to the model to represent energy. Each model element requires a unique `name`, which serves as its identifier in equations and follows the same naming rules as R variables (e.g., no spaces or special characters). We here simply choose `energy`. An optional `label` can be supplied for use in plots and diagrams (e.g., `label = "Energy Level"`); when omitted, the name is used. 

Every stock also needs an *initial condition*: the value of the stock at the start of the simulation. This is set via the `eqn` argument, where here, we initialize energy at .3:

```{r}
sfm <- stock(sfm, name = energy, eqn = .3, label = "Energy")
```



```{r include=FALSE}
if (recreate_figs) {
  sfm_list <- list()
  sfm_list[["A"]] <- sfm
}
```


Plotting the stock-and-flow model yields its stock-and-flow diagram, which now consists of only one stock:

```{r}
plot(sfm)
```

To assess its dynamics, we simulate the model over time and visualise the resulting timeseries:

```{r}
sfm |>
  simulate() |>
  plot()
```

Above, we use the pipe operator `|>` for better legibility. It simply passes the result of an expression to the next expression as its first argument. Across the entirety of the simulation, energy remains at its initial state. Stocks without flows are static, as there is no process specifying how they change. To deplete energy, we introduce an outflow representing work. For simplicity, we specify that work occurs at a constant rate over time, say `.5`. Rather than defining the flow's `eqn` to be `.5` directly, we add a constant to the model, so that it can easily be changed later. This also helps to keep track of how parametrized the model is. 

```{r}
sfm <- constant(sfm, work_rate, eqn = .5, label = "Work Rate")
```

As shown above, `eqn` is a generic argument used for all variable types. `work_rate` can now be used as a variable in the equation for the outflow from energy:

```{r}
sfm <- flow(sfm, work,
  eqn = work_rate, from = energy,
  label = "Depletion from work"
)
```


```{r include=FALSE}
if (recreate_figs) {
  sfm_list[["B"]] <- sfm
}
```


Note that `eqn` accepts any valid R expression, including functions (e.g., `sqrt()`, `min()`) and arithmetic operators (e.g., `*`, `+`). Expressions must evaluate to a scalar, whereas vectors and matrices are not supported. Equations can reference only other variables defined in the model, as shown above; they cannot access objects from the user's R environment.

Flows require more than an `eqn`, and also need to be connected to a stock, at least as either an inflow (`to`) or an outflow (`from`). Note that by definition, outflows are subtracted from the stock, and as such do not need a minus sign in `eqn` to indicate that they decrease the stock. We simulate the model to check whether energy indeed depletes from work:

```{r}
sfm |>
  simulate() |>
  plot()
```


As shown in the plot, energy decreases with a constant rate at each time step, creating linear decay over time. As the outflow is specified as `.5`, energy decreases by `.5` per unit of time (1 week), producing a negative energy state. To rectify this implausible behaviour, a naive solution may be to include a logical statement such as `ifelse(energy < 0, 0, energy)`. However, this computational trick would mask model misspecification. Ideally, stocks should remain within bounds as a result of the model's equations and parameters. For instance, we can prevent negative energy by making `work` proportional to the amount of available energy: `work_rate * energy`. 

To assess whether this produces more plausible model behaviour, we modify the outflow using `update()`:


```{r}
sfm <- update(sfm, work, eqn = work_rate * energy)

sfm |>
  simulate() |>
  plot()
```


```{r include=FALSE}
if (recreate_figs) {
  sfm_list[["C"]] <- sfm
}
```


Energy now follows an exponential decay pattern, where work now depletes energy until it is zero, but not beyond this point. By letting the flow rate depend on the level of the stock itself, we have introduced a *feedback loop* to the system. Positive feedback loops amplify change, whereas negative feedback loops bring the system back to a target state. Here, `work` consists of a negative feedback loop that pulls energy to zero: the higher energy is, the more its outflow decreases it.

To allow energy to recover, we introduce an inflow, again specified as a simple constant rate:

```{r}
sfm <- constant(sfm, recovery_rate, eqn = .3, label = "Recovery Rate") |>
  flow(recovery, eqn = recovery_rate, to = energy, label = "Recovery")

sfm |>
  simulate() |>
  plot()
```



```{r include=FALSE}
if (recreate_figs) {
  sfm_list[["D"]] <- sfm
}
```

Flows can differ in their rates, meaning some processes are slower or faster than others. Here, we have specified recovery to evolve more slowly than work. As a result of the new inflow, energy now stabilizes at a fixed level, as energy recovery and depletion balance out. But what if the ability to recover is not static, but erodes over time? Put differently, what if the recovery rate is not a constant, but a stock? To implement this idea, we change the type of `recovery_rate`, and update its initial value:

```{r}
sfm <- change_type(sfm, recovery_rate, new_type = stock) |>
  update(recovery_rate, eqn = 1)
```

We specify with a new outflow that the recovery rate erodes over time as a function of the available recovery rate and the amount worked.

```{r}
sfm <- flow(sfm, erosion,
  eqn = recovery_rate * work,
  from = recovery_rate, label = "Recovery Erosion"
)
sfm |>
  simulate() |>
  plot()
```


The plot shows how the erosion of the ability to recover produces a characteristic burnout pattern: a steep initial rise followed by a collapse of energy. Counter-intuitively, though the recovery rate only decreases over time, energy first rises. However, this becomes clear once we inspect the balance between the inflow and outflow. To track this balance, we introduce an auxiliary: an intermediate variable computed at each time step that does not itself feed into any stock, but can be used in flow equations or to monitor other dynamic quantities: 

```{r}
# Add auxiliary to keep track of net flow to energy (inflow - outflow)
sfm <- aux(sfm, net_flow, eqn = recovery - work, label = "Net flow to energy")
sfm |>
  simulate() |>
  plot()
```



```{r include=FALSE}
if (recreate_figs) {
  sfm_list[["E"]] <- sfm
}
```


```{r include=FALSE, eval = has_internet(), error = TRUE}
if (recreate_figs) {
  # Create subplot figure
  pl_list <- list()
  max_letter <- max(names(sfm_list))

  VARS_example <- c(
    "energy", 
    "work", "recovery",
    "net_flow"
  )

  # Recovery = recovery_rate so plot as one line
  sfm_list[["E"]] <- sfm_list[["E"]] |>
    update(recovery, label = "Recovery / recovery rate")

  COLORS_example <- COLORS[seq_along(VARS_example)]
  line_widths <- rep(line_width, length(VARS_example))
  line_widths[VARS_example == "energy"] <- line_width_stock

  titles <- paste0(
    paste0(
      "<i>", LETTERS[seq_along(sfm_list)], ". ",
      c(
        "Static Stock",
        "Constant Outflow",
        "Stock-Dependent<br>Outflow",
        "Constant Inflow",
        "Erosion of Recovery Rate"
        # "Auxiliary variable tracking net flow to energy"
      ), "</i>"
    )
  )
  titles <- stats::setNames(titles, names(sfm_list))


  annot_ <- utils::modifyList(annot, list(
    font = list(size = font_size_subplot + 4, family = font_family)
  ))


  for (letter in names(sfm_list)) {
    sim <- simulate(sfm_list[[letter]],
      save_at = save_at_fine, only_stocks = FALSE
    )

    vars_in_sim <- unique(sim$df$variable)

    vars_ <- VARS_example[VARS_example %in% vars_in_sim] 
    colors_ <- COLORS_example[VARS_example %in% vars_in_sim]
    colors_ <- adjustcolor(colors_, alpha.f = .9)
    line_widths_ <- line_widths[VARS_example %in% vars_in_sim]

    pl_list[[letter]] <- plot(sim,
      vars = vars_,
      colors = colors_,
      line_width = line_widths_,
      main = "",
      showlegend = (letter == max_letter)
    ) |>
      rm_grid() |>
      plotly::layout(yaxis = list(
        nticks = 3
      ), xaxis = list(
        tickvals = list(0, 7.5, 15)
      ))
  }

}
```



```{r, include=FALSE, eval = has_internet(), error = TRUE}
if (recreate_figs) {
  
  x1 <- 0.1
  x2 <- 0.37
  x3 <- .78
  y1 <- 1.06
  y2 <- 0.48

  pl <- plotly::subplot(
    plotly::subplot(pl_list[1:4],
      nrows = 2,
      shareX = TRUE,
      shareY = FALSE,
      titleY = FALSE,
      titleX = FALSE,
      margin = 0.075
    ), plotly::subplot(pl_list[5]),
    nrows = 1,
    margin = 0.025
  ) |>
    plotly::layout(
      yaxis4 = list(range = c(0, 1)),
      yaxis2 = list(range = c(0, .4)),
      font = annot_font,
      # legend = annot_legend,
      # legend = utils::modifyList(annot_legend, list(x = 1.4, y = 0.8)),
      legend = utils::modifyList(annot_legend, list(
        y = -.16,
                traceorder = "normal",
        font = list(size = font_size_legend)
      )),
      annotations = list(
        utils::modifyList(annot_xlab, list(x = 0.5, y = -0.05))
      )
    ) |>
    plotly::layout(annotations = list(
      utils::modifyList(annot_, list(x = x1, y = y1, text = titles[["A"]])),
      utils::modifyList(annot_, list(x = x2, y = y1, text = titles[["B"]])),
      utils::modifyList(annot_, list(x = x1, y = y2 - .005, text = titles[["C"]])),
      utils::modifyList(annot_, list(x = x2, y = y2 + .005, text = titles[["D"]])),
      utils::modifyList(annot_, list(x = x3, y = y1, text = titles[["E"]]))
      # utils::modifyList(annot_, list(x = .5, y = -.5, text = "F. Stock-and-Flow Diagram"))
    ))

  export_plot(pl, file.path(filepath_figs, "build_sim.pdf"),
    width = 7, height = 4
  )
}
```


```{r include=FALSE, eval = has_internet(), error = TRUE}
if (recreate_figs) {
  # Version 2: all plots in one column

  y1 <- 1.01
  y2 <- 0.81
  y3 <- 0.63
  y4 <- 0.45
  y5 <- 0.275

  heights <- c(rep(.2, 4), .3)
  heights <- heights / sum(heights)

  titles <- paste0(
    paste0(
      "<i>", LETTERS[seq_along(sfm_list)], ". ",
      c(
        "Static Stock",
        "Constant Outflow",
        "Stock-Dependent Outflow",
        "Constant Inflow",
        "Erosion of Recovery Rate"
        # "Auxiliary variable tracking net flow to energy"
      ), "</i>"
    )
  )
  titles <- stats::setNames(titles, names(sfm_list))

  pl <- plotly::subplot(pl_list,
    nrows = length(pl_list),
    shareX = TRUE, shareY = FALSE,
    titleY = FALSE, titleX = FALSE,
    margin = 0.0275,
     heights = heights
  ) |>
    plotly::layout(
    yaxis = list(range = c(0.1, 0.5)),
    # yaxis2 = list(showline = TRUE),
    yaxis3 = list(range = c(0, 0.35)),
    # yaxis4 = list(showline = TRUE),
    # yaxis5 = list(showline = TRUE),

      font = annot_font,
      legend = utils::modifyList(annot_legend, list(
        y = -.09,
        traceorder = "normal",
        font = list(size = font_size_legend)
      )),
      annotations = list(
        utils::modifyList(annot_xlab, list(x = 0.5, y = -0.04))
      )
    ) |>
    plotly::layout(annotations = list(
      utils::modifyList(annot_, list(x = 0.5, y = y1, text = titles[["A"]])),
      utils::modifyList(annot_, list(x = 0.5, y = y2, text = titles[["B"]])),
      utils::modifyList(annot_, list(x = 0.5, y = y3, text = titles[["C"]])),
      utils::modifyList(annot_, list(x = 0.5, y = y4, text = titles[["D"]])),
      utils::modifyList(annot_, list(x = 0.5, y = y5, text = titles[["E"]]))
      # utils::modifyList(annot_, list(x = .5, y = -.5, text = "F. Stock-and-Flow Diagram"))
    ))

  export_plot(pl, file.path(filepath_figs, "build_sim_column.pdf"),
    width = 2.5, height = 8
  )
}
```


```{r include = FALSE}
if (recreate_figs) {

  # {
  #   viz_str0 <- "

  #   digraph sfm {

  #     graph [layout = dot, rankdir = LR, center=true, outputorder='edgesfirst', pad=.1, nodesep= .15]

  #     # Rank groupings
  #     {rank=same; 'net_flow1'; 'net_flow'; 'invisnode1'; 'energy'; 'invisnode2'; 'recovery_rate'}

  #     # Invisible edge to fix order
  #     'net_flow1' -> 'net_flow' -> 'invisnode1' -> 'energy' -> 'invisnode2' -> 'recovery_rate' [style=invis]

  #     # Define stock nodes
  #     'recovery_rate' [id='recovery_rate',label='Recovery Rate',tooltip = 'eqn = 1',shape=box,style=filled,fillcolor='#09B0ACBF',fontsize=18,fontname='Times New Roman']

  #    # Define invisible node
  #     'invisnode1' [id='invisnode1',label='Invisible Node',shape=box,style=invis,fontsize=12,fontname='Times New Roman', width=.2, height=.1, fixedsize=true]
  #     'invisnode2' [id='invisnode2',label='Invisible Node',shape=box,style=invis,fontsize=12,fontname='Times New Roman', width=.2, height=.1, fixedsize=true]

  #     'energy' [id='energy',label='Energy',tooltip = 'eqn = 0.3',shape=box,style=filled,fillcolor='#09B0ACBF',fontsize=18,fontname='Times New Roman']

  #     # Define flow nodes (intermediate nodes for flows)
  #     'work' [id='work',label='Depletion from work', tooltip = 'eqn = work_rate * energy', shape = plaintext, fontsize=16, fontname='Times New Roman', width=0.6, height=0.3]
  #               'recovery' [id='recovery',label='Recovery', tooltip = 'eqn = recovery_rate', shape = plaintext, fontsize=16, fontname='Times New Roman', width=0.6, height=0.3]
  #               'erosion' [id='erosion',label='Erosion', tooltip = 'eqn = recovery_rate * work', shape = plaintext, fontsize=16, fontname='Times New Roman', width=0.6, height=0.3]

  #     # Define external cloud nodes
  #     'Cloud1' [label='', tooltip = 'Unspecified source',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
  #               'Cloud2' [label='', tooltip = 'Unspecified source',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
  #               'Cloud3' [label='', tooltip = 'Unspecified sink',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]

  #     # Define flow edges (stock -> flow_node -> stock)
  #     'energy' -> 'work' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1, splines=false]
  #     'work' -> 'Cloud1' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1, splines=ortho]
  #     'Cloud3' -> 'recovery' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1, splines=false]
  #     'recovery' -> 'energy' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1, splines=ortho]
  #     'recovery_rate' -> 'erosion' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1, splines=false]
  #     'erosion' -> 'Cloud2' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1, splines=ortho]

  #     # Define dependency edges
  #     'energy' -> 'work' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
  #     'recovery_rate' -> 'recovery' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
  #     'recovery_rate' -> 'erosion' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
  #     'work' -> 'erosion' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
  #     'recovery' -> 'net_flow' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
  #     'work' -> 'net_flow' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']

  #     # Define auxiliary nodes
  #     # 'net_flow' [id='net_flow',xlabel='',label='',tooltip = 'eqn = recovery - work',shape=circle,fontsize=16,fontname='Times New Roman', width=0.15, height=0.15, fixedsize=true, style=filled, fillcolor='grey90']
  #     'net_flow' [id='net_flow',xlabel='',label='',tooltip = 'eqn = recovery - work',shape=circle,fontsize=16,fontname='Times New Roman', width=0.15, height=0.15, fixedsize=true, style=filled, fillcolor='grey90']
  #     'net_flow1' [id='net_flow1',label='Net flow to energy',tooltip = 'eqn = recovery - work',shape=plaintext,fontsize=16,fontname='Times New Roman', width=0.6, height=0.2, fixedsize=true]
  #   }
  #   "

  # }

  # cat(plot(sfm, minlen = 1, nodesep = .15, pad = .1, order = c("net_flow", "energy", "recovery_rate"), align = c("net_flow", "energy", "recovery_rate"))$x$diagram)

  {
    viz_str <- '

    digraph sfm {

      graph [layout = dot, rankdir = LR, center=true, outputorder="edgesfirst", pad=0.1, nodesep=0.15]

      # Shared across all nodes (persists until overridden)
      node [fontsize=18,fontname="Times New Roman",fontcolor="black"]

      # Define stock nodes
      node [shape=box,style=filled,fillcolor="#83d3d4"]
      "energy" [id="energy",label=<Energy<BR/><FONT POINT-SIZE="13" COLOR="black">eqn = 0.3</FONT>>, tooltip = "Stock: Energy\nName: energy\nInitial value: 0.3\nInflows: Recovery\nOutflows: Depletion from work"]
        "recovery_rate" [id="recovery_rate",label=<Recovery Rate<BR/><FONT POINT-SIZE="13" COLOR="black">eqn = 1</FONT>>, tooltip = "Stock: Recovery Rate\nName: recovery_rate\nInitial value: 1\nInflows: —\nOutflows: Recovery Erosion"]

      # Define flow nodes (intermediate nodes for flows)
      node [style = "",shape=plaintext, fontsize=16, width=0.6, height=0.3]
      "erosion" [id="erosion",label=<Recovery Erosion<BR/><FONT POINT-SIZE="13" COLOR="black">eqn = recovery_rate * work</FONT>>, tooltip = "Flow: Recovery Erosion\nName: erosion\nEquation: recovery_rate * work\nFrom: Recovery Rate\nTo: outside model boundary"]
        "recovery" [id="recovery",label=<Recovery<BR/><FONT POINT-SIZE="13" COLOR="black">eqn = recovery_rate</FONT>>, tooltip = "Flow: Recovery\nName: recovery\nEquation: recovery_rate\nFrom: outside model boundary\nTo: Energy"]
        "work" [id="work",label=<Depletion from work<BR/><FONT POINT-SIZE="13" COLOR="black">eqn = work_rate * energy</FONT>>, tooltip = "Flow: Depletion from work\nName: work\nEquation: work_rate * energy\nFrom: Energy\nTo: outside model boundary"]

      # Define external cloud nodes
      node [shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
      "Cloud1" [label="", tooltip = "Outside model boundary\nSink of: Recovery Erosion"]
        "Cloud2" [label="", tooltip = "Outside model boundary\nSink of: Depletion from work"]
        "Cloud3" [label="", tooltip = "Outside model boundary\nSource of: Recovery"]

      # Define auxiliary nodes
      node [shape=circle,fontsize=16, width=0.15, height=0.15, fixedsize=true, style=filled, fillcolor="grey90"]

      "net_flow" [id="net_flow",xlabel="",label="", tooltip = "Auxiliary: Net flow to energy\nName: net_flow\nEquation: recovery - work"]

      "net_flow1" [id="net_flow1",label=<Net flow to energy<BR/><FONT POINT-SIZE="13" COLOR="black">eqn = recovery - work</FONT>>,tooltip = "eqn = recovery - work",shape=plaintext, width=0.6, height=0.2, fillcolor="none", style=""]

      # Define constant nodes



      # Define flow edges (stock -> flow_node)
      edge [style = "", arrowhead="none", color="black:#f48153:black", penwidth=1.1, minlen=1, splines=false, tailport="e", headport="w"]
      "recovery_rate" -> "erosion"
        "Cloud3" -> "recovery"
        "energy" -> "work"

      # Define flow edges (flow_node -> stock)
      edge [style = "", arrowhead="normal", color="black:#f48153:black", arrowsize=1.5, penwidth=1.1, minlen=1, splines=ortho, tailport="e", headport="w"]
      "erosion" -> "Cloud1"
        "recovery" -> "energy"
        "work" -> "Cloud2"

      # Define dependency edges
      edge [style = "", color="#999999", arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport = "_", headport="_"]
      "recovery_rate" -> "erosion"
        "work" -> "erosion"
        "recovery_rate" -> "recovery"
        "energy" -> "work"
        "recovery" -> "net_flow"
        "work" -> "net_flow"

      # Ordering hints (invisible edges)
      edge [style=invis, constraint=true]
      "net_flow" -> "energy"
      "energy" -> "recovery_rate"


      # Rank groupings

      {rank=same; "net_flow"; "energy"; "recovery_rate" }


### Additions
 # Rank groupings
      {rank=same; "net_flow1"; "net_flow"; "invisnode1"; "energy"; "invisnode2"; "recovery_rate"}


      # Invisible edge to fix order
      "net_flow1" -> "net_flow" -> "invisnode1" -> "energy" -> "invisnode2" -> "recovery_rate" [style=invis]

     # Define invisible node
      "invisnode1" [id="invisnode1",label="Invisible Node",shape=box,style=invis,fontsize=12,fontname="Times New Roman", width=.2, height=.1, fixedsize=true]
      "invisnode2" [id="invisnode2",label="Invisible Node",shape=box,style=invis,fontsize=12,fontname="Times New Roman", width=.2, height=.1, fixedsize=true]

    }
    '
  }

  pl <- DiagrammeR::grViz(viz_str)
  pl

  export_plot(pl, file.path(filepath_figs, "build_sfm.pdf"))
}
```

The net flow to energy is initially positive, as recovery exceeds energy depletion from work. As erosion progressively reduces the recovery rate, this inflow weakens, eventually falling below the outflow from work. The net flow to energy is negative, and energy begins to decline, leading to the observed collapse. This complex behaviour is produced by a simple stock-and-flow model consisting of two stocks and three flows:

```{r}
plot(sfm)
```

Throughout the modelling process, variables can be renamed with `change_name()` and removed with `discard()`. Moreover, we can easily check for common mistakes in the model specification, such as references to undefined variables or flows which are not connected to any stock:

```{r}
summary(sfm)
```


In summary, stock-and-flow models consist of one or more stocks, each requiring an inflow and/or outflow to change over time. Without limiting processes, stocks may continue to increase indefinitely; without restorative processes, stocks may deplete past the point of recovery. Although inflows and outflows connected to the same stock could be combined into a single net flow (as often done in differential equation models), stock-and-flow modelling generally keeps flows disaggregated. Separating inflows and outflows encourages more precise thinking about what processes increase and decrease stocks, and what distinct information and rates of change govern each flow. Disaggregation also supports more targeted interventions, for example by identifying whether to limit inflows or promote outflows. 

For convenience, stock-and-flow models may further be supplemented with constants and auxiliaries. Constants are static parameters that do not change over the time course of the simulation, whereas auxiliaries are dynamic, meaning they are computed anew at each step. To illustrate the difference, a constant with `eqn = runif(1)` will be fixed to a random number at the beginning of the simulation, whereas an auxiliary with the same `eqn` will draw a new number each time step. Throughout the tutorial, we use the term ``variable'' for any part of the system, be that a stock, flow, constant, or auxiliary. Though this usage may differ from other scientific fields, we here choose to adhere to system dynamics terminology.

## Variable types

```{r echo = FALSE}
variable_types_table <- data.frame(
  Characteristic = c(
    "Role in system",
    "Varies within time horizon",
    "A process taking place over time",
    "Can be captured at any given moment in time",
    "`eqn` denotes",
    "Allowed dependencies in `eqn`",
    "Examples"
  ),
  Stock = c(
    "Defines the state of the system; accumulates the effects of flow(s) over time",
    "✓",
    "✗",
    "✓",
    "Initial condition",
    "Constants and (initial values of) other stocks",
    "Emotions, beliefs, stress, trust, resources"
  ),
  Flow = c(
    "Increases or decreases a stock",
    "✓",
    "✓",
    "✗",
    "Flow rate computed at every time step",
    "Any other variable",
    "Coping, learning, emotion regulation"
  ),
  Constant = c(
    "Specifies static quantity",
    "✗",
    "✗",
    "✓",
    "Fixed value",
    "Other constants and (initial values of) stocks",
    "Rates, capacities, thresholds"
  ),
  Auxiliary = c(
    "Provides intermediate computations for convenience; keeps track of changing quantities",
    "✓",
    "Possibly",
    "Possibly",
    "Value computed at every time step",
    "Any other variable",
    "Performance indices, ratios, sums of stocks"
  )
)

kableExtra::kbl(variable_types_table,
  format = "markdown",
  col.names = c("Characteristic", "Stock", "Flow", "Constant", "Auxiliary"),
  caption = "Characteristics of Variable Types in Stock-and-Flow Models"
) |>
  kableExtra::kable_styling()
```

The following flowchart can be used to determine a variable's type:

```{r include = FALSE}
{
  viz_str <- "
digraph flowchart {

  graph [layout = dot,
         rankdir = TB,
         bgcolor = white,
         nodesep = 0.45,
         ranksep = 0.5,
         splines = ortho,
         fontname = '%(font_family)s',
         pad = '0.2']

  node [fontsize = 9.5,
        fontname = '%(font_family)s',
        fontcolor = 'white',
        penwidth = 0]

  edge [fontname = '%(font_family)s',
        fontsize = 8.5,
        fontcolor = '#555555',
        color = '#666666',
        penwidth = 1.3,
        arrowsize = 0.6]

  # --- DECISION NODES (rounded boxes with border) ---
  q1 [label = 'Can the variable be represented\\nas a continuous quantity that\\nbe lower or higher?',
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.15,0.1']

  q2 [label = 'Does the variable change within\\nthe chosen time horizon?',
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.15,0.1']

  q3 [label = 'Is the variable a behaviour, action,\\nor process that takes place over time?',
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.15,0.1']

  q4 [label = 'Does the variable accumulate\\nthe effects of flows?',
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.15,0.1']

  q5 [label = 'Does the variable\\'s effect\\naccumulate in a stock?', # Does it directly increase or decrease a stock?
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.15,0.1']

  q6 [label = 'Should it?',
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.12,0.08']

  # --- TERMINAL NODES (no border, darker fill) ---
  constant [label = 'Constant',
            shape = box, style = 'filled,rounded',
            fillcolor = '#0f5c5a', color = 'none',
            fontsize = 10, margin = '0.15,0.08']

  stock [label = 'Stock',
         shape = box, style = 'filled,rounded',
         fillcolor = '#0f5c5a', color = 'none',
         fontsize = 10, margin = '0.15,0.08']

  flow [label = 'Flow',
        shape = box, style = 'filled,rounded',
        fillcolor = '#0f5c5a', color = 'none',
        fontsize = 10, margin = '0.15,0.08']

  auxiliary [label = 'Auxiliary',
             shape = box, style = 'filled,rounded',
             fillcolor = '#0f5c5a', color = 'none',
             fontsize = 10, margin = '0.15,0.08']

  # --- ACTION NODES ---
  reformulate [label = 'Reformulate variable',
               shape = box, style = 'dashed,rounded,filled',
            fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
            fontsize = 8, penwidth = 1.2, color = '#17807c',
            margin = '0.1,0.06']

  add_flow [label = 'Add missing\\nflows',
            shape = box, style = 'dashed,rounded,filled',
            fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
            fontsize = 8, penwidth = 1.2, color = '#17807c',
            margin = '0.1,0.06']

  add_stock [label = 'Add missing\\nstock',
             shape = box, style = 'dashed,rounded,filled',
             fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
             fontsize = 8, penwidth = 1.2, color = '#17807c',
             margin = '0.1,0.06']

  # --- EDGES ---
  q1 -> reformulate [xlabel = 'No', minlen = 2]
  reformulate -> q1 [minlen = 2, style = dashed, penwidth = 1]
  q1 -> q2          [xlabel = 'Yes']

  q2 -> constant [xlabel = 'No', minlen = 2]
  q2 -> q3       [xlabel = 'Yes']

  q3 -> q4 [xlabel = <<font>No, it can be captured<br/>at a single moment in time</font>>]
  q3 -> q5 [xlabel = '  Yes']

  q4 -> stock [xlabel = 'Yes']
  q4 -> q6    [xlabel = 'No']

  q5 -> flow  [xlabel = 'Yes']
  q5 -> q6    [xlabel = 'No']

  q6 -> auxiliary [xlabel = 'No']
  q6 -> add_flow  [xlabel = 'Yes', style = dashed, color = '#2a9d8f', penwidth = 1]
  q6 -> add_stock [xlabel = 'Yes', style = dashed, color = '#2a9d8f', penwidth = 1]
  add_flow  -> stock [style = dashed, color = '#2a9d8f', penwidth = 1]
  add_stock -> flow  [style = dashed, color = '#2a9d8f', penwidth = 1]

  # --- RANK CONSTRAINTS ---
  { rank = same; q1; reformulate }
  { rank = same; q2; constant }
  { rank = same; q4; q5 }
  { rank = same; stock; q6; flow }
  { rank = same; add_flow; auxiliary; add_stock }
}
"
}

viz_str <- sprintf_arg(
  viz_str,
  list(
    font_family = font_family
  )
)

pl <- DiagrammeR::grViz(viz_str)
pl
```



## Overview of package functionality


```{r echo=FALSE}
functions_table <- data.frame(
  Function = c("`stockflow()`", "`stock()`", "`flow()`", "`constant()`", "`aux()`", "`lookup()`", "`update()`", "`simulate()`", "`plot()`", "`summary()`", "`as.data.frame()`", "`sim_settings()`", "`meta()`", "`export_model()`"),
  Purpose = c("Create empty model or load template", "Add or modify a stock", "Add or modify a flow", "Add or modify a constant", "Add or modify an auxiliary", "Add or modify a lookup function", "Add or modify any variable (generic)", "Simulate model", "Plot model diagram or simulation", "Run model diagnostics", "Get model properties in a dataframe", "Modify simulation specifications", "Modify model metadata", "Export models to other formats")
)

kableExtra::kbl(functions_table,
  format = "markdown",
  col.names = c("Function", "Purpose"),
  caption = "Main functions in *sdbuildR*"
) |>
  kableExtra::kable_styling()
```


In some cases, it may be useful to refer to global simulation variables in the model's equations:

```{r echo=FALSE}
functions_table <- data.frame(
  Variable = c(
    "`times`", "`t`", "`dt`"
  ),
  Description = c(
    "Vector with simulation times",
    "Current time in the ODE",
    "Time step of the simulation"
  ),
  Use.case = c(
    "Use in equations, e.g., `pulse(times, 5, width = dt)`",
    "Use in equations, e.g., `input(t)`",
    "Use in equations, e.g., `pulse(times, 5, width = dt)`"
  )
)

kableExtra::kbl(functions_table,
  format = "markdown",
  col.names = c(
    "Variable", "Description", "Use case"
  ),
  caption = "Global simulation variables in *sdbuildR*"
) |>
  kableExtra::kable_styling()
```


### Simulation specifications

We may want to observe the system over a longer time period, or with a different time step.

```{r}
sfm <- sfm |>
  sim_settings(
    start = 0,
    stop = 250,
    dt = 0.001
  )
```


Simulation settings can be set directly on the model object as above, or passed to `simulate()`:

```{r}
sim <- simulate(sfm, start = 0, stop = 250)
```

`dt` refers to the time step of the simulation, which determines how often the model's equations are evaluated. A smaller `dt` can increase the accuracy of the simulation, but also increases computational time and the size of the resulting dataframe. To reduce the saved output, we may save fewer timepoints, for instance, every 0.1 days:

```{r}
sfm <- sim_settings(sfm, save_at = 0.1)
```

Or specific time points:
```{r}
sfm <- sim_settings(sfm, save_at = c(1, 50, 100))
```

Alternatively, we can specify the number of time points to save with `save_n`:

```{r}
sfm <- sim_settings(sfm, save_n = 100)
```

Similarly, we may change the numerical method used to solve the model. The default method is `"euler"`, which is the simplest numerical integration method. For more complex models or when higher accuracy is needed, consider other methods like `"rk4"`:

```{r}
sfm <- sim_settings(sfm, method = "rk4")
```

All available simulation methods can found with:

```{r}
sim_methods()
```

Note that some methods may not be available in Julia and vice versa.

In case the simulation contains stochastic elements, we may want to set a seed to ensure that the simulation is reproducible. For example, the initial value of energy could be a random number:

```{r}
sfm <- stock(sfm, energy, eqn = runif(1, 0, 1))
```

The seed needs to be an integer:

```{r}
sfm <- sim_settings(sfm, seed = 123)
```

The seed can also be removed to ensure variation in the simulation. This can be useful to for example test the sensitivity of the model to initial condition variation. 

```{r}
sfm <- sim_settings(sfm, seed = NULL)
```


### Renaming variables

Variable names can easily be changed:

```{r}
sfm <- change_name(sfm, recovery_rate, new_name = recovery_store)
```

This will ensure that all references to `recovery_rate` are changed to `recovery_store`. 

### Allowed variable names

When creating variables or changing variable names, a warning may be issued that the name was modified to be syntactically valid and unique. For example:

```{r error = TRUE, warning = TRUE}
sfm <- change_name(sfm, recovery_rate, new_name = t)
```

The name `t` is not usable, as this already refers to the current time step. Similarly, names cannot contain spaces or special characters:

```{r error = TRUE}
sfm <- change_name(sfm, recovery_rate, new_name = recovery - rate)
```

Names also cannot be duplicated:

```{r error = TRUE, warning = TRUE}
sfm <- change_name(sfm, energy, new_name = recovery)
```

### Removing variables

To remove a variable from the model, use `discard()`:

```{r}
sfm <- discard(sfm, net_flow)
```

Note that this cannot be undone!

### Lookup functions

Lookup functions (`lookup`), also known as table or graphical functions, are interpolation functions used to create custom input-output functions, where we define the desired output (y) for a specified input (x). They are defined by a set of x- and y-domain points. The interpolation method defines the behaviour of the lookup function between x-points, and the extrapolation method defines the behaviour outside of the x-points. For example, a simple lookup function called `"graph"` may look like this:

```{r}
sfm <- stockflow() |>
  lookup(graph,
    xpts = c(0, 1, 2), ypts = c(0.5, 1, 1),
    interpolation = "linear", extrapolation = "nearest"
  )
```

The function can now be used in any equation in the model like so:

```{r}
sfm <- constant(sfm, x, eqn = graph(1))
```


### Custom functions 

New functions can be defined such that they can be used anywhere in the model. For example, if the `logistic()` function did not exist, you could create it yourself:

```{r}
sfm <- stockflow() |>
  custom_func(f, eqn = function(x, slope = 1, midpoint = .5) 1 / (1 + exp(-slope * (x - midpoint))))
```

This will create a function `f()` that can be used in any equation in the model like so: 

```{r}
sfm <- constant(sfm, x, eqn = f(0))
```




### Documenting

To document meta-properties of the model, use `meta()`. For example, the model's `name`, subtitle (`caption`), or `author`. `meta()` accepts any key-value pair, so custom metadata can also be added.

```{r}
sfm <- meta(sfm, author = "Kyra Evers", affiliation = "University of Amsterdam")
```
