---
title: "Building Stock-and-Flow Models"
output: rmarkdown::html_vignette
bibliography: references/references.bib
vignette: >
  %\VignetteIndexEntry{Building Stock-and-Flow Models}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: inline
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6, # Set default figure width to 6 inches
  fig.height = 3, # Set default figure height to 3 inches
  fig.align = "center", # Center all figures
  out.width = "80%" # Display figures at 80% of container width
  # fig.retina = 2,         # Optimize for retina displays
  # dpi = 96               # Standard DPI for web
)


# Allow LaTeX-style inline markup in prose. The `document` hook rewrites the
# knitted markdown before pandoc sees it, so \textit/\texttt/\code render in
# HTML. Applied to prose only; avoid these literal strings inside code output.
knitr::knit_hooks$set(document = function(doc) {
  doc <- gsub("\\\\textit\\{([^}]*)\\}", "*\\1*", doc) # italic
  doc <- gsub("\\\\textbf\\{([^}]*)\\}", "**\\1**", doc) # bold
  doc <- gsub("\\\\texttt\\{([^}]*)\\}", "`\\1`", doc) # monospace / backticks
  doc <- gsub("\\\\code\\{([^}]*)\\}", "`\\1`", doc) # monospace / backticks
  doc
})
```

```{r, echo=TRUE, message=FALSE}
library(sdbuildR)
```

```{r, echo = FALSE, cache=FALSE}
# Disable WebGL: many plotly widgets per HTML page can exceed the browser WebGL
# context limit and render blank. SVG always renders.
options(sdbuildR.webgl = FALSE)
```


```{r, echo = FALSE}
font_family <- "stix-two-text"
options(sdbuildR.font_family = font_family)
# getOption("sdbuildR.font_family")
```

```{r libraries, include=FALSE}
library(knitr)
library(kableExtra)
```


```{r include=FALSE}
# Set this to TRUE to recreate the figures in the paper
recreate_figs <- interactive()
```


Stock-and-flow models represent systems as states (stocks) that accumulate over time with processes (flows) that change these variables. In this vignette, we will demonstrate how to create stock-and-flow models from scratch using sdbuildR. It covers the basics of stock-and-flow modelling in the context of psychology with an example of burnout. Note that this vignette serves as online supplemental material A accompanying the paper *Formalizing Psychological Theory with sdbuildR: A Stock-and-Flow Modelling Tutorial in R* by Evers et al. (under review). To reproduce the figures in the paper, please see the bottom of the corresponding .Rmd file.

## Stock-and-Flow Models

Stock-and-flow models conceptualize systems in terms of quantities that accumulate (i.e., stocks) and the processes (i.e., flows) that change them over time. Stocks are like the amount of water in a bathtub, whereas flows are like the water flowing in and out of the tub. Inflows -- water from the tap -- raise the stock, while outflows -- water through the drain -- lower it. Without an outflow, the water remains in the bathtub; without an inflow, the bathtub stays empty. Stocks can have multiple inflows and outflows, but each flow can only originate in one stock and terminate in one stock. Flow rates are expressed in units over time (e.g., litre per minute). The net rate of change in the water level is the difference between the inflows and outflows. The current water level (i.e., the value of the stock) is the result of how much water has flowed in and out of the bath tub. In this way, a stock functions as a memory of past activity by integrating flows over time. When a flow directly connects two stocks, it is a \textit{material} flow that conserves the quantity it carries: the source stock decreases exactly as much as the target stock increases. \textit{Informational} flows, such as knowledge, beliefs, and perceptions, are not conserved, as sharing information does not deplete it [@Sterman2000BusinessWorld, p. 201]. Because psychological systems predominantly involve informational flows, we do not discuss material flows further. This structure is the foundation of stock-and-flow models, where stocks represent the state of a system, and flows represent the processes that alter that state over time.

```{r, echo=FALSE}
sfm <- stockflow() |>
  stock(Stock) |>
  flow(Inflow, to = Stock) |>
  flow(Outflow, from = Stock)


pl <- plot(sfm, show_eqn = FALSE)
# cat(pl$x$diagram)

{
  viz_str <- '
   digraph sfm {

      graph [layout = dot, rankdir = LR, center=true, outputorder="edgesfirst", pad=0.9, nodesep=0.3, splines = true, concentrate = false]

      # Shared across all nodes (persists until overridden)
      node [fontsize=18,fontname="Times New Roman",fontcolor="black"]

      # Define stock nodes
      node [shape=box,style=filled,fillcolor="#83d3d4"]
      "Stock" [id="Stock",label="Stock", tooltip = "Stock: Stock\nInitial value: 0\nInflows: Inflow\nOutflows: Outflow"]

      # Define flow nodes (intermediate nodes for flows)
      node [style = "",shape=plaintext, fontsize=16, width=0.6, height=0.3]
      "Inflow" [id="Inflow",label="Inflow", tooltip = "Flow: Inflow\nRate: 0\nFrom: outside model boundary\nTo: Stock"]
        "Outflow" [id="Outflow",label="Outflow", tooltip = "Flow: Outflow\nRate: 0\nFrom: Stock\nTo: outside model boundary"]

      # Define external cloud nodes
      node [shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
      "Cloud1" [label=<Unspecified sink<BR/><FONT POINT-SIZE="13" COLOR="black">(outside of model boundary)</FONT><BR/><BR/><BR/><BR/><FONT COLOR="white">.</FONT>>, tooltip = "Outside model boundary\nSink of: Outflow"]
        "Cloud2" [label=<Unspecified source<BR/><FONT POINT-SIZE="13" COLOR="black">(outside of model boundary)</FONT><BR/><BR/><BR/><BR/><FONT COLOR="white">.</FONT>>, tooltip = "Outside model boundary\nSource of: Inflow"]

      # Define auxiliary nodes



      # Define constant nodes



      # Define flow edges (stock -> flow_node)
      edge [style = "", arrowhead="none", color="black:#f48153:black", penwidth=1.1, minlen=1, tailport="e", headport="w"]
      "Cloud2" -> "Inflow"
        "Stock" -> "Outflow"

      # Define flow edges (flow_node -> stock)
      edge [style = "", arrowhead="normal", color="black:#f48153:black", arrowsize=1.5, penwidth=1.1, minlen=1, tailport="e", headport="w"]
      "Inflow" -> "Stock"
        "Outflow" -> "Cloud1"

      # Define dependency edges
      edge [style = "", color="#999999", arrowsize=0.8, penwidth=1, constraint=false, tailport = "_", headport="_"]

      # Rank groupings


    }
  '
}

pl <- DiagrammeR::grViz(viz_str)
pl
```

```{r, include=FALSE}
if (recreate_figs) {
  pl_sfm_inflow_outflow <- pl
}
```


Stock-and-flow models are easiest to understand through a worked example. We will create a simple model of people waiting in a queue to introduce the syntax of \texttt{sdbuildR}.

We recommend constructing models incrementally and iteratively, as this cultivates a deeper understanding of how each modification influences system behaviour. To do so, we create a stock-and-flow model object called \code{sfm}:

```{r, fig.height=2}
sfm <- stockflow()
print(sfm)
```

The object contains default simulation settings such as the total duration, the timestep (\code{dt}) specifying the temporal resolution of the simulation, and a solver (\code{euler}) indicating the numerical technique used to generate output from the model [for more details, see @deSolve]. We update the simulation duration to ten hours:

```{r}
sfm <- sfm |>
  sim_settings(stop = 10, time_units = "hours", only_stocks = FALSE)
```

We further add a model name and description to the model metadata:  
  
```{r}
sfm <- sfm |>
  meta(
    name = "Queue Model",
    description = "A simple model of a queue of people waiting for service"
  )
```

As we build the model iteratively, the object \code{sfm} will continuously be overwritten.
\code{sfm} contains no model variables yet. Throughout the tutorial, we use the term "variable" for any part of the system, be that a stock, flow, constant, or auxiliary. Though this usage may differ from other scientific fields, we here choose to adhere to system dynamics terminology [@ford_system_2019; @Sterman2000BusinessWorld]. We added a stock to the model called \code{people}. Every stock needs an \textit{initial condition}: its value at the start of the simulation. We initialize \code{people} at 0:

```{r, fig.height=2}
sfm <- sfm |>
  stock(people, eqn = 0, label = "People Waiting in Queue")
```


Plotting the model shows its stock-and-flow diagram:

```{r, fig.height=1}
plot(sfm)
```



```{r include=FALSE}
if (recreate_figs) {
  sfm_list <- list()
  sfm_list[["A"]] <- sfm
}
```

To assess the model's dynamics, we simulate it over time and visualize the resulting timeseries:

```{r, fig.height=5}
sfm |>
  simulate() |>
  plot(animation = "time")
```


Above, we use the pipe operator \code{|>} to pass the result of an expression to the next expression as its first argument (e.g., \code{x |> f(y)} is interpreted as \code{f(x, y)}). As shown in Figure \ref{fig:build_queue_sim}A, \code{people} remains at the same value across the entirety of the simulation. Stocks without flows are static, as there is no process specifying how they change. 
 
We first added an inflow of \code{arrivals} at a constant rate of one person per hour. 
 
```{r, fig.height=2}
sfm <- sfm |>
  flow(arrivals, eqn = 1, to = people, label = "Arrivals")

plot(sfm)
```


```{r include=FALSE}
if (recreate_figs) {
  sfm_list[["B"]] <- sfm
}
```

A stock with a constant inflow increases linearly:

```{r, fig.height=5}
sfm |>
  simulate() |>
  plot(animation = "time")
```

To decrease the stock, we add an outflow of \code{service} which removes people from the stock at a constant rate of two people per hour.

```{r}
sfm <- sfm |>
  flow(service, eqn = 2, from = people, label = "Service")

plot(sfm)
```


```{r include=FALSE}
if (recreate_figs) {
  sfm_list[["C"]] <- sfm
}
```

As the outflow rate exceeds the inflow rate, the stock becomes negative.

```{r, fig.height=5}
sfm |>
  simulate() |>
  plot(animation = "time")
```

To rectify this implausible behaviour, a naive solution may be to include a logical statement such as \code{ifelse(people < 0, 0, people)}. However, this computational trick would mask model misspecification. Ideally, stocks should remain within bounds due to plausible equations and parameters. Below, we prevent negative people by making \code{service} proportional to the number of people in the queue: \code{0.5 * people}. In this way, when \code{people} is zero, the outflow is also zero.

```{r}
sfm <- update(sfm, service, eqn = .5 * people)
plot(sfm)
```


```{r include=FALSE}
if (recreate_figs) {
  sfm_list[["D"]] <- sfm
}
```


```{r, fig.height=5}
sfm |>
  simulate() |>
  plot(animation = "time")
```

This formulation introduces a \textit{negative feedback loop} to the system [@Meadows2008ThinkingPrimer]: the greater the number of people, the greater the outflow. Positive feedback loops amplify change, whereas negative feedback loops bring the system back to a target state [@Sterman2000BusinessWorld]. The number of people stabilizes at a fixed level when the outflow meets the constant inflow.

Lastly, we add a second outflow from the stock representing people leaving as they are tired of waiting. This lowers the level at which the stock stabilizes.

```{r}
sfm <- sfm |>
  flow(leave,
    eqn = people^10, from = people,
    label = "Tired of Waiting"
  )

plot(sfm)
```


```{r include=FALSE}
if (recreate_figs) {
  sfm_list[["E"]] <- sfm
}
```


```{r, fig.height=5}
sfm |>
  simulate() |>
  plot(animation = "time")
```

The queue model demonstrates a principle key to understanding stock-and-flow modelling. Variables are \textit{continuous quantities} that change smoothly over time. As a result, the queue model produces fractional people, and does not track individuals but rather people on aggregate. This remains a contested issue [@Tafreshi2016QuantificationPractice; @franz_are_2022], where some treat it as a pragmatic simplification and others as a serious empirical and philosophical commitment in need of justification [@Michell1997QuantitativePsychology]. If variables cannot reasonably be treated as continuous quantities, other modelling frameworks may be more appropriate (e.g., Markov models for transitions between discrete states).

#### Constants and Auxiliaries

Stock-and-flow models can further be supplemented with two other variable types: constants and auxiliaries.
Constants are static parameters that do not change over the time course of the simulation. In contrast, auxiliaries are dynamic, meaning they are computed anew at each step. They are used for intermediate computations in flow equations or to monitor other dynamic quantities. To illustrate the difference, a constant defined as \code{runif(1)} will be fixed to a random number at the beginning of the simulation, whereas an equivalently defined auxiliary will draw a new number each time step.

Constants and auxiliaries help to make equations more interpretable and easier to revise later. For instance, we may explicitly denote the service rate with a constant:


```{r}
sfm <- sfm |>
  constant(service_rate, eqn = .5, label = "Service Rate") |>
  update(service, eqn = service_rate * people)

order_vars <- c("leave", "service", "service_rate")
plot(sfm, show_constants = TRUE, align = order_vars, order = order_vars, spacing = .5)
```

Similarly, we can add an auxiliary to track satisfaction: the fraction of people leaving the queue who are served, rather than abandoning it out of frustration:

```{r}
sfm <- sfm |>
  aux(satisfaction, eqn = service / (service + leave), label = "Satisfaction")
```


```{r}
order_vars <- list(c("leave", "service", "service_rate"), c("satisfaction"))
plot(sfm, show_constants = TRUE, show_aux = TRUE, align = order_vars, order = order_vars, spacing = .5)
```

```{r, fig.height=5}
sfm |>
  simulate() |>
  plot(animation = "time")
```


```{r, include = FALSE}
order_vars <- list(c("satisfaction", "leave", "service", "service_rate"), c("people"))

pl <- plot(sfm, show_constants = TRUE, show_aux = TRUE, align = order_vars, order = order_vars, spacing = .5, wrap_width = 30)
pl


# pl$x$diagram |> cat()

{
  viz_str <- '
digraph sfm {

      graph [layout = dot, rankdir = LR, center=true, outputorder="edgesfirst", pad=0.1, nodesep=0.5, splines = true, concentrate = false]

      # Shared across all nodes (persists until overridden)
      node [fontsize=18,fontname="stix-two-text",fontcolor="black"]

      # Define stock nodes
      node [shape=box,style=filled]
      "people" [id="people",fillcolor="#83D3D4",label=<People Waiting in Queue<BR/><FONT POINT-SIZE="13" COLOR="black">Initial value = 0</FONT>>, tooltip = "Stock: People Waiting in Queue\nName: people\nInitial value: 0\nInflows: Arrivals\nOutflows: Tired of Waiting, Service"]

      # Define flow nodes (intermediate nodes for flows)
      node [style = "",shape=plaintext, fontsize=16, width=0.6, height=0.3]
      "arrivals" [id="arrivals",label=<Arrivals<BR/><FONT POINT-SIZE="13" COLOR="black">Rate = 1</FONT>>, tooltip = "Flow: Arrivals\nName: arrivals\nRate: 1\nFrom: outside model boundary\nTo: People Waiting in Queue"]
        "leave" [id="leave",label=<Tired of Waiting<BR/><FONT POINT-SIZE="13" COLOR="black">Rate = people^10</FONT>>, tooltip = "Flow: Tired of Waiting\nName: leave\nRate: people^10\nFrom: People Waiting in Queue\nTo: outside model boundary"]
        "service" [id="service",label=<Service<BR/><FONT POINT-SIZE="13" COLOR="black">Rate = service_rate * people</FONT>>, tooltip = "Flow: Service\nName: service\nRate: service_rate * people\nFrom: People Waiting in Queue\nTo: outside model boundary"]

      # Define external cloud nodes
      node [shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
      "Cloud1" [label="", tooltip = "Outside model boundary\nSink of: Tired of Waiting"]
        "Cloud2" [label="", tooltip = "Outside model boundary\nSink of: Service"]
        "Cloud3" [label="", tooltip = "Outside model boundary\nSource of: Arrivals"]

      # Define auxiliary nodes
      node [shape=circle,fontsize=16, width=0.15, height=0.15, fixedsize=true, style=filled]
      "satisfaction" [id="satisfaction",fillcolor="#E5E5E5",
      # label=<Satisfaction<BR/><FONT POINT-SIZE="13" COLOR="black">Equation = service / (service + leave)</FONT>>,
      label=<Satisfaction<BR/><FONT POINT-SIZE="13" COLOR="black">Equation = service / (service + leave)</FONT><BR/><BR/><BR/><FONT COLOR="white">.</FONT>>,

      xlabel="", tooltip = "Auxiliary: Satisfaction\nName: satisfaction\nEquation: service / (service + leave)"]

      # Define constant nodes
      node [shape=diamond,fontsize=16,width=0.15, height=0.15, fixedsize=true, style=filled]
      "service_rate" [id="service_rate",fillcolor="#E5E5E5",xlabel=<<I>Service Rate</I><BR/><FONT POINT-SIZE="13" COLOR="black">Value = 0.5</FONT>>,label="", tooltip = "Constant: Service Rate\nName: service_rate\nValue: 0.5"]

      # Define flow edges (stock -> flow_node)
      edge [style = "", arrowhead="none", penwidth=1.1, minlen=1, tailport="e", headport="w"]
      "Cloud3" -> "arrivals" [color="black:#F48153:black"]
        "people" -> "leave" [color="black:#F48153:black"]
        "people" -> "service" [color="black:#F48153:black"]

      # Define flow edges (flow_node -> stock)
      edge [style = "", arrowhead="normal", arrowsize=1.5, penwidth=1.1, minlen=1, tailport="e", headport="w"]
      "arrivals" -> "people" [color="black:#F48153:black"]
        "leave" -> "Cloud1" [color="black:#F48153:black"]
        "service" -> "Cloud2" [color="black:#F48153:black"]

      # Define dependency edges
      edge [style = "", color="#999999", arrowhead="open", arrowsize=0.8, penwidth=1, constraint=false, tailport = "_", headport="_"]
      "people" -> "leave" [headport = "n", tailport = "n"]
        "service_rate" -> "service"
        "people" -> "service" [headport = "s", tailport = "s"]
        "service" -> "satisfaction"
        "leave" -> "satisfaction"

      # Ordering hints (invisible edges)
      edge [style=invis, constraint=true]
      "satisfaction" -> "leave"
      "leave" -> "service"
      "service" -> "service_rate"


      # Rank groupings
            {rank=same; "service_rate"; "service" }
      {rank=same; "satisfaction"; "leave"; "service"; "service_rate" }

    }
'
}

pl <- DiagrammeR::grViz(viz_str)
pl
```


```{r include=FALSE, eval = interactive(), error = TRUE}
if (recreate_figs) {
  pl_queue_sfm <- pl
}
```



```{r, include=FALSE}
# sfm |>
#   export_model() |>
#   cat()
```




## Variable types

```{r echo = FALSE}
variable_types_table <- data.frame(
  Characteristic = c(
    "Role in system",
    "Varies within time horizon",
    "A process taking place over time",
    "Can be captured at any given moment in time",
    "`eqn` denotes",
    "Allowed dependencies in `eqn`",
    "Examples"
  ),
  Stock = c(
    "Defines the state of the system; accumulates the effects of flow(s) over time",
    "✓",
    "✗",
    "✓",
    "Initial condition",
    "Constants and (initial values of) other stocks",
    "Emotions, beliefs, stress, trust, resources"
  ),
  Flow = c(
    "Increases or decreases a stock",
    "✓",
    "✓",
    "✗",
    "Flow rate computed at every time step",
    "Any other variable",
    "Coping, learning, emotion regulation"
  ),
  Constant = c(
    "Specifies static quantity",
    "✗",
    "✗",
    "✓",
    "Fixed value",
    "Other constants and (initial values of) stocks",
    "Rates, capacities, thresholds"
  ),
  Auxiliary = c(
    "Provides intermediate computations for convenience; keeps track of changing quantities",
    "✓",
    "Possibly",
    "✓",
    "Value computed at every time step",
    "Any other variable",
    "Performance indices, ratios, sums of stocks"
  )
)

kableExtra::kbl(variable_types_table,
  format = "markdown",
  col.names = c("Characteristic", "Stock", "Flow", "Constant", "Auxiliary"),
  caption = "Characteristics of Variable Types in Stock-and-Flow Models"
) |>
  kableExtra::kable_styling()
```

The following flowchart can be used to determine a variable's type:

```{r include = FALSE}
{
  viz_str <- "
digraph flowchart {

  graph [layout = dot,
         rankdir = TB,
         bgcolor = white,
         nodesep = 0.45,
         ranksep = 0.5,
         splines = ortho,
         pad = '0.2']

  node [fontsize = 9.5,
        fontcolor = 'white',
        penwidth = 0]

  edge [fontsize = 8.5,
        fontcolor = '#555555',
        color = '#666666',
        penwidth = 1.3,
        arrowsize = 0.6]

  # --- DECISION NODES (rounded boxes with border) ---
  q1 [label = 'Can the variable be represented\\nas a continuous quantity that\\nbe lower or higher?',
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.15,0.1']

  q2 [label = 'Does the variable change within\\nthe chosen time horizon?',
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.15,0.1']

  q3 [label = 'Is the variable a behaviour, action,\\nor process that takes place over time?',
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.15,0.1']

  q4 [label = 'Does the variable accumulate\\nthe effects of flows?',
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.15,0.1']

  q5 [label = 'Does the variable\\'s effect\\naccumulate in a stock?', # Does it directly increase or decrease a stock?
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.15,0.1']

  q6 [label = 'Should it?',
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.12,0.08']

  # --- TERMINAL NODES (no border, darker fill) ---
  constant [label = 'Constant',
            shape = box, style = 'filled,rounded',
            fillcolor = '#0f5c5a', color = 'none',
            fontsize = 10, margin = '0.15,0.08']

  stock [label = 'Stock',
         shape = box, style = 'filled,rounded',
         fillcolor = '#0f5c5a', color = 'none',
         fontsize = 10, margin = '0.15,0.08']

  flow [label = 'Flow',
        shape = box, style = 'filled,rounded',
        fillcolor = '#0f5c5a', color = 'none',
        fontsize = 10, margin = '0.15,0.08']

  auxiliary [label = 'Auxiliary',
             shape = box, style = 'filled,rounded',
             fillcolor = '#0f5c5a', color = 'none',
             fontsize = 10, margin = '0.15,0.08']

  # --- ACTION NODES ---
  reformulate [label = 'Reformulate variable',
               shape = box, style = 'dashed,rounded,filled',
            fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
            fontsize = 8, penwidth = 1.2, color = '#17807c',
            margin = '0.1,0.06']

  add_flow [label = 'Add missing\\nflows',
            shape = box, style = 'dashed,rounded,filled',
            fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
            fontsize = 8, penwidth = 1.2, color = '#17807c',
            margin = '0.1,0.06']

  add_stock [label = 'Add missing\\nstock',
             shape = box, style = 'dashed,rounded,filled',
             fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
             fontsize = 8, penwidth = 1.2, color = '#17807c',
             margin = '0.1,0.06']

  # --- EDGES ---
  q1 -> reformulate [xlabel = 'No', minlen = 2]
  reformulate -> q1 [minlen = 2, style = dashed, penwidth = 1]
  q1 -> q2          [xlabel = 'Yes']

  q2 -> constant [xlabel = 'No', minlen = 2]
  q2 -> q3       [xlabel = 'Yes']

  q3 -> q4 [xlabel = <<font>No, it can be captured<br/>at a single moment in time</font>>]
  q3 -> q5 [xlabel = '  Yes']

  q4 -> stock [xlabel = 'Yes']
  q4 -> q6    [xlabel = 'No']

  q5 -> flow  [xlabel = 'Yes']
  q5 -> q6    [xlabel = 'No']

  q6 -> auxiliary [xlabel = 'No']
  q6 -> add_flow  [xlabel = 'Yes', style = dashed, color = '#2a9d8f', penwidth = 1]
  q6 -> add_stock [xlabel = 'Yes', style = dashed, color = '#2a9d8f', penwidth = 1]
  add_flow  -> stock [style = dashed, color = '#2a9d8f', penwidth = 1]
  add_stock -> flow  [style = dashed, color = '#2a9d8f', penwidth = 1]

  # --- RANK CONSTRAINTS ---
  { rank = same; q1; reformulate }
  { rank = same; q2; constant }
  { rank = same; q4; q5 }
  { rank = same; stock; q6; flow }
  { rank = same; add_flow; auxiliary; add_stock }
}
"
}


pl <- DiagrammeR::grViz(viz_str)
pl
```



## Overview of package functionality


```{r echo=FALSE}
functions_table <- data.frame(
  Function = c("`stockflow()`", "`stock()`", "`flow()`", "`constant()`", "`aux()`", "`lookup()`", "`update()`", "`simulate()`", "`plot()`", "`summary()`", "`as.data.frame()`", "`sim_settings()`", "`meta()`", "`export_model()`"),
  Purpose = c("Create empty model or load template", "Add or modify a stock", "Add or modify a flow", "Add or modify a constant", "Add or modify an auxiliary", "Add or modify a lookup function", "Add or modify any variable (generic)", "Simulate model", "Plot model diagram or simulation", "Run model diagnostics", "Get model properties in a dataframe", "Modify simulation specifications", "Modify model metadata", "Export models to other formats")
)

kableExtra::kbl(functions_table,
  format = "markdown",
  col.names = c("Function", "Purpose"),
  caption = "Main functions in *sdbuildR*"
) |>
  kableExtra::kable_styling()
```

### Global simulation variables

In some cases, it may be useful to refer to global simulation variables in the model's equations:

```{r echo=FALSE}
functions_table <- data.frame(
  Variable = c(
    "`times`", "`t`", "`dt`"
  ),
  Description = c(
    "Vector with simulation times",
    "Current time in the ODE",
    "Time step of the simulation"
  ),
  Use.case = c(
    "Use in equations, e.g., `pulse(times, 5, width = dt)`",
    "Use in equations, e.g., `input(t)`",
    "Use in equations, e.g., `pulse(times, 5, width = dt)`"
  )
)

kableExtra::kbl(functions_table,
  format = "markdown",
  col.names = c(
    "Variable", "Description", "Use case"
  ),
  caption = "Global simulation variables in *sdbuildR*"
) |>
  kableExtra::kable_styling()
```


### Simulation specifications

We may want to simulate the system over a longer time period, or with a different time step.

```{r}
sfm <- sfm |>
  sim_settings(
    start = 0,
    stop = 250,
    dt = 0.001
  )
```


Simulation settings can be set directly on the model object as above, or passed to `simulate()`:

```{r}
sim <- simulate(sfm, start = 0, stop = 250)
```

`dt` refers to the time step of the simulation, which determines how often the model's equations are evaluated. A smaller `dt` can increase the accuracy of the simulation, but also increases computational time and the size of the resulting dataframe. To reduce the saved output, we may save fewer timepoints, for instance, every 0.1 days:

```{r}
sfm <- sim_settings(sfm, save_by = 0.1)
```

Or specific time points:
```{r}
sfm <- sim_settings(sfm, save_times = c(1, 50, 100))
```

Alternatively, we can specify the number of time points to save with `save_length`:

```{r}
sfm <- sim_settings(sfm, save_length = 100)
```

Similarly, we may change the numerical method used to solve the model. The default method is `"euler"`, which is the simplest numerical integration method. For more complex models or when higher accuracy is needed, consider other methods like `"rk4"`:

```{r}
sfm <- sim_settings(sfm, method = "rk4")
```

All available simulation methods can found with:

```{r}
sim_methods()
```

Note that some methods may not be available in Julia and vice versa.

In case the simulation contains stochastic elements, we can set a seed to ensure that the simulation is reproducible. For example, the initial value of people could be a random number:

```{r}
sfm <- stock(sfm, people, eqn = sample(0:10, 1))
```

The seed needs to be an integer:

```{r}
sfm <- sim_settings(sfm, seed = 123)
```

The seed can also be removed to ensure variation across simulations. 

```{r}
sfm <- sim_settings(sfm, seed = NULL)
```


### Renaming variables

Variable names can easily be changed:

```{r}
sfm <- change_name(sfm, people, new_name = customers)
```

This will ensure that all references to `people` are changed to `customers`. 

### Allowed variable names

When creating variables or changing variable names, a warning may be issued that the name was modified to be syntactically valid and unique. For example:

```{r error = TRUE, warning = TRUE}
sfm <- change_name(sfm, customers, new_name = t)
```

The name `t` is not usable, as this already refers to the current time step. Similarly, names cannot contain spaces or special characters:

```{r error = TRUE}
sfm <- change_name(sfm, t_1, new_name = a - b)
```

Names also cannot be duplicated:

```{r error = TRUE, warning = TRUE}
sfm <- change_name(sfm, service, new_name = served)
```

### Removing variables

To remove a variable from the model, use `discard()`:

```{r}
sfm <- discard(sfm, served)
```

Note that this cannot be undone!

### Lookup functions

Lookup functions (`lookup`), also known as table or graphical functions, are interpolation functions used to create custom input-output functions, where we define the desired output (y) for a specified input (x). They are defined by a set of x- and y-domain points. The interpolation method defines the behaviour of the lookup function between x-points, and the extrapolation method defines the behaviour outside of the x-points. For example, a simple lookup function called `"graph"` may look like this:

```{r}
sfm <- stockflow() |>
  lookup(graph,
    xpts = c(0, 1, 2), ypts = c(0.5, 1, 1),
    interpolation = "linear", extrapolation = "nearest"
  )
```

The function can now be used in any equation in the model like so:

```{r}
sfm <- constant(sfm, x, eqn = graph(1))
```


### Custom functions 

New functions can be defined such that they can be used anywhere in the model. For example, if the `logistic()` function did not exist, you could create it yourself:

```{r}
sfm <- stockflow() |>
  custom_func(f, eqn = function(x, slope = 1, midpoint = .5) 1 / (1 + exp(-slope * (x - midpoint))))
```

This will create a function `f()` that can be used in any equation in the model like so: 

```{r}
sfm <- constant(sfm, x, eqn = f(0))
```




### Documenting

To document meta-properties of the model, use `meta()`. For example, the model's `name`, subtitle (`caption`), or `author`. `meta()` accepts any key-value pair, so custom metadata can also be added.

```{r}
sfm <- meta(sfm, author = "Kyra Evers", affiliation = "University of Amsterdam")
```



```{r include=FALSE, eval = interactive(), error = TRUE}
# Plotting parameters
filepath_base <- this.path::this.dir()
filepath_figs <- file.path(filepath_base, "figs_tutorial2026")

if (recreate_figs) {
  dir.create(filepath_figs, showWarnings = FALSE)
}

VARS <- c(
  "people", "arrivals", "service", "leave"
)
PALETTE <- c("Geyser", "Temps", "Spectral", "Zissou 1", "Roma")[5]
COLORS <- grDevices::hcl.colors(n = 50, palette = PALETTE)
# scales::show_col(COLORS)

inflow_col <- adjustcolor(COLORS[10], alpha.f = 1)
outflow_col <- adjustcolor(COLORS[3], alpha.f = 1)

inflow_fill <- adjustcolor(inflow_col, alpha.f = .5)
outflow_fill <- adjustcolor(outflow_col, alpha.f = .5)

# scales::show_col( grDevices::hcl.colors(n = 50, palette = "Greens"))
increase_col <- adjustcolor(
  grDevices::hcl.colors(n = 50, palette = "Greens")[20],
  alpha.f = 1
)
decrease_col <- adjustcolor(COLORS[2], alpha.f = 1)
decrease_col_strong <- adjustcolor(COLORS[1], alpha.f = 1)
# scales::show_col(c(increase_col, decrease_col))
increase_fill <- adjustcolor(increase_col, alpha.f = .8)
decrease_fill <- adjustcolor(decrease_col, alpha.f = .4)

COLORS_alpha <- adjustcolor(COLORS, alpha.f = .25)

COLORS <- COLORS[c(37, 9, 32, 1, 47, 17)]
stock_col <- adjustcolor(COLORS[1], alpha.f = .75)
flow_col <- COLORS[6]

# Prepare colors for queue
COLORS_queue <- c(stock_col, inflow_col, outflow_col, decrease_col_strong)
names(COLORS_queue) <- VARS

font_size <- 16
font_size_xlab <- round(font_size * 1.4)
font_size_legend <- round(font_size * 1.1)
font_size_subplot <- font_size
xlab <- "Time (hours)"

# Finer-grained simulation
save_length_fine <- 500 # .1 # .01

# Annotation template
annot <- list(xanchor = "center", yanchor = "middle", showarrow = FALSE, xref = "paper", yref = "paper", font = list(size = font_size_subplot))
annot_font <- list(
  family = font_family,
  size = font_size
)
annot_legend <- list(
  orientation = "h", # orientation
  x = 0.5, # Center horizontally
  y = -0.15, # Below the plot
  xanchor = "center",
  yanchor = "top",
  font = list(size = font_size_legend)
)
annot_xlab <- list(
  text = xlab,
  font = list(
    size = font_size_xlab
  ),
  xref = "paper",
  yref = "paper",
  xanchor = "center",
  yanchor = "top",
  x = 0.5, y = -.1,
  showarrow = FALSE
)
annot_ylab <- utils::modifyList(annot_xlab, list(
  x = -.1, y = 0.5,
  textangle = -90, yanchor = "middle", xanchor = "center"
))

rm_grid <- function(pl) {
  plotly::layout(pl,
    xaxis = list(showgrid = FALSE, showline = TRUE),
    yaxis = list(showgrid = FALSE, showline = TRUE)
  )
}
```


```{r, include=FALSE,eval = interactive(), error = TRUE}
fig_sfm_inflow_outflow <- function() {
  export_plot(pl_sfm_inflow_outflow,
    file.path(
      filepath_figs,
      "build_sfm_inflow_outflow.pdf"
    ),
    font_family = font_family
  )
}



fig_queue_sfm <- function() {
  export_plot(pl_queue_sfm, file.path(filepath_figs, "build_queue_sfm.pdf"),
    font_family = font_family
  )
}


fig_queue_sim <- function() {
  # Adjust the first sfm to have a zero inflow
  sfm_list[["A"]] <- sfm_list[["A"]] |>
    flow(arrivals, eqn = 0, to = people)

  # Create subplot figure
  pl_list <- list()
  counter <- 1
  max_letter <- max(names(sfm_list))

  vars_stock <- c("people")
  vars_flows <- c("arrivals", "service", "leave")

  for (letter in names(sfm_list)) {
    sim <- simulate(sfm_list[[letter]],
      save_length = save_length_fine, only_stocks = FALSE
    )

    vars_in_sim <- unique(sim$df$variable)
    vars_flows_ <- vars_flows[vars_flows %in% vars_in_sim]
    colors_flows_ <- COLORS_queue[vars_flows_]
    if (letter == "A") {
      colors_flows_ <- "white"
    }
    order_flows_ <- vars_flows_
    if (letter == "C") {
      order_flows_ <- rev(order_flows_)
    }

    pl_list[[counter]] <- plot(sim,
      vars = vars_stock,
      colors = COLORS_queue[vars_stock],
      margin = 0,
      main = "", xlab = "",
      show_legend = (letter == max_letter)
    ) |>
      rm_grid() |>
      plotly::layout(
        yaxis = list(
          nticks = 3
        )
      )

    counter <- counter + 1

    # Add flows
    pl_list[[counter]] <- plot(sim,
      vars = vars_flows_,
      colors = colors_flows_,
      order = order_flows_,
      margin = 0,
      main = "", xlab = "",
      show_legend = (letter == max_letter)
    ) |>
      rm_grid() |>
      plotly::layout(
        yaxis = list(
          nticks = 3
        )
      )

    if (letter == "A") {
      pl_list[[counter]] <- pl_list[[counter]] |>
        plotly::layout(
          xaxis = list(showline = FALSE, zeroline = FALSE, showticklabels = FALSE),
          yaxis = list(showline = FALSE, zeroline = FALSE, showticklabels = FALSE)
        )
    }


    counter <- counter + 1
  }

  # All plots in one column/row

  # # 1 row
  # x1 <- 0.22
  # x2 <- .44
  # x3 <- .66
  # x4 <- .88
  # y1 <- y2 <- y3 <- y4 <- 1.01
  ylab_pos <- 0.5

  # 1 column
  x1 <- x2 <- x3 <- x4 <- 0.5
  y1 <- 1
  y2 <- 0.82
  y3 <- 0.6
  y4 <- 0.38
  y5 <- 0.17

  titles <- paste0(
    paste0(
      "<i>", LETTERS[seq_along(sfm_list)], ". ",
      c(
        "Static Stock",
        "Constant Inflow",
        "Constant Outflow Larger Than Inflow",
        "Stock-Dependent Outflow",
        "Additional Outflow"
      ), "</i>"
    )
  )
  titles <- stats::setNames(titles, names(sfm_list))

  heights <- c(.2, .25, .25, .25, .2)
  heights <- heights / sum(heights)
  pl <- plotly::subplot(pl_list,
    # nrows = 1,
    nrows = 5,
    # widths = c(.4, .1, .4),
    heights = heights,
    shareX = FALSE, shareY = FALSE,
    titleY = FALSE, titleX = FALSE,
    margin = 0.05
  ) |>
    plotly::layout(
      xaxis = list(showticklabels = FALSE),
      xaxis1 = list(showticklabels = FALSE),
      xaxis2 = list(showticklabels = FALSE),
      xaxis3 = list(showticklabels = FALSE),
      xaxis4 = list(showticklabels = FALSE),
      xaxis5 = list(showticklabels = FALSE),
      xaxis6 = list(showticklabels = FALSE),
      xaxis7 = list(showticklabels = FALSE),
      xaxis8 = list(showticklabels = FALSE),
      yaxis3 = list(range = c(0, 10.1)),
      yaxis4 = list(range = c(0, 1.1)),
      yaxis7 = list(range = c(0, 2.1)),
      yaxis8 = list(range = c(0, 1.1)),
      yaxis9 = list(range = c(0, 1.1)),
      yaxis10 = list(range = c(0, 1.1)),
      font = annot_font,
      legend = utils::modifyList(annot_legend, list(
        y = -.12, x = 0.5
        # font = list(size = font_size_legend)
      )),
      annotations = list(
        utils::modifyList(annot_xlab, list(x = 0.5, y = -0.06)),
        utils::modifyList(annot_ylab, list(text = "People", x = -0.085, y = ylab_pos)),
        utils::modifyList(annot_ylab, list(
          text = "People / Hour", x = 1.04,
          textangle = 90,
          y = ylab_pos
        ))
      )
    ) |>
    plotly::layout(annotations = list(
      utils::modifyList(annot_, list(x = x1, y = y1, text = titles[["A"]])),
      utils::modifyList(annot_, list(x = x2, y = y2, text = titles[["B"]])),
      utils::modifyList(annot_, list(x = x3, y = y3, text = titles[["C"]])),
      utils::modifyList(annot_, list(x = x4, y = y4, text = titles[["D"]])),
      utils::modifyList(annot_, list(x = 0.5, y = y5, text = titles[["E"]]))
    ))


  pl


  export_plot(pl, file.path(filepath_figs, "build_queue_sim.pdf"),
    font_family = font_family,
    width = 5, height = 6
  )


  fig_stock_flow_phases <- function() {
    # Time points for the four phases of stock change (rise and fall)
    t0 <- 0 # start (inflow == outflow, stock flat)
    t1 <- 2 # equal -> rise: inflow > outflow
    t2 <- 4
    t3 <- 6 # rise  -> fall: outflow > inflow
    t4 <- 8


    sfm <- stockflow() |>
      stock(Stock, eqn = 1) |>
      custom_func(func1, eqn = pulse(times, start = !!t1, width = !!t2 - !!t1, height = .5)) |>
      custom_func(func2, eqn = pulse(times, start = !!t3, width = !!t4 - !!t3, height = .5)) |>
      flow(Inflow, eqn = 1 + func1(t), to = Stock) |>
      flow(Outflow, eqn = 1 + func2(t), from = Stock) |>
      sim_settings(only_stocks = FALSE)

    sim <- simulate(sfm, stop = 10)
    df_stock <- as.data.frame(sim, vars = "Stock")
    df_in <- as.data.frame(sim, vars = "Inflow")
    df_out <- as.data.frame(sim, vars = "Outflow")
    df_change <- df_in
    df_change$value <- df_change$value - df_out$value


    size_marker <- 8
    line_width_thick <- 5

    # Stock value at each time point (robust to non-exact time matches)
    stock_at <- function(tt) df_stock$value[which.min(abs(df_stock$time - tt))]
    y0 <- stock_at(t0)
    y1 <- stock_at(t1)
    y2 <- stock_at(t2)
    y3 <- stock_at(t3)
    y4 <- stock_at(t4)

    # --- Top panel: stock with a change indicator per phase ----------------------
    fig1 <- plotly::plot_ly(df_stock,
      x = ~time, y = ~value, type = "scatter",
      show_legend = FALSE,
      mode = "lines", name = "Stock", line = list(color = stock_col, width = line_width_thick)
    )

    # Add arrows to indicate the direction of change in each phase
    arrowwidth <- 2
    fig1 <- fig1 |>
      plotly::layout(annotations = list(
        x = t2,
        y = y2,
        text = "",
        xref = "x",
        yref = "y",
        axref = "x",
        ayref = "y",
        showarrow = TRUE,
        arrowsize = 1.25,
        arrowhead = 2,
        arrowwidth = arrowwidth,
        ax = t2,
        ay = y1,
        arrowcolor = increase_col
      )) |>
      plotly::layout(annotations = list(
        x = t3,
        y = y4,
        text = "",
        xref = "x",
        yref = "y",
        axref = "x",
        ayref = "y",
        showarrow = TRUE,
        arrowsize = 1.25,
        arrowhead = 2,
        arrowwidth = arrowwidth,
        ax = t3,
        ay = y3,
        arrowcolor = decrease_col
      ))

    fig1


    # --- Bottom panel: inflow & outflow with the difference shaded ----------------
    fig2 <- plotly::plot_ly()

    # Net flow
    fig2 <- fig2 |>
      plotly::add_trace(
        data = df_change, x = ~time, y = ~value, type = "scatter", mode = "lines",
        show_legend = FALSE, fill = "tozeroy",
        fillcolor = ifelse(df_change$value >= 0, increase_fill, decrease_fill),
        name = "Net Flow", line = list(width = 0)
      ) |>
      plotly::add_trace(
        data = df_change, x = ~time, y = ~value, type = "scatter", mode = "lines",
        show_legend = FALSE,
        name = "Net Flow", line = list(color = flow_col, width = line_width_thick)
      )


    # --- Combine panels ----------------------------------------------------------
    sp_margin <- 0.12 # vertical gap between the two panels (paper)
    top_margin <- 80 # 40
    pad_axis <- .1
    pad_axis2 <- .25
    fig <- plotly::subplot(fig1, fig2,
      nrows = 2, shareX = FALSE, titleY = TRUE,
      heights = c(.5, .5),
      margin = sp_margin
    )

    # --- Phase labels above each panel -------------------------------------------
    annot_ <- utils::modifyList(annot, list(
      font = list(size = font_size_subplot - 2)
    ))
    top_y <- 1.0 # top labels sit just above the top panel
    bot_y <- (0.5 - sp_margin) + 0.01 # bottom labels sit just above bottom panel
    top_text_y <- c(bot_y + c(.265, .35, .58, .35, .265))
    bot_text_y <- c(bot_y - c(.195, .065, .195, .39, .195))
    top_angles <- c(0, -43, 0, 43, 0)
    bot_angles <- c(0, 0, 0, 0, 0)
    # bot_y    <- (0.5 - sp_margin) - 0.1

    # phase mid-points on the (shared) data x-axis, used to centre each label
    centers <- c((t0 + t1) / 2, (t1 + t2) / 2, (t2 + t3) / 2, (t3 + t4) / 2, (t4 + 10) / 2)
    centers_top <- centers
    centers_top[c(2, 4)] <- centers_top[c(2, 4)] + c(-.17, .17) # shift the bottom labels for the rise/fall phases
    top_text <- paste0("<i>", c("Static", "Increases", "Static", "Decreases", "Static"), "</i>")
    bot_text <- paste0("<i>", c(
      "Inflow ≡ Outflow", "Inflow > Outflow", "Inflow ≡ Outflow",
      "Inflow < Outflow", "Inflow ≡ Outflow"
    ), "</i>")

    phase_label <- function(text, x, y, angle = 0) {
      list(
        x = x, y = y, text = text, xref = "x", yref = "paper",
        textangle = angle,
        xanchor = "center", yanchor = "bottom", showarrow = FALSE, font = annot_
      )
    }


    label_anns <- c(
      lapply(seq_along(top_text), function(i) phase_label(top_text[i], centers_top[i], y = top_text_y[i], angle = top_angles[i])),
      lapply(seq_along(bot_text), function(i) phase_label(bot_text[i], centers[i], y = bot_text_y[i], angle = bot_angles[i]))
    )

    fig <- fig |>
      plotly::layout(
        # unname(): Map() returns a named list, which plotly misparses
        annotations = unname(c(
          label_anns
        )),
        margin = list(t = top_margin)
      ) |>
      plotly::layout(
        xaxis = list(title = "", showline = TRUE, zerolinewidth = 0, showgrid = FALSE),
        xaxis2 = list(
          title = "", showline = TRUE, showgrid = FALSE,
          zerolinecolor = "#ffff",
          zerolinewidth = 0
        ),
        yaxis = list(
          title = "Units", showgrid = FALSE,
          range = c(1 - pad_axis, 2 + pad_axis),
          tickvals = c(1, 2), ticktext = c("1", "2")
        ),
        yaxis2 = list(
          title = "Units / Time",
          zerolinewidth = 0, showgrid = FALSE,
          showline = TRUE,
          range = c(-.5 - pad_axis2, .5 + pad_axis2),
          tickvals = c(-.5, .5), ticktext = c("-0.5", "0.5")
        ),
        font = annot_font
      )


    # if (recreate_figs) {

    pl_stock_flow_phases <- fig

    # }


    # fig2 <- fig |>
    #   plotly::layout(
    #     annotations = list(
    #       utils::modifyList(annot_xlab, list(text = "<i>Timeseries of Stock</i>", x = 0.5, y = top_y + .13)),
    #       utils::modifyList(annot_xlab, list(
    #         text = "<i>Timeseries of Change in Stock</i>", x = 0.5,
    #         y = bot_y + .11
    #       )),
    #       phase_label("<i>Area = Increase\nin stock</i>",
    #         x = centers[2],
    #         y = bot_y - .2
    #       ),
    #       phase_label("<i>Area = Decrease\nin stock</i>", x = centers[4], y = bot_y - .325),
    #       utils::modifyList(annot_xlab, list(text = "Time", x = 0.5, y = -0.06))
    #     )
    #   )


    fig1 <- pl_stock_flow_phases |>
      plotly::layout(
        annotations = list(
          utils::modifyList(annot_xlab, list(text = "<i>A. Stock-and-Flow Diagram</i>", x = 0.5, y = top_y + .22)),
          utils::modifyList(annot_xlab, list(text = "<i>B. Timeseries of Stock</i>", x = 0.5, y = top_y + .13)),
          utils::modifyList(annot_xlab, list(
            text = "<i>C. Timeseries of Change in Stock</i>", x = 0.5,
            y = bot_y + .11
          )),
          phase_label("<i>Area = Increase\nin stock</i>",
            x = centers[2],
            y = bot_y - .2
          ),
          phase_label("<i>Area = Decrease\nin stock</i>", x = centers[4], y = bot_y - .325),
          utils::modifyList(annot_xlab, list(text = "Time", x = 0.5, y = -0.06))
        )
      )

    export_plot(fig1, file.path(filepath_figs, "build_stock_flow_phases.pdf"),
      font_family = font_family,
      width = 6, height = 4.2
    )
  }
}
```


```{r, include=FALSE,eval=interactive(), error = TRUE}
if (recreate_figs) {
  fig_sfm_inflow_outflow()
  fig_queue_sfm()
  fig_queue_sim()
  fig_stock_flow_phases()
}
```
