---
title: "Ensemble simulations"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Ensemble simulations}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 10,
  fig.height = 4,
  fig.align = "center", # Center all figures
  out.width = "80%" # Display figures at 80% of container width
  # fig.retina = 2,         # Optimize for retina displays
  # dpi = 96               # Standard DPI for web
)
```

```{r, echo=TRUE, message=FALSE}
library(sdbuildR)
```


```{r, echo = FALSE}
# Disable WebGL: many plotly widgets per HTML page can exceed the browser WebGL
# context limit and render blank. SVG always renders.
options(sdbuildR.webgl = FALSE)
```


```{r, echo = FALSE}
font_family <- "stix-two-text"
options(sdbuildR.font_family = font_family)
# getOption("sdbuildR.font_family")
```

```{r, echo = FALSE, eval = interactive()}
# # Check if the Julia environment is set up, and if not, install it (for local pkgdown::build_site())
# if (!nzchar(system.file("Manifest.toml", package = "sdbuildR"))) {
#   install_julia_env()
# }
```

After having built a stock-and-flow model, you may want to explore how different parameter values affect the model's behaviour. Running multiple simulations with varying parameters is also called an ensemble, which provides insight into the range of possible outcomes and uncertainty associated with your model. In this vignette, we will explore how to set up and run ensemble simulations using the sdbuildR package. 

## Setting up the model

For this example, we will use Crielaard et al.'s (2022) model of eating behaviour, including the stocks hunger, eating, and compensatory behaviour (i.e., disordered eating behaviour such as purging and overexercising). For more details, see Crielaard et al. (2022). We can load this example from the model library and look what is inside:

```{r}
sfm <- stockflow("crielaard2022")
print(sfm)
```

Simulations run in R by default. If you want to use Julia for faster execution or additional parallelization options, you can activate the Julia environment for sdbuildR and change the simulation language to Julia (see below). For guidance on installing and setting up the Julia environment, see [this vignette](https://kcevers.github.io/sdbuildR/articles/julia-setup.html).

Without changing the parameters, we can run a single simulation to see how the model behaves:

```{r include=FALSE}
# Add seed to ensure the difference between trajectories is immediately visible
sfm <- sim_settings(sfm, seed = 1)
```

```{r}
sfm |>
  simulate() |>
  plot()
```

As the model has random initial conditions, another run will be different:

```{r include=FALSE}
# Add seed to ensure the difference between trajectories is immediately visible
sfm <- sim_settings(sfm, seed = 3)
```

```{r}
sfm |>
  simulate() |>
  plot()
```

To explore this more systematically, we can run an ensemble simulation using the `ensemble()` function. 

## Running ensemble simulations

Ensemble simulations create multiple runs of the model, which only makes sense if the model either has some random elements or if parameters are being varied. Our model already has random initial conditions, but if it did not, we could create these:

```{r}
sfm <- update(sfm, c(Food_intake, Hunger, Compensatory_behaviour),
  eqn = runif(1)
)
```

With random initial conditions, multiple runs of the same model will be different. As running ensemble simulations can be quite memory intensive, it is highly recommended to reduce the size of the returned timeseries. This will save memory and speed up the simulation. For example, we can save only 50 evenly-spaced time points per run with `save_length`:

```{r}
sfm <- sim_settings(sfm, save_length = 50)
```

With random initial conditions and reduced output, the model is now ready for running ensemble simulations. We complete 100 runs using the `ensemble()` function:

```{r}
sims <- ensemble(sfm, n = 100)
```

```{r}
plot(sims)
```

The plot shows the mean and confidence interval of the stocks (mean with 95% confidence interval). We can also plot the individual runs, for which we first have to rerun the simulation with `save_sims = TRUE`:

```{r}
sfm <- sim_settings(sfm, save_sims = TRUE)
# or pass `save_sims = TRUE` directly to `ensemble()`:
sims <- ensemble(sfm, n = 30, save_sims = TRUE)
```


```{r message=FALSE, warning=FALSE}
plot(sims, which = "sims")
```

We can change which simulations we plot by specifying the `sim` argument:

```{r }
plot(sims, which = "sims", sim = 15:30)
```

### Selecting variables to save

By default, only the stocks are saved. We can save all variables by updating the simulation settings:

```{r}
sfm <- sim_settings(sfm, only_stocks = FALSE)
```

Alternatively, we can specify which variables to save with the `vars` argument. For example, to save only the stocks Hunger and Food_intake, we can run:

```{r}
sfm <- sim_settings(sfm, vars = c("Hunger", "Food_intake"))
```

```{r}
sims <- ensemble(sfm, n = 100)
```

```{r}
head(sims, direction = "wide", which = "sims", n = 1)
```

Resetting `vars` to `NULL` will save all variables again:

```{r}
sfm <- sim_settings(sfm, vars = NULL, only_stocks = TRUE)
```

### Selecting summary statistics

By default, `ensemble()` computes as summary statistics the mean and median, together with a 95% interval (the 2.5% and 97.5% quantiles), across runs at each time point. The quantile levels can be changed as follows:

```{r}
sims <- ensemble(sfm,
  n = 100,
  quantiles = c(0.1, 0.9)
)
head(sims)
```

Here, quantiles appear as columns `quant1`, `quant2` which correspond to `sims$quantiles`.

To change which summary statistics are computed, use `central` (defining the central tendency) and `spread` (defining the measure of dispersion). For example, to summarize each run by its median together with a standard deviation:

```{r}
sims <- ensemble(sfm,
  n = 100,
  central = "median",
  spread = "sd"
)
head(sims)
```


All available summary statistics can be computed with:

```{r}
sims <- ensemble(sfm,
  n = 100,
  central = c("mean", "median"),
  spread = c("quantile", "sd", "range")
)
head(sims, n = 1)
```

Plots can then switch between summary statistics:

```{r}
plot(sims, central = "median", spread = "quantile")
plot(sims, central = "median", spread = "range")
plot(sims, central = "none", spread = "range")
```



### Computational efficiency

#### Parallel simulations (R)

By default, R ensemble simulations run sequentially. To run simulations in parallel, use the `future` package to control parallel execution:

```{r, eval=FALSE}
# Set up parallel execution with 4 workers
future::plan(future::multisession, workers = 4)

# Run 100 simulations in parallel
sims <- ensemble(sfm, n = 100)

# Restore sequential execution
future::plan(future::sequential)
```

The `workers` argument specifies how many parallel processes to use; adjust this based on your system's capabilities. On Windows, `multisession` is the recommended backend. On POSIX systems (Linux, macOS), you can also use `multicore` for potentially better performance, though `multisession` will work on those systems as well. 

#### Ensemble simulations (Julia)

Ensemble simulations can also be run in Julia, which is typically faster than R for large simulations. To use the Julia backend, first follow the instructions in the [Julia setup vignette](https://kcevers.github.io/sdbuildR/articles/julia-setup.html) to install and set up the Julia environment for sdbuildR.

Activate the Julia environment for sdbuildR:

```{r}
use_julia()
```

Then, set the simulation language to Julia:

```{r}
sfm <- sim_settings(sfm, language = "julia")
```


We can also enable parallel execution in Julia by setting the number of threads:

```{r, eval=FALSE}
use_julia(nthreads = 4)
```

To stop using threaded simulations, run:

```{r, eval=FALSE}
use_julia(restart = TRUE)
```

## Specifying conditions

Instead of generating an ensemble with random initial conditions, we can also specify exact constant and initial stock values to vary. For example, we could vary the `a2` parameter, which determines how strongly having eaten increases compensatory behaviour.

```{r}
sims <- ensemble(sfm,
  n = 100,
  conditions = list(a2 = c(0.2, 0.4, 0.6, 0.8))
)
```


```{r, fig.height = 6}
plot(sims)
```


### Varying multiple parameters

We can also vary multiple parameters at once. For example, we can vary both `a2` and `a1`, where the latter influences how strongly food intake leads to more food intake. `n` now specifies the number of simulations per condition. By default, `cross = TRUE`, which means that all possible combinations of parameters are simulated.

```{r}
sims <- ensemble(sfm,
  conditions = list(
    a2 = c(0.2, 0.8),
    a1 = c(1.3, 1.5)
  ),
  n = 100
)
```

```{r, fig.height = 6}
plot(sims)
```

The plot shows similarity within columns but differences between columns. As `a1` differs between columns, it appears that `a1` has a larger effect than `a2`. 

To view the parameter combination corresponding to each condition, view `conditions` in `sims`: 

```{r}
sims$conditions
```

To explore the effect of each parameter interactively, use `condition_display = "slider"` or `condition_display = "dropdown"`:

```{r, fig.height = 6}
plot(sims, condition_display = "slider")
```

### Crossed vs non-crossed designs

To generate a non-crossed designed, set `cross = FALSE`. In this case, the length of each vector in `conditions` needs to be the same.

```{r}
sims <- ensemble(sfm,
  conditions = list(
    a2 = c(0.4, 0.5, 0.6),
    a1 = c(1.3, 1.4, 1.5)
  ),
  n = 100, cross = FALSE, save_sims = TRUE
)
```

```{r}
plot(sims, nrows = 1)
```

We can select specific conditions to compare, where here we plot the first fifteen simulations of the first two conditions:

```{r}
plot(sims, sim = 1:15, condition = 1:2, which = "sims", nrows = 1)
```

### Accessing simulation results

The results of the ensemble simulation are stored in the `sims` object.
 <!-- which is a list containing, among others:
- `summary`: summary statistics across all simulations per condition
- `df`: individual simulation data (if `save_sims = TRUE`)
- `init`: initial values of stocks
- `constants`: parameter values used
- `conditions`: matrix showing parameter combinations for each condition -->

You can access the summary statistics per condition and per time point, such as the mean and confidence intervals, using:

```{r}
head(sims, n = 1)
# or
# sims |> as.data.frame() |> head()
```

By default, simulations are returned in long format, but can also be shaped in wide format as well: 

```{r}
head(sims, n = 1, direction = "wide")
```

With `save_sims = TRUE`, the individual simulation runs can be accessed by specifying `which = "sims"`. The dataframe contains the value of each variable, for each time point, for each simulation, for each condition. 

```{r}
head(sims, n = 1, which = "sims", direction = "long")
```


Finally, to access the parameters (i.e., constants) of each simulation per condition, run:

```{r}
head(sims, n = 1, which = "sims", type = "constant")
```

To view their summary statistics, run:

```{r}
head(sims, n = 1, which = "summary", type = "constant")
```

## Close Julia session

```{r}
use_julia(stop = TRUE)
```

