---
title: "Importing and Exporting Models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Importing and Exporting Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(sdbuildR)
```

sdbuildR can exchange models with several external tools. This vignette covers
three workflows:

1. **Importing** from [Insight Maker](https://insightmaker.com/)
2. **Importing** from [deSolve](https://cran.r-project.org/package=deSolve)
3. **Exporting** to deSolve or [Psychomodels](https://www.psychomodels.org/)

---

## Importing from Insight Maker

[Insight Maker](https://insightmaker.com/) is a free, browser-based system
dynamics tool with a large library of public models. `import_insightmaker()`
converts any public stock-and-flow model into an `sdbuildR` object.

### From a URL

Pass the URL of any public Insight Maker model:

```{r eval=FALSE}
sfm <- import_insightmaker(
  URL = "https://insightmaker.com/insight/43tz1nvUgbIiIOGSGtzIzj/Romeo-Juliet"
)
```

### From a downloaded file

Models can be downloaded as `.InsightMaker` files (go to
**Share → Export → Download Insight Maker file**) or as ModelJSON (`.json`)
files:

```{r eval=FALSE}
sfm <- import_insightmaker(file = "my_model.InsightMaker")
sfm <- import_insightmaker(file = "my_model.json")
```

After importing, use the model as usual:

```{r eval=FALSE}
summary(sfm)
sim <- simulate(sfm)
plot(sim)
```

---

## Importing from deSolve

[deSolve](https://cran.r-project.org/package=deSolve) is the standard R
package for solving ODEs. If you have an existing deSolve model you can import
it into sdbuildR to access features like ensemble simulations, stock-and-flow
diagrams, and export to Psychomodels.

### Supported pattern

`import_desolve()` parses the canonical deSolve model convention:

```r
model <- function(t, state, parameters) {
  with(as.list(c(state, parameters)), {
    # intermediate calculations become auxiliary variables
    aux_var <- <expression>
    # d<VarName> assignments become net flows for stock VarName
    dVarName <- <expression>
    list(c(dVarName, ...))
  })
}
```

**Requirements:**

- The function must have exactly the arguments `(t, state, parameters)`.
- Variables must be accessed by name inside a `with(as.list(c(state, parameters)), {...})` block.
- Derivatives must be written as `d<VarName> <- <expression>` where `VarName`
  matches a name in `init`.
- `times` must be evenly spaced — use `seq()`.

### Example: SIR epidemic model

```{r}
sir_model <- function(t, state, parameters) {
  with(as.list(c(state, parameters)), {
    SI <- beta * S * I / N # intermediate calculation → aux variable
    IR <- gamma * I
    dS <- -SI
    dI <- SI - IR
    dR <- IR
    list(c(dS, dI, dR))
  })
}

sfm <- import_desolve(
  model  = sir_model,
  params = c(beta = 0.3, gamma = 0.1, N = 1000),
  init   = c(S = 990, I = 10, R = 0),
  times  = seq(0, 100, by = 0.1),
  name   = "SIR epidemic"
)

summary(sfm)
```

```{r}
sfm_plot <- sim_specs(sfm, save_at = 5)
sim <- simulate(sfm_plot)
plot(sim)
```

The imported model is a full `sdbuildR` object. You can modify it with
`update()`, add unit tests with `unit_test()`, run ensembles, and export it to
any supported format.

---

## Exporting models

`export_model()` converts an `sdbuildR` object to another format. The `format`
argument is case-insensitive, so `"deSolve"` and `"desolve"` are equivalent.

```r
export_model(sfm, format = "desolve")
export_model(sfm, format = "psychomodels")
```

Use `file = "path"` to write directly to disk; omit `file` to return the
result in memory.

---

### Export to deSolve

`format = "desolve"` produces a ready-to-run deSolve model.

```{r}
sfm <- sdbuildR("SIR")
ds <- export_model(sfm, format = "desolve")
```

When `file = NULL`, the return value is a named list:

| Element  | Description                                  |
|----------|----------------------------------------------|
| `model`  | R function in deSolve `function(t, state, parameters)` format |
| `params` | Named numeric vector of constants            |
| `init`   | Named numeric vector of initial stock values |
| `times`  | Numeric time sequence                        |
| `method` | Integration method (character)               |

```{r}
str(ds[c("params", "init", "times", "method")])
```

You can run the exported model directly with deSolve:

```{r}
library(deSolve)
out <- deSolve::ode(
  y = ds$init, times = ds$times,
  func = ds$model, parms = ds$params,
  method = ds$method
)
head(out)
```

Or write the full script to a file for sharing (`.R` extension is appended
automatically if not present):

```{r eval=FALSE}
export_model(sfm, format = "desolve", file = "sir_desolve.R")
```

#### Round-trip: deSolve → sdbuildR → deSolve

Models can travel in both directions:

```{r}
sir_fn <- function(t, state, parameters) {
  with(as.list(c(state, parameters)), {
    dS <- -beta * S * I / N
    dI <- beta * S * I / N - gamma * I
    dR <- gamma * I
    list(c(dS, dI, dR))
  })
}

sfm2 <- import_desolve(
  model  = sir_fn,
  params = c(beta = 0.3, gamma = 0.1, N = 1000),
  init   = c(S = 990, I = 10, R = 0),
  times  = seq(0, 100, by = 0.1)
)

ds2 <- export_model(sfm2, format = "desolve")
# ds2$model is ready for deSolve::ode()
```

---

### Export to Psychomodels

[Psychomodels](https://www.psychomodels.org/) is a repository of computational
models in psychology. `format = "psychomodels"` generates the JSON record
expected by the platform. When `file = NULL`, returns a JSON character string;
when `file` is provided, writes a `.json` file (extension appended if absent).

```{r eval=FALSE}
json <- export_model(
  sfm,
  format          = "psychomodels",
  publication_doi = "10.0000/my-paper",
  include_latex   = TRUE # appends LaTeX equations to the explanation field
)
```

Write directly to a `.json` file for upload:

```{r eval=FALSE}
export_model(sfm,
  format = "psychomodels",
  publication_doi = "10.0000/my-paper",
  file = "sir_psychomodels.json"
)
```

Key arguments (all optional):

| Argument              | Description |
|-----------------------|-------------|
| `title`               | Model title (defaults to `sfm$meta$name`) |
| `description`         | Short description |
| `explanation`         | Long-form explanation |
| `publication_doi`     | DOI of the linked publication |
| `include_latex`       | Append LaTeX equations (`TRUE`/`FALSE`) |
| `programming_language`| Programming language string |
| `psychology_discipline`| Discipline id(s) as comma-separated string |
| `file`                | File path; if `NULL`, returns JSON string |
