---
title: "Formalizing Job Demands-Resources Theory"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Formalizing Job Demands-Resources Theory}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  include = TRUE,
  eval = TRUE,
  comment = "#>",
  fig.width = 12, # 1
  fig.height = 4, # 3
  fig.align = "center", # Center all figures
  out.width = "80%" # Display figures at 80% of container width
)
```

This is the online supplemental material accompanying the paper *Formalizing Psychological Theories with sdbuildR: A System Dynamics Modelling Tutorial in R* by Evers et al. (submitted). It contains the complete code to build, test, and simulate a formal model of Job Demands-Resources (JD-R) theory. To reproduce the figures in the paper, please see the corresponding .Rmd file.

To follow the entire tutorial, it is necessary to install Julia as well as set up the Julia environment. See [this vignette](https://kcevers.github.io/sdbuildR/articles/julia-setup.html) for guidance.

```{r libraries, results='hide', message=FALSE}
library(sdbuildR)
```

```{r eval = TRUE, include=FALSE}
# Set this to TRUE to recreate the figures in the paper
recreate_figs <- TRUE # FALSE # TRUE
```

```{r include=FALSE}
# Helper function
sprintf_arg <- function(template, values) {
  for (nm in names(values)) {
    template <- gsub(paste0("%\\(", nm, "\\)s"), values[[nm]], template)
  }
  return(template)
}
font_family <- "Times New Roman"

if (recreate_figs) {
  # Plotting parameters
  filepath_base <- dirname(rstudioapi::getActiveDocumentContext()$path)
  filepath_figs <- file.path(filepath_base, "figs_jdr")

  dir.create(filepath_figs)

  VARS <- c("E", "P", "X", "D", "R", "r_X_D")
  PALETTE <- c("Geyser", "Temps", "Spectral", "Zissou 1", "Roma")[5]
  COLORS <- grDevices::hcl.colors(n = 50, palette = PALETTE)

  COLORS <- COLORS[c(36, 9, 30, 1, 47, 17)]
  stock_col <- adjustcolor(COLORS[1], alpha.f = .75)
  flow_col <- COLORS[6]

  health_path <- adjustcolor(COLORS[4], alpha.f = .75)
  motivation_path <- adjustcolor(COLORS[1], alpha.f = .85)
  health_path_fill <- adjustcolor(COLORS[4], alpha.f = 1)
  motivation_path_fill <- adjustcolor(COLORS[1], alpha.f = 1)
  performance_path <- adjustcolor(COLORS[2], alpha.f = .75)
  performance_path_fill <- adjustcolor(COLORS[2], alpha.f = 1)

  font_size <- 16
  font_size_xlab <- round(font_size * 1.4)
  font_size_legend <- round(font_size * 1.1)
  font_size_subplot <- round(font_size * 1.5)
  xlab <- "Time (days)"

  # Finer-grained simulation
  dt <- .1
}
```

```{r}
{
viz_str <- "
digraph flowchart {

  graph [layout = dot,
         rankdir = TB,
         bgcolor = white,
         nodesep = 0.45,
         ranksep = 0.5,
         splines = ortho,
         fontname = '%(font_family)s',
         pad = '0.2']

  node [fontsize = 9.5,
        fontname = '%(font_family)s',
        fontcolor = 'white',
        penwidth = 0]

  edge [fontname = '%(font_family)s',
        fontsize = 8.5,
        fontcolor = '#555555',
        color = '#666666',
        penwidth = 1.3,
        arrowsize = 0.6]

  # --- DECISION NODES (rounded boxes with border) ---
  q1 [label = 'Can the variable be represented\\nas a continuous quantity that\\ncan increase and decrease?',
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.15,0.1']

  q2 [label = 'Does the variable change within\\nthe chosen time horizon?',
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.15,0.1']

  q3 [label = 'Is the variable a behaviour, action,\\nor process that takes place over time?',
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.15,0.1']

  q4 [label = 'Does the variable accumulate\\nthe effects of flows?',
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.15,0.1']

  q5 [label = 'Does the variable\\'s effect\\naccumulate in a stock?', # Does it directly increase or decrease a stock?
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.15,0.1']

  q6 [label = 'Should it?',
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.12,0.08']

  # --- TERMINAL NODES (no border, darker fill) ---
  constant [label = 'Constant',
            shape = box, style = 'filled,rounded',
            fillcolor = '#0f5c5a', color = 'none',
            fontsize = 10, margin = '0.15,0.08']

  stock [label = 'Stock',
         shape = box, style = 'filled,rounded',
         fillcolor = '#0f5c5a', color = 'none',
         fontsize = 10, margin = '0.15,0.08']

  flow [label = 'Flow',
        shape = box, style = 'filled,rounded',
        fillcolor = '#0f5c5a', color = 'none',
        fontsize = 10, margin = '0.15,0.08']

  auxiliary [label = 'Auxiliary',
             shape = box, style = 'filled,rounded',
             fillcolor = '#0f5c5a', color = 'none',
             fontsize = 10, margin = '0.15,0.08']

  # --- ACTION NODES ---
  reformulate [label = 'Reformulate variable',
               shape = box, style = 'dashed,rounded,filled',
            fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
            fontsize = 8, penwidth = 1.2, color = '#17807c',
            margin = '0.1,0.06']

  add_flow [label = 'Add missing\\nflows',
            shape = box, style = 'dashed,rounded,filled',
            fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
            fontsize = 8, penwidth = 1.2, color = '#17807c',
            margin = '0.1,0.06']

  add_stock [label = 'Add missing\\nstock',
             shape = box, style = 'dashed,rounded,filled',
             fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
             fontsize = 8, penwidth = 1.2, color = '#17807c',
             margin = '0.1,0.06']

  # --- EDGES ---
  q1 -> reformulate [xlabel = 'No', minlen = 2]
  reformulate -> q1 [minlen = 2, style = dashed, penwidth = 1]
  q1 -> q2          [xlabel = 'Yes']

  q2 -> constant [xlabel = 'No', minlen = 2]
  q2 -> q3       [xlabel = 'Yes']

  q3 -> q4 [xlabel = <<font>No, it can be captured<br/>at a single moment in time</font>>]
  q3 -> q5 [xlabel = '  Yes']

  q4 -> stock [xlabel = 'Yes']
  q4 -> q6    [xlabel = 'No']

  q5 -> flow  [xlabel = 'Yes']
  q5 -> q6    [xlabel = 'No']

  q6 -> auxiliary [xlabel = 'No']
  q6 -> add_flow  [xlabel = 'Yes', style = dashed, color = '#2a9d8f', penwidth = 1]
  q6 -> add_stock [xlabel = 'Yes', style = dashed, color = '#2a9d8f', penwidth = 1]
  add_flow  -> stock [style = dashed, color = '#2a9d8f', penwidth = 1]
  add_stock -> flow  [style = dashed, color = '#2a9d8f', penwidth = 1]

  # --- RANK CONSTRAINTS ---
  { rank = same; q1; reformulate }
  { rank = same; q2; constant }
  { rank = same; q4; q5 }
  { rank = same; stock; q6; flow }
  { rank = same; add_flow; auxiliary; add_stock }
}
"
}

viz_str <- sprintf_arg(
  viz_str,
  list(
    font_family = font_family
  )
)

pl <- DiagrammeR::grViz(viz_str)
pl
```




## full
```{r}
sfm <- sdbuildR() |>
  meta(name = "Burnout") |>
  sim_specs(stop = 16, time_units = "weeks", only_stocks = FALSE) |>
  stock(energy, eqn = .3, label = "Energy") |>
  flow(working, 
  eqn = .3 * energy,
  # eqn = .4 * energy / (.2 + energy),
  from = energy) |>
  stock(recovery_rate, eqn = 1, label = "Recovery Rate") |>
  flow(erosion, 
  eqn = recovery_rate * working, 
  # eqn = recovery_rate * working / (1 + energy), 
  from = recovery_rate, label = "Erosion") |>
  flow(recovery,
  #  eqn = recovery_rate * energy, 
   eqn = recovery_rate, 
   to = energy) 

sim <- simulate(sfm)
plot(sim)

```


## step by step
```{r}
sfm <- sdbuildR() |>
  meta(name = "Burnout") |>
  sim_specs(stop = 16, time_units = "weeks", only_stocks = FALSE)
```

```{r}
sfm <- sfm |>
  stock(energy, eqn = .3, label = "Energy")

sim <- simulate(sfm)
plot(sim)
```

```{r}
sfm <- sfm |>
  constant(work_rate, eqn = .3)
```

```{r}
sfm <- sfm |>
  flow(working, eqn = work_rate,
  from = energy, label = "Depletion from work")

sim <- simulate(sfm)
plot(sim)
```

```{r}
sfm <- sfm |>
  update(working, eqn = work_rate * energy)

sim <- simulate(sfm)
plot(sim)
```

```{r}
sfm <- sfm |>
  constant(recovery_rate, eqn = .1) |>
  flow(recovery, eqn = recovery_rate, to = energy, label = "Recovery")

sim <- simulate(sfm)
plot(sim)
```


What if the ability to recover is not static, but erodes over time?

```{r}
sfm <- sfm |>
  change_type(recovery_rate, new_type = "stock") 
```


```{r}
sfm <- sfm |>
  update(recovery_rate, eqn = 1) |>
  flow(erosion, eqn = recovery_rate * working, 
    # eqn = capacity * working / (1 + energy), 
  from = recovery_rate, label = "Erosion")

sim <- simulate(sfm)
plot(sim)
```

Counterintuitively, as restoration capacity erodes, energy first increases. This is because recovery first still exceeds energy depletion from work. 


```{r include=FALSE}
if (recreate_figs) {
  sfm <- sdbuildR() |>
    stock(Stock, label = "Stock A", eqn = Constant) |>
    constant(Constant, eqn = 10) |>
    auxiliary(Auxiliary, eqn = Constant * 2) |>
    flow(Inflow, to = Stock, eqn = Auxiliary) |>
    flow(Outflow, from = Stock, eqn = Stock) |>
    stock(StockB, label = "Stock B") |>
    flow(InflowB,
      from = Stock, to = StockB, eqn = Auxiliary,
      label = "Outflow"
    )

  pl <- plot(sfm,
    wrap_width = 9.5,
    font_size = 20, minlen = 1.5,
    stock_col = stock_col,
    flow_col = flow_col,
    font_family = font_family,
    show_constants = TRUE, show_aux = TRUE, show_dependencies = TRUE
  )
  pl

  export_plot(pl, file.path(filepath_figs, "sfm.pdf"))
}
```


```{r include=FALSE}
if (recreate_figs) {

  {
    viz_str <- "
  digraph JDR_Model {

    # Graph attributes
    graph [layout = neato,
           rankdir = TB,
           bgcolor = white,
           overlap = true,
           outputorder='nodesfirst']

    # Default node attributes
    node [shape = circle,
          style = 'filled',
          fontcolor = 'white',
          color = none,
          font_family = '%(font_family)s',
          width = 1.4,
          height = 1.4, fixedsize = true,
          fontsize = 16,
          penwidth = 1.5]

    edge [fontsize = 22, labeldistance = 3.35]

    # Motivation path
    'proactive_behaviour' [label = 'Proactive\\nBehaviour', fillcolor = '%(motivation_path)s', pos = '0,2!']
    'node1' [label = '', fillcolor = '%(motivation_path)s', pos = '1.8,2!', width = .0001, height = .0001]
    'node2' [label = '', fillcolor = '%(health_path)s', pos = '4.2,2!', width = .0001, height = .0001]
    'resources' [label = 'Job\\nResources', fillcolor = '%(motivation_path)s', pos = '1.8,3.5!']
    'engagement' [label = 'Work\\nEngagement', fillcolor = '%(motivation_path)s', pos = '1.8,0.5!']
    'performance' [label = 'Job\\nPerformance', fillcolor = '%(performance_path)s', pos = '3,-1.45!']

    # Health deterioriation path
    'demands' [label = 'Job\\nDemands', fillcolor = '%(health_path)s', pos = '4.2,3.5!']
    'exhaustion' [label = 'Exhaustion', fillcolor = '%(health_path)s', pos = '4.2,0.5!']
    'undermining' [label = 'Self-\\nundermining', fillcolor = '%(health_path)s', pos = '6,2!']

    # Edges
    'proactive_behaviour' -> 'resources' [color = '%(motivation_path)s', fillcolor = '%(motivation_path_fill)s', penwidth = 2, headlabel = '+', fontcolor = '%(motivation_path)s', labeldistance = 4.4]
      'resources' -> 'node1' [arrowhead=none, color = '%(motivation_path)s', fillcolor = '%(motivation_path_fill)s', penwidth = 2, headlabel = '', fontcolor = '%(motivation_path)s']
      'node1' -> 'engagement' [color = '%(motivation_path)s', fillcolor = '%(motivation_path_fill)s', penwidth = 2, taillabel = '+', fontcolor = '%(motivation_path)s', labeldistance = 3.6]
    'engagement' -> 'performance' [color = '%(motivation_path)s', fillcolor = '%(motivation_path_fill)s', penwidth = 2, taillabel = '+', fontcolor = '%(motivation_path)s', labeldistance = 3]
    'engagement' -> 'proactive_behaviour' [color = '%(motivation_path)s', fillcolor = '%(motivation_path_fill)s', penwidth = 2, headlabel = '+', fontcolor = '%(motivation_path)s', labeldistance = 4.25]
    'proactive_behaviour' -> 'demands' [color = '%(motivation_path)s', fillcolor = '%(motivation_path_fill)s', penwidth = 2, headlabel = '−', fontcolor = '%(motivation_path)s', tailport='e', labeldistance = 3]

    'resources' -> 'node2' [color = '%(motivation_path)s', fillcolor = '%(motivation_path_fill)s', penwidth = 2, headlabel = '−', fontcolor = '%(motivation_path)s', labeldistance = 3]
    'demands' -> 'node1' [color = '%(health_path)s', fillcolor = '%(health_path_fill)s', penwidth = 2, headlabel = '+', fontcolor = '%(health_path)s']
    'demands' -> 'node2' [arrowhead=none, color = '%(health_path)s', fillcolor = '%(health_path_fill)s', penwidth = 2, headlabel = '', fontcolor = '%(health_path)s']
    'node2' -> 'exhaustion' [color = '%(health_path)s', fillcolor = '%(health_path_fill)s', penwidth = 2, headlabel = '+', fontcolor = '%(health_path)s', labeldistance = 2.9]
    'undermining' -> 'demands' [color = '%(health_path)s', fillcolor = '%(health_path_fill)s', penwidth = 2, taillabel = '+', fontcolor = '%(health_path)s', labeldistance = 3.75]

    'exhaustion' -> 'undermining' [color = '%(health_path)s', fillcolor = '%(health_path_fill)s', penwidth = 2, labeldistance = 3.75, taillabel = '+', fontcolor = '%(health_path)s']
    'exhaustion' -> 'engagement' [color = '%(health_path)s', fillcolor = '%(health_path_fill)s', penwidth = 2, labeldistance = 3.75, taillabel = '−', fontcolor = '%(health_path)s']
    'exhaustion' -> 'performance' [color = '%(health_path)s', fillcolor = '%(health_path_fill)s', penwidth = 2, headlabel = '−', fontcolor = '%(health_path)s', labeldistance = 3.75]

  }
  "
  }
  viz_str <- sprintf_arg(
    viz_str,
    list(
      health_path = health_path,
      motivation_path = motivation_path,
      health_path_fill = health_path_fill,
      motivation_path_fill = motivation_path_fill,
      performance_path = performance_path,
      performance_path_fill = performance_path_fill,
      font_family = font_family # "Georgia"
    )
  )


  pl <- DiagrammeR::grViz(viz_str)
  pl

  export_plot(pl, file.path(filepath_figs, "JDR.pdf"))
}
```


```{r} 
sfm <- sdbuildR() |>
stock(c("E", "R", "D", "X"), label = c("Engagement", "Job Resources", "Job Demands", "Energy")) |>
flow("E_R", to = "E", label = "Motivation", eqn = R * D) |>#  "Boost hypothesis: Job demands amplify the impact of job resources on work engagement (and more generally on work motivation)."
flow("A_to_R", to = "R", label = "Proactive Behaviour", eqn = E) |> # "Engaged employees are intrinsically motivated to stay engaged and therefore start to proactively optimize their job (increase resources and optimize demands)"
flow("A_from_D", from = "D", label = "Proactive Behaviour", eqn = E) |> # "Engaged employees are intrinsically motivated to stay engaged and therefore start to proactively optimize their job (increase resources and optimize demands)"
flow("to_D", to = "D", label = "Undermining", eq = D - X) |> # "Proposition 8: Job demands and strain may lead to maladaptive self-regulation cognitions and behaviors (self-undermining)"
flow("X_X", from = "X", label = "Effort", eqn = D - R) |> # "Buffer hypothesis: Job resources weaken the impact of job demands on strain."
flow("E_X", from = "E", label = "Depletion", eqn = -X) |> 
aux(P, label = "Job Performance", eqn = E + X) 

plot(sfm, show_constants = TRUE, show_aux = TRUE, show_dependencies = TRUE)
```

## Simulating Stock-and-Flow Models in R


** why this value
b contrls slope
setting a = b gives same height

```{r}


a = .01; plot(x, a / (a + x), type = "l", col = "red", ylim = c(0, 1))
a = .05; lines(x, a / (a + x), type = "l", col = "orange")
a = .1; lines(x, a / (a + x), type = "l", col = "green")
a = .2; lines(x, a / (a + x), type = "l", col = "yellow")
a = .5; lines(x, a / (a + x), type = "l", col = "blue")
a = 1; lines(x, a / (a + x), type = "l", col = "purple")

```

```{r}
sfm <- sdbuildR() |>
  meta(name = "Job Resources and Demands Theory") |>
  # Add a seed for reproducibility:
  sim_specs(seed = 123) |>
  sim_specs(time_units = day, stop = 365 / 2, 
  only_stocks = FALSE) |>
  # Make save_at larger to reduce image file size
  sim_specs(save_at = .1, dt  =.01)
```

```{r, fig.width = 1, fig.height=1}
sfm <- sfm |>
  stock(E, .5, label = "Engagement") |>
  stock(R, .7, label = "Job Resources")

plot(sfm)
```

```{r eval=FALSE, include=FALSE}
if (recreate_figs) {
  {
    viz_str <- "digraph sfm {

      graph [layout = dot, rankdir = LR, center=true, outputorder='edgesfirst', pad=.1, nodesep= .3]

      # Rank groupings
  {rank=same; E; R}

     		'R' [label='Job Resources',tooltip = 'eqn = 0.7',shape=box,style=filled,fillcolor='#35B7ADBF',fontsize=18,fontname='Times New Roman']

		'E' [label='Engagement',tooltip = 'eqn = 0.5',shape=box,style=filled,fillcolor='#35B7ADBF',fontsize=18,fontname='Times New Roman']

    }"
  }
  pl <- DiagrammeR::grViz(viz_str)
  pl


  export_plot(pl, file.path(filepath_figs, "sfm_JDR_A.pdf"))
}
```

```{r build-jdr-a, fig.width = 1, fig.height = 3}
sim <- simulate(sfm)
plot(sim)
```

```{r include=FALSE}
if (recreate_figs) {
  sfm_list <- list()
  sfm_list[["A"]] <- sfm
}
```

```{r}
diagnose(sfm)
```

```{r build-jdr-b}
sfm <- sfm |>
  constant(r_E_R, 
  eqn = 1.25, 
  label = "Motivation Rate") |>
  flow(E_R, r_E_R * R, to = E, label = "Motivation")

sim <- simulate(sfm)
plot(sim)
```

```{r include=FALSE}
if (recreate_figs) {
  sfm_list[["B"]] <- sfm
}
```

```{r}

micmen <- function(x, a = 2, b = 1) {
 a * x/(b + x)
}

hill <- function(x, a = 2, b = 1) {
 x^a/(b + x^a)
}

curve(hyperbolic(x), from = 0, to = 5, ylim = c(0, 2))
curve(micmen(x), add = T, col = "blue")


```


```{r build-jdr-c}
sfm <- sfm |>
  flow(E_R,
  eqn = r_E_R * R * exp(-3 * E),
  #  eqn = r_E_R * R / (r_E_R + E)
   ) |>
    constant(r_E_X,
  eqn = 0.1,
  #  eqn = 0.2, 
   label = "Disengagement Rate") |>
   flow(from_E, eqn = r_E_X * E, from = E, label = "Decay")

sim <- simulate(sfm)
plot(sim)
```

```{r include=FALSE}
if (recreate_figs) {
  sfm_list[["C"]] <- sfm
}
```

```{r build-jdr-d}
sfm <- sfm |>
  constant(c(r_A, K_R),
    eqn = c(0.2, 1),
    label = c("Proactive Behaviour Rate", "Resource Capacity")
  ) |>
  flow(A_to_R,
    eqn = r_A * E * (1 - R / K_R), to = R,
    label = "Proactive Behaviour"
  )

sim <- simulate(sfm)
plot(sim)
```

```{r include=FALSE}
if (recreate_figs) {
  sfm_list[["D"]] <- sfm
}
```


<!-- **jdr specifies resources decay due to effort -->
<!-- ```{r build-jdr-e}
sfm <- sfm |>
  constant(r_R, eqn = 0.05, label = "Resource Decay Rate") |>
  flow(from_R, eqn = r_R * R, from = R, label = "Decay")

sim <- simulate(sfm)
plot(sim)
``` -->

```{r build-jdr-e}
sfm <- sfm |>
  stock(D, eqn = .2, label = "Job Demands") |>
  stock(X, eqn = .5, label = "Energy") |>
  aux(P, eqn = E + X, label = "Job Performance")

sim <- simulate(sfm)
plot(sim)
```


```{r include=FALSE}
if (recreate_figs) {
  sfm_list[["E"]] <- sfm
}
```

<!-- code not included in paper -->

** Michaelis-Menten function: f(x) = ax/(b + x)

```{r build-jdr-f}
sfm <- sfm |>
  constant(r_X_D,
    eqn = 0.4,
    label = "Fatigue from Demand Rate"
  ) |>
  constant(K_X, eqn = 1, label = "Energy Capacity") |>
  flow(X_D,
    eqn = r_X_D * X * D / (1 + R),
    from = X, label = "Effort", doc = "Buffer Hypothesis: Job resources weaken the impact of job demands on energy"
  ) |>
  flow(E_R,
    eqn = r_E_R * X * R * (1 + D) * (1 - E / K_E),
    doc = "Boost Hypothesis: Job demands amplify the impact of job resources on work engagement; made dependent on energy levels as engagement is partially defined as vigour"
  )


sim <- simulate(sfm, only_stocks = FALSE)

# Specify variables to plot
vars <- c("E", "D", "R", "X", "P")
plot(sim, vars = vars)
```

```{r include=FALSE}
if (recreate_figs) {
  sfm_list[["F"]] <- sfm
}
```

```{r build-jdr-g}
sfm <- sfm |>
  constant(r_R, eqn = 0.05, label = "Resource Decay Rate") |>
  flow(from_R, eqn = r_R * R * (1 - X/K_X), from = R, label = "Decay from low energy") |>

  constant(r_X_X, eqn = 0.15, label = "Restoration Rate") |>
  flow(X_X,
    eqn = r_X_X * X * (1 - X / K_X),
    to = X, label = "Restoration"
  ) |>
  constant(r_E_X, eqn = 0.1, label = "Disengagement Rate") |>
  flow(E_X,
    eqn = r_E_X * E * (1 - X / K_X), from = E,
    label = "Disengagement from Exhaustion"
  )

sim <- simulate(sfm, only_stocks = FALSE)
plot(sim, vars = vars)
```

```{r include=FALSE}
if (recreate_figs) {
  sfm_list[["G"]] <- sfm
}
```


```{r build-jdr-h}
sfm <- sfm |>
  constant(r_U, eqn = 0.15, label = "Self-undermining Rate") |>
  constant(K_D, eqn = 1, label = "Demand Capacity") |>
  flow(A_from_D,
    eqn = r_A * E * D, from = D,
    label = "Proactive Behaviour"
  ) |>
  flow(U,
    eqn = r_U * (1 - X / K_X), to = D,
    label = "Self-undermining"
  )

sim <- simulate(sfm, only_stocks = FALSE)
plot(sim, vars = vars)
```

```{r include=FALSE}
if (recreate_figs) {
  sfm_list[["H"]] <- sfm
}
```


```{r build-jdr-i}
sfm <- sfm |>
  constant(r_D, eqn = 0.2, label = "Demand Regulation Rate") |>
  flow(D_D,
    eqn = r_D * (1 - D / K_D), to = D,
    label = "Demand Regulation"
  )

sim <- simulate(sfm, only_stocks = FALSE)
plot(sim, vars = vars)
```

```{r include=FALSE}
if (recreate_figs) {
  sfm_list[["I"]] <- sfm
}
```

```{r}
plot(sfm, minlen = 1.5, font_size = 22)
```

```{r eval=FALSE, include=FALSE}
if (recreate_figs) {
  {
    viz_str <- "digraph sfm {

      graph [layout = dot, rankdir = LR, center=true, outputorder='edgesfirst', pad=.1, nodesep= .3]

      # Rank groupings
  {rank=same; X; P}


      # Define auxiliary nodes
      'P' [label='Job Performance',tooltip = 'eqn = E + X',shape=plaintext,fontsize=20,fontname='Times New Roman', width=0.6, height=0.3]

      # Define stock nodes
      'X' [label='Energy',tooltip = 'eqn = 0.5',shape=box,style=filled,fillcolor='#35B7ADBF',fontsize=22,fontname='Times New Roman']

		'E' [label='Engagement',tooltip = 'eqn = 0.5',shape=box,style=filled,fillcolor='#35B7ADBF',fontsize=22,fontname='Times New Roman']

		'R' [label='Job Resources',tooltip = 'eqn = 0.7',shape=box,style=filled,fillcolor='#35B7ADBF',fontsize=22,fontname='Times New Roman']

 		'D' [label='Job Demands',tooltip = 'eqn = 0.2',shape=box,style=filled,fillcolor='#35B7ADBF',fontsize=22,fontname='Times New Roman']


      # Define flow nodes (intermediate nodes for flows)
      'flow_E_R' [label='Motivation', tooltip = 'eqn = r_E_R * X * R * (1 + D) * (1 - E/K_E)', shape = plaintext, fontsize=20, fontname='Times New Roman', width=0.6, height=0.3]
		'flow_A_to_R' [label='Proactive Behaviour', tooltip = 'eqn = r_A * E * (K_R - R)/K_R', shape = plaintext, fontsize=20, fontname='Times New Roman', width=0.6, height=0.3]
		'flow_from_R' [label='Decay', tooltip = 'eqn = r_R * R', shape = plaintext, fontsize=20, fontname='Times New Roman', width=0.6, height=0.3]
		'flow_X_D' [label='Effort', tooltip = 'eqn = r_X_D * X * D / (1 + R)', shape = plaintext, fontsize=20, fontname='Times New Roman', width=0.6, height=0.3]
		'flow_X_X' [label='Restoration', tooltip = 'eqn = r_X_X * X * (1 - X/K_X)', shape = plaintext, fontsize=20, fontname='Times New Roman', width=0.6, height=0.3]
		'flow_E_X' [label='Disengagement from Exhaustion', tooltip = 'eqn = r_E_X * E * (1 - X/K_X)', shape = plaintext, fontsize=20, fontname='Times New Roman', width=0.6, height=0.3]
		'flow_A_from_D' [label='Proactive Behaviour', tooltip = 'eqn = r_A * E * D', shape = plaintext, fontsize=20, fontname='Times New Roman', width=0.6, height=0.3]
		'flow_U' [label='Self-undermining', tooltip = 'eqn = r_U * (1 - X/K_X)', shape = plaintext, fontsize=20, fontname='Times New Roman', width=0.6, height=0.3]
		'flow_D_D' [label='Demand Regulation', tooltip = 'eqn = r_D * (1 - D/K_D)', shape = plaintext, fontsize=20, fontname='Times New Roman', width=0.6, height=0.3]

      # Define external cloud nodes
      'Cloud1' [label='', tooltip = 'Unspecified source',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
		'Cloud2' [label='', tooltip = 'Unspecified source',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
		'Cloud3' [label='', tooltip = 'Unspecified source',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
		'Cloud4' [label='', tooltip = 'Unspecified source',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
		'Cloud5' [label='', tooltip = 'Unspecified sink',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
		'Cloud6' [label='', tooltip = 'Unspecified sink',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
		'Cloud7' [label='', tooltip = 'Unspecified sink',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
		'Cloud8' [label='', tooltip = 'Unspecified sink',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
		'Cloud9' [label='', tooltip = 'Unspecified sink',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]

      # Define flow edges (stock -> flow_node -> stock)
      'Cloud5' -> 'flow_E_R' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1.5, splines=false]
		'flow_E_R' -> 'E' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]
		'Cloud6' -> 'flow_A_to_R' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1.5, splines=false]
		'flow_A_to_R' -> 'R' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]
		'R' -> 'flow_from_R' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1.5, splines=false]
		'flow_from_R' -> 'Cloud1' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]
		'X' -> 'flow_X_D' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1.5, splines=false]
		'flow_X_D' -> 'Cloud2' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]
		'Cloud7' -> 'flow_X_X' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1.5, splines=false]
		'flow_X_X' -> 'X' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]
		'E' -> 'flow_E_X' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1.5, splines=false]
		'flow_E_X' -> 'Cloud3' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]
		'D' -> 'flow_A_from_D' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1.5, splines=false]
		'flow_A_from_D' -> 'Cloud4' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]
		'Cloud8' -> 'flow_U' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1.5, splines=false]
		'flow_U' -> 'D' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]
		'Cloud9' -> 'flow_D_D' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1.5, splines=false]
		'flow_D_D' -> 'D' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]

      # Define dependency edges
      'E' -> 'P' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
		'X' -> 'P' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
		'X' -> 'flow_E_R' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
		'R' -> 'flow_E_R' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
		'D' -> 'flow_E_R' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
		'E' -> 'flow_E_R' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
		'E' -> 'flow_A_to_R' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
		'R' -> 'flow_A_to_R' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
		'R' -> 'flow_R_X' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
		'X' -> 'flow_X_D' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
		'D' -> 'flow_X_D' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
		'R' -> 'flow_X_D' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
		'X' -> 'flow_X_X' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
		'E' -> 'flow_E_X' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
		'X' -> 'flow_E_X' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
		'E' -> 'flow_A_from_D' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
		'D' -> 'flow_A_from_D' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
		'X' -> 'flow_U' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
		'D' -> 'flow_D_D' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']


    }"
  }
  pl <- DiagrammeR::grViz(viz_str)
  pl

  export_plot(pl, file.path(filepath_figs, "sfm_JDR_I.pdf"))
}
```


```{r include=FALSE}
if (recreate_figs) {
  # Create subplot figure
  pl_list <- list()
  max_letter <- max(names(sfm_list))
  for (letter in names(sfm_list)) {
    sim <- simulate(sfm_list[[letter]] |> sim_specs(save_at = .01),
      only_stocks = FALSE
    )

    vars_in_sim <- intersect(vars, unique(sim$df$variable))

    vars_ <- VARS[VARS %in% vars_in_sim]
    colors_ <- COLORS[VARS %in% vars_in_sim]

    pl_list[[letter]] <- plot(sim,
      vars = vars_,
      colors = colors_,
      main = "",
      showlegend = (letter == max_letter)
    )
  }

  x1 <- 0.14
  x2 <- 0.5
  x3 <- .86
  y1 <- 1.05
  y2 <- 0.67
  y3 <- .2775

  pl <- plotly::subplot(pl_list,
    nrows = 3,
    shareX = TRUE,
    shareY = FALSE,
    titleY = FALSE,
    titleX = FALSE,
    margin = c(0, 0.05, 0.05, 0)
  ) |>
    plotly::layout(
      font = list(family = font_family, size = font_size),
      legend = list(
        orientation = "h", # orientation
        x = 0.5, # Center horizontally
        y = -0.16, # Below the plot
        xanchor = "center",
        yanchor = "top",
        font = list(size = font_size_legend)
      )
    ) |>
    plotly::layout(
      annotations = list(
        list(
          text = xlab,
          font = list(
            family = font_family,
            size = font_size_xlab
          ),
          xref = "paper",
          yref = "paper",
          xanchor = "center", yanchor = "top",
          x = 0.5, y = -.075,
          showarrow = FALSE
        )
      )
    ) |>
    plotly::layout(annotations = list(
      list(x = x1, y = y1, text = "A", size = font_size_subplot, showarrow = FALSE, xref = "paper", yref = "paper"),
      list(x = x2, y = y1, text = "B", size = font_size_subplot, showarrow = FALSE, xref = "paper", yref = "paper"),
      list(x = x3, y = y1, text = "C", size = font_size_subplot, showarrow = FALSE, xref = "paper", yref = "paper"),
      list(x = x1, y = y2, text = "D", size = font_size_subplot, showarrow = FALSE, xref = "paper", yref = "paper"),
      list(x = x2, y = y2, text = "E", size = font_size_subplot, showarrow = FALSE, xref = "paper", yref = "paper"),
      list(x = x3, y = y2, text = "F", size = font_size_subplot, showarrow = FALSE, xref = "paper", yref = "paper"),
      list(x = x1, y = y3, text = "G", size = font_size_subplot, showarrow = FALSE, xref = "paper", yref = "paper"),
      list(x = x2, y = y3, text = "H", size = font_size_subplot, showarrow = FALSE, xref = "paper", yref = "paper"),
      list(x = x3, y = y3, text = "I", size = font_size_subplot, showarrow = FALSE, xref = "paper", yref = "paper")
    ))

  pl

  export_plot(pl, file.path(filepath_figs, "build_JDR.pdf"),
    width = 5.1, height = 5
  )
}
```

We save this version of the stock-and-flow model as `sfm0` for later use:

```{r}
sfm0 <- sfm |> sim_specs(save_at = .1)
```

Note that this is identical to the version stored in the model library, which can be loaded using `sdbuildR()`:

```{r}
sfm0 <- sdbuildR("JDR") |> sim_specs(save_at = .1)
```


To assess the dependence of the burnout phenomenon on the model's initial conditions, we can randomly initialize the stocks and re-run the simulation. A more systematic approach will be covered in a later section.

```{r}
sfm <- sfm0 |>
  update(c(D, R, E, X), eqn = runif(1, 0.01, 1)) |>
  sim_specs(only_stocks = FALSE)

sim1 <- simulate(sfm, seed = 1)
pl1 <- plot(sim1, vars = vars)

sim2 <- simulate(sfm, seed = 2)
pl2 <- plot(sim2, vars = vars, showlegend = FALSE)

sim3 <- simulate(sfm, seed = 3)
pl3 <- plot(sim3, vars = vars, showlegend = FALSE)

sim4 <- simulate(sfm, seed = 4)
pl4 <- plot(sim4, vars = vars, showlegend = FALSE)

pl <- plotly::subplot(pl1, pl2, pl3, pl4, nrows = 2)
pl
```



## Interrogating Your Model

### Verification Tests

```{r}
sfm <- sfm0

# Job demands and energy should negatively correlate
sfm <- unit_test(sfm, expr = cor(D, X) < -.2)
verify(sfm)

# Job resources and energy should positive correlate
sfm <- unit_test(sfm, expr = cor(R, X) > .2)
verify(sfm)

# Job performance and work engagement should positively correlate
sfm <- unit_test(sfm, expr = cor(E, P) > .2)
verify(sfm)

# Job performance and energy should positively correlate
sfm <- unit_test(sfm, expr = cor(P, X) > .2)
verify(sfm)
```

```{r}
# Behaviours should always be positive
sfm <- unit_test(sfm, expr = all(U >= 0))
sfm <- unit_test(sfm, expr = all(A_to_R >= 0))
sfm <- unit_test(sfm, expr = all(A_from_D >= 0))
verify(sfm)
```


```{r}
# # All flows should be non-negative
# as.data.frame(sfm, type = "flows", properties = "name")
# sfm <- unit_test(sfm, expr = all(E_R >= 0))

```


With zero job demands, engagement, energy, we would expect job performance to be low.

```{r}
sfm <- unit_test(sfm, expr = mean(P) < 0.2, 
                conditions = list(D = 0, r_D = 0, r_U = 0))
verify(sfm)
```

Counterintuitively, when job demands are set to zero, energy rises, which leads to very high job performance. 

```{r}
v <- verify(sfm, return_sims = TRUE)
v$sim_index
plot(v$sims[[2]])
# sfm2 <- sfm0 |> update(c(D, r_D, r_U), eqn = 0)

# sim <- simulate(sfm2, only_stocks = FALSE)
# plot(sim, vars = vars)
```

Adjust model to make job performance dependent on demands:

```{r}
sfm <- sfm |> update(P, eqn = D * (E + X))
```

Rerun unit tests:

```{r}
verify(sfm)
```

Update saved model:

```{r}
sfm0 <- sfm0 |> update(P, eqn = D * (E + X))
```

We include some additional verification tests checks that are omitted from the paper for brevity. First, we would expect that when resources are zero, engagement decays to zero, which is indeed what we find.

```{r}
sfm <- unit_test(sfm, expr = tail(E, 1) < 0.01, 
  conditions = list(R = 0, r_A = 0))
verify(sfm)

sfm2 <- sfm0 |> update(c(R, r_A), eqn = 0)

sim <- simulate(sfm2, only_stocks = FALSE)
plot(sim, vars = vars)
```

When engagement is zero, job performance is not necessarily zero (see beginning peak).

```{r}
sfm <- unit_test(sfm, expr = all(head(P, 100) > 0.2), 
  conditions = list(E = 0, r_E_R = 0))
verify(sfm)

sfm2 <- sfm0 |> update(c(E, r_E_R), eqn = 0)

sim <- simulate(sfm2, only_stocks = FALSE)
plot(sim, vars = vars)
```

Similarly, when energy is set to zero, job performance is not necessarily zero (see beginning peak).

```{r}
sfm <- unit_test(sfm, expr = all(head(P, 100) > 0.2), 
  conditions = list(X = 0, r_E_X = 0), 
  label = "Performance should not be zero in the beginning when energy is zero")
verify(sfm)


# Set energy to zero
sfm2 <- sfm0 |> update(c(X, r_E_R), eqn = 0)

sim <- simulate(sfm2, only_stocks = FALSE)
plot(sim, vars = vars)
```

As an extreme condition test, we can initialize all stocks at zero. In this case, only job demands rises.

```{r}
sfm2 <- sfm0 |> update(c(E, R, X, D), eqn = 0)

sim <- simulate(sfm2, only_stocks = FALSE)
plot(sim, vars = vars)
```

As a robustness check, we initialize all stocks at high values, which the model is able to handle.

```{r}
sfm <- unit_test(sfm, expr = all(is.finite(c(E, R, X, D))), 
  conditions = list(E = 5, R = 5, X = 5, D = 5))
verify(sfm)

sfm2 <- sfm0 |> update(c(E, R, X, D),
  eqn = 5
)

sim <- simulate(sfm2, only_stocks = FALSE)
plot(sim, vars = vars)
```

<!-- 
### Interpretability

To add units, the Julia environment needs to be set up:

```{r, eval = FALSE}
use_julia()
```

Change the simulation engine to Julia:

```{r}
sfm <- sim_specs(sfm0, language = "Julia")
```

```{r}
sfm <- custom_unit(sfm, "UWES", doc = "Utrecht Work Engagement Scale") |>
  update(E, units = "UWES")
```

```{r}
sfm <- discard(sfm, "UWES") |>
  custom_unit("uE") |>
  update(E, units = "uE")
```

We can add similar placeholder units for all other stocks, as well as for performance:

```{r}
sfm <- custom_unit(sfm, c("uE", "uR", "uX", "uD", "uP"),
  doc = c(
    "Engagement", "Job Resources", "Energy",
    "Job Demands", "Job Performance"
  )
) |>
  update(c(R, X, D, P), units = c("uR", "uX", "uD", "uP"))
```

### Dimensional Consistency

To ensure dimensional consistency, we add units for all other variables in the model.

```{r, eval = FALSE}
sfm <- update(sfm, r_E_R, units = "uE/uR/day") |>
  update(E_R, units = "uE/day") |>
  update(K_E, units = "uE") |>
  update(c(r_A, K_R), units = c("uR/uE/day", "uR")) |>
  update(A_to_R, units = "uR/day") |>
  update(r_R, units = "1/day") |>
  update(R_X, units = "uR/day") |>
  update(r_X_D, units = "1/day") |>
  update(K_X, units = "uX") |>
  update(X_D, units = "uX/day") |>
  update(r_X_X, units = "1/day") |>
  update(X_X, units = "uX/day") |>
  update(r_E_X, units = "1/day") |>
  update(E_X, units = "uE/day") |>
  update(r_U, units = "uD/day") |>
  update(K_D, units = "uD") |>
  update(A_from_D, units = "uD/day") |>
  update(U, units = "uD/day") |>
  update(c(r_D, D_D), units = "uD/day")
```

Some equations need to be adjusted. In order to make job performance dimensionally consistent, two new constants are added:

```{r, eval = FALSE}
sfm <- sfm |>
  constant(sfm, r_P_E,
    eqn = 1,
    label = "Influence Engagement on Performance", units = "uP/(uE*uD)"
  ) |>
  constant(r_P_X,
    eqn = 1,
    label = "Influence Energy on Performance", units = "uP/(uX*uD)"
  ) |>
  update(P, eqn = D * (r_P_E * E + r_P_X * X))
```

In order to give $r_{X_D}$ the unit $\frac{1}{\text{day}}$, we need to add an additional factor, $r_{R_D}$ with the units $\frac{uR}{uD}$, indicating how many unit resources are needed per unit demand to buffer energy depletion.

```{r, eval = FALSE}
sfm <- sfm |> constant(r_R_D,
  eqn = 1,
  label = "Resources Needed per Unit Demand for Buffer", units = "uR/uD"
)
```

We use the unit function `u()` to add one to resources:

```{r, eval = FALSE}
sfm <- update(sfm, X_D, eqn = r_X_D * r_R_D * X * D / (u("1uR") + R))
```

Proactive behaviour now increases resources and decreases demands with the same constant $r_A$, creating dimensional inconsistency. We create a new constant for the outflow from job demands:

```{r, eval = FALSE}
sfm <- sfm |>
  constant(r_A_D,
    eqn = .1,
    label = "Proactive Behaviour Rate", units = "1/(day*uE)"
  ) |>
  flow(A_from_D, eqn = r_A_D * E * D)
```

We verify whether our model is dimensionally consistent by simulation:

```{r}
sim <- simulate(sfm, only_stocks = FALSE)
plot(sim, vars = vars)
```

Because units can slow down simulations, we only modify our original model with the missing variables identified by the dimensional consistency test:

```{r, eval = FALSE}
sfm0 <- sfm0 |>
  constant(r_P_E, eqn = 1, label = "Influence Engagement on Performance") |>
  constant(r_P_X, eqn = 1, label = "Influence Energy on Performance") |>
  update(P, eqn = D * (r_P_E * E + r_P_X * X)) |>
  constant(r_R_D,
    eqn = 1,
    label = "Resources Needed per Unit Demand for Buffer"
  ) |>
  update(X_D, eqn = r_X_D * r_R_D * X * D / (1 + R)) |>
  constant(r_A_D, eqn = .1, label = "Proactive Behaviour Rate") |>
  flow(A_from_D, eqn = r_A_D * E * D)

# Check that simulation runs without errors
sim <- simulate(sfm0)
```

 -->

### Robustness to Functional Form Differences

```{r}
# Grab parameter values from dataframe
df <- as.data.frame(sfm)
r_E_X <- as.numeric(df[df[["name"]] == "r_E_X", "eqn"])

  #  flow(from_E, eqn = r_E_X * E, from = E, label = "Decay")

# Range of engagement to evaluate functional form over
E <- seq(0, 1, by = .01)

# Create dataframe with two functional forms
df <- rbind(
  data.frame(
    E = E,
    # Add a minus sign as it is an outflow
    value = r_E_X * E,
    type = "Linear"
  ),
  data.frame(
    E = E,
    value = r_E_X * E / (r_E_X + E),
    type = ""
  )
)

# Pick two colours for linear and exponential functional form
colors <- grDevices::hcl.colors(n = 50, palette = "Roma")[c(20, 8)]

# Plot functional form
pl <- plotly::plot_ly() |>
  plotly::add_trace(
    data = df, x = ~E,
    y = ~value,
    colors = colors,
    color = ~type,
    legendgroup = ~type,
    type = "scatter",
    mode = "lines"
  ) |>
  plotly::add_trace(
    x = 0.2, y = -0.5, type = "scatter", mode = "markers",
    marker = list(size = 10, color = colors[2]),
    showlegend = FALSE
  ) |>
  plotly::layout(
    xaxis = list(title = "Work Engagement (E)"),
    yaxis = list(title = "Effect on Depletion Work Engagement (E)"),
    font = list(family = "Times New Roman")
  )

pl
``` 

Simulate two different models. 

```{r} 
sfm2 <- sfm0 |> update(from_E,
  eqn = r_E_X * E / (r_E_X + E)
)

sim1 <- simulate(sfm0, only_stocks = FALSE)
pl1 <- plot(sim1, 
# xlim = c(0, 100), 
vars = c("D", "E", "P"), showlegend = FALSE)

sim2 <- simulate(sfm2, only_stocks = FALSE)
pl2 <- plot(sim2, 
# xlim = c(0, 100), 
vars = c("D", "E", "P"))

pl <- plotly::subplot(pl1, pl2, nrows = 1) |>
  plotly::layout(title = "Linear versus Exponential Functional Form")
pl
```


```{r include=FALSE}
if (recreate_figs) {
  colors_ <- grDevices::hcl.colors(n = 50, palette = PALETTE)[c(24, 42)]
  scales::show_col(c(colors_, COLORS))

  pl_func <- plotly::plot_ly() |>
    plotly::add_trace(
      data = df,
      x = ~X,
      y = ~value,
      color = ~type,
      legendgroup = ~type,
      colors = colors_,
      type = "scatter",
      mode = "lines"
    ) |>
    plotly::add_trace(
      x = 0.2, y = -0.5, type = "scatter", mode = "markers",
      showlegend = FALSE,
      marker = list(
        # legendgroup = "Logistic",
        size = 10,
        color = colors_[2]
      )
    ) |>
    plotly::layout(
      title = "",
      xaxis = list(
        title = "",
        font = list(size = font_size)
      ),
      yaxis = list(
        title = "",
        font = list(size = font_size)
      ),
      font = list(family = font_family, size = font_size_xlab * 2)
    )
  pl_func

  show_vars <- c("D", "E", "X", "P")
  vars_ <- VARS[VARS %in% show_vars]
  colors_ <- COLORS[VARS %in% show_vars]

  pl1 <- plot(sim1,
    xlim = c(0, 65), vars = vars_, colors = colors_,
    showlegend = FALSE, main = "", xlab = ""
  )

  pl2 <- plot(sim2, xlim = c(0, 65), vars = vars_, colors = colors_, main = "", xlab = "")

  pl <- plotly::subplot(pl_func,
    plotly::subplot(pl1, pl2,
      nrows = 1,
      shareY = TRUE, shareX = TRUE
    ),
    shareX = FALSE,
    shareY = FALSE,
    margin = c(.1, .1, .1, .1),
    nrows = 2
  ) |>
    plotly::layout(
      # font = list(size = font_size_xlab),
      margin = list(t = 50, b = 50, l = 100, r = 50),
      legend = list(
        orientation = "h", # orientation
        x = 0.5, # Center horizontally
        y = -0.16, # Below the plot
        xanchor = "center",
        yanchor = "top",
        font = list(size = font_size_legend)
      )
    ) |>
    plotly::layout(
      annotations = list(
        list(
          text = xlab,
          font = list(
            size = font_size_xlab
          ),
          xref = "paper",
          yref = "paper",
          xanchor = "center", yanchor = "top",
          x = 0.5, y = -0.075,
          showarrow = FALSE
        )
      )
    ) |>
    plotly::layout(annotations = list(
      list(
        x = 0.5, y = .5, text = "Energy (X)",
        font = list(size = font_size_xlab),
        showarrow = FALSE, xref = "paper", yref = "paper"
      ),
      list(
        x = -0.28, y = 1.06, text = "Effect on\nWork Engagement (E)",
        textangle = 270,
        font = list(size = font_size_xlab),
        showarrow = FALSE, xref = "paper", yref = "paper"
      )
    ))
  pl


  export_plot(pl,
    file.path(filepath_figs, "functional_forms_JDR.pdf"),
    width = 4.3, height = 5
  )
}
```


### Uncertainty

We implement a Cox–Ingersoll–Ross model for job demands:

```{r}
sfm <- sfm0 |>
    constant(D_mean, eqn = 0.5) |>
    constant(D_sigma, eqn = 1) |>
    aux(D_deterministic, eqn = r_D * (D_mean - D)) |>
    aux(D_stochastic, eqn = D_sigma * sqrt(D) * rnorm(1)) |>
    update(D_D, eqn = D_deterministic + D_stochastic)

sim <- simulate(sfm, only_stocks = FALSE, seed = 123)
plot(sim, vars = vars)
```


```{r include=FALSE}
if (recreate_figs) {
  show_vars <- c("E", "R", "X", "D", "P")
  vars_ <- VARS[VARS %in% show_vars]
  colors_ <- COLORS[VARS %in% show_vars]

  pl <- plot(sim, vars = vars_, colors = colors_, xlab = "", main = "")

  pl <- pl |>
    plotly::layout(
      legend = list(
        orientation = "h", # orientation
        x = 0.5, # Center horizontally
        y = -0.16, # Below the plot
        xanchor = "center",
        yanchor = "top",
        font = list(size = font_size_legend)
      )
    ) |>
    plotly::layout(
      annotations = list(
        list(
          text = xlab,
          font = list(
            family = font_family,
            size = font_size_xlab
          ),
          xref = "paper",
          yref = "paper",
          xanchor = "center", yanchor = "top",
          x = 0.5, y = -0.075,
          showarrow = FALSE
        )
      )
    )
  pl

  export_plot(pl,
    file.path(filepath_figs, "noise_JDR.pdf"),
    width = 4, height = 4
  )
}
```

### Removing Variables

**undermining is crucial

```{r}
sim1 <- simulate(sfm0)
pl1 <- plot(sim1)

sfm2 <- discard(sfm0, U)
sim2 <- simulate(sfm2)
pl2 <- plot(sim2, showlegend = FALSE)

pl <- plotly::subplot(pl1, pl2, nrows = 1,
shareY = TRUE)
pl
```



### Challenging the Model Boundary

As job performance is defined by `D * (E + X)`, we update `E_R` by replacing the terms `D` and `X`.
We replace **

```{r, fig.height=3}
# Make performance feedback to the system
sfm2 <- sfm0 |>
  constant(r_P, eqn = 1, label = "Effect Job Performance on Work Engagement") |>
  # flow(motivation_feedback, eqn = r_P * P, to = E) 
  # update(E_R, eqn = (r_P * P + r_E_R * X * R * (1 + D)) * (K_E - E) / K_E) **
    # eqn = r_E_R * X * R * (1 + D) * (1 - E / K_E),
    # update(E_R, eqn = r_E_R * R * r_P * P * (1 - E / K_E))
    flow(P_to_E, eqn = r_P * P * exp(-3 * E), to = E)

sim1 <- simulate(sfm0, only_stocks = FALSE)
pl1 <- plot(sim1, vars = vars)

sim2 <- simulate(sfm2, only_stocks = FALSE)
pl2 <- plot(sim2, vars = vars, showlegend = FALSE)

pl <- plotly::subplot(pl1, pl2, nrows = 1, shareY = TRUE) |>
  plotly::layout(title = "Model Boundary")
pl
```



```{r include=FALSE}
if (recreate_figs) {
  {
    viz_str <- "
  digraph JDR_Model {

    # Graph attributes
    graph [layout = neato,
           rankdir = TB,
           bgcolor = white,
           overlap = true,
           outputorder='nodesfirst']

    # Default node attributes
    node [shape = circle,
          style = 'filled',
          fontcolor = 'white',
          color = none,
          font_family = '%(font_family)s',
          width = 1.4,
          height = 1.4, fixedsize = true,
          fontsize = 16,
          penwidth = 1.5]

    edge [fontsize = 22, labeldistance = 3.35]

    'engagement' [label = 'Work\\nEngagement', fillcolor = '%(motivation_path)s', pos = '0,0!']
    'performance' [label = 'Job\\nPerformance', fillcolor = '%(performance_path)s', pos = '3, 0!']

    # Edges
    'engagement' -> 'performance' [color = '%(motivation_path)s', fillcolor = '%(motivation_path_fill)s', penwidth = 2, headlabel = '+', fontcolor = '%(motivation_path)s', labeldistance = 3]

  }
  "
  }
  viz_str <- sprintf_arg(
    viz_str,
    list(
      health_path = health_path,
      motivation_path = motivation_path,
      health_path_fill = health_path_fill,
      motivation_path_fill = motivation_path_fill,
      performance_path = performance_path,
      performance_path_fill = performance_path_fill,
      font_family = font_family # "Georgia"
    )
  )


  pl <- DiagrammeR::grViz(viz_str)
  pl

  export_plot(pl, file.path(filepath_figs, "feedback_loop_A_JDR.pdf"))


  {
    viz_str <- "
  digraph JDR_Model {

    # Graph attributes
    graph [layout = neato,
           rankdir = TB,
           bgcolor = white,
           overlap = true,
           outputorder='nodesfirst']

    # Default node attributes
    node [shape = circle,
          style = 'filled',
          fontcolor = 'white',
          color = none,
          font_family = '%(font_family)s',
          width = 1.4,
          height = 1.4, fixedsize = true,
          fontsize = 16,
          penwidth = 1.5]

    edge [fontsize = 22, labeldistance = 3.35]

    'engagement' [label = 'Work\\nEngagement', fillcolor = '%(motivation_path)s', pos = '0,0!']
    'performance' [label = 'Job\\nPerformance', fillcolor = '%(performance_path)s', pos = '3, 0!']

    # Edges
    'engagement' -> 'performance' [color = '%(motivation_path)s', fillcolor = '%(motivation_path_fill)s', penwidth = 2, headlabel = '+', fontcolor = '%(motivation_path)s', labeldistance = 3]
    'performance' -> 'engagement' [color = '%(performance_path)s', fillcolor = '%(performance_path_fill)s', penwidth = 2, headlabel = '+', fontcolor = '%(performance_path)s', labeldistance = 3]

  }
  "
  }
  viz_str <- sprintf_arg(
    viz_str,
    list(
      health_path = health_path,
      motivation_path = motivation_path,
      health_path_fill = health_path_fill,
      motivation_path_fill = motivation_path_fill,
      performance_path = performance_path,
      performance_path_fill = performance_path_fill,
      font_family = font_family # "Georgia"
    )
  )


  pl <- DiagrammeR::grViz(viz_str)
  pl

  export_plot(pl, file.path(filepath_figs, "feedback_loop_B_JDR.pdf"))
}
```


```{r include=FALSE}
if (recreate_figs) {
  # Model boundary
  show_vars <- c("E", "R", "X", "D", "P")
  vars_ <- VARS[VARS %in% show_vars]
  colors_ <- COLORS[VARS %in% show_vars]

  sim1 <- simulate(sfm0 |> sim_specs(save_at = dt), only_stocks = FALSE)
  pl1 <- plot(sim1, vars = vars_, colors = colors_, main = "")

  sim2 <- simulate(sfm2 |> sim_specs(save_at = dt), only_stocks = FALSE)
  pl2 <- plot(sim2, vars = vars_, colors = colors_, main = "", showlegend = FALSE)


  pl_boundary <- plotly::subplot(pl1, pl2,
    nrows = 1,
    shareX = TRUE,
    shareY = TRUE,
    titleY = FALSE,
    titleX = FALSE,
    margin = c(0, 0.05, 0.05, 0)
  ) |>
    plotly::layout(
      margin = list(t = 150, b = 50, l = 50, r = 50),
      font = list(family = font_family, size = font_size),
      legend = list(
        orientation = "h", # orientation
        x = 0.5, # Center horizontally
        y = -0.3, # Below the plot
        xanchor = "center",
        yanchor = "top",
        font = list(size = font_size_legend)
      )
    ) |>
    plotly::layout(
      annotations = list(
        list(
          text = xlab,
          font = list(
            family = font_family,
            size = font_size_xlab
          ),
          xref = "paper",
          yref = "paper",
          xanchor = "center", yanchor = "top",
          x = 0.5, y = -.14,
          showarrow = FALSE
        )
      )
    )

  pl_boundary

  export_plot(pl_boundary,
    file.path(filepath_figs, "boundary_ts_JDR.pdf"),
    width = 4.8, height = 4
  )
}
```



## Learning From Your Model

If resources allow, set up threaded simulation:

```{r, eval = FALSE}
use_julia(nthreads = 4)
```

or: 

```{r}
future::plan(future::multisession, workers = 16)
```

If not possible, regular simulation can still be used for the analyses below, albeit with longer runtimes.

### Establishing Boundary Conditions of Phenomena

```{r, fig.height=4}
# Simulation study
sfm0 <- sfm0 |> sim_specs(save_n = 100, language = "R",
   seed = NULL) |>
  # Random initial conditions
  update(c(D, R, E, X), eqn = runif(1, 0.01, 1))
sfm <- sfm0

sims <- ensemble(sfm, n = 1000)
plot(sims, vars = c("E", "D"))
```

#### Parameter Variation

Vary $r_{X_D}$: the rate of energy depletion from demand.

```{r, fig.height = 8}
n <- 100
sims <- ensemble(sfm,
  n = n,
  # conditions = list(r_X_D = c(0.175, 0.1, 0.05)),
  # conditions = list(r_X_D = c(0.2, 0.175, 0.125)),
  # conditions = list(r_X_X = c(1, 0.5, 0.1)),
  conditions = list(r_U = c(3, 2, 1)),
  return_sims = TRUE
)

plot(sims,
  type = "sims", i = i:n, alpha = .25,
  nrows = 1, central_tendency = FALSE, 
  vars = c("E", "D")
)
```


```{r include=FALSE}
if (recreate_figs) {
  sfm_ens_par <- sfm
}
```

### "What If" Scenarios: Developing Interventions

```{r}
# Early intervention
sfm <- sfm0 |>
  constant(start, eqn = 21) |> # **14
  constant(input, eqn = pulse(times, start,
  #  height = .1,  #**
   height = .025,
   width = 14)) |>
  change_type(r_X_D, new_type = stock) |>
  flow(intervention, from = r_X_D, eqn = r_X_D * input(t), label = "Intervention")

# Specify seed, as not all initial conditions will show improvement from intervention
sim <- simulate(sfm, seed = 123)
plot(sim)
```

```{r}
# Simulation study
n <- 100
sims <- ensemble(sfm, n = n, return_sims = TRUE)
plot(sims,
  type = "sims", i = 1:n, alpha = .25, central_tendency = FALSE,
  vars = c("E", "D", "r_X_D")
)
```


```{r include=FALSE}
if (recreate_figs) {
  sfm_int <- sfm
}
```


```{r eval=FALSE, include=FALSE}
if (recreate_figs) {
  show_vars <- c("E", "D")
  vars_ <- VARS[VARS %in% show_vars]
  colors_ <- COLORS[VARS %in% show_vars]

  # Ensemble
  sfm <- sfm_ens_par |> sim_specs(save_at = dt)
  sims <- ensemble(sfm, n = 1000)

  pl_ens <- plot(sims,
    vars = vars_, colors = colors_,
    showlegend = FALSE,
    main = "", sub = ""
  )

  # Parameter variation
  show_vars <- c("E", "D")
  vars_ <- VARS[VARS %in% show_vars]
  colors_ <- COLORS[VARS %in% show_vars]

  sims <- ensemble(sfm,
    n = 100,
    conditions = list(r_X_D = c(0.175, 0.15, 0.1)), return_sims = TRUE
  )

  pl_par <- plot(sims,
    type = "sims", i = 1:100,
    vars = vars_, colors = colors_, alpha = .25,
    main = "", sub = "",
    nrows = 3, central_tendency = FALSE,
    showlegend = FALSE, j_labels = FALSE, xlab = ""
  )
  pl_par

  # Intervention
  show_vars <- c("E", "D", "r_X_D")
  vars_ <- VARS[VARS %in% show_vars]
  colors_ <- COLORS[VARS %in% show_vars]

  sfm <- sfm_int |> sim_specs(save_at = dt)
  sims <- ensemble(sfm, n = 100, only_stocks = FALSE, return_sims = TRUE)

  pl_int <- plot(sims,
    type = "sims", i = 1:100,
    vars = vars_, colors = colors_, alpha = .25,
    main = "", sub = "", central_tendency = FALSE,
    wrap_width = 40
  )
  pl_int

  # Plot thicker line for r_X_D
  col_ <- COLORS[VARS == "r_X_D"]
  pl_int <- pl_int |>
    plotly::add_trace(
      data = sims$df[sims$df$variable == "r_X_D", , drop = FALSE],
      x = ~time,
      y = ~value,
      line = list(
        color = col_,
        linewidth = 7
      ),
      showlegend = FALSE,
      split = ~ interaction(variable, i), # ensures each line is treated separately
      type = "scatter",
      mode = "lines"
    )

  pl <- plotly::subplot(
    plotly::subplot(pl_ens, pl_int,
      shareX = TRUE,
      shareY = TRUE,
      titleY = FALSE,
      titleX = FALSE,
      margin = c(0.025, 0.05, 0.025, 0.025),
      nrows = 2
    ) |>
      plotly::layout(annotations = list(
        list(
          x = 0.325, y = 1.05, text = "(A) Ensemble", size = font_size_subplot,
          showarrow = FALSE, xref = "paper", yref = "paper"
        ),
        list(
          x = 0.325, y = 0.49, text = "(E) Intervention", size = font_size_subplot,
          showarrow = FALSE, xref = "paper", yref = "paper"
        )
      )),
    pl_par |>
      plotly::layout(annotations = list(
        list(
          x = .725, y = 1.05, text = "(B) Burnout",
          size = font_size_subplot, showarrow = FALSE, xref = "paper", yref = "paper"
        ),
        list(
          x = .725, y = 0.69, text = "(C) Bistability",
          size = font_size_subplot, showarrow = FALSE, xref = "paper", yref = "paper"
        ),
        list(
          x = .725, y = 0.3, text = "(D) Healthy",
          size = font_size_subplot, showarrow = FALSE, xref = "paper", yref = "paper"
        )
      )),
    shareX = TRUE,
    shareY = TRUE,
    titleY = FALSE,
    titleX = FALSE,
    margin = c(0.05, 0.025, 0.025, 0.025),
    nrows = 1, widths = c(.5, .5)
  ) |>
    plotly::layout(
      title = list(text = ""),
      font = list(family = font_family, size = font_size),
      legend = list(
        orientation = "h", # orientation
        x = 0.5, # Center horizontally
        y = -0.12, # Below the plot
        xanchor = "center",
        yanchor = "top",
        font = list(size = font_size_legend)
      )
    ) |>
    plotly::layout(
      annotations = list(
        list(
          text = xlab,
          font = list(
            family = font_family,
            size = font_size_xlab
          ),
          xref = "paper",
          yref = "paper",
          xanchor = "center", yanchor = "top",
          x = 0.5, y = -0.06,
          showarrow = FALSE
        )
      )
    )
  pl


  export_plot(pl, file.path(filepath_figs, "learning_JDR.pdf"),
    width = 5, height = 5
  )

  sfm <- sfm |> sim_specs(save_at = 1)
}
```


We increase the simulation length and only save the last timepoint to compute the effectiveness of the intervention.

```{r}
n <- 1000
sfm <- sfm |> sim_specs(stop = 1000, save_n = 1)
sims <- ensemble(sfm, n = n, return_sims = TRUE)
df <- as.data.frame(sims, direction = "wide", type = "sims")
tab <- table(round(df$X, 4)) |>
  prop.table() |>
  as.data.frame()
colnames(tab) <- c("Energy", "Proportion")
print(tab)
```

A later intervention is not effective.

```{r}
sfm <- update(sfm, "start", eqn = 28)
sims <- ensemble(sfm, n = n, return_sims = TRUE)
df <- as.data.frame(sims, direction = "wide", type = "sims")
tab <- table(round(df$X, 4)) |>
  prop.table() |>
  as.data.frame()
colnames(tab) <- c("Energy", "Proportion")
print(tab)
```


### Deriving Testable Hypotheses

```{r}
# What treatment works after burnout has occurred?
sfm <- sfm0 |>
  # Prolong simulation to see long-term effects of treatment
  sim_specs(stop = 500) |>
  constant(start, eqn = 200) |>
  constant(input, eqn = step(times, start))

sfm1 <- sfm |> flow(reduce_demands, from = D, eqn = D * input(t))
sfm2 <- sfm |> flow(add_resources, to = R, eqn = .1 * input(t)) # ** prev: .05 * input(t)
sfm3 <- sfm |> flow(add_engagement, to = E, eqn = .1 * input(t))
sfm4 <- sfm |> flow(add_energy, to = X, eqn = .3 * input(t))

sim1 <- simulate(sfm1)
sim2 <- simulate(sfm2)
sim3 <- simulate(sfm3)
sim4 <- simulate(sfm4)

pl1 <- plot(sim1)
pl2 <- plot(sim2, showlegend = FALSE)
pl3 <- plot(sim3, showlegend = FALSE)
pl4 <- plot(sim4, showlegend = FALSE)

pl <- plotly::subplot(pl1, pl2, pl3, pl4,
  nrows = 2,
  margin = c(0.05, 0.1, 0.05, 0.05)
)
pl
```


```{r include=FALSE}
if (recreate_figs) {
  # Create subplot figure
  sim1 <- simulate(sfm1 |> sim_specs(save_at = 1), only_stocks = FALSE)
  sim2 <- simulate(sfm2 |> sim_specs(save_at = 1), only_stocks = FALSE)
  sim3 <- simulate(sfm3 |> sim_specs(save_at = 1), only_stocks = FALSE)
  sim4 <- simulate(sfm4 |> sim_specs(save_at = 1), only_stocks = FALSE)

  show_vars <- c("E", "R", "X", "D", "P")
  vars_ <- VARS[VARS %in% show_vars]
  colors_ <- COLORS[VARS %in% show_vars]

  pl1 <- plot(sim1, vars = vars_, colors = colors_, main = "")
  pl2 <- plot(sim2, vars = vars_, colors = colors_, main = "", showlegend = FALSE)
  pl3 <- plot(sim3, vars = vars_, colors = colors_, main = "", showlegend = FALSE)
  pl4 <- plot(sim4, vars = vars_, colors = colors_, main = "", showlegend = FALSE)

  x1 <- .25
  y1 <- 1.04
  x2 <- .78
  y2 <- .495

  pl <- plotly::subplot(pl1, pl2, pl3, pl4,
    nrows = 2,
    shareX = TRUE,
    shareY = TRUE,
    titleY = FALSE,
    titleX = FALSE,
    margin = c(0.05, 0.05, 0.05, 0.05)
  ) |>
    plotly::layout(
      font = list(family = font_family, size = font_size),
      legend = list(
        orientation = "h", # orientation
        x = 0.5, # Center horizontally
        y = -0.16, # Below the plot
        xanchor = "center",
        yanchor = "top",
        font = list(size = font_size_legend)
      )
    ) |>
    plotly::layout(
      annotations = list(
        list(
          text = xlab,
          font = list(
            family = font_family,
            size = font_size_xlab
          ),
          xref = "paper",
          yref = "paper",
          xanchor = "center", yanchor = "top",
          x = 0.5, y = -.075,
          showarrow = FALSE
        )
      )
    ) |>
    plotly::layout(annotations = list(
      list(x = x1, y = y1, text = "(A) Reducing job demands", size = font_size_subplot, showarrow = FALSE, xref = "paper", yref = "paper", xanchor = "center", yanchor = "middle"),
      list(x = x2, y = y1, text = "(B) Increasing job resources", size = font_size_subplot, showarrow = FALSE, xref = "paper", yref = "paper", xanchor = "center", yanchor = "middle"),
      list(x = x1, y = y2, text = "(C) Increasing work engagement", size = font_size_subplot, showarrow = FALSE, xref = "paper", yref = "paper", xanchor = "center", yanchor = "middle"),
      list(x = x2, y = y2, text = "(D) Increasing energy", size = font_size_subplot, showarrow = FALSE, xref = "paper", yref = "paper", xanchor = "center", yanchor = "middle")
    ))
  pl


  export_plot(pl, file.path(filepath_figs, "hypotheses_JDR.pdf"),
    width = 4.8, height = 5
  )
}
```

## Session Information

```{r echo=TRUE}
sessionInfo()
```

