---
title: "Formalizing Job Demands-Resources Theory"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Formalizing Job Demands-Resources Theory}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  include = TRUE,
  eval = TRUE,
  comment = "#>",
  fig.width = 12, # 1
  fig.height = 4, # 3
  fig.align = "center", # Center all figures
  out.width = "80%" # Display figures at 80% of container width
)
```

This is the online supplemental material accompanying the paper *Formalizing Psychological Theories with sdbuildR: A System Dynamics Modelling Tutorial in R* by Evers et al. (submitted). It contains the complete code to build, test, and simulate a formal model of Job Demands-Resources (JD-R) theory. To reproduce the figures in the paper, please see the corresponding .Rmd file.

To follow the entire tutorial, it is necessary to install Julia as well as set up the Julia environment. See [this vignette](https://kcevers.github.io/sdbuildR/articles/julia-setup.html) for guidance.

```{r libraries, results='hide', message=FALSE}
library(sdbuildR)
library(kableExtra)
```

```{r eval = TRUE, include=FALSE}
# Set this to TRUE to recreate the figures in the paper
recreate_figs <- TRUE # FALSE # TRUE
```

```{r include=FALSE}
# Helper function
sprintf_arg <- function(template, values) {
  for (nm in names(values)) {
    template <- gsub(paste0("%\\(", nm, "\\)s"), values[[nm]], template)
  }
  return(template)
}
font_family <- "Times New Roman"
# font_family <- "Lato"

if (recreate_figs) {
  # Plotting parameters
  filepath_base <- dirname(rstudioapi::getActiveDocumentContext()$path)
  filepath_figs <- file.path(filepath_base, "figs_jdr")

  dir.create(filepath_figs)

  VARS <- c("engagement", "performance", "energy", "demands", "resources", "effort_rate")
  PALETTE <- c("Geyser", "Temps", "Spectral", "Zissou 1", "Roma")[5]
  COLORS <- grDevices::hcl.colors(n = 50, palette = PALETTE)

  flow_col1 <- adjustcolor(COLORS[17], alpha.f = 1)
  flow_col2 <- adjustcolor(COLORS[9], alpha.f = 1)

  COLORS_alpha <- adjustcolor(COLORS, alpha.f = .25)

  COLORS <- COLORS[c(37, 9, 32, 1, 47, 17)]
  stock_col <- adjustcolor(COLORS[1], alpha.f = .75)
  flow_col <- COLORS[6]

  health_path <- adjustcolor(COLORS[4], alpha.f = .75)
  motivation_path <- adjustcolor(COLORS[1], alpha.f = .85)
  health_path_fill <- adjustcolor(COLORS[4], alpha.f = .25)
  motivation_path_fill <- adjustcolor(COLORS[1], alpha.f = .25)
  performance_path <- adjustcolor(COLORS[2], alpha.f = .75)
  performance_path_fill <- adjustcolor(COLORS[2], alpha.f = .25)

  font_size <- 16
  font_size_xlab <- round(font_size * 1.4)
  font_size_legend <- round(font_size * 1.1)
  font_size_subplot <- font_size #round(font_size * 1.5)
  xlab <- "Time (days)"

  # Finer-grained simulation
  dt <- .1

  # Annotation template
  annot <- list(xanchor = "center", yanchor = "middle", showarrow = FALSE, xref = "paper", yref = "paper", font = list(size = font_size_subplot, family = font_family))
  annot_font <-list(family = font_family, size = font_size)
  annot_legend <- list(
        orientation = "h", # orientation
        x = 0.5, # Center horizontally
        y = -0.15, # Below the plot
        xanchor = "center",
        yanchor = "top",
        # traceorder = "normal",
        # tracegroupgap = 0,
        # itemwidth = 15,
        font = list(size = font_size_legend)
      )
  annot_xlab <- list(
          text = xlab,
          font = list(
            family = font_family,
            size = font_size_xlab
          ),
          xref = "paper",
          yref = "paper",
          xanchor = "center", 
          yanchor = "top",
          x = 0.5, y = -.1,
          showarrow = FALSE
        )

  rm_grid <- function(pl){
    plotly::layout(pl,
      xaxis = list(showgrid = FALSE, showline = TRUE),
      yaxis = list(showgrid = FALSE, showline = TRUE)
    )
  }

}
```



```{r include=FALSE}
if (recreate_figs) {
  sfm <- sdbuildR() |>
    stock(Stock, label = "Stock", eqn = 1) |>
    constant(pulse_func1, eqn = pulse(times, start = 20, height = .1, width = 5)) |>
    constant(pulse_func2, eqn = pulse(times, start = 40, height = .1, width = 5)) |>
    constant(pulse_func3, eqn = pulse(times, start = 60, height = -.1, width = 5)) |>
    constant(pulse_func4, eqn = pulse(times, start = 80, height = -.1, width = 5)) |>
    flow(Inflow, to = Stock, eqn = 1 + pulse_func1(t) + pulse_func3(t)) |>
    flow(Outflow, from = Stock, eqn = 1 + pulse_func2(t) + pulse_func4(t))

  pl <- plot(sfm,
    wrap_width = 9.5,
    font_size = 20, minlen = 1.5,
    stock_col = stock_col,
    flow_col = flow_col,
    font_family = font_family,
    show_constants = FALSE, show_aux = FALSE, show_dependencies = FALSE
  )
  pl

  export_plot(pl, file.path(filepath_figs, "sfm.pdf"))

  # Simulation
  sim <- simulate(sfm, save_at = .01, only_stocks = FALSE)
  pl_sim <- plot(sim,
    vars = rev(c("Stock", "Inflow", "Outflow")),
    xlab = "", main = "",
    colors = rev(c(stock_col, flow_col1, flow_col2)),
    font_family = font_family,
    font_size = font_size_subplot,
    showlegend = TRUE
  ) |>
    plotly::layout(
      font = annot_font,
      legend = utils::modifyList(annot_legend, list(
        y = -0.2)
    ) |>
    plotly::layout(
      annotations = list(
        utils::modifyList(annot_xlab, list(x = 0.5))
      )
    ) |> rm_grid()
  pl_sim
  
  export_plot(pl_sim, file.path(filepath_figs, "sfm_sim.pdf"),
    width = 4, height = 3
  )

}
```


# Step 1b. Key Variables

```{r}
def_table <- data.frame(
  Variable = c(
    "Job Demands", "Job Resources", "Work Engagement", "Exhaustion", "Proactive Behaviour", "Self-Undermining Behaviour", "Job Performance"
  ),
  Definition = c(
"The physical, psychological, social, or organizational aspects of the job that require sustained physical, cognitive, and/or emotional effort and are therefore associated with certain physiological and/or psychological costs", 
"The physical, psychological, social, or organizational aspects of the job that have motivating potential, that are functional in achieving work goals, that regulate the impact of job demands, and that stimulate learning and personal growth",
"a positive, fulfilling, work-related state of mind that is characterized by vigor, dedication, and absorption. Vigor refers to high levels of energy and mental resilience while working, the willingness to invest effort in one’s work, and persistence even in the face of difficulties. Dedication implies being strongly involved in one’s work and experiencing a sense of significance, enthusiasm, and challenge. Absorption refers to being fully concentrated and happily engrossed in one’s work, whereby time passes quickly. Thus, work engagement is characterized by a high level of energy and strong identification with one’s work, whereas burnout is characterized by the opposite: a low level of energy and poor identification with one’s work",
"Depletion of energy resources; also used interchangeable or as part of job strain",
"Also called job crafting; employees’ personal initiative to change their job demands and job resources in order to better align the design of the job with their own abilities and preferences",
"Employees’ dysfunctional behaviors (e.g., poor communication, conflict behaviors) that create obstacles and may undermine performance.",
"Undefined in the literature; refers to the extent to which an individual performs well at their job (e.g., fulfilling responsibilities)"

  )
)

kableExtra::kbl(def_table,
  format = "markdown",
  col.names = c("Variable", "Definition"),
  caption = "Definition of key variables in Job-Demands Resources Theory from Bakker et al. (2025)"
) |>
  kableExtra::kable_styling()
```

# Step 3b. Formulating a Stock-and-Flow Diagram




% **gaps revealed
% **
% We can immediately see that engagement only has an inflow that can increase engagement, namely motivation.


```{r}
{
viz_str <- "
digraph flowchart {

  graph [layout = dot,
         rankdir = TB,
         bgcolor = white,
         nodesep = 0.45,
         ranksep = 0.5,
         splines = ortho,
         fontname = '%(font_family)s',
         pad = '0.2']

  node [fontsize = 9.5,
        fontname = '%(font_family)s',
        fontcolor = 'white',
        penwidth = 0]

  edge [fontname = '%(font_family)s',
        fontsize = 8.5,
        fontcolor = '#555555',
        color = '#666666',
        penwidth = 1.3,
        arrowsize = 0.6]

  # --- DECISION NODES (rounded boxes with border) ---
  q1 [label = 'Can the variable be represented\\nas a continuous quantity that\\nbe lower or higher?',
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.15,0.1']

  q2 [label = 'Does the variable change within\\nthe chosen time horizon?',
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.15,0.1']

  q3 [label = 'Is the variable a behaviour, action,\\nor process that takes place over time?',
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.15,0.1']

  q4 [label = 'Does the variable accumulate\\nthe effects of flows?',
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.15,0.1']

  q5 [label = 'Does the variable\\'s effect\\naccumulate in a stock?', # Does it directly increase or decrease a stock?
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.15,0.1']

  q6 [label = 'Should it?',
      shape = box, style = 'filled,rounded',
      fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
      color = '#0f5c5a',
      penwidth = 2, margin = '0.12,0.08']

  # --- TERMINAL NODES (no border, darker fill) ---
  constant [label = 'Constant',
            shape = box, style = 'filled,rounded',
            fillcolor = '#0f5c5a', color = 'none',
            fontsize = 10, margin = '0.15,0.08']

  stock [label = 'Stock',
         shape = box, style = 'filled,rounded',
         fillcolor = '#0f5c5a', color = 'none',
         fontsize = 10, margin = '0.15,0.08']

  flow [label = 'Flow',
        shape = box, style = 'filled,rounded',
        fillcolor = '#0f5c5a', color = 'none',
        fontsize = 10, margin = '0.15,0.08']

  auxiliary [label = 'Auxiliary',
             shape = box, style = 'filled,rounded',
             fillcolor = '#0f5c5a', color = 'none',
             fontsize = 10, margin = '0.15,0.08']

  # --- ACTION NODES ---
  reformulate [label = 'Reformulate variable',
               shape = box, style = 'dashed,rounded,filled',
            fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
            fontsize = 8, penwidth = 1.2, color = '#17807c',
            margin = '0.1,0.06']

  add_flow [label = 'Add missing\\nflows',
            shape = box, style = 'dashed,rounded,filled',
            fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
            fontsize = 8, penwidth = 1.2, color = '#17807c',
            margin = '0.1,0.06']

  add_stock [label = 'Add missing\\nstock',
             shape = box, style = 'dashed,rounded,filled',
             fillcolor = '#e8f4f3', fontcolor = '#0f5c5a',
             fontsize = 8, penwidth = 1.2, color = '#17807c',
             margin = '0.1,0.06']

  # --- EDGES ---
  q1 -> reformulate [xlabel = 'No', minlen = 2]
  reformulate -> q1 [minlen = 2, style = dashed, penwidth = 1]
  q1 -> q2          [xlabel = 'Yes']

  q2 -> constant [xlabel = 'No', minlen = 2]
  q2 -> q3       [xlabel = 'Yes']

  q3 -> q4 [xlabel = <<font>No, it can be captured<br/>at a single moment in time</font>>]
  q3 -> q5 [xlabel = '  Yes']

  q4 -> stock [xlabel = 'Yes']
  q4 -> q6    [xlabel = 'No']

  q5 -> flow  [xlabel = 'Yes']
  q5 -> q6    [xlabel = 'No']

  q6 -> auxiliary [xlabel = 'No']
  q6 -> add_flow  [xlabel = 'Yes', style = dashed, color = '#2a9d8f', penwidth = 1]
  q6 -> add_stock [xlabel = 'Yes', style = dashed, color = '#2a9d8f', penwidth = 1]
  add_flow  -> stock [style = dashed, color = '#2a9d8f', penwidth = 1]
  add_stock -> flow  [style = dashed, color = '#2a9d8f', penwidth = 1]

  # --- RANK CONSTRAINTS ---
  { rank = same; q1; reformulate }
  { rank = same; q2; constant }
  { rank = same; q4; q5 }
  { rank = same; stock; q6; flow }
  { rank = same; add_flow; auxiliary; add_stock }
}
"
}

viz_str <- sprintf_arg(
  viz_str,
  list(
    font_family = font_family
  )
)

pl <- DiagrammeR::grViz(viz_str)
pl
```


## Stock-and-Flow Models

### Single stock example

```{r}
sfm <- sdbuildR() |>
  meta(name = "Burnout") |>
  sim_specs(stop = 16, time_units = "weeks", only_stocks = FALSE)
```

```{r}
sfm <- sfm |>
  stock(energy, eqn = .3, label = "Energy")

sim <- simulate(sfm)
plot(sim)
```


```{r include=FALSE}
if (recreate_figs) {
  sfm_list <- list()
  sfm_list[["A"]] <- sfm
}
```



```{r}
diagnose(sfm)
```


```{r}
sfm <- sfm |>
  constant(work_rate, eqn = .5)
```

```{r}
sfm <- sfm |>
  flow(work, eqn = work_rate,
  from = energy, label = "Depletion from work")

sim <- simulate(sfm)
plot(sim)
```


```{r include=FALSE}
if (recreate_figs) {
  sfm_list[["B"]] <- sfm
}
```


```{r}
sfm <- sfm |>
  update(work, eqn = work_rate * energy)

sim <- simulate(sfm)
plot(sim)
```



```{r include=FALSE}
if (recreate_figs) {
  sfm_list[["C"]] <- sfm
}
```

```{r}
sfm <- sfm |>
  constant(recovery_rate, eqn = .3, label = "Recovery Rate") |>
  flow(recovery, eqn = recovery_rate, to = energy, label = "Recovery")

sim <- simulate(sfm)
plot(sim)
```



```{r include=FALSE}
if (recreate_figs) {
  sfm_list[["D"]] <- sfm
}
```

What if the ability to recover is not static, but erodes over time?

```{r}
sfm <- sfm |>
  change_type(recovery_rate, new_type = "stock") 
```


```{r}
sfm <- sfm |>
  update(recovery_rate, eqn = 1) |>
  flow(erosion, eqn = recovery_rate * work, 
    # eqn = capacity * work / (1 + energy), 
  from = recovery_rate, label = "Erosion")

sim <- simulate(sfm)
plot(sim)
```


Counterintuitively, as recovery capacity erodes, energy first increases. This is because recovery first still exceeds energy depletion from work. 

```{r}
sfm <- sfm |>
  aux(net_change, eqn = recovery - work, label = "Net change in energy")

sim <- simulate(sfm)
plot(sim)
```


```{r include=FALSE}
if (recreate_figs) {
  sfm_list[["E"]] <- sfm
}
```



```{r include=FALSE}
if (recreate_figs) {
  # Create subplot figure
  pl_list <- list()
  max_letter <- max(names(sfm_list))
  VARS_example <- c("energy", "work", "recovery", "recovery_rate", "net_change")

  # COLORS_example <- grDevices::hcl.colors(n = 50, palette = PALETTE)
  COLORS_example <- COLORS[seq_along(VARS_example)]

  titles <- paste0(
    paste0("<i>", LETTERS[seq_along(sfm_list)], ". ", 
    c("Static stock", 
    "Constant outflow", 
    "Stock-dependent<br>outflow", 
    "Constant inflow",
    "Erosion of recovery rate" 
    # "Auxiliary variable tracking net change in energy"
    ), "</i>")
  )
  titles <- stats::setNames(titles, names(sfm_list))

  for (letter in names(sfm_list)) {
    sim <- simulate(sfm_list[[letter]], 
    save_at = .01, only_stocks = FALSE
    )

    # vars_in_sim <- intersect(vars, unique(sim$df$variable))
    vars_in_sim <- unique(sim$df$variable)

    vars_ <- VARS_example[VARS_example %in% vars_in_sim] |> rev()
    colors_ <- COLORS_example[VARS_example %in% vars_in_sim] |> rev()

    pl_list[[letter]] <- plot(sim,
      vars = vars_,
      colors = colors_,
      main = "", 
      # ylim = ylim,
      showlegend = (letter == max_letter)
    ) |> rm_grid() |>
    plotly::layout(yaxis = list(
      nticks = 3
      ), xaxis = list(
      tickvals = list(0, 7.5, 15)
      ))
  }

  # x1 <- 0.18
  # x2 <- 0.8
  # x3 <- .5
  # y1 <- 1.03
  # y2 <- 0.77
  # y3 <- 0.45

  x1 <- 0.1
  x2 <- 0.37
  x3 <- .8
  y1 <- 1.08
  y2 <- 0.49
  # y3 <- 0.45

  annot_ <- utils::modifyList(annot, list(
    font = list(size = font_size_subplot + 4, family = font_family))
  )

  pl <- plotly::subplot(
    plotly::subplot(pl_list[1:4],
    nrows = 2,
    shareX = TRUE,
    shareY = FALSE,
    titleY = FALSE,
    titleX = FALSE,
    margin = 0.075
  ), plotly::subplot(pl_list[5]), 
  nrows = 1,
    # shareX = TRUE
    # shareY = FALSE,
    # titleY = FALSE,
    # titleX = FALSE,
    margin = 0.03
  ) |>
    plotly::layout(
      font = annot_font,
      # legend = annot_legend,
      # legend = utils::modifyList(annot_legend, list(x = 1.4, y = 0.8)),
      legend = utils::modifyList(annot_legend, list(y = -.2, 
      font = list(size = font_size_legend))),
      annotations = list(
        utils::modifyList(annot_xlab, list(x = 0.5, y = -0.075))
      )
    ) |>
    plotly::layout(annotations = list(
      utils::modifyList(annot_, list(x = x1, y = y1, text = titles[["A"]])),
      utils::modifyList(annot_, list(x = x2, y = y1, text = titles[["B"]])),
      utils::modifyList(annot_, list(x = x1, y = y2, text = titles[["C"]])),
      utils::modifyList(annot_, list(x = x2, y = y2, text = titles[["D"]])),
      utils::modifyList(annot_, list(x = x3, y = y1, text = titles[["E"]]))
    ))

  pl

  export_plot(pl, file.path(filepath_figs, "build_sfm.pdf"),
    width = 7, height = 4.5
  )
}
```



```{r} 
sfm <- sdbuildR() |>
  stock(c("E", "R", "D", "X"), label = c("Engagement", "Job Resources", "Job Demands", "Energy")) |>
  flow("motivation", to = "E", label = "Motivation", eqn = R * D) |>#  "Boost hypothesis: Job demands amplify the impact of job resources on work engagement (and more generally on work motivation)."
  flow("proactive_R", to = "R", label = "Proactive Behaviour", eqn = E) |> # "Engaged employees are intrinsically motivated to stay engaged and therefore start to proactively optimize their job (increase resources and optimize demands)"
  flow("proactive_D", from = "D", label = "Proactive Behaviour", eqn = E) |> # "Engaged employees are intrinsically motivated to stay engaged and therefore start to proactively optimize their job (increase resources and optimize demands)"
  flow("undermining", to = "D", label = "Undermining", eq = D - X) |> # "Proposition 8: Job demands and strain may lead to maladaptive self-regulation cognitions and behaviors (self-undermining)"
  flow("effort", from = "X", label = "Effort", eqn = D - R) |> # "Buffer hypothesis: Job resources weaken the impact of job demands on strain."
  flow("R_X", from = "R", label = "Depletion from Exhaustion", eqn = -X) |> 
  flow("E_X", from = "E", label = "Depletion from Exhaustion", eqn = -X) |> 
  aux(P, label = "Job Performance", eqn = E + X) 

plot(sfm, show_constants = TRUE, show_aux = TRUE, show_dependencies = TRUE)
```

## Simulating Stock-and-Flow Models in R

```{r}
sfm <- sdbuildR() |>
  meta(name = "Job Demands and Resources (JD-R) Theory") |>
  # Add a seed for reproducibility:
  sim_specs(seed = 123) |>
  sim_specs(time_units = day, stop = 365 / 2, 
  only_stocks = FALSE) |>
  # Make save_at larger to reduce image file size
  sim_specs(save_at = .1, dt  =.01)
```

```{r, fig.width = 1, fig.height=1}
sfm <- sfm |>
  stock(engagement, eqn = .7, label = "Work Engagement") |>
  stock(resources, eqn = .5, label = "Job Resources")
```

Both stocks should remain static.

```{r, fig.width = 1, fig.height = 3}
sim <- simulate(sfm)
plot(sim)
```

We now implement the positive feedback loop between engagement and job resources. As a first functional form, we try a linear effect of resources on motivation, and of motivation on engagement:

```{r, fig.width = 1, fig.height=1}
sfm <- sfm |> 
  constant(motivation_rate, eqn = .3, label = "Motivation Rate") |>
  flow(motivation, motivation_rate * resources, to = engagement, label = "Motivation") |>
  constant(proactive_rate, eqn = 0.2, label = "Proactive Behaviour Rate") |>
  flow(proactive, eqn = proactive_rate * engagement, to = resources,
   label = "Proactive behaviour" )

sim <- simulate(sfm)
plot(sim)
```


```{r include=FALSE}
if (recreate_figs) {
  func_form_list <- list()
  func_form_list[["A"]] <- sfm
}
```

Both resources and engagement grow exponentially to extremely high levels, which is not very realistic. We could add a negative feedback loop to limit this growth, for example by adding a (linear) decay flow from engagement and resources:

```{r, fig.width = 1, fig.height=1}
sfm <- sfm |> 
  constant(engagement_decay_rate, eqn = 0.2, label = "Engagement Decay Rate") |>
  flow(engagement_decay, eqn = engagement_decay_rate * engagement, from = engagement, label = "Decay") |>
  constant(resource_decay_rate, eqn = 0.1, label = "Resource Decay Rate") |>
  flow(resource_decay, eqn = resource_decay_rate * resources, from = resources, label = "Decay")
```


```{r}
plot(sfm)
```

```{r} 
sim <- simulate(sfm)
plot(sim)
```


```{r include=FALSE}
if (recreate_figs) {
  func_form_list[["B"]] <- sfm
}
```

However, the two negative feedback loops are not sufficient to counter the exponential growth in resources and engagement. The linear inflows represents the idea that motivation and proactive behaviour are proportional to the level of resources and engagement, respectively. There is no upper limit to their beneficial effects. 

Instead, we can implement a more realistic functional form which follows an S-curve, such as a Hill function. This function has two parameters: the slope, which determines how steeply the function increases, and the midpoint, which determines at what level of the cause variable the effect is half of its maximum. With a slope $>1$, the function is S-shaped, which captures the idea that low levels of the cause variable have neglible effects, and its effect saturates at high levels. The Hill function is only defined for positive values of the cause variable.


```{r, echo=FALSE}
# Functional forms
linear <- function(x, a) {a * x}
hill <- function(x, midpoint, slope) {x^slope / (midpoint^slope + x^slope)}
x <- seq(0, 1, by = .01)

# Linear function
slopes <- c(0.1, 0.5, 1)
df_linear <- data.frame(
  x = rep(x, length(slopes)),
  value = unlist(lapply(slopes, function(a) linear(x, a = a))),
  type = rep(slopes, each = length(x))
)

# Order legend increasingly by slope
df_linear$type <- factor(df_linear$type, levels = as.character(sort(slopes)))

# Hill function
midpoint <- 0.5
slopes <- c(1, 3, 10)
df_hill <- data.frame(
  x = rep(x, length(slopes)),
  value = unlist(lapply(slopes, function(slope) hill(x, midpoint = midpoint, slope = slope))),
  type = rep(slopes, each = length(x))
)

# Order legend increasingly by slope
df_hill$type <- factor(df_hill$type, levels = as.character(sort(slopes)))

# Pick colours for linear and hill functional form
colors_linear <- grDevices::hcl.colors(n = 50, palette = "Roma")[c(20, 15, 8)]
colors_hill <- grDevices::hcl.colors(n = 50, palette = "Roma")[c(45, 35, 28)]

# Plot functional form
pl1 <- plotly::plot_ly() |>
  plotly::add_trace(
    data = df_linear, x = ~x,
    y = ~value,
    colors = colors_linear,
    color = ~type,
    legendgroup = ~type,
    type = "scatter",
    mode = "lines"
  )

pl2 <- plotly::plot_ly() |>
  plotly::add_trace(
    data = df_hill, x = ~x,
    y = ~value,
    colors = colors_hill,
    color = ~type,
    legendgroup = ~type,
    type = "scatter",
    mode = "lines"
  )

pl_func_form <- plotly::subplot(pl1, pl2, 
      nrows = 1,
    shareX = TRUE,
    shareY = TRUE,
    titleY = TRUE,
    titleX = TRUE,
    margin = c(0.1, 0.05, 0.05, 0.05)
) |>
  plotly::layout(title = "", 
  yaxis = list(title = "Effect on receiving variable"),
  font = annot_font,
  legend = utils::modifyList(annot_legend, list(
        y = -0.2, 
        title = list(text = "Slope", font = list(size = font_size_legend)),
        traceorder = "normal",
        tracegroupgap = 0,
        itemwidth = 15,
        font = list(size = font_size_legend)
      )),
      annotations = list(
        utils::modifyList(annot_xlab, 
        list(text = "Influencing variable", y = -.05)
        )
      )
    )
pl_func_form

```


```{r, include=FALSE}
# Functional forms
logistic_eq <- function(x, midpoint, slope) {slope * x * (1 - x / midpoint)}
probit <- function(x, mean, sd) {pnorm(x, mean = mean, sd = sd)}
hill <- function(x, midpoint, slope) {x^slope / (midpoint^slope + x^slope)}
x <- seq(0, 1, by = .01)
midpoint <- 0.5
ylim <- c(-0.3, 1.1)

# Logistic function
slopes <- c(1, 5, 8)
df_logistic <- data.frame(
  x = rep(x, length(slopes)),
  value = unlist(lapply(slopes, function(a) logistic_eq(x, midpoint = midpoint, slope = a))),
  type = rep(slopes, each = length(x))
)

# Order legend increasingly by slope
df_logistic$type <- factor(df_logistic$type, levels = as.character(sort(slopes)))

# Probit function
slopes <- c(0.1, 0.3, 1)
df_probit <- data.frame(
  x = rep(x, length(slopes)),
  value = unlist(lapply(slopes, function(sd) probit(x, mean = midpoint, sd = sd))),
  type = rep(slopes, each = length(x))
)

# Order legend increasingly by slope
df_probit$type <- factor(df_probit$type, levels = as.character(sort(slopes)))

# Hill function
slopes <- c(1, 3, 10)
df_hill <- data.frame(
  x = rep(x, length(slopes)),
  value = unlist(lapply(slopes, function(slope) hill(x, midpoint = midpoint, slope = slope))),
  type = rep(slopes, each = length(x))
)

# Order legend increasingly by slope
df_hill$type <- factor(df_hill$type, levels = as.character(sort(slopes)))

# Simulate three models with different functional forms
sfm0_rev <- sdbuildR() |> stock(Stock, eqn = 0.1) |> sim_specs(only_stocks = FALSE, stop = 10)
sfm1_rev <- sfm0_rev |> flow(Inflow, eqn = 1 * Stock * (1 - Stock / 0.5), to = Stock)
sfm2_rev <- sfm0_rev |> flow(Inflow, eqn = pnorm(Stock, mean = 0.5, sd = 0.3), to = Stock)
sfm3_rev <- sfm0_rev |> flow(Inflow, eqn = Stock^3 / (0.5^3 + Stock^3), to = Stock)
vars <- c("Stock", "Inflow")

pl1 <- plot(simulate(sfm1_rev), vars = vars)
pl2 <- plot(simulate(sfm2_rev), showlegend = FALSE, vars = vars)
pl3 <- plot(simulate(sfm3_rev), showlegend = FALSE, vars = vars)
pl_sim_rev <- plotly::subplot(pl1, pl2, pl3, nrows = 1, shareX = TRUE, shareY = FALSE)


# Pick colours for linear and hill functional form
colors_logistic <- grDevices::hcl.colors(n = 100, palette = "Roma")[c(1, 15, 30)]
colors_probit <- grDevices::hcl.colors(n = 100, palette = "Roma")[c(40, 55, 68)]
colors_hill <- grDevices::hcl.colors(n = 100, palette = "Roma")[c(72, 83, 100)]

# cl <- grDevices::hcl.colors(n = 100, palette = "Roma")
# scales::show_col(cl)

titles_top <- paste0("<i>", c("Logistic function", "Probit function", "Hill function"), "</i>")
titles_bottom <- paste0("<i>", c(
  "Inflow =<br>s * Stock * (1 - Stock / K)<br>s (slope) = 1, K = 0.5",
  "Inflow =<br>pnorm(Stock, mu, s)<br>s (slope) = 0.3, mu = 0.5",
  "Inflow =<br>Stock^s / (K^s + Stock^s)<br>s (slope) = 3, K = 0.5"
), "</i>")

x1 <- 0.16
x2 <- 0.50
x3 <- 0.86
y1 <- 1.15
y2 <- 0.42

annot <- list(
  size = font_size_subplot - 4,
  xanchor = "center",
  yanchor = "middle",
  showarrow = FALSE,
  xref = "paper",
  yref = "paper"
)

# Plot functional form
pl1 <- plotly::plot_ly() |>
  plotly::add_trace(
    data = df_logistic, x = ~x,
    y = ~value,
    colors = colors_logistic,
    color = ~type,
    legendgroup = ~type,
    type = "scatter",
    mode = "lines"
  ) |>
  plotly::layout(
    yaxis = list(range = ylim)
    # margin = list(t = 0, b = 0, l = 0, r = 0)
  )

pl2 <- plotly::plot_ly() |>
  plotly::add_trace(
    data = df_probit, x = ~x,
    y = ~value,
    colors = colors_probit,
    color = ~type,
    legendgroup = ~type,
    type = "scatter",
    mode = "lines"
  ) |>
  plotly::layout(
    yaxis = list(range = ylim)
    # margin = list(t = 0, b = 0, l = 0, r = 0)
  )

pl3 <- plotly::plot_ly() |>
  plotly::add_trace(
    data = df_hill, x = ~x,
    y = ~value,
    colors = colors_hill,
    color = ~type,
    legendgroup = ~type,
    type = "scatter",
    mode = "lines"
  ) |>
  plotly::layout(
    yaxis = list(range = ylim)
    # margin = list(t = 0, b = 0, l = 0, r = 0)
  )

sim1_rev <- plot(simulate(sfm1_rev), vars = vars,
 xlab = "", showlegend = TRUE, main = "")
sim2_rev <- plot(simulate(sfm2_rev), vars = vars, xlab = "", showlegend = FALSE, main = "")
sim3_rev <- plot(simulate(sfm3_rev), vars = vars, xlab = "", showlegend = FALSE, main = "")

# Route bottom row traces to legend2
sim1_rev <- plotly::style(sim1_rev, legend = "legend2")
sim2_rev <- plotly::style(sim2_rev, legend = "legend2")
sim3_rev <- plotly::style(sim3_rev, legend = "legend2")

pl_func_form_rev <- plotly::subplot(pl1, pl2, pl3,
      nrows = 1,
    shareX = TRUE,
    shareY = TRUE,
    titleY = TRUE,
    titleX = FALSE,
    margin = c(0.05, 0.05, 0.05, 0.05)
) |>
  plotly::layout(title = "", yaxis = list(title = "Effect on receiving variable")) |>
    plotly::layout(
      font = list(family = font_family, size = font_size),
      legend = list(
        orientation = "h",
        x = 0.5, y = 0,
        xanchor = "center",
        yanchor = "top",
        traceorder = "normal",
        tracegroupgap = 0,
        itemwidth = 15,
        title = list(text = "Slope", font = list(size = font_size_legend)),
        font = list(size = font_size_legend)
      )
    ) |>
    plotly::layout(annotations = list(
      utils::modifyList(annot, list(x = x1, y = y1, text = titles_top[[1]])),
      utils::modifyList(annot, list(x = x2, y = y1, text = titles_top[[2]])),
      utils::modifyList(annot, list(x = x3, y = y1, text = titles_top[[3]]))
    ))

pl_bottom_rev <- plotly::subplot(
  sim1_rev, sim2_rev, sim3_rev,
  nrows = 1,
  shareX = TRUE,
  shareY = FALSE,
  titleY = FALSE,
  titleX = FALSE,
  margin = c(0.05, 0.05, 0.05, 0.05)
) |>
  plotly::layout(title = "", yaxis = list(title = "Value"))

pl_spacer_rev <- plotly::plot_ly(
  x = 0,
  y = 0,
  type = "scatter",
  mode = "markers",
  visible = FALSE,
  showlegend = FALSE,
  hoverinfo = "skip"
) |>
  plotly::layout(
    xaxis = list(
      visible = FALSE,
      showgrid = FALSE,
      zeroline = FALSE,
      showticklabels = FALSE
    ),
    yaxis = list(
      visible = FALSE,
      showgrid = FALSE,
      zeroline = FALSE,
      showticklabels = FALSE
    )
    # margin = c(0.05, 0.05, 0.05, 0.05)
  )

pl_final_rev <- plotly::subplot(
  pl_spacer_rev,
  pl_func_form_rev,
  pl_spacer_rev,
  pl_bottom_rev,
  pl_spacer_rev,
  nrows = 5,
  titleY = TRUE,
  heights = c(0.001, 0.35, 0.25, 0.35, 0.01),
  margin = c(0.05, 0.05, 0.05, 0.05)
) |>
  plotly::layout(
    font = list(family = font_family, size = font_size),
    legend = list(
      orientation = "v",
      x = 1.02,
      y = 1,
      xanchor = "left",
      yanchor = "top",
      traceorder = "normal",
      tracegroupgap = 0,
      itemwidth = 15,
      title = list(text = "Slope", font = list(size = font_size_legend)),
      font = list(size = font_size_legend)
    ),
    legend2 = list(
      orientation = "v",
      x = 1.02,
      y = y2 - .25,
      xanchor = "left",
      yanchor = "center",
      traceorder = "normal",
      tracegroupgap = 0,
      itemwidth = 15,
      font = list(size = font_size_legend)
    ),
    annotations = list(
      utils::modifyList(annot, list(
        text = "<i>Functional forms</i>",
        x = 0.5,
        y = 1.03,
        yanchor = "top",
        font = list(size = font_size_xlab)
      )),
      utils::modifyList(annot, list(
        text = "<i>Timeseries</i>",
        x = 0.5,
        y = y2 + 0.06,
        yanchor = "bottom",
        font = list(size = font_size_xlab)
      )),
      utils::modifyList(annot, list(x = x1, y = y2, text = titles_bottom[[1]])),
      utils::modifyList(annot, list(x = x2, y = y2, text = titles_bottom[[2]])),
      utils::modifyList(annot, list(x = x3, y = y2, text = titles_bottom[[3]])),
      utils::modifyList(annot, list(
        text = "Influencing variable",
        x = 0.5,
        y = y2 + 0.2,
        yanchor = "top",
        font = list(size = font_size_xlab - 2)
      )),
      utils::modifyList(annot, list(
        text = "Time",
        x = 0.5,
        y = 0.03,
        yanchor = "top",
        font = list(size = font_size_xlab - 2)
      ))
    )
  )
pl_final_rev


export_plot(pl_final_rev, file.path(filepath_figs, "revision_functional_forms.pdf"),
  width = 6, height = 5
)
```


```{r, fig.width = 1, fig.height=1}
sfm <- sfm |> 
  custom_func(hill, label = "Hill function", 
    eqn = "function(x, slope, midpoint = 0.5) {x^slope / (midpoint^slope + x^slope)}",
    doc = "S-shaped functional form used to capture saturation effects. The midpoint determines the inflection point of the curve where the function reaches half its maximum value. The slope determines the steepness of the curve, where slope > 1 creates an S-shaped curve.") |>
  constant(m_slope, eqn = 3, label = "Medium Slope") |>
  update(motivation, eqn = motivation_rate * hill(resources, m_slope)) |>
  update(proactive, eqn = proactive_rate * hill(engagement, m_slope))

sim <- simulate(sfm)
plot(sim)
```


```{r include=FALSE}
if (recreate_figs) {
  func_form_list[["D"]] <- sfm
}
```


However, this formalization of the positive feedback loop still requires the outflows from resources and engagement in order to stabilize at a steady state.


```{r, fig.width = 1, fig.height=1}
sfm2 <- sfm |> 
  discard(c(engagement_decay, resource_decay))

sim <- simulate(sfm2)
plot(sim)
```



```{r include=FALSE}
if (recreate_figs) {
  func_form_list[["C"]] <- sfm2
  # Sort list by name
  func_form_list <- func_form_list[order(names(func_form_list))]
}
```



```{r include=FALSE}
if (recreate_figs) {
  # Create subplot figure
  pl_list <- list()
  max_letter <- max(names(func_form_list))
  VARS_example <- c("engagement", "resources")

  # COLORS_example <- grDevices::hcl.colors(n = 50, palette = PALETTE)
  COLORS_example <- COLORS[seq_along(VARS_example)]

  titles <- paste0("<i>",
    paste0(LETTERS[1:6], ". ", c(
      "Linear functional form", 
      "Linear feedback loop", 
       "Add outflows", 
      "S-shaped functional form", 
      "S-shaped feedback loop", 
      "Add outflows"
    )), "</i>"
  )


  for (letter in names(func_form_list)) {
    sim <- simulate(func_form_list[[letter]], 
    save_at = .01, only_stocks = FALSE
    )

    vars_in_sim <- unique(sim$df$variable)
    vars_ <- VARS_example[VARS_example %in% vars_in_sim]
    colors_ <- COLORS_example[VARS_example %in% vars_in_sim]

    pl_list[[letter]] <- plot(sim,
      vars = vars_,
      colors = colors_,
      main = "",
      # ylim = ylim,
      showlegend = (letter == max_letter)
    ) |> plotly::layout(xaxis = list(
      # nticks = 3
      tickvals = list(0, 150)
      ))

  }

  # Route simulation traces to legend2 and fix margins/automargin for equal widths
  pl_list <- lapply(pl_list, function(p) {
    plotly::style(p, legend = "legend2") |>
      plotly::layout(
        margin = list(t = 0, b = 0, l = 0, r = 0),
        yaxis = list(automargin = FALSE)
      ) |> rm_grid()
  })

  x1 <- 0.22
  x2 <- 0.1
  x3 <- 0.37
  x4 <- 0.78
  x5 <- 0.62
  x6 <- 0.88
  y1 <- 1
  y2 <- 0.28

  annot <- list(size = font_size_subplot - 4, xanchor = "center", yanchor = "middle", showarrow = FALSE, xref = "paper", yref = "paper")

  inner_pl <- plotly::subplot(pl_list,
    nrows = 1,
    shareX = TRUE,
    shareY = FALSE,
    titleY = FALSE,
    titleX = FALSE,
    margin = c(0.05, 0.05, 0.05, 0.05),
    widths = c(1/4, 1/4, 1/4, 1/4)
  ) 

  pl <- plotly::subplot(
    pl_func_form,
    inner_pl,
    nrows = 2, titleY = TRUE,
    # heights = c(1/3, 2/3),
    heights = c(2/3, 1/3),
    # heights = c(6/10, 4/10),
    # heights = c(0.5,0.5),
    margin = c(0.1, 0.1, 0.1, 0.1)
  ) |>
  plotly::layout(
    font = list(family = font_family, size = font_size),
    legend = list(
      orientation = "h",
      x = 0.5, y = 0.32,
      xanchor = "center",
      yanchor = "top",
      traceorder = "normal",
      tracegroupgap = 0,
      itemwidth = 15,
      title = list(text = "Slope", font = list(size = font_size_legend)),
      font = list(size = font_size_legend)
    ),
    legend2 = list(
      orientation = "h",
      x = 0.5, y = -0.16,
      xanchor = "center",
      yanchor = "top",
      font = list(size = font_size_legend)
    )
  ) |>
    plotly::layout(
      annotations = list(
        utils::modifyList(annot_xlab, list(y = -0.075))
      )
    ) |>
    plotly::layout(annotations = list(
      utils::modifyList(annot, list(x = x1, y = y1, text = titles[[1]])),
      utils::modifyList(annot, list(x = x2, y = y2, text = titles[[2]])),
      utils::modifyList(annot, list(x = x3, y = y2, text = titles[[3]])),
      utils::modifyList(annot, list(x = x4, y = y1, text = titles[[4]])),
      utils::modifyList(annot, list(x = x5, y = y2, text = titles[[5]])),
      utils::modifyList(annot, list(x = x6, y = y2, text = titles[[6]]))
    ))

  pl

  export_plot(pl, file.path(filepath_figs, "JDR_feedback_loop.pdf"),
    width = 6, height = 5
  )
}
```

To implement the health impairment pathway, we add two stocks -- job demands and energy -- as well as an auxiliary for job performance. 

```{r }
sfm <- sfm |>
  stock(demands, eqn = .2, label = "Job Demands") |>
  stock(energy, eqn = .7, label = "Energy") |>
  aux(performance, eqn = engagement + energy, label = "Job Performance")

sim <- simulate(sfm)
plot(sim)
```

We next implement the boost and buffer hypothesis.

```{r }
sfm <- sfm |>
  constant(effort_rate, eqn = 0.5, label = "Effort Rate") |>
  flow(effort, 
  eqn = effort_rate / (1 + engagement) * energy * demands / (1 + resources),
    from = energy, label = "Effort", 
    doc = "Decreases in energy are induced by job demands, but buffered by resources (i.e., buffer hypothesis). Additionally, engaged employees optimize job demands, so the effect of demands on energy is smaller when engagement is higher. These effects are multiplied by energy to avoid depleting energy past zero."
  ) |>
  flow(motivation,
    eqn = motivation_rate * energy * hill(resources, m_slope) * demands,
    doc = "Resources are motivating such that they increase engagement. This effect is amplified by job demands (i.e., boost hypothesis), which is multiplicative according to Bakker et al. (2025). Furthermore, the effect is multiplied by energy because engagement is partially defined as vigour, and we suppose energy is required to sustain motivation."
  )

sim <- simulate(sfm)

# Specify variables to plot
vars <- c("engagement", "demands", "resources", "energy", "performance")
plot(sim, vars = vars)
```


Now that energy has been added to the model, we can implement the draining effect of exhaustion on resources and engagement. To do so, we modify the outflow from resources and engagement to get larger when energy is low:

```{r}
sfm <- sfm |>
  flow(resource_decay, eqn = resource_decay_rate * resources / (1 + energy),
  doc = "Resource decay is larger when energy is low, reflecting the idea that exhaustion depletes resources"
  ) |> 
  flow(engagement_decay,
     eqn = engagement_decay_rate * engagement / (1 + energy),
     doc = "Disengagement is larger when energy is low, reflecting the idea that exhaustion depletes engagement"
  ) 
  
sim <- simulate(sfm)
plot(sim, vars = vars)
```
  

JD-R theory only specifies how energy can be depleted, not how it can recover. This omission leaves a substantial gap in our understanding of burnout, as sufficient recovery could counter the energy-depleting effects of job demands, thereby preventing burnout.

We next allow energy to recover with a new inflow:

```{r}  
sfm <- sfm |>
  constant(recovery_rate, eqn = 0.5, label = "Restoration Rate") |>
  constant(s_slope, eqn = 5, label = "Steep Slope") |>
  flow(recovery, eqn = recovery_rate * energy * exp(-s_slope * energy), 
    to = energy, label = "Restoration",
    doc = "Energy is restored as a Ricker function, meaning that energy recovery is zero when energy is zero (some energy is needed to recover), recovery is highest at mid-levels of energy, and drops off at higher levels of energy (where recovery is not needed)"
  ) 

sim <- simulate(sfm)
plot(sim, vars = vars)
```


```{r}
sfm <- sfm |>
  constant(work_rate, eqn = 0.5, label = "Demand Reduction Rate") |>
  flow(work, eqn = work_rate * energy * demands * (1 + engagement), 
    from = demands,
    label = "Work",
    doc = "Job demands are reduced by expending energy. Engaged employees optimize job demands and are thus more effective at reducing demands. However, by specifying (1 + engagement), it captures the idea that engagement is not *necessary* for reducing demands (whereas * engagement would turn the outflow from demands to zero when engagement is zero). The effect is multiplied by demands to prevent reducing demands past zero."
  ) |>
  # constant(e_slope, eqn = 10, label = "Extreme Slope") |>
  constant(e_slope, eqn = 10, label = "Extreme Slope") |>
  constant(undermining_rate, 
  # eqn = 5, 
  eqn = 4,
  label = "Self-undermining Rate") |>
  flow(undermining, eqn = undermining_rate * energy * exp(-e_slope * energy), to = demands,
    label = "Self-undermining",
    doc = "Self-undermining behaviour is implemented as a steep Ricker function: it is zero when energy is zero (some energy is needed to self-undermine), increases steeply at low levels of energy where self-undermining is highest, and drops off quickly to represent the idea that only exhausted individuals self-undermine. The effect is independent of demands because self-undermining behaviour creates new demands even when there are no existing demands."
  )

sim <- simulate(sfm)
plot(sim, vars = vars)
```

As of now, job resources are only cultivated through proactive behaviour, and job demands are only created by self-undermining behaviour. However, JD-R theory implies that both job resources and demands are partially exogenously driven. Both are defined as "physical, psychological, social, or organizational aspects of the job" (Bakker et al., 2025, p. 32). We capture these exogenous drivers with a new inflow to job resources and demands. As a functional form, we suppose that exogenously provided resources (demands) drop off exponentially at higher levels of resources (demands):

```{r }
sfm <- sfm |>
  constant(exo_resource_rate, eqn = .1, label = "New resource rate") |>
  flow(exo_resources,
    eqn = exo_resource_rate * exp(-s_slope * resources),
    to = resources,
    label = "Exogenous support",
    doc = "Exogenously provided resources drop off exponentially at higher levels of resources, reflecting the idea that employers are less likely to provide resources when employees are already well-resourced"
  )

sim <- simulate(sfm)
plot(sim)
```

In the absence of more information about the functional form of exogenous demands, we use the same functional form as for exogenous resources:

```{r}
sfm <- sfm |>
  constant(exo_demand_rate, eqn = .1, label = "New task rate") |>
  flow(exo_demands, eqn = exo_demand_rate * exp(-s_slope * demands), to = demands,
    label = "Exogenous tasks",
    doc = "Exogenously provided demands drop off exponentially at higher levels of demands, reflecting the idea that employers are less likely to assign new tasks when employees are already at high workloads"
  )

sim <- simulate(sfm)
plot(sim, vars = vars)
```

We have now implemented the first iteration of the full model, which includes both the health impairment and motivational pathways, as well as the boost and buffer hypotheses.


```{r include=FALSE}
if (recreate_figs) {
  sfm_list <- list()
  sfm_list[["A"]] <- sfm
}
```


```{r}
plot(sfm, minlen = 1.5, font_size = 22)
```

```{r eval=FALSE, include=FALSE}
if (recreate_figs) {
# copy cat(viz_str) from plot.sdbuildR() with minlen=1.5
  {
    viz_str <- "
    
    digraph sfm {

      graph [layout = dot, rankdir = LR, center=true, outputorder='edgesfirst', pad=.1, nodesep= .3]

      # Rank groupings
{rank=same; E; P; X; R; D}

# Enforce top-to-bottom order
E -> P -> X -> R -> D [style=invis]

      # Define stock + auxiliary nodes
       'R' [id='R',label='Job Resources',tooltip = 'eqn = runif(1, 0.01, 2)',shape=box,style=filled,fillcolor='#35B7ADBF',fontsize=18,fontname='Times New Roman']
       'D' [id='D',label='Job Demands',tooltip = 'eqn = runif(1, 0.01, 2)',shape=box,style=filled,fillcolor='#35B7ADBF',fontsize=18,fontname='Times New Roman']  
       'P' [id='P',label='Job Performance',tooltip = 'eqn = E + X',shape=plaintext,fontsize=16,fontname='Times New Roman', width=0.6, height=0.3]
       'X' [id='X',label='Energy',tooltip = 'eqn = runif(1, 0.01, 2)',shape=box,style=filled,fillcolor='#35B7ADBF',fontsize=18,fontname='Times New Roman']
      'E' [id='E',label='Engagement',tooltip = 'eqn = runif(1, 0.01, 2)',shape=box,style=filled,fillcolor='#35B7ADBF',fontsize=18,fontname='Times New Roman']   
                
      'E' -> 'P' [color='#999999', arrowsize=0.8, penwidth=1, splines=false, constraint=false, tailport='_']
      'X' -> 'P' [color='#999999', arrowsize=0.8, penwidth=1, splines=false, constraint=false, tailport='_']


      # Define flow nodes (intermediate nodes for flows)
      'motivation' [id='motivation',label='Motivation', tooltip = 'eqn = motivation_rate * X * hill(R, m_slope) * D', shape = plaintext, fontsize=16, fontname='Times New Roman', width=0.6, height=0.3]
                'proactive' [id='proactive',label='Proactive behaviour', tooltip = 'eqn = proactive_rate * hill(E, m_slope)', shape = plaintext, fontsize=16, fontname='Times New Roman', width=0.6, height=0.3]
                'engagement_decay' [id='engagement_decay',label='Decay', tooltip = 'eqn = engagement_decay_rate * E / (1 + X)', shape = plaintext, fontsize=16, fontname='Times New Roman', width=0.6, height=0.3]
                'resource_decay' [id='resource_decay',label='Decay', tooltip = 'eqn = resource_decay_rate * R / (1 + X)', shape = plaintext, fontsize=16, fontname='Times New Roman', width=0.6, height=0.3]
                'effort' [id='effort',label='Effort', tooltip = 'eqn = effort_rate / (1 + E) * X * D / (1 + R)', shape = plaintext, fontsize=16, fontname='Times New Roman', width=0.6, height=0.3]
                'X_X' [id='X_X',label='Restoration', tooltip = 'eqn = r_X_X * X * exp(-s_slope * X)', shape = plaintext, fontsize=16, fontname='Times New Roman', width=0.6, height=0.3]
                'work' [id='work',label='Work', tooltip = 'eqn = r_D_E * X * D * (1 + E)', shape = plaintext, fontsize=16, fontname='Times New Roman', width=0.6, height=0.3]
                'U' [id='U',label='Self-undermining', tooltip = 'eqn = r_U * X * exp(-e_slope * X)', shape = plaintext, fontsize=16, fontname='Times New Roman', width=0.6, height=0.3]
                'R_R' [id='R_R',label='Exogenous support', tooltip = 'eqn = r_R_new * exp(-s_slope * R)', shape = plaintext, fontsize=16, fontname='Times New Roman', width=0.6, height=0.3]
                'D_D' [id='D_D',label='Exogenous tasks', tooltip = 'eqn = r_D * exp(-s_slope * D)', shape = plaintext, fontsize=16, fontname='Times New Roman', width=0.6, height=0.3]

      # Define external cloud nodes
      'Cloud1' [label='', tooltip = 'Unspecified source',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
                'Cloud2' [label='', tooltip = 'Unspecified source',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
                'Cloud3' [label='', tooltip = 'Unspecified source',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
                'Cloud4' [label='', tooltip = 'Unspecified source',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
                'Cloud5' [label='', tooltip = 'Unspecified sink',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
                'Cloud6' [label='', tooltip = 'Unspecified sink',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
                'Cloud7' [label='', tooltip = 'Unspecified sink',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
                'Cloud8' [label='', tooltip = 'Unspecified sink',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
                'Cloud9' [label='', tooltip = 'Unspecified sink',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
                'Cloud10' [label='', tooltip = 'Unspecified sink',shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]

      # Define flow edges (stock -> flow_node -> stock)
      'Cloud5' -> 'motivation' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1.5, splines=false]
      'motivation' -> 'E' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]
      'Cloud6' -> 'proactive' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1.5, splines=false]
      'proactive' -> 'R' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]
      'E' -> 'engagement_decay' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1.5, splines=false]
      'engagement_decay' -> 'Cloud1' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]
      'R' -> 'resource_decay' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1.5, splines=false]
      'resource_decay' -> 'Cloud2' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]
      'X' -> 'effort' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1.5, splines=false]
      'effort' -> 'Cloud3' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]
      'Cloud7' -> 'X_X' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1.5, splines=false]
      'X_X' -> 'X' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]
      'D' -> 'work' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1.5, splines=false]
      'work' -> 'Cloud4' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]
      'Cloud8' -> 'U' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1.5, splines=false]
      'U' -> 'D' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]
      'Cloud9' -> 'R_R' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1.5, splines=false]
      'R_R' -> 'R' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]
      'Cloud10' -> 'D_D' [arrowhead='none', color='black:#CCC568:black', penwidth=1.1, minlen=1.5, splines=false]
      'D_D' -> 'D' [arrowhead='normal', color='black:#CCC568:black', arrowsize=1.5, penwidth=1.1, minlen=1.5, splines=ortho]

      # Define dependency edges
      'X' -> 'motivation' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
      'R' -> 'motivation' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
      'D' -> 'motivation' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
      'E' -> 'proactive' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
      'E' -> 'engagement_decay' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
      'X' -> 'engagement_decay' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
      'R' -> 'resource_decay' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
      'X' -> 'resource_decay' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']

      'E' -> 'effort' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
      'X' -> 'effort' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
      'D' -> 'effort' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
      'R' -> 'effort' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
      'X' -> 'X_X' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
      'X' -> 'work' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
      'D' -> 'work' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
      'E' -> 'work' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
      'X' -> 'U' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
      'R' -> 'R_R' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']
      'D' -> 'D_D' [color='#999999', arrowsize=0.8, penwidth=1, splines=true, constraint=false, tailport='_']

    }
    "
  

  pl <- DiagrammeR::grViz(viz_str)
  pl
  }

  export_plot(pl, file.path(filepath_figs, "sfm_JDR_full.pdf"))
}
```

We save this version of the stock-and-flow model as `sfm0` for later use:

```{r}
sfm0 <- sfm |> sim_specs(save_at = .1)
```

Note that this is identical to the version stored in the model library, which can be loaded using `sdbuildR()`:

**uncomment

```{r}
# sfm0 <- sdbuildR("JDR") |> sim_specs(save_at = .1)
```


To assess the dependence of the burnout phenomenon on the model's initial conditions, we can randomly initialize the stocks and re-run the simulation. A more systematic approach will be covered in a later section.

```{r}
sfm <- sfm0 |>
  update(c(demands, resources, energy, engagement), eqn = runif(1, 0.01, 2)) |>
  sim_specs(only_stocks = FALSE)

sim1 <- simulate(sfm, seed = 1)
pl1 <- plot(sim1, vars = vars)

sim2 <- simulate(sfm, seed = 2)
pl2 <- plot(sim2, vars = vars, showlegend = FALSE)

sim3 <- simulate(sfm, seed = 6)
pl3 <- plot(sim3, vars = vars, showlegend = FALSE)

pl <- plotly::subplot(pl1, pl2, pl3, nrows = 1)
pl
```




```{r include=FALSE}
if (recreate_figs) {
  sfm_list[["B"]] <- sim_specs(sfm, seed = 1)
  sfm_list[["C"]] <- sim_specs(sfm, seed = 2)
  sfm_list[["D"]] <- sim_specs(sfm, seed = 6)
}
```




```{r include=FALSE}
if (recreate_figs) {
  # Create subplot figure
  pl_list <- list()
  min_letter <- min(names(sfm_list))
  max_letter <- max(names(sfm_list))
  VARS_example <- vars
  COLORS_example <- COLORS[seq_along(VARS_example)]

  for (letter in names(sfm_list)) {
    sim <- simulate(sfm_list[[letter]], 
    save_at = .01, only_stocks = FALSE
    )

    vars_in_sim <- unique(sim$df$variable)
    vars_ <- VARS_example[VARS_example %in% vars_in_sim]
    colors_ <- COLORS_example[VARS_example %in% vars_in_sim]

    pl_list[[letter]] <- plot(sim,
      vars = vars_,
      colors = colors_,
      main = "",
      # ylim = ylim,
      showlegend = (letter == max_letter)
    ) |> rm_grid()

    if (letter != min_letter) {
      # Remove x-axis labels and ticks
      pl_list[[letter]] <- pl_list[[letter]] |> 
      plotly::layout(xaxis = list(
        tickvals = list(0, 150),
        showticklabels = FALSE
      )) |> rm_grid()
    }

  }
  inner_pl <- plotly::subplot(pl_list[c(2,3,4)],
    nrows = 1,
    shareX = TRUE,
    shareY = TRUE,
    titleY = FALSE,
    titleX = FALSE,
    margin = c(0.025, 0.025, 0.025, 0.025)
  ) 

  pl <- plotly::subplot(
    inner_pl,
    plotly::subplot(pl_list[1]),
    nrows = 2, titleY = TRUE, shareY = TRUE,
    heights = c(1/3, 2/3),
    margin = c(0.025, 0.025, 0.025, 0.025)
  ) |>
    plotly::layout(
      font = annot_font,
      legend = utils::modifyList(annot_legend, list(
        y = -0.15,
        traceorder = "normal",
        tracegroupgap = 0,
        itemwidth = 8,
        entrywidth = 100,          # removes left/right padding between entries
        entrywidthmode = "pixels",
        borderpad = 0           # removes padding inside legend border
      ))
    ) |> 
    plotly::layout(
      annotations = list(
        utils::modifyList(annot_xlab, list(y = -.075))
      )
    ) 

  pl

  export_plot(pl, file.path(filepath_figs, "sim_JDR_full.pdf"),
    width = 5, height = 5.5
  )
}
```


## Interrogating Your Model

### Verification Tests

```{r}
sfm <- sfm0

# Job demands and energy should negatively correlate
sfm <- unit_test(sfm, expr = cor(demands, energy) < -.2)
verify(sfm)

# Job resources and energy should positive correlate
sfm <- unit_test(sfm, expr = cor(resources, energy) > .2)
verify(sfm)

# Job performance and work engagement should positively correlate
sfm <- unit_test(sfm, expr = cor(engagement, performance) > .2)
verify(sfm)

# Job performance and energy should positively correlate
sfm <- unit_test(sfm, expr = cor(performance, energy) > .2)
verify(sfm)
```

```{r}
# Behaviours should always be positive
sfm <- unit_test(sfm, expr = all(undermining >= 0))
sfm <- unit_test(sfm, expr = all(proactive >= 0))
sfm <- unit_test(sfm, expr = all(work >= 0))
verify(sfm)
```


```{r}
# # All flows should be non-negative
# as.data.frame(sfm, type = "flows", properties = "name")
# sfm <- unit_test(sfm, expr = all(E_R >= 0))

```

With conditions:

```{r}
sfm <- unit_test(sfm, expr = all(motivation == 0),
                conditions = list(motivation_rate = 0))
verify(sfm)
```


With zero job demands, we would expect job performance to be low.

```{r}
sfm <- unit_test(sfm, expr = mean(performance) < 0.2, 
                conditions = list(demands = 0, exo_demand_rate = 0, undermining_rate = 0))
verify(sfm)
```

Counterintuitively, when job demands are set to zero, energy rises, which leads to very high job performance. 

```{r}
v <- verify(sfm, return_sims = TRUE)
v$sim_index
plot(v$sims[[2]])
# sfm2 <- sfm0 |> update(c(demands, r_D, undermining_rate), eqn = 0)

# sim <- simulate(sfm2, only_stocks = FALSE)
# plot(sim, vars = vars)
```

Adjust model to make job performance dependent on demands:

```{r}
sfm <- sfm |> update(performance, eqn = demands * (engagement + energy))
```

Rerun unit tests:

```{r}
verify(sfm)
```

Update saved model:

```{r}
sfm0 <- sfm0 |> update(performance, eqn = demands * (engagement + energy))
```

We include some additional verification tests checks that are omitted from the paper for brevity. First, we would expect that when resources are zero, engagement decays to zero, which is indeed what we find.

```{r}
sfm <- unit_test(sfm, expr = tail(engagement, 1) < 0.01, 
  conditions = list(resources = 0, exo_resource_rate = 0))
verify(sfm)

sfm2 <- sfm0 |> update(c(resources, exo_resource_rate), eqn = 0)

sim <- simulate(sfm2, only_stocks = FALSE)
plot(sim, vars = vars)
```

When engagement is zero, job performance is not necessarily zero (see beginning peak).

```{r}
sfm <- unit_test(sfm, expr = all(head(performance, 100) > 0.2), 
  conditions = list(engagement = 0, motivation_rate = 0))
verify(sfm)

sfm2 <- sfm0 |> update(c(engagement, motivation_rate), eqn = 0)

sim <- simulate(sfm2, only_stocks = FALSE)
plot(sim, vars = vars)
```

Similarly, when energy is set to zero, job performance is not necessarily zero (see beginning peak).

```{r}
sfm <- unit_test(sfm, expr = all(head(performance, 100) > 0.2), 
  conditions = list(energy = 0, engagement_decay_rate = 0), 
  label = "Performance should not be zero in the beginning when energy is zero")
verify(sfm)


# Set energy to zero
sfm2 <- sfm0 |> update(c(energy, engagement_decay_rate), eqn = 0)

sim <- simulate(sfm2, only_stocks = FALSE)
plot(sim, vars = vars)
```

As an extreme condition test, we can initialize all stocks at zero. In this case, only job demands rises.

```{r}
sfm2 <- sfm0 |> update(c(engagement, resources, energy, demands), eqn = 0)

sim <- simulate(sfm2, only_stocks = FALSE)
plot(sim, vars = vars)
```

As a robustness check, we initialize all stocks at high values, which the model is able to handle.

```{r}
sfm <- unit_test(sfm, expr = all(is.finite(c(engagement, resources, energy, demands))), 
  conditions = list(engagement = 5, resources = 5, energy = 5, demands = 5))
verify(sfm)

sfm2 <- sfm0 |> update(c(engagement, resources, energy, demands),
  eqn = 5
)

sim <- simulate(sfm2, only_stocks = FALSE)
plot(sim, vars = vars)
```

<!-- 
### Robustness to Functional Form Differences

```{r}
# Grab parameter values from dataframe
df <- as.data.frame(sfm)
engagement_decay_rate <- as.numeric(df[df[["name"]] == "engagement_decay_rate", "eqn"])

# Range of engagement to evaluate functional form over
E <- seq(0, 1, by = .01)

# Create dataframe with two functional forms
df <- rbind(
  data.frame(
    E = E,
    # Add a minus sign as it is an outflow
    value = engagement_decay_rate * E,
    type = "Linear"
  ),
  data.frame(
    E = E,
    value = engagement_decay_rate * E / (engagement_decay_rate + E),
    type = ""
  )
)

# Pick two colours for linear and exponential functional form
colors <- grDevices::hcl.colors(n = 50, palette = "Roma")[c(20, 8)]

# Plot functional form
pl <- plotly::plot_ly() |>
  plotly::add_trace(
    data = df, x = ~E,
    y = ~value,
    colors = colors,
    color = ~type,
    legendgroup = ~type,
    type = "scatter",
    mode = "lines"
  ) |>
  plotly::add_trace(
    x = 0.2, y = -0.5, type = "scatter", mode = "markers",
    marker = list(size = 10, color = colors[2]),
    showlegend = FALSE
  ) |>
  plotly::layout(
    xaxis = list(title = "Work Engagement (E)"),
    yaxis = list(title = "Effect on Depletion Work Engagement (E)"),
    font = list(family = "Times New Roman")
  )

pl
``` 

Simulate two different models. 

```{r} 
sfm2 <- sfm0 |> update(from_E,
  eqn = engagement_decay_rate * E / (engagement_decay_rate + E)
)

sim1 <- simulate(sfm0, only_stocks = FALSE)
pl1 <- plot(sim1, 
# xlim = c(0, 100), 
vars = c("demands", "engagement", "performance"), showlegend = FALSE)

sim2 <- simulate(sfm2, only_stocks = FALSE)
pl2 <- plot(sim2, 
# xlim = c(0, 100), 
vars = c("demands", "engagement", "performance"))

pl <- plotly::subplot(pl1, pl2, nrows = 1) |>
  plotly::layout(title = "Linear versus Exponential Functional Form")
pl
```


```{r include=FALSE}
if (recreate_figs) {
  colors_ <- grDevices::hcl.colors(n = 50, palette = PALETTE)[c(24, 42)]
  scales::show_col(c(colors_, COLORS))

  pl_func <- plotly::plot_ly() |>
    plotly::add_trace(
      data = df,
      x = ~X,
      y = ~value,
      color = ~type,
      legendgroup = ~type,
      colors = colors_,
      type = "scatter",
      mode = "lines"
    ) |>
    plotly::add_trace(
      x = 0.2, y = -0.5, type = "scatter", mode = "markers",
      showlegend = FALSE,
      marker = list(
        # legendgroup = "Logistic",
        size = 10,
        color = colors_[2]
      )
    ) |>
    plotly::layout(
      title = "",
      xaxis = list(
        title = "",
        font = list(size = font_size)
      ),
      yaxis = list(
        title = "",
        font = list(size = font_size)
      ),
      font = list(family = font_family, size = font_size_xlab * 2)
    ) |> rm_grid()
  pl_func

  show_vars <- c("demands", "engagement", "energy", "performance")
  vars_ <- VARS[VARS %in% show_vars]
  colors_ <- COLORS[VARS %in% show_vars]

  pl1 <- plot(sim1,
    xlim = c(0, 65), vars = vars_, colors = colors_,
    showlegend = FALSE, main = "", xlab = ""
  ) |> rm_grid()

  pl2 <- plot(sim2, xlim = c(0, 65), vars = vars_, colors = colors_, main = "", xlab = "") |> rm_grid()

  pl <- plotly::subplot(pl_func,
    plotly::subplot(pl1, pl2,
      nrows = 1,
      shareY = TRUE, shareX = TRUE
    ),
    shareX = FALSE,
    shareY = FALSE,
    margin = c(.1, .1, .1, .1),
    nrows = 2
  ) |>
    plotly::layout(
      # font = list(size = font_size_xlab),
      margin = list(t = 50, b = 50, l = 100, r = 50),
      legend = list(
        orientation = "h", # orientation
        x = 0.5, # Center horizontally
        y = -0.16, # Below the plot
        xanchor = "center",
        yanchor = "top",
        font = list(size = font_size_legend)
      )
    ) |>
    plotly::layout(
      annotations = list(
        list(
          text = xlab,
          font = list(
            size = font_size_xlab
          ),
          xref = "paper",
          yref = "paper",
          xanchor = "center", yanchor = "top",
          x = 0.5, y = -0.075,
          showarrow = FALSE
        )
      )
    ) |>
    plotly::layout(annotations = list(
      list(
        x = 0.5, y = .5, text = "Energy (X)",
        font = list(size = font_size_xlab),
        showarrow = FALSE, xref = "paper", yref = "paper"
      ),
      list(
        x = -0.28, y = 1.06, text = "Effect on\nWork Engagement (E)",
        textangle = 270,
        font = list(size = font_size_xlab),
        showarrow = FALSE, xref = "paper", yref = "paper"
      )
    ))
  pl


  # export_plot(pl,
  #   file.path(filepath_figs, "functional_forms_JDR.pdf"),
  #   width = 4.3, height = 5
  # )
}
```
 -->

### Uncertainty





To illustrate the impact of aleatory uncertainty in our JD-R model, we substitute the deterministic formulation of demand influx for a stochastic process. This more closely aligns with JD-R theory, which posits job demands may fluctuate rapidly on a daily basis due to environmental volatility \parencite{Bakker2024JobQuestions, Downes2021IncorporatingMeta-Analysis}. To implement this, we adopt the Cox-Ingersoll-Ross model \parencite{cox_theory_1985}, a classical stochastic process. It consists of both a deterministic and stochastic component. The former ensures `D` reverts to its mean `D_mean` with a rate `r_D`. The latter adds normally distributed noise at each step, and scales this with the amount of demands. As long as `r_D` and `D_mean` are positive, this ensures that the inflow to demands cannot become negative, as noise is reduced to zero when demands are zero.

```{r}
sfm <- sfm0 |>
    constant(demand_mean, eqn = 1) |>
    constant(demand_sigma, eqn = 1) |>
    constant(exo_demand_rate, eqn = .25) |>
    aux(D_deterministic, eqn = exo_demand_rate * (demand_mean - demands)) |>
    aux(D_stochastic, eqn = demand_sigma * sqrt(demands) * rnorm(1)) |>
    update(exo_demands, eqn = D_deterministic + D_stochastic)

sim <- simulate(sfm, only_stocks = FALSE, seed = 123, save_at = .1)
plot(sim, vars = vars)
```


```{r include=FALSE}
if (recreate_figs) {
  show_vars <- c("engagement", "resources", "energy", "demands", "performance")
  vars_ <- VARS[VARS %in% show_vars]
  colors_ <- COLORS[VARS %in% show_vars]

  pl <- plot(sim, vars = vars_, colors = colors_, xlab = "", main = "") |>    
    rm_grid()

  pl <- pl |>
    plotly::layout(
      legend = utils::modifyList(annot_legend, list(
        y = -0.16
      ))
    ) |>
    plotly::layout(
      annotations = list(
        utils::modifyList(annot_xlab, list(y = -0.075))
      )
    )
  pl

  export_plot(pl,
    file.path(filepath_figs, "noise_JDR.pdf"),
    width = 4, height = 4
  )
}
```

### Removing Variables

```{r}
sim1 <- simulate(sfm0)
pl1 <- plot(sim1)

sfm2 <- discard(sfm0, undermining)
sim2 <- simulate(sfm2)
pl2 <- plot(sim2, showlegend = FALSE)

pl <- plotly::subplot(pl1, pl2, nrows = 1, shareY = TRUE)
pl
```



### Challenging the Model Boundary

As of now, job performance is merely an output of the model. Job performance depends on other variables:

```{r}
dependencies(sfm0, name = "performance")
```

But no other variable depends on job performance:

```{r}
dependencies(sfm0, name = "performance", reverse = TRUE)
```

However, it is plausible that job performance also feeds back into work engagement (e.g., through feelings of pride). We can easily test this by adding a feedback loop from job performance to work engagement.

```{r, fig.height=3}
# Make performance feedback to the system
sfm2 <- sfm0 |>
  constant(performance_effect, eqn = .1, 
  label = "Effect Job Performance on Work Engagement") |>
  flow(pride, eqn = performance_effect * performance, to = engagement) #|>
# ** 
  # update(c(demands, resources, engagement, energy), eqn = c(0.2, 0.5, 0.7, 0.7))

sim1 <- simulate(sfm0, stop = 500, only_stocks = FALSE)
pl1 <- plot(sim1, vars = vars)

sim2 <- simulate(sfm2, stop = 500, only_stocks = FALSE)
pl2 <- plot(sim2, vars = vars, showlegend = FALSE)

pl <- plotly::subplot(pl1, pl2, nrows = 2, shareY = TRUE) |>
  plotly::layout(title = "Model Boundary")
pl
```



```{r include=FALSE}
if (recreate_figs) {
  # Model boundary
  # show_vars <- c("engagement", "resources", "energy", "demands", "performance")
  show_vars <- c("engagement", "demands", "performance")
  vars_ <- VARS[VARS %in% show_vars]
  colors_ <- COLORS[VARS %in% show_vars]

  sim1 <- simulate(sfm0, save_at = dt, 
  stop = 500, only_stocks = FALSE)
  pl1 <- plot(sim1, vars = vars_, colors = colors_, main = "")

  sim2 <- simulate(sfm2, save_at = dt, 
  stop = 500, only_stocks = FALSE)
  pl2 <- plot(sim2, vars = vars_, colors = colors_, main = "", showlegend = FALSE)


  pl_boundary <- plotly::subplot(pl1, pl2,
    nrows = 2,
    shareX = TRUE,
    shareY = TRUE,
    titleY = FALSE,
    titleX = FALSE,
    margin = c(0, 0.05, 0.05, 0)
  ) |>
    plotly::layout(
      margin = list(t = 50, b = 50, l = 50, r = 50),
      font = annot_font,
      legend = utils::modifyList(annot_legend, list(y = -0.2)),
      annotations = list(
        utils::modifyList(annot_xlab, list(y = -0.12))
      )
    ) |> rm_grid() |>
    # Remove grid lines
    plotly::layout(
      xaxis2 = list(showgrid = FALSE, showline = TRUE),
      yaxis2 = list(showgrid = FALSE, showline = TRUE)
    )

  pl_boundary

  export_plot(pl_boundary,
    file.path(filepath_figs, "challenging_boundary_ts_JDR.pdf"),
    width = 4.8, height = 5
  )
}
```


```{r include=FALSE}
if (recreate_figs) {
  # Model boundary
  # show_vars <- c("engagement", "resources", "energy", "demands", "performance")
  show_vars <- c("engagement", "demands", "performance")
  vars_ <- VARS[VARS %in% show_vars]
  colors_ <- COLORS[VARS %in% show_vars]

  sim1 <- simulate(sfm0, save_at = dt, 
  stop = 500, only_stocks = FALSE)
  pl1 <- plot(sim1, vars = vars_, colors = colors_, main = "", xlab = "")

  sim2 <- simulate(sfm2, save_at = dt, 
  stop = 500, only_stocks = FALSE)
  pl2 <- plot(sim2, vars = vars_, colors = colors_, main = "", xlab = "")

  plot_func <- function(pl){
    pl |>
    plotly::layout(
      margin = list(autoexpand = FALSE, t = 0, b = 100, l = 50, r = 0),
      margin = c(0, 0, 0, 0), 
      font = annot_font,
      legend = utils::modifyList(annot_legend, list(y = -0.35,
      margin = list(t = 0, b = 0, l = 0, r = 0, pad = 0), # remove legend margins
      font = list(size = font_size_legend + 1),
  tracegroupgap = 0,    # remove gap between legend items
  # itemwidth = 30,        # shrink the line swatch width (default 30)
  entrywidth = 0,
  entrywidthmode = "pixels",
  x = 0.5
    )),
      annotations = list(
        utils::modifyList(annot_xlab, list(y = -0.14))
      )
    ) |> rm_grid()
  }
  
  export_plot(plot_func(pl1),
    file.path(filepath_figs, "challenging_boundary_ts_JDR_A.pdf"),
    width = 4.75, height = 2.5
  )

  export_plot(plot_func(pl2),
    file.path(filepath_figs, "challenging_boundary_ts_JDR_B.pdf"),
    width = 4.75, height = 2.5
  )
}
```


## Learning From Your Model

### Parallelization

If resources allow, set up threaded simulation:

```{r, eval = FALSE}
use_julia(nthreads = parallelly::availableCores() - 1)
```

or: 

```{r}
# future::plan(future::multisession, workers = parallelly::availableCores() - 1)
```

If not possible, regular simulation can still be used for the analyses below, albeit with longer runtimes.

### Establishing Boundary Conditions of Phenomena

```{r, fig.height=4}
# Simulation study
sfm0 <- sfm0 |> sim_specs(save_n = 100, 
language = "julia",
   seed = 123) |>
  # Random initial conditions
  update(c(demands, resources, engagement, energy), eqn = runif(1, 0.01, 2))
sfm <- sfm0

sims <- ensemble(sfm, n = 1000)
plot(sims, vars = c("engagement", "demands"))
```

#### Parameter Variation

Vary `effort_rate`: the rate of energy depletion from demand.

```{r, fig.height = 8}
n <- 250
# effort_rates <- c(0.6, 0.325, 0.1)
# effort_rates <- c(0.6, 0.35, 0.1)
# effort_rates <- c(0.65, 0.325, 0.32, 0.31, 0.1)
# effort_rates <- c(0.65, 0.325, 0.1)
# effort_rates <- c(0.65, 0.35, 0.1)
effort_rates <- c(0.65, 0.35, 0.1)
sims <- ensemble(sfm0,
  n = n,
  conditions = list(effort_rate = effort_rates),
  return_sims = TRUE, seed = 123
)

plot(sims,
  type = "sims", i = 1:n, alpha = .25,
  nrows = 1, central_tendency = FALSE, 
  vars = c("engagement", "demands")
)
```



```{r, fig.height = 8}
# n <- 100
# effort_rates <- c(0.36, 0.35, 0.34, 0.33, 0.32, 0.31, .3)
# effort_rates <- c(0.31, .3, .305, .29, .28, .27, .25, .2)
# # effort_rates <- c(.25, .225, .2)
# effort_rates <- c(0.38, .36, .35, .34, .325, .31, 3)
# # # effort_rates <- c(0.275)
# # effort_rates <- c(0.5)
# sims <- ensemble(sfm0 |> 
# update(undermining_rate, eqn = 3) |>
# # update(motivation_rate, eqn = 0.25) |>
# # update(exo_demand_rate, eqn = .1) |>
# # update(recovery_rate, eqn = 0.5) |> 
# # update(work_rate, eqn = .5) |>
# update(e_slope, eqn = 10),
#   n = n,
#   conditions = list(effort_rate = effort_rates
#   # recovery_rate = c(0.4, 0.5, 0.6),
#   # undermining_rate = c(1, 1.5, 1.75)
#   ),
#   return_sims = TRUE, seed = 123, stop = 500, save_at = 100
# )

# plot(sims,
#   type = "sims",
#    i = 1:n,
#    alpha = .25,
#   # j = 1:9,
#   # j = 10:18,
#   # j = 19:27,
#   # j = 2, 
#   # i = 1:10,
#   nrows = 3, 
#   central_tendency = FALSE, 
#   vars = c("engagement", "demands")
# )

# sims$conditions




# undermining_rate = 4
#  [5,]  5        0.34              0.5                4

#  [17,] 17        0.34              0.6                5'
# [27,] 27        0.10              0.6                6
```





```{r include=FALSE}
if (recreate_figs) {
  sfm_ens_par <- sfm
  n_ens_par <- n
}
```

### "What If" Scenarios: Developing Interventions

```{r}
sfm <- sfm0 |>
  constant(start, eqn = 21) |>
  constant(input, eqn = pulse(times, start,
   height = .1, width = 14)) |>
  change_type(effort_rate, new_type = stock) |>
  flow(intervention, from = effort_rate, 
  eqn = effort_rate * input(t), 
  label = "Intervention")

# Specify seed, as not all initial conditions will show improvement from intervention
sim <- simulate(sfm, seed = 12345)
plot(sim)
```

```{r}
# Simulation study
n <- 100
sims <- ensemble(sfm, n = n, return_sims = TRUE)
plot(sims,
  type = "sims", i = 1:n, alpha = .25, central_tendency = FALSE,
  vars = c("energy", "demands", "effort_rate")
)
```


```{r include=FALSE}
if (recreate_figs) {
  sfm_int <- sfm
}
```



```{r eval=FALSE, include=FALSE}
if (recreate_figs) {
  show_vars <- c("engagement", "demands")
  vars_ <- VARS[VARS %in% show_vars]
  colors_ <- COLORS[VARS %in% show_vars]

  # Ensemble
  sfm <- sfm_ens_par |> sim_specs(save_at = dt)
  sims <- ensemble(sfm, n = 1000)

  pl_ens <- plot(sims,
    vars = vars_, colors = colors_,
    showlegend = FALSE,
    main = "", sub = ""
  ) |> rm_grid()

  # Parameter variation
  show_vars <- c("engagement", "demands")
  vars_ <- VARS[VARS %in% show_vars]
  colors_ <- COLORS[VARS %in% show_vars]

  sims <- ensemble(sfm,
    n = n_ens_par,
    conditions = list(effort_rate = effort_rates),
     return_sims = TRUE
  )

  pl_par <- plot(sims,
    type = "sims", i = 1:n_ensemble,
    vars = vars_, colors = colors_, alpha = .25,
    main = "", sub = "",
    nrows = 3, central_tendency = FALSE,
    showlegend = FALSE, j_labels = FALSE, xlab = ""
  ) |>    
    rm_grid()
  pl_par

  # Intervention
  show_vars <- c("engagement", "demands", "effort_rate")
  vars_ <- VARS[VARS %in% show_vars]
  colors_ <- COLORS[VARS %in% show_vars]

  sfm <- sfm_int |> sim_specs(save_at = dt)
  sims <- ensemble(sfm, n = 100, only_stocks = FALSE, return_sims = TRUE)

  pl_int <- plot(sims,
    # type = "sims", i = 1:100,alpha = .25,central_tendency = FALSE,
    type = "summary",
    vars = vars_, colors = colors_, 
    main = "", sub = "", 
    wrap_width = 40
  ) |> rm_grid()
    
  pl_int

  # Plot thicker line for effort_rate
  col_ <- COLORS[VARS == "effort_rate"]
  pl_int <- pl_int |>
    plotly::add_trace(
      data = sims$df[sims$df$variable == "effort_rate", , drop = FALSE],
      x = ~time,
      y = ~value,
      line = list(
        color = col_,
        linewidth = 7
      ),
      showlegend = FALSE,
      split = ~ interaction(variable, i), # ensures each line is treated separately
      type = "scatter",
      mode = "lines"
    )

  pl <- plotly::subplot(
    plotly::subplot(pl_ens, pl_int,
      shareX = TRUE,
      shareY = TRUE,
      titleY = FALSE,
      titleX = FALSE,
      margin = c(0.025, 0.05, 0.025, 0.025),
      nrows = 2
    ) |>
      plotly::layout(annotations = list(
        utils::modifyList(annot, list(x = .325, y = 1.05, text = "<i>A. Ensemble</i>")),
        utils::modifyList(annot, list(x = .325, y = .47, text = "<i>E. Intervention</i>"))
      )),
    pl_par |>
      plotly::layout(annotations = list(
        utils::modifyList(annot, list(x = .725, y = 1.05, text = "<i>B. Burnout</i>")),
        utils::modifyList(annot, list(x = .725, y = 0.69, text = "<i>C. Bistability</i>")),
        utils::modifyList(annot, list(x = .725, y = 0.3, text = "<i>D. Healthy</i>"))
      )),
    shareX = TRUE,
    shareY = TRUE,
    titleY = FALSE,
    titleX = FALSE,
    margin = c(0.05, 0.025, 0.025, 0.025),
    nrows = 1, widths = c(.5, .5)
  ) |>
    plotly::layout(
      title = list(text = ""),
      font = annot_font,
      legend = utils::modifyList(annot_legend, list(y = -0.12)),
    ) |>
    plotly::layout(
      annotations = list(
        utils::modifyList(annot_xlab, list(y = -0.06))
      )
    )
  pl


  export_plot(pl, file.path(filepath_figs, "learning_JDR.pdf"),
    width = 5, height = 5
  )

  sfm <- sfm |> sim_specs(save_at = 1)
}
```




```{r eval=FALSE, include=FALSE}
if (recreate_figs) {
  show_vars <- c("engagement", "demands")
  vars_ <- VARS[VARS %in% show_vars]
  colors_ <- COLORS[VARS %in% show_vars]

plot_func <- function(pl){ 
    rm_grid(pl) |>
    plotly::layout(
      title = list(text = ""),
      font = annot_font
    )
  }

  # Ensemble
  sfm <- sfm_ens_par |> sim_specs(save_at = dt)
  sims <- ensemble(sfm, n = 1000)

  pl_ens <- plot(sims,
    vars = vars_, colors = colors_,
    showlegend = FALSE,
    main = "", sub = ""
  )
  
  export_plot(plot_func(pl_ens), file.path(filepath_figs, "learning_JDR_A.pdf"),
    width = 5, height = 5
  )

  # Parameter variation
  show_vars <- c("engagement", "demands")
  vars_ <- VARS[VARS %in% show_vars]
  colors_ <- COLORS[VARS %in% show_vars]

for (k in 1:3){
  sims <- ensemble(sfm,
    n = n_ens_par,
    conditions = list(effort_rate = effort_rates[k]),
     return_sims = TRUE
  )

  pl_par <- plot(sims,
    type = "sims", i = 1:n_ensemble,
    vars = vars_, colors = colors_, alpha = .25,
    main = "", sub = "",
    nrows = 1,
    central_tendency = FALSE,
    showlegend = FALSE, j_labels = FALSE, xlab = ""
  ) 
  pl_par

   
  export_plot(plot_func(pl_int), file.path(filepath_figs, sprintf("learning_JDR_%s.pdf", LETTERS[1 + k])),
    width = 5, height = 5
  )
}


  # Intervention
  show_vars <- c("engagement", "demands", "effort_rate")
  vars_ <- VARS[VARS %in% show_vars]
  colors_ <- COLORS[VARS %in% show_vars]

  sfm <- sfm_int |> sim_specs(save_at = dt)
  sims <- ensemble(sfm, n = 100, only_stocks = FALSE, return_sims = TRUE)

  pl_int <- plot(sims,
    # type = "sims", i = 1:100,alpha = .25,central_tendency = FALSE,
    type = "summary",
    vars = vars_, colors = colors_, 
    main = "", sub = "", 
    wrap_width = 40
  ) 
  
  # Plot thicker line for effort_rate
  col_ <- COLORS[VARS == "effort_rate"]
  pl_int <- pl_int |>
    plotly::add_trace(
      data = sims$df[sims$df$variable == "effort_rate", , drop = FALSE],
      x = ~time,
      y = ~value,
      line = list(
        color = col_,
        linewidth = 7
      ),
      showlegend = FALSE,
      split = ~ interaction(variable, i), # ensures each line is treated separately
      type = "scatter",
      mode = "lines"
    ) 
  
  export_plot(plot_func(pl_int), file.path(filepath_figs, "learning_JDR_E.pdf"),
    width = 5, height = 5
  )

 pl_legend <- plot_func(pl_int) |>
    plotly::layout(
      legend = utils::modifyList(annot_legend, list(y = -0.12))
    ) |>
    plotly::layout(
      annotations = list(
        utils::modifyList(annot_xlab, list(y = -0.06))
      )
    )
  export_plot(plot_func(pl_legend), file.path(filepath_figs, "learning_JDR_F.pdf"),
    width = 5, height = 5
  )


  # pl <- plotly::subplot(
  #   plotly::subplot(pl_ens, pl_int,
  #     shareX = TRUE,
  #     shareY = TRUE,
  #     titleY = FALSE,
  #     titleX = FALSE,
  #     margin = c(0.025, 0.05, 0.025, 0.025),
  #     nrows = 2
  #   ) |>
  #     plotly::layout(annotations = list(
  #       utils::modify_list(annot, list(x = .325, y = 1.05, text = "A. Ensemble")),
  #       utils::modify_list(annot, list(x = .325, y = .47, text = "E. Intervention"))
  #     )),
  #   pl_par |>
  #     plotly::layout(annotations = list(
  #       utils::modify_list(annot, list(x = .725, y = 1.05, text = "B. Burnout")),
  #       utils::modify_list(annot, list(x = .725, y = 0.69, text = "C. Bistability")),
  #       utils::modify_list(annot, list(x = .725, y = 0.3, text = "D. Healthy"))
  #     )),
  #   shareX = TRUE,
  #   shareY = TRUE,
  #   titleY = FALSE,
  #   titleX = FALSE,
  #   margin = c(0.05, 0.025, 0.025, 0.025),
  #   nrows = 1, widths = c(.5, .5)
  # ) 
  # pl



  sfm <- sfm |> sim_specs(save_at = 1)
}
```



We increase the simulation length and only save the last timepoint to compute the effectiveness of the intervention.

```{r}
n <- 1000
sfm <- sfm |> sim_specs(stop = 1000, save_n = 1)
sims <- ensemble(sfm, n = n, return_sims = TRUE)
df <- as.data.frame(sims, direction = "wide", type = "sims")
tab <- table(round(df$energy, 4)) |>
  prop.table() |>
  as.data.frame()
colnames(tab) <- c("Energy", "Proportion")
print(tab)
```

A later intervention is not as effective.

```{r}
sfm <- update(sfm, "start", eqn = 28)
sims <- ensemble(sfm, n = n, return_sims = TRUE)
df <- as.data.frame(sims, direction = "wide", type = "sims")
tab <- table(round(df$energy, 4)) |>
  prop.table() |>
  as.data.frame()
colnames(tab) <- c("Energy", "Proportion")
print(tab)
```

### Informing Experimental and Statistical Design

Fit cross-lagged panel model to simulated data.

```{r}
library(dplyr)
library(tidyr)
library(lavaan)
library(semPlot)

vars0 <- c("energy", "demands", "engagement", "resources", "performance")

# View constant values
as.data.frame(sfm, type = "constant", properties = "eqn")

# Simulation study
sfm <- sfm0 |> 
  # Bistable parameters
  update(effort_rate, eqn = runif(1, 0.05, 0.6)) |>
  update(motivation_rate, eqn = runif(1, .2, 1)) |>
  update(m_slope, eqn = runif(1, 2, 4)) |>
  update(engagement_decay_rate, eqn = runif(1, 0.05, 0.3)) |>
  update(proactive_rate, eqn = runif(1, 0.1, 0.5)) |>
  update(resource_decay_rate, eqn = runif(1, 0.1, 0.4)) |>
  update(exo_resource_rate, eqn = runif(1, 0.01, 0.2)) |>
  update(recovery_rate, eqn = runif(1, 0.3, 0.7)) |>
  update(s_slope, eqn = runif(1, 4, 6)) |>
  update(work_rate, eqn = runif(1, 0.25, .75)) |>
  update(e_slope, eqn = runif(1, 8, 12)) |>
  update(exo_demand_rate, eqn = runif(1, 0.05, 0.2)) |>
  update(undermining_rate, eqn = runif(1, 3, 7)) |>

  sim_specs(
    save_at = 1,
    # save_at = c(20, 60, 100), stop = 100,
    # save_at = seq(0, 200, by = 1),
     stop = 200,
  language = "julia", only_stocks = FALSE,
  # language = "R",
    seed = 123 # ensemble seed
    ) |>
  # Random initial conditions
  update(c(demands, resources, engagement, energy), eqn = runif(1, 0.01, 2))

n <- 10000

sims <- ensemble(sfm, n = n, return_sims = TRUE)
plot(sims, vars = vars0, type = "sims", i = 1:100)
```


```{r}
# Run cross-lagged panel analysis between X and D for different lags
wave1 <- 20
lags <- seq(10, as.numeric(sfm$sim_specs$stop) - wave1, by = 1)
df <- as.data.frame(sims, direction = "wide", type = "sims")
dfX <- df[, c("i", "time", "energy")] 
dfD <- df[, c("i", "time", "demands")] 
dfX_wave1 <- dfX[dfX$time == wave1, setdiff(names(dfX), "time")]
colnames(dfX_wave1) <- gsub("energy", "energy_wave1", colnames(dfX_wave1))
dfD_wave1 <- dfD[dfD$time == wave1, setdiff(names(dfD), "time")]
colnames(dfD_wave1) <- gsub("demands", "demands_wave1", colnames(dfD_wave1))

model <- '
  # wave 1 -> wave 2
  energy_wave2  ~ 1 + energy_wave1 + demands_wave1
  demands_wave2  ~ 1 + energy_wave1 + demands_wave1

  # concurrent covariances
  energy_wave1  ~~ demands_wave1
  energy_wave2  ~~ demands_wave2
'
```


```{r}
# Plot SEM plot for one lag
lag <- chosen_lag <- 90

# Create wide format data for lavaan
dfX_wave2 <- dfX[dfX$time == (wave1 + lag), setdiff(names(dfX), "time")]
colnames(dfX_wave2) <- gsub("energy", "energy_wave2", colnames(dfX_wave2))
dfD_wave2 <- dfD[dfD$time == (wave1 + lag), setdiff(names(dfD), "time")]
colnames(dfD_wave2) <- gsub("demands", "demands_wave2", colnames(dfD_wave2))

data_wide <- dfX_wave1 |>
  dplyr::left_join(dfX_wave2, by = "i") |>
  dplyr::left_join(dfD_wave1, by = "i") |>
  dplyr::left_join(dfD_wave2, by = "i")

# Fit lavaan model
m1 <- lavaan::sem(model, data = data_wide, 
# robust to non-normality
estimator = "MLR")
          
# # summary(m1, standardized = TRUE) 
# parameterEstimates(m1, standardized = TRUE, boot.ci.type = "bca.simple")

# # Fit statistics:
# fitMeasures(m1, c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr", "bic"))

# # chisq should be non-significant, but is often significant in large samples
# # rmsea < 0.01, 0.05, and 0.08 to indicate excellent, good, and mediocre fit, respectively.
# # srmr < .08 is good, < .10 is acceptable
# # lower bic is better


# modificationindices(m1) %>%
#   as_tibble() %>%
#   filter(mi>4) %>%
#   arrange(-mi)

# # Plot
# semPaths(m1, whatLabels = "std",
# #  layout = "spring", 
# #  edge.color = "black", sizeMan = 5, sizeLat = 7, nCharNodes = 0
# )


# Standardized parameter estimates (with standardized SEs/tests/CIs)
m1_est <- chosen_lag_est <- lavaan::standardizedSolution(
  m1,
  type = "std.all",
  se = TRUE,
  pvalue = TRUE,
  ci = TRUE
)

label_col <- "est.std"
label_digits <- 2

semPlot::semPaths(m1,
  whatLabels = "std",
  edge.label.cex = 1.2,
  edge.color = "black",
  sizeMan = 5, sizeLat = 7, nCharNodes = 0
  # layout = "spring"
)

```

```{r}
# Storage
results <- vector("list", length(lags))

# Loop over lags
for (lag in lags) {

  # Create wide format data for lavaan
  dfX_wave2 <- dfX[dfX$time == (wave1 + lag), setdiff(names(dfX), "time")] 
  colnames(dfX_wave2) <- gsub("energy", "energy_wave2", colnames(dfX_wave2))
  dfD_wave2 <- dfD[dfD$time == (wave1 + lag), setdiff(names(dfD), "time")]
  colnames(dfD_wave2) <- gsub("demands", "demands_wave2", colnames(dfD_wave2))

  data_wide <- dfX_wave1 |>
    dplyr::left_join(dfX_wave2, by = "i") |>
    dplyr::left_join(dfD_wave1, by = "i") |>
    dplyr::left_join(dfD_wave2, by = "i")

  # Fit lavaan model
  m1 <- lavaan::sem(model, data = data_wide, estimator = "MLR", # robust to non-normality
          #  se = "bootstrap", bootstrap = 5000
           )
  # summary(m1, standardized = TRUE) 

  # Store standardized parameter estimates (including standardized SEs and CIs)
  out <- lavaan::standardizedSolution(
    m1,
    type = "std.all",
    se = TRUE,
    pvalue = TRUE,
    ci = TRUE
  ) 
  out <- out[out[["op"]] == "~", c("lhs", "rhs", 
  "est.std", "se", "ci.lower", "ci.upper")]
  out[["lag"]] <- lag
  out[["est"]] <- out[["est.std"]]
  out[["est.std"]] <- NULL
  out <- out[, c("lhs", "rhs", "est", "se", "ci.lower", "ci.upper", "lag")]
  results[[as.character(lag)]] <- out

}

# Plot
results_df <- as.data.frame(do.call(rbind, results))
results_df[["path"]] <- paste(results_df[["lhs"]], results_df[["rhs"]], sep = " ~ ")
results_df <- results_df[, c("path", "est", "se", "ci.lower", "ci.upper", "lag")]
rownames(results_df) <- NULL
head(results_df, n = 10)
```

```{r}
# colors <- c("energy_wave2 ~ energy_wave1" = "#1b9e77", 
# "energy_wave2 ~ demands_wave1" = "#d95f02", 
# "demands_wave2 ~ energy_wave1" = "#7570b3", 
# "demands_wave2 ~ demands_wave1" = "#e7298a")

# Relabel dataframe
label_map <- c(
  "demands_wave2 ~ energy_wave1" = "Demands (wave 2) ~ Energy (wave 1)",
  "energy_wave2 ~ demands_wave1" = "Energy (wave 2) ~ Demands (wave 1)",
  "demands_wave2 ~ demands_wave1" = "Demands (wave 2) ~ Demands (wave 1)",
  "energy_wave2 ~ energy_wave1" = "Energy (wave 2) ~ Energy (wave 1)"
)

results_df$path <- as.factor(label_map[results_df$path])

# Sort paths for maximum lag
results_max_lag <- results_df[results_df$lag == max(results_df$lag), ]
results_max_lag[order(results_max_lag$est, decreasing = TRUE), ]

# Chosen lag
results_df[results_df$lag == chosen_lag, ]
chosen_lag_est

# colors <- c("Energy (wave 2) ~ Energy (wave 1)" = "#1b9e77", 
# "Energy (wave 2) ~ Demands (wave 1)" = "#d95f02", 
# "Demands (wave 2) ~ Energy (wave 1)" = "#7570b3",
# "Demands (wave 2) ~ Demands (wave 1)" = "#e7298a")

colors <- c("Energy (wave 2) ~ Energy (wave 1)" = COLORS[VARS == "effort_rate"],
"Energy (wave 2) ~ Demands (wave 1)" = COLORS[VARS == "performance"], 
"Demands (wave 2) ~ Energy (wave 1)" = COLORS[VARS == "engagement"],
"Demands (wave 2) ~ Demands (wave 1)" = COLORS[VARS == "resources"]
) |> rev()

# colors <- COLORS[1:4]

# Plot lag-dependent parameter estimates with confidence intervals
pl <- plotly::plot_ly(results_df, x = ~lag, y = ~est, 
color = ~path, split = ~path, legendgroup = ~path,
    colors = colors,
    type = 'scatter', mode = 'lines',
        line = list(),
        showlegend = TRUE) |> plotly::layout(
  shapes = list(
    list(
      type = "rect",
      x0 = chosen_lag - 0.5, x1 = chosen_lag + 0.5,
      y0 = 0, y1 = 1,
      yref = "paper",
       fillcolor = "#DDD9C3",
      # opacity = 0.7,
      line = list(width = 0),
        layer = "below"
      # line = list(color = COLORS[VARS == "demands"], 
      # width = 2
      # dash = "dash"
      )
    )
  ) |>
   plotly::add_ribbons(x = ~lag, ymin = ~ci.lower, 
   ymax = ~ci.upper, color = ~path, split = ~path,
    line = list(color = 'transparent'),
            showlegend = FALSE) |>
    plotly::layout(
      showlegend = TRUE,
      title = "",
      xaxis = list(
        title = "Time between waves (days)",
        font = list(size = font_size)
      ),
      yaxis = list(
        title = "Parameter estimate (standardized)",
        font = list(size = font_size)
      ),
      font = annot_font
    ) |> rm_grid() |>
    plotly::layout(
      xaxis = list(tickvals = list(10, 50, 90, 130, 170)),
      font = annot_font,
      # legend = utils::modifyList(annot_legend, list(x = 0.5, y = -0.2))
      legend = utils::modifyList(annot_legend, list(x = 1.7, y = 0.6))
    )  


pl

if (recreate_figs) {
  export_plot(pl,
    file.path(filepath_figs, "JDR_SEM_across_lags.pdf"),
    width = 6, height = 3
    # width = 5, height = 5
  )
}

```

## Session Information

```{r echo=TRUE}
sessionInfo()
```

