---
title: "Setting up Julia"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Setting up Julia}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

sdbuildR uses Julia for ensemble simulations and working with units in models. This is supported by the R package [JuliaConnectoR](https://github.com/stefan-m-lenz/JuliaConnectoR/). This guide will help you install Julia and configure it to work with sdbuildR.

## Step 1: Install Julia

Download and install Julia from [https://julialang.org/install/](https://julialang.org/install/). If you already have Julia installed, you can go to step 2.

We recommend using **juliaup**, Julia's official version manager, which makes it easy to install and switch between Julia versions. Some useful commands are listed below.

```bash
# See installed versions and current default
juliaup status

# Install latest stable version
juliaup add release

# Install specific version
juliaup add 1.11.3

# Set default version
juliaup default 1.11.3

# Update juliaup and all Julia versions
juliaup update

# Remove a version
juliaup remove 1.10.0

# List all available versions
juliaup list
```

Alternatively, you can download a standalone Julia installer from [https://julialang.org/downloads/](https://julialang.org/downloads/).

## Step 2: Check if Julia is accessible

After installing Julia, check if R can find it:

```{r eval=FALSE}
# Check if Julia is on your PATH
Sys.which("julia")

# Check if JULIA_BINDIR is set
Sys.getenv("JULIA_BINDIR")

# Verify JuliaConnectoR can find Julia
JuliaConnectoR::juliaSetupOk()
```

If both `Sys.which("julia")` and `Sys.getenv("JULIA_BINDIR")` return an empty string (`""`), you'll need to add Julia to your PATH. If the Julia installation can be found, try to start a Julia session:

```{r eval=FALSE}
# Verify JuliaConnectoR can start Julia
JuliaConnectoR::startJuliaServer()
JuliaConnectoR::stopJulia()
```

If this runs successfully, you can skip to step 6. If this throws an error, go to the next step.

## Step 3: Find your Julia installation path

You need to locate the **bin directory** of your Julia installation (not the Julia executable itself).

### From Julia

If you can start Julia from your terminal/command prompt, run this inside Julia:

```julia
juliaBase.julia_cmd()[1]
```

This returns something like:

- Windows: `"C:\\Users\\YourName\\.julia\\juliaup\\julia-1.11.3+0.x64.w64.mingw32\\bin\\julia.exe"`
- macOS: `"/Applications/Julia-1.11.app/Contents/Resources/julia/bin/julia"`
- Linux: `"/usr/bin/julia"`

**Important**: You need the bin directory containing the executable (remove `julia.exe` or `julia` from the end).


### From your terminal/command prompt

**Windows:**
```bash
where julia
```

**macOS/Linux:**
```bash
which julia
# or
whereis julia
```

Take note of the path, removing the filename to get just the `bin` directory.

## Step 4: Add Julia to your PATH permanently

To make Julia accessible to R across all sessions, add it to your `.Renviron` file.

### Open .Renviron

Run this in R to open your `.Renviron` file:

```{r eval=FALSE}
# Install usethis if needed
if (!require("usethis")) install.packages("usethis")

# Open .Renviron for editing
usethis::edit_r_environ()
```

This will open `.Renviron` in your text editor. If the file doesn't exist, it will be created.

### Add Julia to PATH

Add one of the following lines to `.Renviron`, replacing the path with your actual Julia bin directory:

**Windows (use forward slashes):**
```
PATH="C:/Users/YourName/.julia/juliaup/julia-1.11.3+0.x64.w64.mingw32/bin;${PATH}"
```

**macOS:**
```
PATH="/Applications/Julia-1.11.app/Contents/Resources/julia/bin:${PATH}"
```

**Linux:**
```
PATH="/usr/bin:${PATH}"
```

If using **juliaup** (recommended), you can point to the **juliaup** directory instead of a specific version:

**Windows:**
```
PATH="C:/Users/YourName/.julia/juliaup;${PATH}"
```

**macOS/Linux:**
```
PATH="~/.julia/juliaup/bin:${PATH}"
```

This allows **juliaup** to manage which Julia version is used.


### Save and restart R.

1. Save the `.Renviron` file
2. Restart R (Session → Restart R in RStudio, or close and reopen R)

## Step 5: Verify the setup

After restarting R, verify that Julia is accessible:

```{r eval=FALSE}
# Should show the path to julia executable
Sys.which("julia")

# Check Julia version
system("julia --version")

# Verify JuliaConnectoR can find Julia
JuliaConnectoR::juliaSetupOk()
```

If `Sys.which("julia")` returns a path and `juliaSetupOk()` returns `TRUE`, R can find your Julia installation!

Check whether a Julia session can be started from R:

```{r eval=FALSE}
# Verify JuliaConnectoR can start Julia
JuliaConnectoR::startJuliaServer()
JuliaConnectoR::stopJulia()
```

## Troubleshooting

### Julia still not found after restarting R

1. Double-check the path you added to `.Renviron`
2. Make sure you're pointing to the `bin` directory, not the Julia executable
3. On Windows, use forward slashes (`/`) not backslashes (`\`)
4. Verify you saved `.Renviron` and restarted R

### View your current PATH

To see all directories in your PATH:

```{r eval=FALSE}
# View PATH split by separator for readability
strsplit(Sys.getenv("PATH"), .Platform$path.sep)[[1]]
```

Look for your Julia bin directory in this list.

## Advanced: Using multiple Julia versions

If you have multiple Julia versions installed, the first one found in your PATH will be used by default.

### Switch versions with juliaup

If using juliaup, you can change the default version without modifying `.Renviron`:

```bash
# From terminal/command prompt
juliaup default 1.11.3
```

## Step 6: Using Julia with sdbuildR

To set up the Julia environment for sdbuildR, including all necessary packages, run `use_julia()`. First time use can take quite a while (i.e., 10-30 minutes). 

```{r eval=FALSE}
library(sdbuildR)
```

```{r eval=FALSE}
use_julia()
```

The Julia environment needs to be set up for *each* new R session, but should now be much quicker.

### Using a specific version with sdbuildR

By default, sdbuildR uses your default Julia version. To use a different version:

```{r eval=FALSE}
use_julia(JULIA_HOME = "C:/Users/YourName/.julia/juliaup/julia-1.11.3+0.x64.w64.mingw32/bin")
```

Note: This only affects the current R session and is not permanent.

### Close Julia session

To close the Julia session:

```{r eval=FALSE}
use_julia(stop = TRUE)
```

