---
title: "Get Started With sdbuildR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get Started With sdbuildR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=TRUE, message=FALSE}
library(sdbuildR)
```


```{r, echo = FALSE, cache=FALSE}
# Disable WebGL: many plotly widgets per HTML page can exceed the browser WebGL
# context limit and render blank. SVG always renders.
options(sdbuildR.webgl = FALSE)
```


```{r, echo = FALSE}
font_family <- "stix-two-text"
options(sdbuildR.font_family = font_family)
# getOption("sdbuildR.font_family")
```

sdbuildR is an R package for building, simulating, and testing stock-and-flow models. This page gives a quick overview of its main features.

### Load models from the model library

Dozens of example models can be loaded using `stockflow()`. Here we load the SIR (Susceptible-Infected-Recovered) model, a classic model in epidemiology:

```{r SIR_print}
sfm <- stockflow("sir")
print(sfm)
```

Plot the stock-and-flow diagram for a structural overview:

```{r SIR_diagram, fig.width=8, fig.height=2}
plot(sfm)
```

Simulate and plot the model's behaviour over time:

```{r SIR_plot, fig.width=8, fig.height=4}
sim <- simulate(sfm)
plot(sim)
```

### Build a model from scratch

The equivalent stock-and-flow model can also be built from scratch. We initialise an empty model with `stockflow()` and add three stocks. 

```{r scratch_build}
sfm <- stockflow() |>
  stock(Susceptible, eqn = 99999) |>
  stock(Infected, eqn = 1) |>
  stock(Recovered, eqn = 0)
```

Above, we use the pipe operator `|>` for better legibility. It simply passes the result of an expression to the next expression as its first argument. 

```{r}
plot(sfm)
```

Next, we add two flows that move population from one stock to another:

```{r scratch_build3}
sfm <- sfm |>
  flow(New_infections,
    eqn = "Infection_rate * Susceptible * Infected",
    from = "Susceptible", to = "Infected"
  ) |>
  flow(New_recoveries,
    eqn = "Recovery_rate * Infected",
    from = "Infected", to = "Recovered"
  )

plot(sfm)
```

The flows refer to four constants, which still need to be defined:

```{r scratch_build2}
sfm <- sfm |>
  constant(Total_population,
    eqn = "Susceptible + Infected + Recovered"
  ) |>
  constant(Contact_rate, eqn = 2) |>
  constant(Recovery_rate, eqn = 0.1) |>
  constant(Infection_rate,
    eqn = "Contact_rate / Total_population"
  )
```


Simulation settings such as the duration and time unit used to label the axis are configured with `sim_settings()`.

```{r}
sfm <- sfm |>
  sim_settings(start = 0, stop = 20, time_units = "weeks") |>
  # Add model name
  meta(name = "Susceptible-Infected-Recovered (SIR)")
```

```{r scratch_sim, fig.width=8, fig.height=4}
sim <- simulate(sfm)
plot(sim)
```

See the [Build](https://kcevers.github.io/sdbuildR/articles/build.html) vignette for a full guide on constructing and modifying models.

### Ensemble simulations

Running multiple simulations (i.e., an ensemble) provides insight into a model's behavioural variability. Here, we initialize all stocks to a random value:

```{r ensemble_setup}
sfm_ens <- sfm |>
  update(c(Susceptible, Infected, Recovered),
    eqn = runif(1, 1, 1000)
  ) |>
  # Save fewer values for computational efficiency
  sim_settings(stop = 50, save_by = 1)
```

```{r ensemble_plot, fig.width=8, fig.height=4}
sims <- ensemble(sfm_ens, n = 100)
plot(sims)
```

Ensemble simulations can be run in parallel in R for speed, or in Julia for even faster performance.

See the [Ensemble](https://kcevers.github.io/sdbuildR/articles/ensemble.html) vignette for more, including varying multiple parameters, parallel execution, and accessing simulation data.

### Unit tests

Stock-and-flow models can easily become complex, producing surprising results as the model is developed. Unit tests assert that a model behaves as expected. An expression that evaluates to `TRUE` means the test passes. Here we check that the susceptible population never becomes negative, and that the total population is conserved:

```{r unit_tests}
sfm <- sfm |>
  unit_test(expr = all(Susceptible >= 0)) |>
  unit_test(
    expr = all(abs(Susceptible + Infected + Recovered - Total_population) < 1e-8),
    label = "Population is conserved"
  )

verify(sfm)
```

See the [Unit tests](https://kcevers.github.io/sdbuildR/articles/unit-tests.html) vignette for more, including conditional tests, visualizing and extracting test results, and debugging failed tests.

## Learn more

* [Build](https://kcevers.github.io/sdbuildR/articles/build.html): Build, modify, and simulate stock-and-flow models.
* [Ensemble Simulations](https://kcevers.github.io/sdbuildR/articles/ensemble.html): Explore a model's behaviour across parameter ranges and initial conditions.
* [Unit Tests](https://kcevers.github.io/sdbuildR/articles/unit-tests.html): Verify models behave as intended with unit tests.
* [Job Demands-Resources Theory](https://kcevers.github.io/sdbuildR/articles/jdr.html): An example of formalizing psychological theory with sdbuildR.
* [Julia Setup](https://kcevers.github.io/sdbuildR/articles/julia-setup.html): Speed up simulations with Julia.
* [Import/Export](https://kcevers.github.io/sdbuildR/articles/import-export.html): Import models from deSolve or Insight Maker, and export to other formats.
