---
title: "Get started with sdbuildR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with sdbuildR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=TRUE, message=FALSE}
library(sdbuildR)
```

sdbuildR is an R package for building, simulating, and testing stock-and-flow models. This page gives a quick overview of the main features — load a model, build one from scratch, run ensemble simulations, and verify model behaviour with unit tests.

### Load from the model library

Dozens of example models can be loaded using `sdbuildR()`. Here we load the SIR (Susceptible-Infected-Recovered) model, a classic model in epidemiology:

```{r SIR_print}
sfm <- sdbuildR("SIR")
print(sfm)
```

Plot the stock-and-flow diagram for a structural overview:

```{r SIR_diagram, fig.width=8, fig.height=2}
plot(sfm)
```

Simulate and plot the model's behaviour over time:

```{r SIR_plot, fig.width=8, fig.height=4}
sim <- simulate(sfm)
plot(sim)
```

### Build a model from scratch

The equivalent stock-and-flow model can also be built from scratch. We initialise an empty model with `sdbuildR()` and add three stocks. 

```{r scratch_build}
sfm <- sdbuildR() |>
  stock(susceptible, eqn = 99999, label = "Susceptible") |>
  stock(infected, eqn = 1, label = "Infected") |>
  stock(recovered, eqn = 0, label = "Recovered")
```

Above, we use the pipe operator `|>` for better legibility. It simply passes the result of an expression to the next expression as its first argument. Next, we add four constants to the model:

```{r scratch_build2}
sfm <- sfm |>
  constant(total_population, eqn = "susceptible + infected + recovered", 
          label = "Total population") |>
  constant(contact_rate, eqn = 2, label = "Contact rate") |> 
  constant(recovery_rate, eqn = 0.1, label = "Recovery rate") |> 
  constant(infection_rate, eqn = "contact_rate / total_population", 
          label = "Infection rate")
```

Finally, we add two flows that move population from one stock to another:

```{r scratch_build3}
sfm <- sfm |>
  flow(new_infections, eqn = "infection_rate * susceptible * infected", 
      from = "susceptible", to = "infected", label = "New infections") |>
  flow(new_recoveries, eqn = "recovery_rate * infected", 
      from = "infected", to = "recovered", label = "New recoveries") 

print(sfm)
```

Simulation settings such as the time range and time step (`dt`) are configured with `sim_settings()`.

```{r}
sfm <- sfm |>
  sim_settings(start = 0, stop = 20, time_units = "weeks") |>
  # Add model name
  meta(name = "Susceptible-Infected-Recovered (SIR)")
```

```{r scratch_sim, fig.width=8, fig.height=4}
sim <- simulate(sfm)
plot(sim)
```

See the [Build](https://kcevers.github.io/sdbuildR/articles/build.html) vignette for a full guide on constructing and modifying models.

### Ensemble simulations

Running multiple simulations (i.e., an ensemble) provides insight into a model's behavioural variability. Here, we initialize all stocks to a random value:

```{r ensemble_setup}
sfm_ens <- sdbuildR("SIR") |>
  update(c(susceptible, infected, recovered), eqn = runif(1, 1, 1000)) |>
  # Save fewer values for computational speed
  sim_settings(stop = 50, dt = 0.25, save_at = 1)
```

```{r ensemble_plot, fig.width=8, fig.height=4}
sims <- ensemble(sfm_ens, n = 20)
plot(sims)
```

Ensemble simulations can be run in parallel in R for speed, or in Julia for even faster performance.

See the [Ensemble](https://kcevers.github.io/sdbuildR/articles/ensemble.html) vignette for more, including varying multiple parameters, parallel execution, and accessing simulation data.

### Unit tests

Stock-and-flow models can easily become complex, producing unexpected behaviours as the model is developed. Unit tests assert that a model behaves as expected. An expression that evaluates to `TRUE` means the test passes. Here we check that the susceptible population never becomes negative, and that the total population is conserved:

```{r unit_tests}
sfm <- sdbuildR("SIR") |>
  unit_test(expr = all(susceptible >= 0)) |>
  unit_test(
    expr = all(abs(susceptible + infected + recovered - total_population) < 1e-8),
    label = "Population is conserved"
  )

verify(sfm)
```

See the [Unit tests](https://kcevers.github.io/sdbuildR/articles/unit-tests.html) vignette for more, including conditional tests, ensemble-based verification, and debugging failed tests.

## Learn more

* [Build](https://kcevers.github.io/sdbuildR/articles/build.html): Build, modify, and simulate stock-and-flow models.
* [Ensemble](https://kcevers.github.io/sdbuildR/articles/ensemble.html): Explore a model's behaviour across parameter ranges and initial conditions.
* [Unit tests](https://kcevers.github.io/sdbuildR/articles/unit-tests.html): Verify models behave as intended with unit tests.
* [Job Demands-Resources Theory](https://kcevers.github.io/sdbuildR/articles/jdr.html): An example of formalizing psychological theory with sdbuildR.
* [Julia setup](https://kcevers.github.io/sdbuildR/articles/julia-setup.html): Speed up simulations with Julia.
* [Import/Export](https://kcevers.github.io/sdbuildR/articles/import-export.html): Import models from deSolve or Insight Maker, and export to other formats.

## Citation

To cite sdbuildR, please use:

```{r include=TRUE}
citation("sdbuildR")
```