---
title: "Get started with sdbuildR"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Quick start

```{r libraries, include=FALSE}
library(sdbuildR) 
```

Stock-and-flow models can be created in three ways. Firstly, dozens of example models can be loaded using `xmile()`. Plot the stock-and-flow diagram to get an overview of the model:

```{r SIR_diagram, fig.width=8, fig.height=2}
sfm = xmile("SIR")
plot(sfm)
```

The model can be simulated using `simulate()`:

```{r SIR_plot, fig.width=8, fig.height=4}
sim = simulate(sfm)
plot(sim)
```

Secondly, an [Insight Maker](https://insightmaker.com/) model can be imported to R using its URL: 

```{r message=FALSE, warning=FALSE}
URL = "https://insightmaker.com/insight/5LxQr0waZGgBcPJcNTC029/Crielaard-2022"
sfm = insightmaker_to_sfm(URL = URL)
```


```{r Crielaard2022_plot, fig.height=4, fig.width=8, message=FALSE, warning=FALSE}
sim = simulate(sfm)
```


```{r}
# plot(sim)
```

And lastly, a stock-and-flow model can be created from scratch. We make use of the convenient pipe operator `%\>%`, which simply passes the result of an expression to the next expression as a first argument. We first initialize an empty stock and flow model with `xmile()`, and then use `build()` to create a logistic model of population growth. The simulation specifications, such as the start time, stop time, and simulation time step (`dt`), are set with `sim_specs()`. Custom units such as people can be added with `model_units()`.

```{r logistic_plot, fig.width=8, fig.height=4}
sfm = xmile() %>%
  header(name = "Population growth") %>%
  build("X", "stock", eqn = ".01", label = "Population size") %>%
  build("change", "flow", eqn = "r * (1 - X/K)", to = "X", 
        label = "Births and Deaths") %>%
  build("r", "constant", eqn = "0.1", label = "Growth rate") %>%
  build("K", "constant", eqn = "1", label = "Carrying capacity") %>%
  sim_specs(stop = 200) 
plot(simulate(sfm))
```

An overview of the model components and simulation specifications can be accessed with `summary()`:

```{r summary}
summary(sfm)
```

To quickly view all model variable properties, use `as.data.frame()`:

```{r as.data.frame}
as.data.frame(sfm)
```

## Features

* **Accessibility**: Get started with stock-and-flow models with limited knowledge.
* **Flexibility**: Modify your models easily and safely.
* **Insight Maker integration**: Import models from Insight Maker.
* **Unit support**: Use standard or custom units to verify and interpret your model. 
* **Julia backend**: Leverage Julia for high-performance simulations.

All package capabilities are described in the vignettes:

* `vignette("build")` Learn how to build, modify, and debug stock-and-flow models.
* `vignette("units")` Learn why and how to use (custom) units.
<!-- 3. **Delays**: Not all dynamics are instantaneous. Learn about delays (fixed, smooth, material and information) and how to implement them with sdbuildR. -->
<!-- 4. **Sensitivity**:  -->

## Limitations

<!-- Unlike other System Dynamics software, sdbuildR makes a distinction between static and dynamic variables. This means that static variables - the initial values of stocks, constants, and graphical functions - cannot depend on dynamic variables - auxiliaries and flows. If this is the case, an error will be thrown with the problematic dependencies and the model will not be simulated.  -->

Non-negative stocks and flows are only minimally supported, where flows are not adjusted in case of non-negative stocks. Note that setting stocks and flows to be non-negative is not recommended [@sterman2000]. If a stock or flow can logically never be negative, for example in the case of animals or deaths, this should ideally arise from the model equations and parameters itself. If they unintentionally turn negative, it is likely due to model misspecification and including hard-coded constraints can lead to unexpected results [@sterman2000].

In addition, sdbuildR does not support vectorized operations. 

## Other System Dynamics software

sdbuildR is heavily based on common System Dynamics software such as [Vensim](https://vensim.com/), [Powersim](https://powersim.com/), [Stella](https://www.iseesystems.com/), and [Insight Maker](https://insightmaker.com/). To translate xmile models to R, see the R package [readsdr](https://github.com/jandraor/readsdr). To build stock-and-flow models with the R package [deSolve](https://cran.r-project.org/web/packages/deSolve/index.html), the book [System Dynamics Modelling with R](https://link.springer.com/book/10.1007/978-3-319-34043-2) by Jim Duggan will prove useful. In Python, stock-and-flow models are supported by [PySD](https://pysd.readthedocs.io/en/master/).

## Troubleshooting

sdbuildR is under active development, leveraging Julia for high-performance simulations via custom R-to-Julia translations. While thoroughly tested, the package may have bugs, particularly in complex model translations. We encourage users to report issues on GitHub (https://github.com/KCEvers/sdbuildR/issues) - your input helps the package improve! Use `debugger(sfm)` to diagnose model errors, and use the vignettes for guidance.
