---
title: "Theory Building in Psychology (Summer School 2026)"
output: rmarkdown::html_vignette
bibliography: references/references.bib
vignette: >
  %\VignetteIndexEntry{Theory Building in Psychology (Summer School 2026)}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: inline
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE, # Hide code by default
  fig.width = 6, # Set default figure width to 6 inches
  fig.height = 3, # Set default figure height to 3 inches
  fig.align = "center", # Center all figures
  out.width = "80%" # Display figures at 80% of container width
  # fig.retina = 2,         # Optimize for retina displays
  # dpi = 96               # Standard DPI for web
)
```

<style type="text/css">
body {
  line-height: 1.55;
}

h2 {
  margin-top: 2.75rem;
  margin-bottom: 1rem;
  padding-top: 0.15rem;
}

h3 {
  margin-top: 2rem;
  margin-bottom: 0.75rem;
}

h2 + h3 {
  margin-top: 1.4rem;
}

p,
ul,
ol {
  margin-bottom: 0.95rem;
}

pre,
.sourceCode,
.figure,
.html-widget,
.plotly {
  margin-top: 0.85rem;
  margin-bottom: 1.4rem;
}
</style>

```{r, echo=FALSE, message=FALSE}
library(sdbuildR)
```

```{r, echo = FALSE}
# Disable WebGL: many plotly widgets per HTML page can exceed the browser WebGL
# context limit and render blank. SVG always renders.
options(sdbuildR.webgl = FALSE)
```


```{r, echo = FALSE}
font_family <- "stix-two-text"
options(sdbuildR.font_family = font_family)
# getOption("sdbuildR.font_family")
frame_ms <- 200
```


```{r include=FALSE}
# Helper function
sprintf_arg <- function(template, values) {
  for (nm in names(values)) {
    template <- gsub(paste0("%\\(", nm, "\\)s"), values[[nm]], template)
  }
  return(template)
}
# font_family <- "Times New Roman"

# Plotting parameters
PALETTE <- c("Geyser", "Temps", "Spectral", "Zissou 1", "Roma")[5]
COLORS <- grDevices::hcl.colors(n = 50, palette = PALETTE)
# scales::show_col(COLORS)

inflow_col <- adjustcolor(COLORS[10], alpha.f = 1)
outflow_col <- adjustcolor(COLORS[3], alpha.f = 1)

inflow_fill <- adjustcolor(inflow_col, alpha.f = .5)
outflow_fill <- adjustcolor(outflow_col, alpha.f = .5)

# scales::show_col( grDevices::hcl.colors(n = 50, palette = "Greens"))
increase_col <- adjustcolor(
  grDevices::hcl.colors(n = 50, palette = "Greens")[20],
  alpha.f = 1)
decrease_col <- adjustcolor(COLORS[2], alpha.f = 1)
decrease_col_strong <- adjustcolor(COLORS[1], alpha.f = 1)
# scales::show_col(c(increase_col, decrease_col))
increase_fill <- adjustcolor(increase_col, alpha.f = .8)
decrease_fill <- adjustcolor(decrease_col, alpha.f = .4)

COLORS_alpha <- adjustcolor(COLORS, alpha.f = .25)

COLORS <- COLORS[c(37, 9, 32, 1, 47, 17)]
stock_col <- adjustcolor(COLORS[1], alpha.f = .75)
stock2_col <- adjustcolor(COLORS[2], alpha.f = .75)
aux_col <- adjustcolor(COLORS[3], alpha.f = .75)
flow_col <- COLORS[6]

line_width <- 1.5
line_width_stock <- line_width + 2

font_size <- 16
font_size_xlab <- round(font_size * 1.4)
font_size_legend <- round(font_size * 1.1)
font_size_subplot <- font_size # round(font_size * 1.5)
xlab <- "Time (days)"

# Finer-grained simulation
save_by_fine <- .1 # .01
save_by_coarse <- 1

# Annotation template
annot <- list(xanchor = "center", yanchor = "middle", showarrow = FALSE, xref = "paper", yref = "paper", font = list(size = font_size_subplot))
annot_font <- list(size = font_size, family = font_family)
annot_legend <- list(
  orientation = "h", # orientation
  x = 0.5, # Center horizontally
  y = -0.15, # Below the plot
  xanchor = "center",
  yanchor = "top",
  # traceorder = "normal",
  # tracegroupgap = 0,
  # itemwidth = 15,
  font = list(size = font_size_legend)
)
annot_xlab <- list(
  text = xlab,
  font = list(
    size = font_size_xlab
  ),
  xref = "paper",
  yref = "paper",
  xanchor = "center",
  yanchor = "top",
  x = 0.5, y = -.1,
  showarrow = FALSE
)

rm_grid <- function(pl) {
  plotly::layout(pl,
    xaxis = list(showgrid = FALSE, showline = TRUE),
    yaxis = list(showgrid = FALSE, showline = TRUE)
  )
}
```


```{r, echo = FALSE}
# Interactive functional-form explorer -----------------------------------
#
# Returns a single self-contained plotly htmlwidget that lets the reader
# toggle between four functional forms (constant, linear, hill with a
# negative slope, ricker) and adjust their parameters with sliders. The
# current configuration is shown as an sdbuildR-style equation, e.g.
#   eqn = ricker(engagement, location = 0.1, shape = 3)
#
# All curves are precomputed in R over parameter grids and embedded in the
# page as JSON, so the widget works in a static (embed-resources) HTML
# document without a Shiny server. Purely multiplicative/additive
# parameters (upper, intercept, constant value) are applied client-side as
# exact scaling of the precomputed unit curves, which keeps the embedded
# data small; the shape-determining parameters index the precomputed grid.
#
functional_forms_widget <- function(
  var = "engagement", 
height = 430,
xlab = tools::toTitleCase(var), ylab = "Motivation",
xlim = c(0, 1), ylim = c(-0.03, 1.08), n_points = 201, 
forms = c("constant", "linear", "hill", "ricker"),
arg_labels = list()
) {
  for (pkg in c("plotly", "htmlwidgets", "htmltools")) {
    if (!requireNamespace(pkg, quietly = TRUE)) {
      stop("Package '", pkg, "' is required.", call. = FALSE)
    }
  }

  forms_was_missing <- missing(forms)
  valid_forms <- c("constant", "linear", "hill", "ricker")
  if (!is.character(forms) || length(forms) == 0 || anyNA(forms)) {
    stop("`forms` must be a non-empty character vector.", call. = FALSE)
  }
  unknown_forms <- setdiff(forms, valid_forms)
  if (length(unknown_forms) > 0) {
    stop(
      "Unknown functional form(s): ", paste(unknown_forms, collapse = ", "),
      ". Use one or more of: ", paste(valid_forms, collapse = ", "), ".",
      call. = FALSE
    )
  }
  forms <- unique(forms)
  initial_form <- if (forms_was_missing) "ricker" else forms[[1]]

  # Per-form slider labels. Override the argument name shown next to each
  # slider via a named list of named vectors, e.g.
  #   arg_labels = list(constant = c(value = "recovery"),
  #                     linear = c(slope = "erosion rate"))
  if (!is.list(arg_labels)) {
    stop("`arg_labels` must be a named list of named character vectors.", call. = FALSE)
  }
  slider_labels <- list(
    constant = list(value = "value"),
    linear = list(slope = "slope"),
    hill = list(slope = "slope", midpoint = "midpoint", upper = "upper"),
    ricker = list(location = "location", shape = "shape", upper = "upper")
  )
  for (.form in names(arg_labels)) {
    if (!.form %in% names(slider_labels)) {
      stop("Unknown form in `arg_labels`: ", .form, call. = FALSE)
    }
    for (.arg in names(arg_labels[[.form]])) {
      if (!.arg %in% names(slider_labels[[.form]])) {
        stop("Unknown argument ", .arg, " for form ", .form, ".", call. = FALSE)
      }
      slider_labels[[.form]][[.arg]] <- arg_labels[[.form]][[.arg]]
    }
  }

  font_family <- getOption("sdbuildR.font_family", default = "stix-two-text")
  if (!is.character(font_family) || length(font_family) != 1 || !nzchar(font_family)) {
    font_family <- "stix-two-text"
  }
  font_family_css <- chartr('"', "'", font_family)
  webfont_css <- tryCatch(
    get("webfont_css", envir = asNamespace("sdbuildR"))(font_family),
    error = function(e) NULL
  )
  if (is.null(webfont_css)) {
    generic_fonts <- c("serif", "sans-serif", "monospace", "cursive", "fantasy", "system-ui")
    is_webfont <- grepl("^[a-z0-9]+(-[a-z0-9]+)*$", font_family) && !font_family %in% generic_fonts
    if (is_webfont) {
      webfont_file <- function(weight, style) {
        sprintf(
          "https://cdn.jsdelivr.net/npm/@fontsource/%s@5/files/%s-latin-%s-%s.woff2",
          font_family, font_family, weight, style
        )
      }
      files <- c(
        normal_400 = webfont_file(400, "normal"),
        normal_700 = webfont_file(700, "normal"),
        italic_400 = webfont_file(400, "italic"),
        italic_700 = webfont_file(700, "italic")
      )
      regular <- files[["normal_400"]]
      webfont_css <- paste(vapply(names(files), function(key) {
        style <- sub("_.*", "", key)
        weight <- sub(".*_", "", key)
        src <- sprintf("url(%s) format('woff2')", files[[key]])
        if (files[[key]] != regular) {
          src <- paste0(src, sprintf(", url(%s) format('woff2')", regular))
        }
        sprintf(
          "@font-face { font-family: '%s'; font-style: %s; font-weight: %s; src: %s; }",
          font_family, style, weight, src
        )
      }, character(1)), collapse = "\n")
    }
  }
  webfont_css <- paste0(if (is.null(webfont_css)) "" else webfont_css, "\n")

  x <- round(seq(xlim[1], xlim[2], length.out = n_points), 4)

  # Parameter grids over which curves are precomputed
  lin_slopes <- round(seq(0, 1, by = 0.05), 2)
  hill_slopes <- seq(-20, 20, by = 1)
  hill_midpoints <- round(seq(ylim[1], ylim[2], by = 0.05), 2)
  ricker_locations <- round(seq(0.05, 1, by = 0.05), 2)
  ricker_shapes <- c(0.25, 0.5, 0.75, 1, 1.5, 2, 3, 4, 5, 7, 10)

  rnd <- function(y) round(y, 3)
  grid_index <- function(values, target) which.min(abs(values - target)) - 1L

  lin_curves <- lapply(lin_slopes, function(s) rnd(s * x))

  # Flattened grids: index = slope_index * n_midpoints + midpoint_index
  # (0-based) on the JS side, so the inner loop must match the JS lookup.
  hill_curves <- list()
  for (s in hill_slopes) {
    for (m in hill_midpoints) {
      hill_curves[[length(hill_curves) + 1]] <- rnd(hill(x, slope = s, midpoint = m))
    }
  }

  ricker_curves <- list()
  for (l in ricker_locations) {
    for (sh in ricker_shapes) {
      ricker_curves[[length(ricker_curves) + 1]] <- rnd(ricker(x, location = l, shape = sh))
    }
  }

  wid <- paste0(
    "ffw-",
    paste(sample(c(letters, 0:9), 8, replace = TRUE), collapse = "")
  )

  # Initial state: matches the example eqn = ricker(engagement, location = .1,
  # shape = 3). Grid-based parameters are stored as 0-based indices.
  init_vals <- list(
    constant = list(value = 0.5),
    linear = list(
      slope = grid_index(lin_slopes, 0.5)
    ),
    hill = list(
      slope = grid_index(hill_slopes, 5),
      midpoint = grid_index(hill_midpoints, 0.5),
      upper = 1
    ),
    ricker = list(
      location = grid_index(ricker_locations, 0.1),
      shape = grid_index(ricker_shapes, 1),
      upper = 1
    )
  )
  initial_y <- switch(initial_form,
    constant = rep(init_vals$constant$value, length(x)),
    linear = {
      v <- init_vals$linear
      lin_curves[[v$slope + 1L]]
    },
    hill = {
      v <- init_vals$hill
      hill_curves[[v$slope * length(hill_midpoints) + v$midpoint + 1L]] * v$upper
    },
    ricker = {
      v <- init_vals$ricker
      ricker_curves[[v$location * length(ricker_shapes) + v$shape + 1L]] * v$upper
    }
  )
  initial_color <- c(
    constant = "#8E8E93", linear = "#0072B2",
    hill = "#D55E00", ricker = "#009E73"
  )[[initial_form]]

  payload <- list(
    id = wid,
    xvar = var,
    x = x,
    labels = slider_labels,
    linear = list(slopes = lin_slopes, curves = lin_curves),
    hill = list(
      slopes = hill_slopes, midpoints = hill_midpoints, curves = hill_curves
    ),
    ricker = list(
      locations = ricker_locations, shapes = ricker_shapes,
      curves = ricker_curves
    ),
    init = list(
      form = initial_form,
      vals = init_vals
    )
  )

  css <- htmltools::tags$style(htmltools::HTML(sprintf(r"---(%s
.ffw-controls { margin: 0 0 4px 0; max-width: 660px;
  font-family: "%s", serif; }
.ffw-tabs { display: inline-flex; border: 1px solid #cfd6dd;
  border-radius: 8px; overflow: hidden; margin-bottom: 10px; }
.ffw-tabs button { border: none; background: #fff; padding: 6px 16px;
  cursor: pointer; font: inherit; font-size: 14px; color: #333;
  border-right: 1px solid #cfd6dd; transition: background .15s, color .15s; }
.ffw-tabs button:last-child { border-right: none; }
.ffw-slider-row { display: flex; align-items: center; gap: 10px;
  margin: 4px 0; }
.ffw-slider-label { width: 80px; font-size: 14px; color: #333;
  text-align: right; }
.ffw-slider-row input[type=range] { flex: 1; max-width: 320px;
  accent-color: var(--ffw-col, #009E73); }
.ffw-slider-value { width: 56px; font-size: 13px; color: #555; }
.ffw-eqn { margin-top: 8px; }
.ffw-eqn code { display: inline-block; background: #f6f8fa;
  border: 1px solid #e4e8ec; border-left: 4px solid var(--ffw-col, #009E73);
  border-radius: 6px; padding: 6px 12px; font-size: 14px; }
)---", webfont_css, font_family_css)))

  controls <- htmltools::tags$div(
    id = wid, class = "ffw-controls",
    htmltools::tags$div(
      class = "ffw-tabs",
      lapply(forms, function(f) {
        htmltools::tags$button(type = "button", `data-form` = f, f)
      })
    ),
    htmltools::tags$div(class = "ffw-sliders"),
    htmltools::tags$div(class = "ffw-eqn", htmltools::tags$code("eqn = "))
  )

  js <- r"---(
function(el, x, data) {
  var D = data;
  var root = document.getElementById(D.id);
  if (!root) return;

  var COLORS = {
    constant: '#8E8E93', linear: '#0072B2',
    hill: '#D55E00', ricker: '#009E73'
  };

  // For grid-based parameters, state.vals holds the 0-based index into the
  // precomputed grid; for continuous parameters, the value itself.
  var state = { form: D.init.form, vals: D.init.vals };

  var SLIDERS = {
    constant: [
      { key: 'value', type: 'cont', min: 0, max: 1, step: 0.01 }
    ],
    linear: [
      { key: 'slope', type: 'grid', values: D.linear.slopes }
    ],
    hill: [
      { key: 'slope', type: 'grid', values: D.hill.slopes },
      { key: 'midpoint', type: 'grid', values: D.hill.midpoints },
      { key: 'upper', type: 'cont', min: 0, max: 1, step: 0.01 }
    ],
    ricker: [
      { key: 'location', type: 'grid', values: D.ricker.locations },
      { key: 'shape', type: 'grid', values: D.ricker.shapes },
      { key: 'upper', type: 'cont', min: 0, max: 1, step: 0.01 }
    ]
  };

  function val(form, key) {
    var cfgs = SLIDERS[form];
    for (var i = 0; i < cfgs.length; i++) {
      if (cfgs[i].key === key) {
        var s = state.vals[form][key];
        return cfgs[i].type === 'grid' ? cfgs[i].values[s] : s;
      }
    }
  }

  function computeY() {
    var f = state.form, v = state.vals[f], base;
    if (f === 'constant') {
      return D.x.map(function() { return v.value; });
    }
    if (f === 'linear') {
      return D.linear.curves[v.slope];
    }
    if (f === 'hill') {
      base = D.hill.curves[v.slope * D.hill.midpoints.length + v.midpoint];
      return base.map(function(y) { return y * v.upper; });
    }
    base = D.ricker.curves[v.location * D.ricker.shapes.length + v.shape];
    return base.map(function(y) { return y * v.upper; });
  }

  function fmt(n) {
    return parseFloat(n.toFixed(3)).toString();
  }

  function equation() {
    var f = state.form, xv = D.xvar, args;
    if (f === 'constant') {
      return 'eqn = ' + fmt(val(f, 'value'));
    }
    if (f === 'linear') {
      var s = val(f, 'slope');
      if (s === 0) return 'eqn = 0';
      return 'eqn = ' + (s < 0 ? '-' : '') + fmt(Math.abs(s)) + ' * ' + xv;
    }
    if (f === 'hill') {
      args = [xv,
        'slope = ' + fmt(val(f, 'slope')),
        'midpoint = ' + fmt(val(f, 'midpoint'))];
      if (val(f, 'upper') !== 1) args.push('upper = ' + fmt(val(f, 'upper')));
      return 'eqn = hill(' + args.join(', ') + ')';
    }
    // ricker: omit arguments left at their sdbuildR defaults
    args = [xv];
    if (val(f, 'location') !== 1) {
      args.push('location = ' + fmt(val(f, 'location')));
    }
    if (val(f, 'upper') !== 1) args.push('upper = ' + fmt(val(f, 'upper')));
    if (val(f, 'shape') !== 1) args.push('shape = ' + fmt(val(f, 'shape')));
    return 'eqn = ricker(' + args.join(', ') + ')';
  }

  function update() {
    Plotly.restyle(el, { y: [computeY()], 'line.color': COLORS[state.form] }, [0]);
    root.querySelector('.ffw-eqn code').textContent = equation();
  }

  var buttons = root.querySelectorAll('.ffw-tabs button');

  function setActive() {
    root.style.setProperty('--ffw-col', COLORS[state.form]);
    buttons.forEach(function(b) {
      var on = b.dataset.form === state.form;
      b.classList.toggle('ffw-active', on);
      b.style.background = on ? COLORS[state.form] : '';
      b.style.color = on ? '#fff' : '';
    });
  }

  function renderSliders() {
    var box = root.querySelector('.ffw-sliders');
    box.innerHTML = '';
    SLIDERS[state.form].forEach(function(cfg) {
      var row = document.createElement('div');
      row.className = 'ffw-slider-row';

      var lab = document.createElement('span');
      lab.className = 'ffw-slider-label';
      var lbls = D.labels[state.form];
      lab.textContent = (lbls && lbls[cfg.key]) || cfg.key;

      var input = document.createElement('input');
      input.type = 'range';
      if (cfg.type === 'grid') {
        input.min = 0;
        input.max = cfg.values.length - 1;
        input.step = 1;
      } else {
        input.min = cfg.min;
        input.max = cfg.max;
        input.step = cfg.step;
      }
      input.value = state.vals[state.form][cfg.key];

      var out = document.createElement('span');
      out.className = 'ffw-slider-value';
      out.textContent =
        fmt(cfg.type === 'grid' ? cfg.values[+input.value] : +input.value);

      input.addEventListener('input', function() {
        var raw = +input.value;
        state.vals[state.form][cfg.key] = raw;
        out.textContent = fmt(cfg.type === 'grid' ? cfg.values[raw] : raw);
        update();
      });

      row.appendChild(lab);
      row.appendChild(input);
      row.appendChild(out);
      box.appendChild(row);
    });
  }

  buttons.forEach(function(btn) {
    btn.addEventListener('click', function() {
      state.form = btn.dataset.form;
      setActive();
      renderSliders();
      update();
    });
  });

  setActive();
  renderSliders();
  update();
}
)---"

  p <- plotly::plot_ly(
    x = x,
    y = initial_y,
    type = "scatter", mode = "lines",
    line = list(width = 3, color = initial_color),
    hovertemplate = paste0(var, " = %{x:.2f}<br>eqn = %{y:.3f}<extra></extra>"),
    height = height
  )
  p <- plotly::layout(
    p,
    xaxis = list(
      title = xlab, range = c(0, 1), zeroline = FALSE, fixedrange = TRUE
    ),
    yaxis = list(
      title = ylab, range = c(-0.03, 1.08),
      zeroline = FALSE, fixedrange = TRUE
    ),
    showlegend = FALSE,
    margin = list(l = 55, r = 15, t = 10, b = 45)
  ) |>
  plotly::layout(font = utils::modifyList(annot_font, list(family = font_family)))
  p <- plotly::config(p, displayModeBar = FALSE)
  p <- htmlwidgets::prependContent(p, css, controls)
  htmlwidgets::onRender(p, js, data = payload)
}
```


```{r, echo = FALSE}
#' Combine animated sdbuildR plots into one subplot with a linked animation
#'
#' plotly::subplot() does not merge animation frames, so this builds each
#' plot, builds the subplot, and reconstructs the frames: for every frame
#' name, the trace updates of all plots are concatenated, with trace indices
#' offset to the subplot's trace order and re-anchored to the subplot's axes.
#' The animation controls (slider + play button) of the first plot are
#' reused, so the sdbuildR styling is preserved. All plots must share the
#' same animation frames (e.g. come from the same simulation).
subplot_linked_animation <- function(..., nrows = 2, heights = NULL,
                                     margin = 0.08, shareX = TRUE,
                                     titleY = TRUE) {
  plots <- list(...)
  builds <- lapply(plots, plotly::plotly_build)

  frame_names <- lapply(builds, function(b) {
    vapply(b$x$frames, function(f) as.character(f$name), character(1))
  })
  if (length(frame_names[[1]]) == 0) {
    stop("The first plot has no animation frames.")
  }
  same <- vapply(frame_names[-1], identical, logical(1), frame_names[[1]])
  if (!all(same)) {
    stop("All plots must have identical animation frames (same time grid).")
  }

  sp <- plotly::subplot(plots,
    nrows = nrows, heights = heights, margin = margin,
    shareX = shareX, titleY = titleY
  )
  sp <- plotly::plotly_build(sp)

  n_traces <- vapply(builds, function(b) length(b$x$data), integer(1))
  offsets <- cumsum(c(0L, utils::head(n_traces, -1L)))

  sp$x$frames <- lapply(seq_along(frame_names[[1]]), function(i) {
    fr <- list(name = frame_names[[1]][i], data = list(), traces = integer(0))
    for (k in seq_along(builds)) {
      bfr <- builds[[k]]$x$frames[[i]]
      idx <- unlist(bfr$traces)
      for (j in seq_along(idx)) {
        tr <- bfr$data[[j]]
        new_idx <- offsets[k] + idx[j]
        base <- sp$x$data[[new_idx + 1L]]
        tr$xaxis <- base$xaxis
        tr$yaxis <- base$yaxis
        fr$data[[length(fr$data) + 1L]] <- tr
        fr$traces <- c(fr$traces, new_idx)
      }
    }
    fr
  })

  # One set of animation controls for the whole subplot, in package style
  sp$x$layout$sliders <- builds[[1]]$x$layout$sliders
  sp$x$layout$updatemenus <- builds[[1]]$x$layout$updatemenus

  # subplot() merges layouts with later plots winning, so a plot with
  # main = "" wipes out the first plot's title; restore the first non-empty one
  titles <- lapply(builds, function(b) b$x$layout$title)
  has_title <- vapply(titles, function(tt) {
    txt <- if (is.list(tt)) tt$text else tt
    !is.null(txt) && nzchar(txt)
  }, logical(1))
  if (any(has_title)) {
    sp$x$layout$title <- titles[[which(has_title)[1]]]
  }
  sp
}

#' Fill the area between a trace and zero, preserved across animation frames
#'
#' Sets fill = "tozeroy" on the matched trace(s) of a (possibly animated)
#' plotly figure, and re-applies the fill in every animation frame so the
#' shaded area grows together with the cumulative-reveal line. By default the
#' fill colour is a translucent version of the trace's own line colour, so
#' the package palette is preserved; pass fillcolor (e.g. increase_fill) to
#' override.
#'
#' @param sp Plotly object (e.g. from subplot_linked_animation()).
#' @param trace Trace name(s) (matched against trace $name) or 1-based indices.
#' @param fillcolor Fill colour; NULL (default) derives it from the line colour.
#' @param alpha Opacity of the derived fill colour. Defaults to 0.25.
fill_to_zero <- function(sp, trace, fillcolor = NULL, alpha = 0.25) {
  sp <- plotly::plotly_build(sp)

  nm <- vapply(sp$x$data, function(tr) {
    if (is.null(tr$name)) "" else as.character(tr$name)
  }, character(1))
  idx <- if (is.character(trace)) which(nm %in% trace) else as.integer(trace)
  if (length(idx) == 0) {
    stop("No trace matches: ", paste(trace, collapse = ", "),
         ". Available: ", paste(unique(nm), collapse = ", "))
  }

  translucent <- function(col, alpha) {
    if (grepl("^rgba?\\(", col)) {
      nums <- as.numeric(strsplit(gsub("rgba?\\(|\\)", "", col), ",")[[1]])
      sprintf("rgba(%d,%d,%d,%s)", nums[1], nums[2], nums[3], alpha)
    } else {
      grDevices::adjustcolor(col, alpha.f = alpha)
    }
  }

  for (i in idx) {
    fc <- fillcolor
    if (is.null(fc)) {
      lc <- sp$x$data[[i]]$line$color
      if (is.null(lc)) lc <- "#1f77b4"
      fc <- translucent(lc, alpha)
    }
    sp$x$data[[i]]$fill <- "tozeroy"
    sp$x$data[[i]]$fillcolor <- fc
    for (f in seq_along(sp$x$frames)) {
      hit <- which(unlist(sp$x$frames[[f]]$traces) == (i - 1L))
      for (h in hit) {
        sp$x$frames[[f]]$data[[h]]$fill <- "tozeroy"
        sp$x$frames[[f]]$data[[h]]$fillcolor <- fc
      }
    }
  }
  sp
}

```





## Formalizing Theory with Stock-and-Flow Models

<!-- 
The **stock** value is the amount of something at a moment in time. A **flow** is the process that changes that amount per unit of time.

A stock represents the accumulated history of its net flow.


Stocks are **states** -->

```{r, echo = FALSE, message=FALSE}
library(sdbuildR)
```



```{r, echo=FALSE}

{
  viz_str <- '
   digraph sfm {

      graph [layout = dot, rankdir = LR, center=true, outputorder="edgesfirst", pad=%(pad)s, nodesep=0.3, splines = true, concentrate = false]

      # Shared across all nodes (persists until overridden)
      node [fontsize=18,fontname="Times New Roman",fontcolor="black"]

      # Define stock nodes
      node [shape=box,style=filled,fillcolor="#83d3d4"]
      "Stock" [id="Stock",label="Stock", tooltip = "Stock: Stock\nInitial value: 0\nInflows: Inflow\nOutflows: Outflow"]

      # Define flow nodes (intermediate nodes for flows)
      node [style = "",shape=plaintext, fontsize=16, width=0.6, height=0.3]
      "Inflow" [id="Inflow",label="Inflow", tooltip = "Flow: Inflow\nEquation: 0\nFrom: outside model boundary\nTo: Stock"]
        "Outflow" [id="Outflow",label="Outflow", tooltip = "Flow: Outflow\nEquation: 0\nFrom: Stock\nTo: outside model boundary"]

      # Define external cloud nodes
      node [shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
      "Cloud1" [label=%(label_sink)s, tooltip = "Outside model boundary\nSink of: Outflow"]
        "Cloud2" [label=%(label_source)s, tooltip = "Outside model boundary\nSource of: Inflow"]

      # Define auxiliary nodes
      
      

      # Define constant nodes
      
      

      # Define flow edges (stock -> flow_node)
      edge [style = "", arrowhead="none", color="black:#f48153:black", penwidth=1.1, minlen=1, tailport="e", headport="w"]
      "Cloud2" -> "Inflow"
        "Stock" -> "Outflow"

      # Define flow edges (flow_node -> stock)
      edge [style = "", arrowhead="normal", color="black:#f48153:black", arrowsize=1.5, penwidth=1.1, minlen=1, tailport="e", headport="w"]
      "Inflow" -> "Stock"
        "Outflow" -> "Cloud1"

      # Define dependency edges
      edge [style = "", color="#999999", arrowsize=0.8, penwidth=1, constraint=false, tailport = "_", headport="_"]
      

      


      # Rank groupings
      

    }
  '

  viz_str <- sprintf_arg(viz_str, list(
    pad = ".9",
    label_source = '<Unspecified source<BR/><FONT POINT-SIZE="13" COLOR="black">(outside of model boundary)</FONT><BR/><BR/><BR/><BR/><FONT COLOR="white">.</FONT>>',
    label_sink = '<Unspecified sink<BR/><FONT POINT-SIZE="13" COLOR="black">(outside of model boundary)</FONT><BR/><BR/><BR/><BR/><FONT COLOR="white">.</FONT>>'
  ))

}

  pl <- DiagrammeR::grViz(viz_str)
  pl
```



## A Concrete Example: A Queue of People Waiting for Service

```{r, fig.height=2}
sfm <- stockflow() |>
  stock(queue, eqn = 0, label = "People Waiting in Queue") |>
  sim_settings(stop = 10, time_units = "hours", only_stocks = FALSE) |>
  meta(name = "")

plot(sfm)
```

```{r}
sfm |> simulate() |> plot()
```


### Inflows Increase a Stock

<!-- 
What **process** causes the state to change? A stock increases with inflows and decreases with outflows.
 -->

```{r, fig.height=2}
sfm <- sfm |>
  flow(arrivals, eqn = 1, to = queue, label = "Arrivals")

plot(sfm)
```



```{r, fig.height=5}
sim <- simulate(sfm)

pl1 <- plot(sim, vars = "queue", animation = "time", control_options = list(frame_ms = frame_ms), 
colors = stock_col,
# showlegend = FALSE, main = "People Waiting in Queue"
showlegend = TRUE
)

pl2 <- plot(sim, vars = "arrivals", animation = "time", control_options = list(frame_ms = frame_ms),
colors = inflow_col,
# showlegend = FALSE, main = ""
showlegend = TRUE
)

pl <- subplot_linked_animation(pl1, pl2, nrows = 2) |>
  fill_to_zero("Arrivals") 
# Update traceorder in legend to normal
pl$x$layout$legend$traceorder <- "normal"

pl
```

<!-- 
Note: stocks **accumulate** over time. 

Note: don't specify origin of source 

Flows are **instantaneous** rates. You have to simulate over time to see how they accumulate into stocks. 

Note: fractional people (simplification and unavoidable with stock-and-flow models in continuous time). -->


```{r}
df <- as.data.frame(sim, direction = "wide")
# Rename columns with variable labels
df_prop <- as.data.frame(sfm, properties = "label")
idx <- match(names(df)[-1], df_prop$name)
names(df)[-1] <- df_prop$label[idx]
names(df)[1] <- "Time"
gt::gt(head(df))
```

<!-- Ask audience question: What happens when we double the flow rate? -->
<!-- Ask audience question: What happens when we set the flow rate to zero? -->



### Outflows Decrease a Stock

What **processes** cause the state to change?

```{r}
sfm <- sfm |>
  flow(service, eqn = .5, from = queue, label = "Service")

plot(sfm)
```


```{r, fig.height=5}
sim <- simulate(sfm)

pl1 <- plot(sim, vars = "queue", animation = "time", control_options = list(frame_ms = frame_ms), 
colors = stock_col,
# showlegend = FALSE, main = "People Waiting in Queue"
showlegend = TRUE
)


pl2 <- plot(sim, vars = c("arrivals", "service"), animation = "time", control_options = list(frame_ms = frame_ms),
colors = c(inflow_col, outflow_col),
ylim = c(0, 1.1),
# showlegend = FALSE, main = ""
showlegend = TRUE
)

pl <- subplot_linked_animation(pl1, pl2, nrows = 2) |>
  fill_to_zero("Arrivals") |>
  fill_to_zero("Service") 
# Update traceorder in legend to normal
pl$x$layout$legend$traceorder <- "normal"

pl
```


```{r}
df <- as.data.frame(sim, direction = "wide")
# Rename columns with variable labels
df_prop <- as.data.frame(sfm, properties = "label")
idx <- match(names(df)[-1], df_prop$name)
names(df)[-1] <- df_prop$label[idx]
names(df)[1] <- "Time"
gt::gt(head(df))
```


### What Happens When the Outflow Is Larger Than the Inflow?

```{r}
sfm <- update(sfm, service, eqn = 2)

plot(sfm)
```


```{r, fig.height=5}
sim <- simulate(sfm)

pl1 <- plot(sim, vars = "queue", animation = "time", control_options = list(frame_ms = frame_ms), 
colors = stock_col,
# showlegend = FALSE, main = "People Waiting in Queue"
showlegend = TRUE
)


pl2 <- plot(sim, 
vars = c("arrivals", "service"),
 animation = "time", control_options = list(frame_ms = frame_ms),
colors = c(inflow_col, outflow_col),
ylim = c(0, 2.1),
# showlegend = FALSE, main = ""
showlegend = TRUE
)

pl <- subplot_linked_animation(pl1, pl2, nrows = 2) |>
  fill_to_zero("Arrivals") |>
  fill_to_zero("Service") 
# Update traceorder in legend to normal
pl$x$layout$legend$traceorder <- "normal"

pl
```



```{r}
df <- as.data.frame(sim, direction = "wide")
# Rename columns with variable labels
df_prop <- as.data.frame(sfm, properties = "label")
idx <- match(names(df)[-1], df_prop$name)
names(df)[-1] <- df_prop$label[idx]
names(df)[1] <- "Time"
gt::gt(head(df))
```


### Making the Outflow Stock-Dependent

```{r}
sfm <- update(sfm, service, eqn = .5 * queue)
plot(sfm)
```


This is a **negative feedback loop**: the more people, the greater the outflow, which reduces the amount of people in the queue.

A **positive** feedback loop **amplifies**, whereas a **negative** feedback loop **dampens** or **balances**.




```{r, fig.height=5}
sim <- simulate(sfm)

pl1 <- plot(sim, vars = "queue", animation = "time", control_options = list(frame_ms = frame_ms), 
colors = stock_col,
# showlegend = FALSE, main = "People Waiting in Queue"
showlegend = TRUE
)


pl2 <- plot(sim, 
vars = c("arrivals", "service"),
 animation = "time", control_options = list(frame_ms = frame_ms),
colors = c(inflow_col, outflow_col),
ylim = c(0, 1.1),
# showlegend = FALSE, main = ""
showlegend = TRUE
)

pl <- subplot_linked_animation(pl1, pl2, nrows = 2) |>
  fill_to_zero("Arrivals") |>
  fill_to_zero("Service") 
# Update traceorder in legend to normal
pl$x$layout$legend$traceorder <- "normal"

pl
```



### Functional Forms

**Functional form**: mathematical relationship specifying how one variable affects another.

```{r}
#| echo: false
functional_forms_widget(forms = c("constant", "linear"), 
ylab = "Service", xlab = "People Waiting in Queue", var = "queue",
arg_labels = list(constant = c(value = "service rate"),
                    linear   = c(slope = "service rate")))
```



### Multiple Outflows

```{r}
sfm <- sfm |>
  flow(leave, eqn = .1 * queue^2, from = queue, 
  label = "Life Is Too Short")

plot(sfm)
```

```{r, fig.height=5}
sim <- simulate(sfm)

pl1 <- plot(sim, vars = "queue", animation = "time", control_options = list(frame_ms = frame_ms), 
colors = stock_col,
# showlegend = FALSE, main = "People Waiting in Queue"
showlegend = TRUE
)


pl2 <- plot(sim, 
vars = c("arrivals", "service", "leave"),
 animation = "time", control_options = list(frame_ms = frame_ms),
colors = c(inflow_col, outflow_col, decrease_col_strong),
ylim = c(0, 1.1),
# showlegend = FALSE, main = ""
showlegend = TRUE
)

pl <- subplot_linked_animation(pl1, pl2, nrows = 2) |>
  fill_to_zero("Arrivals") |>
  fill_to_zero("Service") |>
  fill_to_zero("Life Is Too Short")
# Update traceorder in legend to normal
pl$x$layout$legend$traceorder <- "normal"

pl
```

### Multiple Stocks


```{r}
sfm <- sfm |>
stock(served, eqn = 0, label = "People Served") |>
update(service, to = served)

plot(sfm)
```




```{r, fig.height=5}
sim <- simulate(sfm)

pl1 <- plot(sim, vars = c("queue", "served"), animation = "time", control_options = list(frame_ms = frame_ms), 
colors = c(stock_col, stock2_col),
# showlegend = FALSE, main = "People Waiting in Queue"
showlegend = TRUE
)


pl2 <- plot(sim, 
vars = c("arrivals", "service", "leave"),
 animation = "time", control_options = list(frame_ms = frame_ms),
colors = c(inflow_col, outflow_col, decrease_col_strong),
ylim = c(0, 1.1),
# showlegend = FALSE, main = ""
showlegend = TRUE
)

pl <- subplot_linked_animation(pl1, pl2, nrows = 2) |>
  fill_to_zero("Arrivals") |>
  fill_to_zero("Service") |>
  fill_to_zero("Life Is Too Short")
# Update traceorder in legend to normal
pl$x$layout$legend$traceorder <- "normal"

pl
```


### Add a Constant for Service Rate

```{r}
sfm <- sfm |> 
  constant(service_rate, eqn = .5, label = "Service Rate") |>
  update(service, eqn = service_rate * queue)

order_vars <- c("leave", "service", "service_rate")
plot(sfm, show_constants = TRUE, align = order_vars, order = order_vars, spacing = .5)
```

### Add an Auxiliary for Satisfaction

```{r}
sfm <- sfm |> 
  aux(satisfaction, eqn = service / (service + leave), label = "Satisfaction")

order_vars <- list(c("leave", "service", "service_rate"), c("served", "satisfaction"))
plot(sfm, show_constants = TRUE, show_aux = TRUE, align = order_vars, order = order_vars, spacing = .5)

```




```{r, fig.height=5}
sim <- simulate(sfm)

pl1 <- plot(sim, vars = c("queue", "served"), animation = "time", control_options = list(frame_ms = frame_ms), 
colors = c(stock_col, stock2_col),
# showlegend = FALSE, main = "People Waiting in Queue"
showlegend = TRUE
)


pl2 <- plot(sim, 
vars = c("arrivals", "service", "leave", "satisfaction"),
 animation = "time", control_options = list(frame_ms = frame_ms),
colors = c(inflow_col, outflow_col, decrease_col_strong, aux_col),
ylim = c(0, 1.1),
# showlegend = FALSE, main = ""
showlegend = TRUE
)

pl <- subplot_linked_animation(pl1, pl2, nrows = 2) |>
  fill_to_zero("Arrivals") |>
  fill_to_zero("Service") |>
  fill_to_zero("Life Is Too Short")
# Update traceorder in legend to normal
pl$x$layout$legend$traceorder <- "normal"

pl
```





### Recap

Stock-and-flow models **formalize the processes by which states change** over time. 

<!-- The change in a stock is the difference between inflows and outflows -->



```{r, echo=FALSE}

{
  # pl <- plot(sfm, show_eqn = FALSE, show_tooltip = FALSE)
  # cat(pl$x$diagram)

  viz_str <- '
   digraph sfm {

      graph [layout = dot, rankdir = LR, center=true, outputorder="edgesfirst", pad=%(pad)s, nodesep=0.3, splines = true, concentrate = false]

      # Shared across all nodes (persists until overridden)
      node [fontsize=18,fontname="Times New Roman",fontcolor="black"]

      # Define stock nodes
      node [shape=box,style=filled,fillcolor="#83d3d4"]
      "Stock" [id="Stock",label="Stock", tooltip = "Stock: Stock\nInitial value: 0\nInflows: Inflow\nOutflows: Outflow"]

      # Define flow nodes (intermediate nodes for flows)
      node [style = "",shape=plaintext, fontsize=16, width=0.6, height=0.3]
      "Inflow" [id="Inflow",label="Inflow", tooltip = "Flow: Inflow\nEquation: 0\nFrom: outside model boundary\nTo: Stock"]
        "Outflow" [id="Outflow",label="Outflow", tooltip = "Flow: Outflow\nEquation: 0\nFrom: Stock\nTo: outside model boundary"]

      # Define external cloud nodes
      node [shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
      "Cloud1" [label=%(label_sink)s, tooltip = "Outside model boundary\nSink of: Outflow"]
        "Cloud2" [label=%(label_source)s, tooltip = "Outside model boundary\nSource of: Inflow"]

      # Define auxiliary nodes
      
      

      # Define constant nodes
      
      

      # Define flow edges (stock -> flow_node)
      edge [style = "", arrowhead="none", color="black:#f48153:black", penwidth=1.1, minlen=1, tailport="e", headport="w"]
      "Cloud2" -> "Inflow"
        "Stock" -> "Outflow"

      # Define flow edges (flow_node -> stock)
      edge [style = "", arrowhead="normal", color="black:#f48153:black", arrowsize=1.5, penwidth=1.1, minlen=1, tailport="e", headport="w"]
      "Inflow" -> "Stock"
        "Outflow" -> "Cloud1"

      # Define dependency edges
      edge [style = "", color="#999999", arrowsize=0.8, penwidth=1, constraint=false, tailport = "_", headport="_"]
      

      


      # Rank groupings
      

    }
  '

  viz_str <- sprintf_arg(viz_str, list(
    pad = ".9",
    label_source = '<Unspecified source<BR/><FONT POINT-SIZE="13" COLOR="black">(outside of model boundary)</FONT><BR/><BR/><BR/><BR/><FONT COLOR="white">.</FONT>>',
    label_sink = '<Unspecified sink<BR/><FONT POINT-SIZE="13" COLOR="black">(outside of model boundary)</FONT><BR/><BR/><BR/><BR/><FONT COLOR="white">.</FONT>>'
  ))

}

  pl <- DiagrammeR::grViz(viz_str)
  pl
```






## Applied Examples of Stock-and-Flow Models

This framework can be applied to model a wide range of phenomena.


```{r}
sfm <- stockflow("SIR")
plot(sfm, show_eqn = FALSE)
sfm |> simulate() |> plot()
```


```{r}
sfm <- stockflow("predator_prey") |>
  sim_settings(stop = 50, time_units = "days")
sfm |> simulate() |> plot()
```



```{r}
sfm <- stockflow("fitzhugh_nagumo") |>
  meta(name = "FitzHugh-Nagumo (Neuronal Dynamics)")
sfm |> simulate() |> plot()
```


## Stock-and-Flow Models to Build Psychological Theory

```{r}
stockflow() |>
  stock(Stock, label = "Knowledge") |>
  flow(Inflow, label = "Learning", to = Stock) |>
  flow(Outflow, label = "Forgetting", from = Stock) |>
  plot(show_eqn = FALSE)
```




```{r}
stockflow() |>
  stock(Stock, label = "Self-Esteem") |>
  flow(Inflow, label = "Positive Experiences", to = Stock) |>
  flow(Outflow, label = "Negative Experiences", from = Stock) |>
  plot(show_eqn = FALSE)
```


```{r}
stockflow() |>
  stock(Stock, label = "Fear") |>
  flow(Inflow, label = "Perceived Threat", to = Stock) |>
  flow(Outflow, label = "Coping Mechanisms", from = Stock) |>
  flow(Outflow2, label = "Homeostatic Regulation", from = Stock) |>
  plot(show_eqn = FALSE, wrap_width = 15)
```



## Package Setup

The workshop materials are available at <https://kcevers.github.io/sdbuildR/articles/summerschool2026.html>.

Install the package from GitHub:

```{r, eval = FALSE}
if (!requireNamespace("remotes", quietly = TRUE)) install.packages("remotes")
remotes::install_github("kcevers/sdbuildR")
```


```{r, echo=TRUE, message=FALSE}
library(sdbuildR)
```


```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE
)
```


Load the stock-and-flow model of people waiting in a queue from the package's model library:

```{r}
sfm <- stockflow("queue")
```

`sfm` stands for **s**tock-and-**f**low **m**odel.

Look at the model's structure:

```{r}
print(sfm)
```

Or inspect it in data frame format:

```{r}
as.data.frame(sfm, properties = "eqn")
```

Plot the stock-and-flow diagram:

```{r}
plot(sfm, show_constants = TRUE)
```

Simulate the model and plot the resulting time series:

```{r}
sfm |> simulate() |> plot()
```

The pipe operator `|>` passes the result on its left into the function on its right. For example, `x |> f(y)` is interpreted as `f(x, y)`.
 
The timeseries can also be inspected in data frame format:

```{r}
sfm |> simulate() |> as.data.frame(direction = "wide") |> head()
```



## A Simplified Model of Burnout

Draw the **target phenomenon**:

<!-- On board: draw overshoot and collapse pattern of engagement over six months, time unit day -->



```{r, warning = FALSE, message = FALSE, echo = FALSE, fig.height=5}

fig <- plotly::plot_ly()
fig <- plotly::layout(fig, dragmode="drawopenpath", 
title = 'Burnout') |>
plotly::layout(
xaxis = list(title = "Time (days)", range = c(0, 180), dtick = 30),
yaxis = list(title = "Engagement", range = c(0, 4),
dtick = 1),
font = annot_font
)

fig <- fig |>
  plotly::config(modeBarButtonsToAdd = list(
  "drawopenpath", 
  "eraseshape" ) )

fig
```






### Building a Model from Scratch

Create an empty stock-and-flow model:

```{r}
sfm <- stockflow()
print(sfm)
```

Change name of the model:

```{r}
sfm <- meta(sfm, name = "Burnout")
```


### Defining the Time Horizon and Time Unit

Change simulation settings:

```{r}
sfm <- sim_settings(sfm, 

  # Run simulation for 6 months (~ 180 days)
  stop = round(365/2), time_unit = "days", 

  # Return all variables in output (not just stocks)
  only_stocks = FALSE)

print(sfm)
```


### Adding a Stock

```{r}
sfm <- stock(sfm, name = engagement, eqn = .3)
```

Plot stock-and-flow diagram:

```{r}
sfm |> plot()
```

Simulate and visualise timeseries:

```{r}
sfm |> simulate() |> plot()
```


### Adding an Outflow

```{r}
sfm <- constant(sfm, decay_rate, eqn = .05) |>
       flow(decay, eqn = decay_rate, from = engagement)
```


Plot stock-and-flow diagram:

```{r}
sfm |> plot()
```

Simulate and visualise timeseries:

```{r}
sfm |> simulate() |> plot()
```

Animate the simulation over time:

```{r}
sfm |> simulate() |> plot(animation = "time")
```


### Stock-Dependent Outflow

```{r}
sfm <- update(sfm, decay, eqn = decay_rate * engagement)
```


Plot stock-and-flow diagram:

```{r}
sfm |> plot()
```

Simulate and visualise timeseries:

```{r}
sfm |> simulate() |> plot()
```


### Adding an Inflow

```{r}
sfm <- constant(sfm, enjoyment, eqn = .3) |>
       flow(motivation, eqn = enjoyment, to = engagement)
```


Plot stock-and-flow diagram:

```{r}
sfm |> plot()
```

Simulate and visualise timeseries:

```{r}
sfm |> simulate() |> plot()
```

The stock settles at a steady state where inflow == outflow.

### Dynamic Inflow Rate

What if work enjoyment is not static, but erodes over time? Put differently, what if work enjoyment is not a constant, but a stock?

```{r}
sfm <- change_type(sfm, enjoyment, new_type = "stock")
sfm |> plot()
```


### Adding an Outflow from Enjoyment

```{r}
sfm <- sfm |>
    flow(overcommitment, 
         eqn = enjoyment * new_projects, 
         from = enjoyment) |>
    aux(new_projects, eqn = .1 * engagement)
```


Plot stock-and-flow diagram:

```{r}
sfm |> plot()
```

Simulate and visualise timeseries:

```{r}
sfm |> simulate() |> plot()
```

### Recap

Iteratively, we have built a stock-and-flow model with:

- A stock for engagement
- A constant outflow, updated to a stock-dependent outflow
- A constant inflow, updated to an inflow with a dynamic rate

In other words, the inflow rate changed from an **exogenous to an endogenous variable**; from a constant to a stock that erodes over time.




## Dependence on Initial Condition

If you did not complete all steps above, load the model from the model library:

```{r}
sfm <- stockflow("burnout")
```

Initialize the engagement stock with a random value between 0 and 4:

```{r}
sfm <- update(sfm, engagement, eqn = runif(1, min = 0, max = 4)) 
```

Equations are stored as expressions. Note the initial value of engagement:

```{r}
as.data.frame(sfm, properties = "eqn")
```

This means that each simulation will have a different initial value for engagement:

```{r}
sfm |> simulate() |> plot()
```


## Systematically Investigating Dependence on Initial Conditions

Update simulation settings to prepare for ensemble simulations:

```{r}
sfm <- sfm |>

  # Save only 50 time points 
  sim_settings(save_length = 50, 

  # Save only stocks (not flows or auxiliaries)
  only_stocks = TRUE, 

  # Save individual simulation runs (not just summary statistics)
  save_sims = TRUE)
```

Run 50 simulations with different initial conditions:

```{r}
sims <- ensemble(sfm, n = 50)
```

Summary statistics:

```{r}
plot(sims)
```

Individual simulation runs:

```{r}
plot(sims, which = "sims")
```



## Assignment: Expressing Theoretical Ideas with Equations


### Set-up

Let's return to the simpler model with a static recovery rate:

```{r}
sfm <- sfm |>

 # Remove the outflow and auxiliary
 discard(c(overcommitment, new_projects)) |>

 # Change recovery_rate back to a constant
 change_type(enjoyment, new_type = "constant")

sfm |> plot(show_constants = TRUE)
```

This model has a constant inflow. The functional form of the motivation inflow can be changed to explore different theoretical assumptions about how motivation depends on engagement. 


### Functional Forms


```{r, echo = FALSE}
functional_forms_widget()
```

### Questions

Explore the functional forms above:

1. What theoretical ideas does each functional form express? 


2. How do they differ in terms of the dynamics they produce?


3. Is the behaviour dependent on the initial condition of energy?



## Bonus Material


### Parallel Simulations

If you have a multi-core machine, you can run simulations in parallel to speed up simulations:

```{r}
# Set up parallel execution (3 cores less than the total number of available cores)
n_workers <- max(c(1, parallel::detectCores() - 3))
future::plan(future::multisession, workers = n_workers)
```

If configured, `ensemble()` will automatically run simulations in parallel. Otherwise, it will run sequentially.

```{r}
sfm <- sim_settings(sfm, save_sims = TRUE)
```

Run 100 simulations:

```{r}
sims <- ensemble(sfm, n = 100)
```

Individual simulation runs:

```{r}
plot(sims, which = "sims")
```



### Parameter Sensitivity

Ensemble simulations can also be used to explore the sensitivity of the model to changes in parameter values.

```{r}
# Reload model from library to reset any changes made above
sfm <- stockflow("burnout", version = 2) |>
  update(engagement, eqn = runif(1, min = 0, max = 4)) |>
  sim_settings(save_length = 50, only_stocks = TRUE, save_sims = TRUE)

# Specify a range of values of decay_rate to vary
conditions <- list(decay_rate = c(0.005, 0.01, 0.05, 0.1, 0.2))

# Run 50 simulations for each condition
sims <- ensemble(sfm, n = 50, conditions = conditions)
```

```{r}
plot(sims, which = "sims", condition_display = "slider")
```

If parallel execution was configured above, restore sequential execution:

```{r}
future::plan(future::sequential)
```


