---
title: "Theory Building in Psychology (Summer School 2026)"
output: rmarkdown::html_vignette
bibliography: references/references.bib
vignette: >
  %\VignetteIndexEntry{Theory Building in Psychology (Summer School 2026)}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: inline
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE, # Hide code by default
  fig.width = 6, # Set default figure width to 6 inches
  fig.height = 4, # Set default figure height to 3 inches
  fig.align = "center", # Center all figures
  out.width = "80%", # Display figures at 80% of container width
  cache.path = "summerschool2026_cache/" # keep cache dir next to the .Rmd
  # fig.retina = 2,         # Optimize for retina displays
  # dpi = 96               # Standard DPI for web
)
```

<style type="text/css">
body {
  line-height: 1.55;
}

h2 {
  margin-top: 2.75rem;
  margin-bottom: 1rem;
  padding-top: 0.15rem;
}

h3 {
  margin-top: 2rem;
  margin-bottom: 0.75rem;
}

h2 + h3 {
  margin-top: 1.4rem;
}

p,
ul,
ol {
  margin-bottom: 0.95rem;
}

pre,
.sourceCode,
.figure,
.html-widget,
.plotly {
  margin-top: 0.85rem;
  margin-bottom: 1.4rem;
}
</style>

```{r, echo=FALSE, message=FALSE}
library(sdbuildR)
```

```{r, echo = FALSE, cache=FALSE}
# Disable WebGL: many plotly widgets per HTML page can exceed the browser WebGL
# context limit and render blank. SVG always renders.
options(sdbuildR.webgl = FALSE)
```


```{r, echo = FALSE}
font_family <- "stix-two-text"
options(sdbuildR.font_family = font_family)
# getOption("sdbuildR.font_family")
frame_ms <- 200
```


These materials were developed for a stock-and-flow modelling workshop given at the Summer School [Theory Building in Psychology](https://theorymethodssociety.org/summer_school.html) (University of Amsterdam, July 2026; rendered materials can be accessed [here](https://kcevers.github.io/sdbuildR/articles/summerschool2026.html)). 
As the materials supported a lecture, they contain limited written explanations. For more detailed guidance on stock-and-flow modelling, see the [Build](https://kcevers.github.io/sdbuildR/articles/build.html) vignette.


## Package Setup

First install the package from GitHub:

```{r, echo = TRUE, eval = FALSE}
if (!requireNamespace("remotes", quietly = TRUE)) install.packages("remotes")
remotes::install_github("kcevers/sdbuildR")
```


```{r, echo=TRUE, message=FALSE}
library(sdbuildR)
```


```{r include=FALSE}
# Helper function
sprintf_arg <- function(template, values) {
  for (nm in names(values)) {
    template <- gsub(paste0("%\\(", nm, "\\)s"), values[[nm]], template)
  }
  return(template)
}

# Plotting parameters
PALETTE <- c("Geyser", "Temps", "Spectral", "Zissou 1", "Roma")[5]
COLORS <- grDevices::hcl.colors(n = 50, palette = PALETTE)

inflow_col <- adjustcolor(COLORS[10], alpha.f = 1)
outflow_col <- adjustcolor(COLORS[3], alpha.f = 1)

inflow_fill <- adjustcolor(inflow_col, alpha.f = .5)
outflow_fill <- adjustcolor(outflow_col, alpha.f = .5)

decrease_col_strong <- adjustcolor(COLORS[1], alpha.f = 1)

COLORS_alpha <- adjustcolor(COLORS, alpha.f = .25)

COLORS <- COLORS[c(37, 9, 32, 1, 47, 17)]
stock_col <- adjustcolor(COLORS[1], alpha.f = .75)
stock2_col <- adjustcolor(COLORS[2], alpha.f = .75)
aux_col <- adjustcolor(COLORS[3], alpha.f = .75)
flow_col <- COLORS[6]

font_size <- 16

# Annotation template
annot_font <- list(size = font_size, family = font_family)
```


```{r, echo = FALSE}
# Interactive functional-form explorer -----------------------------------
#
# Returns a single self-contained plotly htmlwidget that lets the reader
# toggle between four functional forms (constant, linear, hill with a
# negative slope, ricker) and adjust their parameters with sliders. The
# current configuration is shown as an sdbuildR-style equation, e.g.
#   eqn = ricker(engagement, location = 0.1, shape = 3)
#
# All curves are precomputed in R over parameter grids and embedded in the
# page as JSON, so the widget works in a static (embed-resources) HTML
# document without a Shiny server. Purely multiplicative/additive
# parameters (upper, intercept, constant value) are applied client-side as
# exact scaling of the precomputed unit curves, which keeps the embedded
# data small; the shape-determining parameters index the precomputed grid.
#
functional_forms_widget <- function(
  var = "engagement",
  height = 430,
  xlab = tools::toTitleCase(var), ylab = "Motivation",
  xlim = c(0, 1), ylim = c(-0.03, 1.08), n_points = 201,
  forms = c("constant", "linear", "hill", "ricker"),
  arg_labels = list()
) {
  for (pkg in c("plotly", "htmlwidgets", "htmltools")) {
    if (!requireNamespace(pkg, quietly = TRUE)) {
      stop("Package '", pkg, "' is required.", call. = FALSE)
    }
  }

  forms_was_missing <- missing(forms)
  valid_forms <- c("constant", "linear", "hill", "ricker")
  if (!is.character(forms) || length(forms) == 0 || anyNA(forms)) {
    stop("`forms` must be a non-empty character vector.", call. = FALSE)
  }
  unknown_forms <- setdiff(forms, valid_forms)
  if (length(unknown_forms) > 0) {
    stop(
      "Unknown functional form(s): ", paste(unknown_forms, collapse = ", "),
      ". Use one or more of: ", paste(valid_forms, collapse = ", "), ".",
      call. = FALSE
    )
  }
  forms <- unique(forms)
  initial_form <- if (forms_was_missing) "ricker" else forms[[1]]

  # Per-form slider labels. Override the argument name shown next to each
  # slider via a named list of named vectors, e.g.
  #   arg_labels = list(constant = c(value = "recovery"),
  #                     linear = c(slope = "erosion rate"))
  if (!is.list(arg_labels)) {
    stop("`arg_labels` must be a named list of named character vectors.", call. = FALSE)
  }
  slider_labels <- list(
    constant = list(value = "value"),
    linear = list(slope = "slope"),
    hill = list(slope = "slope", midpoint = "midpoint", upper = "upper"),
    ricker = list(location = "location", shape = "shape", upper = "upper")
  )
  for (.form in names(arg_labels)) {
    if (!.form %in% names(slider_labels)) {
      stop("Unknown form in `arg_labels`: ", .form, call. = FALSE)
    }
    for (.arg in names(arg_labels[[.form]])) {
      if (!.arg %in% names(slider_labels[[.form]])) {
        stop("Unknown argument ", .arg, " for form ", .form, ".", call. = FALSE)
      }
      slider_labels[[.form]][[.arg]] <- arg_labels[[.form]][[.arg]]
    }
  }

  font_family <- getOption("sdbuildR.font_family", default = "stix-two-text")
  if (!is.character(font_family) || length(font_family) != 1 || !nzchar(font_family)) {
    font_family <- "stix-two-text"
  }
  font_family_css <- chartr('"', "'", font_family)
  webfont_css <- tryCatch(
    get("webfont_css", envir = asNamespace("sdbuildR"))(font_family),
    error = function(e) NULL
  )
  if (is.null(webfont_css)) {
    generic_fonts <- c("serif", "sans-serif", "monospace", "cursive", "fantasy", "system-ui")
    is_webfont <- grepl("^[a-z0-9]+(-[a-z0-9]+)*$", font_family) && !font_family %in% generic_fonts
    if (is_webfont) {
      webfont_file <- function(weight, style) {
        sprintf(
          "https://cdn.jsdelivr.net/npm/@fontsource/%s@5/files/%s-latin-%s-%s.woff2",
          font_family, font_family, weight, style
        )
      }
      files <- c(
        normal_400 = webfont_file(400, "normal"),
        normal_700 = webfont_file(700, "normal"),
        italic_400 = webfont_file(400, "italic"),
        italic_700 = webfont_file(700, "italic")
      )
      regular <- files[["normal_400"]]
      webfont_css <- paste(vapply(names(files), function(key) {
        style <- sub("_.*", "", key)
        weight <- sub(".*_", "", key)
        src <- sprintf("url(%s) format('woff2')", files[[key]])
        if (files[[key]] != regular) {
          src <- paste0(src, sprintf(", url(%s) format('woff2')", regular))
        }
        sprintf(
          "@font-face { font-family: '%s'; font-style: %s; font-weight: %s; src: %s; }",
          font_family, style, weight, src
        )
      }, character(1)), collapse = "\n")
    }
  }
  webfont_css <- paste0(if (is.null(webfont_css)) "" else webfont_css, "\n")

  x <- round(seq(xlim[1], xlim[2], length.out = n_points), 4)

  # Parameter grids over which curves are precomputed
  lin_slopes <- round(seq(0, 1, by = 0.05), 2)
  hill_slopes <- seq(-20, 20, by = 1)
  hill_midpoints <- round(seq(ylim[1], ylim[2], by = 0.05), 2)
  ricker_locations <- round(seq(0.05, 1, by = 0.05), 2)
  ricker_shapes <- c(0.25, 0.5, 0.75, 1, 1.5, 2, 3, 4, 5, 7, 10)

  rnd <- function(y) round(y, 3)
  grid_index <- function(values, target) which.min(abs(values - target)) - 1L

  lin_curves <- lapply(lin_slopes, function(s) rnd(s * x))

  # Flattened grids: index = slope_index * n_midpoints + midpoint_index
  # (0-based) on the JS side, so the inner loop must match the JS lookup.
  hill_curves <- list()
  for (s in hill_slopes) {
    for (m in hill_midpoints) {
      hill_curves[[length(hill_curves) + 1]] <- rnd(hill(x, slope = s, midpoint = m))
    }
  }

  ricker_curves <- list()
  for (l in ricker_locations) {
    for (sh in ricker_shapes) {
      ricker_curves[[length(ricker_curves) + 1]] <- rnd(ricker(x, location = l, shape = sh))
    }
  }

  wid <- paste0(
    "ffw-",
    paste(sample(c(letters, 0:9), 8, replace = TRUE), collapse = "")
  )

  # Initial state: matches the example eqn = ricker(engagement, location = .1,
  # shape = 3). Grid-based parameters are stored as 0-based indices.
  init_vals <- list(
    constant = list(value = 0.5),
    linear = list(
      slope = grid_index(lin_slopes, 0.5)
    ),
    hill = list(
      slope = grid_index(hill_slopes, 5),
      midpoint = grid_index(hill_midpoints, 0.5),
      upper = 1
    ),
    ricker = list(
      location = grid_index(ricker_locations, 0.1),
      shape = grid_index(ricker_shapes, 1),
      upper = 1
    )
  )
  initial_y <- switch(initial_form,
    constant = rep(init_vals$constant$value, length(x)),
    linear = {
      v <- init_vals$linear
      lin_curves[[v$slope + 1L]]
    },
    hill = {
      v <- init_vals$hill
      hill_curves[[v$slope * length(hill_midpoints) + v$midpoint + 1L]] * v$upper
    },
    ricker = {
      v <- init_vals$ricker
      ricker_curves[[v$location * length(ricker_shapes) + v$shape + 1L]] * v$upper
    }
  )
  initial_color <- c(
    constant = "#8E8E93", linear = "#0072B2",
    hill = "#D55E00", ricker = "#009E73"
  )[[initial_form]]

  payload <- list(
    id = wid,
    xvar = var,
    x = x,
    labels = slider_labels,
    linear = list(slopes = lin_slopes, curves = lin_curves),
    hill = list(
      slopes = hill_slopes, midpoints = hill_midpoints, curves = hill_curves
    ),
    ricker = list(
      locations = ricker_locations, shapes = ricker_shapes,
      curves = ricker_curves
    ),
    init = list(
      form = initial_form,
      vals = init_vals
    )
  )

  css <- htmltools::tags$style(htmltools::HTML(sprintf(r"---(%s
.ffw-controls { margin: 0 0 4px 0; max-width: 660px;
  font-family: "%s", serif; }
.ffw-tabs { display: inline-flex; border: 1px solid #cfd6dd;
  border-radius: 8px; overflow: hidden; margin-bottom: 10px; }
.ffw-tabs button { border: none; background: #fff; padding: 6px 16px;
  cursor: pointer; font: inherit; font-size: 14px; color: #333;
  border-right: 1px solid #cfd6dd; transition: background .15s, color .15s; }
.ffw-tabs button:last-child { border-right: none; }
.ffw-slider-row { display: flex; align-items: center; gap: 10px;
  margin: 4px 0; }
.ffw-slider-label { width: 80px; font-size: 14px; color: #333;
  text-align: right; }
.ffw-slider-row input[type=range] { flex: 1; max-width: 320px;
  accent-color: var(--ffw-col, #009E73); }
.ffw-slider-value { width: 56px; font-size: 13px; color: #555; }
.ffw-eqn { margin-top: 8px; }
.ffw-eqn code { display: inline-block; background: #f6f8fa;
  border: 1px solid #e4e8ec; border-left: 4px solid var(--ffw-col, #009E73);
  border-radius: 6px; padding: 6px 12px; font-size: 14px; }
)---", webfont_css, font_family_css)))

  controls <- htmltools::tags$div(
    id = wid, class = "ffw-controls",
    htmltools::tags$div(
      class = "ffw-tabs",
      lapply(forms, function(f) {
        htmltools::tags$button(type = "button", `data-form` = f, f)
      })
    ),
    htmltools::tags$div(class = "ffw-sliders"),
    htmltools::tags$div(class = "ffw-eqn", htmltools::tags$code("eqn = "))
  )

  js <- r"---(
function(el, x, data) {
  var D = data;
  var root = document.getElementById(D.id);
  if (!root) return;

  var COLORS = {
    constant: '#8E8E93', linear: '#0072B2',
    hill: '#D55E00', ricker: '#009E73'
  };

  // For grid-based parameters, state.vals holds the 0-based index into the
  // precomputed grid; for continuous parameters, the value itself.
  var state = { form: D.init.form, vals: D.init.vals };

  var SLIDERS = {
    constant: [
      { key: 'value', type: 'cont', min: 0, max: 1, step: 0.01 }
    ],
    linear: [
      { key: 'slope', type: 'grid', values: D.linear.slopes }
    ],
    hill: [
      { key: 'slope', type: 'grid', values: D.hill.slopes },
      { key: 'midpoint', type: 'grid', values: D.hill.midpoints },
      { key: 'upper', type: 'cont', min: 0, max: 1, step: 0.01 }
    ],
    ricker: [
      { key: 'location', type: 'grid', values: D.ricker.locations },
      { key: 'shape', type: 'grid', values: D.ricker.shapes },
      { key: 'upper', type: 'cont', min: 0, max: 1, step: 0.01 }
    ]
  };

  function val(form, key) {
    var cfgs = SLIDERS[form];
    for (var i = 0; i < cfgs.length; i++) {
      if (cfgs[i].key === key) {
        var s = state.vals[form][key];
        return cfgs[i].type === 'grid' ? cfgs[i].values[s] : s;
      }
    }
  }

  function computeY() {
    var f = state.form, v = state.vals[f], base;
    if (f === 'constant') {
      return D.x.map(function() { return v.value; });
    }
    if (f === 'linear') {
      return D.linear.curves[v.slope];
    }
    if (f === 'hill') {
      base = D.hill.curves[v.slope * D.hill.midpoints.length + v.midpoint];
      return base.map(function(y) { return y * v.upper; });
    }
    base = D.ricker.curves[v.location * D.ricker.shapes.length + v.shape];
    return base.map(function(y) { return y * v.upper; });
  }

  function fmt(n) {
    return parseFloat(n.toFixed(3)).toString();
  }

  function equation() {
    var f = state.form, xv = D.xvar, args;
    if (f === 'constant') {
      return 'eqn = ' + fmt(val(f, 'value'));
    }
    if (f === 'linear') {
      var s = val(f, 'slope');
      if (s === 0) return 'eqn = 0';
      return 'eqn = ' + (s < 0 ? '-' : '') + fmt(Math.abs(s)) + ' * ' + xv;
    }
    if (f === 'hill') {
      args = [xv,
        'slope = ' + fmt(val(f, 'slope')),
        'midpoint = ' + fmt(val(f, 'midpoint'))];
      if (val(f, 'upper') !== 1) args.push('upper = ' + fmt(val(f, 'upper')));
      return 'eqn = hill(' + args.join(', ') + ')';
    }
    // ricker: omit arguments left at their sdbuildR defaults
    args = [xv];
    if (val(f, 'location') !== 1) {
      args.push('location = ' + fmt(val(f, 'location')));
    }
    if (val(f, 'upper') !== 1) args.push('upper = ' + fmt(val(f, 'upper')));
    if (val(f, 'shape') !== 1) args.push('shape = ' + fmt(val(f, 'shape')));
    return 'eqn = ricker(' + args.join(', ') + ')';
  }

  function update() {
    Plotly.restyle(el, { y: [computeY()], 'line.color': COLORS[state.form] }, [0]);
    root.querySelector('.ffw-eqn code').textContent = equation();
  }

  var buttons = root.querySelectorAll('.ffw-tabs button');

  function setActive() {
    root.style.setProperty('--ffw-col', COLORS[state.form]);
    buttons.forEach(function(b) {
      var on = b.dataset.form === state.form;
      b.classList.toggle('ffw-active', on);
      b.style.background = on ? COLORS[state.form] : '';
      b.style.color = on ? '#fff' : '';
    });
  }

  function renderSliders() {
    var box = root.querySelector('.ffw-sliders');
    box.innerHTML = '';
    SLIDERS[state.form].forEach(function(cfg) {
      var row = document.createElement('div');
      row.className = 'ffw-slider-row';

      var lab = document.createElement('span');
      lab.className = 'ffw-slider-label';
      var lbls = D.labels[state.form];
      lab.textContent = (lbls && lbls[cfg.key]) || cfg.key;

      var input = document.createElement('input');
      input.type = 'range';
      if (cfg.type === 'grid') {
        input.min = 0;
        input.max = cfg.values.length - 1;
        input.step = 1;
      } else {
        input.min = cfg.min;
        input.max = cfg.max;
        input.step = cfg.step;
      }
      input.value = state.vals[state.form][cfg.key];

      var out = document.createElement('span');
      out.className = 'ffw-slider-value';
      out.textContent =
        fmt(cfg.type === 'grid' ? cfg.values[+input.value] : +input.value);

      input.addEventListener('input', function() {
        var raw = +input.value;
        state.vals[state.form][cfg.key] = raw;
        out.textContent = fmt(cfg.type === 'grid' ? cfg.values[raw] : raw);
        update();
      });

      row.appendChild(lab);
      row.appendChild(input);
      row.appendChild(out);
      box.appendChild(row);
    });
  }

  buttons.forEach(function(btn) {
    btn.addEventListener('click', function() {
      state.form = btn.dataset.form;
      setActive();
      renderSliders();
      update();
    });
  });

  setActive();
  renderSliders();
  update();
}
)---"

  p <- plotly::plot_ly(
    x = x,
    y = initial_y,
    type = "scatter", mode = "lines",
    line = list(width = 3, color = initial_color),
    hovertemplate = paste0(var, " = %{x:.2f}<br>eqn = %{y:.3f}<extra></extra>"),
    height = height
  )
  p <- plotly::layout(
    p,
    xaxis = list(
      title = xlab, range = c(0, 1), zeroline = FALSE, fixedrange = TRUE
    ),
    yaxis = list(
      title = ylab, range = c(-0.03, 1.08),
      zeroline = FALSE, fixedrange = TRUE
    ),
    showlegend = FALSE,
    margin = list(l = 55, r = 15, t = 10, b = 45)
  ) |>
    plotly::layout(font = utils::modifyList(annot_font, list(family = font_family)))
  p <- plotly::config(p, displayModeBar = FALSE)
  p <- htmlwidgets::prependContent(p, css, controls)
  htmlwidgets::onRender(p, js, data = payload)
}
```


## Conceptual Understanding of Stock-and-Flow Models




```{r, echo=FALSE}
{
  viz_str <- '
   digraph sfm {

      graph [layout = dot, rankdir = LR, center=true, outputorder="edgesfirst", pad=%(pad)s, nodesep=0.3, splines = true, concentrate = false]

      # Shared across all nodes (persists until overridden)
      node [fontsize=18,fontname="%(font_family)s",fontcolor="black"]

      # Define stock nodes
      node [shape=box,style=filled,fillcolor="#83d3d4"]
      "Stock" [id="Stock",label="Stock", tooltip = "Stock: Stock\nInitial value: 0\nInflows: Inflow\nOutflows: Outflow"]

      # Define flow nodes (intermediate nodes for flows)
      node [style = "",shape=plaintext, fontsize=16, width=0.6, height=0.3]
      "Inflow" [id="Inflow",label="Inflow", tooltip = "Flow: Inflow\nEquation: 0\nFrom: outside model boundary\nTo: Stock"]
        "Outflow" [id="Outflow",label="Outflow", tooltip = "Flow: Outflow\nEquation: 0\nFrom: Stock\nTo: outside model boundary"]

      # Define external cloud nodes
      node [shape=doublecircle, fixedsize=true, width = .25, height = .25, orientation=15]
      "Cloud1" [label=%(label_sink)s, tooltip = "Outside model boundary\nSink of: Outflow"]
      "Cloud2" [label=%(label_source)s, tooltip = "Outside model boundary\nSource of: Inflow"]

      # Define auxiliary nodes



      # Define constant nodes



      # Define flow edges (stock -> flow_node)
      edge [style = "", arrowhead="none", color="black:#f48153:black", penwidth=1.1, minlen=1, tailport="e", headport="w"]
      "Cloud2" -> "Inflow"
        "Stock" -> "Outflow"

      # Define flow edges (flow_node -> stock)
      edge [style = "", arrowhead="normal", color="black:#f48153:black", arrowsize=1.5, penwidth=1.1, minlen=1, tailport="e", headport="w"]
      "Inflow" -> "Stock"
        "Outflow" -> "Cloud1"

      # Define dependency edges
      edge [style = "", color="#999999", arrowsize=0.8, penwidth=1, constraint=false, tailport = "_", headport="_"]





      # Rank groupings


    }
  '

  viz_str <- sprintf_arg(viz_str, list(
    pad = ".9",
    font_family = font_family,
    label_source = '<Unspecified source<BR/><FONT POINT-SIZE="13" COLOR="black">(outside of model boundary)</FONT><BR/><BR/><BR/><BR/><FONT COLOR="white">.</FONT>>',
    label_sink = '<Unspecified sink<BR/><FONT POINT-SIZE="13" COLOR="black">(outside of model boundary)</FONT><BR/><BR/><BR/><BR/><FONT COLOR="white">.</FONT>>'
  ))
}

pl <- DiagrammeR::grViz(viz_str)
pl
```



### A Concrete Example: A Queue of People Waiting for Service

In iterative steps, we construct a simple model of a queue of people waiting for service. Either scroll through the steps below by unfolding the section, or use the widget.

<details><summary>Scroll through steps</summary>


```{r, include=FALSE}

# Each named block above advances `sfm` by one step. Map each block to the
# title shown on its panel.
df_queue <- data.frame(
  section = c(
    "queue-stock", "queue-inflow", "queue-outflow",
    "queue-outflow-large", "queue-stock-dependent",
    "queue-multiple-outflows", "queue-multiple-stocks",
    # "queue-constant",
     "queue-aux"
  ),
  title = c(
    "A Stock with No Flows",
    "Inflows Increase a Stock",
    "Outflows Decrease a Stock",
    "What Happens When the Outflow Is Larger Than the Inflow?",
    "Making the Outflow Stock-Dependent",
    "Multiple Outflows",
    "Multiple Stocks",
    # "Add a Constant for Service Rate",
    "Add an Auxiliary for Satisfaction"
  ),
  stringsAsFactors = FALSE
)

df_queue$letter <- LETTERS[seq_len(nrow(df_queue))]
```

```{r queue-stock, echo = FALSE, fig.height=1}
sfm <- stockflow() |>
  stock(queue, eqn = 0, label = "People Waiting in Queue") |>
  sim_settings(stop = 10, time_units = "hours", only_stocks = FALSE) |>
  meta(name = "")
```

```{r, echo = FALSE, fig.height=1}
sfm |> plot()
```

```{r, echo = FALSE}
sfm |>
  simulate() |>
  plot(animation = "time", main = df_queue$title[1])
```


```{r queue-inflow, echo=FALSE}
sfm <- sfm |>
  flow(arrivals, eqn = 1, to = queue, label = "Arrivals")
```

```{r, echo = FALSE}
sfm |> plot()
```

```{r, echo = FALSE}
sfm |>
  simulate() |>
  plot(animation = "time", main = df_queue$title[2])
```


```{r queue-outflow, echo=FALSE}
sfm <- sfm |>
  flow(service, eqn = .5, from = queue, label = "Service")
```

```{r, echo = FALSE}
sfm |> plot()
```

```{r, echo = FALSE}
sfm |>
  simulate() |>
  plot(animation = "time", main = df_queue$title[3])
```


```{r queue-outflow-large, echo=FALSE}
sfm <- update(sfm, service, eqn = 2)
```

```{r, echo = FALSE}
sfm |> plot()
```

```{r, echo = FALSE}
sfm |>
  simulate() |>
  plot(animation = "time", main = df_queue$title[4])
```


```{r queue-stock-dependent, echo=FALSE}
sfm <- update(sfm, service, eqn = .5 * queue)
```

```{r, echo = FALSE}
sfm |> plot()
```

```{r, echo = FALSE}
sfm |>
  simulate() |>
  plot(animation = "time", main = df_queue$title[5])
```


```{r queue-multiple-outflows, echo=FALSE}
sfm <- sfm |>
  flow(leave,
    eqn = queue^10, from = queue,
    label = "Tired of Waiting"
  )
```

```{r, echo = FALSE}
sfm |> plot()
```

```{r, echo = FALSE}
sfm |>
  simulate() |>
  plot(animation = "time", main = df_queue$title[6])
```


```{r queue-multiple-stocks, echo=FALSE}
sfm <- sfm |>
  stock(served, eqn = 0, label = "People Served") |>
  update(service, to = served)
```

```{r, echo = FALSE}
sfm |> plot()
```

```{r, echo = FALSE}
sfm |>
  simulate() |>
  plot(animation = "time", main = df_queue$title[7])
```


```{r queue-aux, echo=FALSE}
sfm <- sfm |>
  aux(satisfaction, eqn = service / (service + leave), label = "Satisfaction")
```

```{r, echo = FALSE}
sfm |> plot()
```

```{r, echo = FALSE}
sfm |>
  simulate() |>
  plot(animation = "time", main = df_queue$title[8])
```


</details>



```{r queue-panels, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
library(bslib)

# Force the grViz SVG to scale to its container so the diagram fits the panel.
# viz.js emits a viewBox, so height:auto keeps the aspect ratio.
diagram_css <- htmltools::tags$style(htmltools::HTML(
  ".sfm-diagram { width: 100%; overflow-x: auto; }
   .sfm-diagram svg { width: 100% !important; height: auto !important; max-width: 100%; }"
))


# One colour per variable, reused across panels so each variable keeps its
# colour as the model grows.
queue_colors <- c(
  queue = stock_col, served = stock2_col,
  arrivals = inflow_col, service = outflow_col,
  leave = decrease_col_strong, satisfaction = aux_col
)

# Retrieve a named block's source text.
get_code <- function(label) paste(knitr::knit_code$get(label), collapse = "\n")

# Rebuild the model step by step, capturing a diagram and a simulation after
# each named block.
build_env <- new.env(parent = knitr::knit_global())
queue_panels <- lapply(seq_len(nrow(df_queue)), function(i) {
  eval(parse(text = get_code(df_queue$section[i])), envir = build_env)
  model <- build_env$sfm

  diagram <- plot(model, show_constants = TRUE, show_aux = TRUE)$x$diagram

  sim <- simulate(model)
  vars <- intersect(
    names(queue_colors),
    names(as.data.frame(sim, direction = "wide"))[-1]
  )
  pl <- plot(sim,
    vars = vars,
    animation = "time", main = df_queue$title[i],
    control_options = list(frame_ms = frame_ms),
    colors = queue_colors[vars],
    show_legend = TRUE,
    webgl = FALSE
  )

  nav_panel(
    df_queue$letter[i],
    pl,
    accordion(
      open = TRUE,
      accordion_panel(
        "Stock-and-flow diagram",
        htmltools::div(
          class = "sfm-diagram",
          DiagrammeR::grViz(diagram, width = "100%", height = "100%")
        )
      )
    )
  )
})

htmltools::tagList(
  diagram_css,
  do.call(navset_card_pill, c(
    queue_panels,
    list(selected = df_queue$letter[1])
  ))
)
```



### Recap


- Stock-and-flow models **formalize the processes by which states change** over time. 

- Stocks are **accumulations** of a quantity, and flows are the **processes that increase or decrease** those accumulations. The change in a stock is the **net flow** (inflow minus outflow) into that stock.

- A model can have multiple stocks, and each stock can have multiple inflows and outflows.

- Stocks are **continuous quantities**, meaning we can end up with fractional values (e.g., 3.5 people in a queue). This is a simplifying assumption.

- When the outflow depends on the stock, the model contains a **negative feedback loop**: the more people in the queue, the greater the outflow, which in turn shrinks the queue. A **positive** feedback loop **amplifies**; a **negative** feedback loop **dampens** or **balances**.




### Functional Forms

A **functional form** is the mathematical relationship specifying how one variable affects another. Above, we modified the service outflow from a constant to a linear function of the queue. This is visualized in the widget below:

```{r, include = TRUE, echo = FALSE, message=FALSE, warning=FALSE}
functional_forms_widget(
  forms = c("constant", "linear"),
  ylab = "Service", xlab = "People Waiting in Queue", var = "queue",
  arg_labels = list(
    constant = c(value = "service rate"),
    linear = c(slope = "service rate")
  )
)
```




## Applied Examples of Stock-and-Flow Models

The stock-and-flow framework can be applied to model a wide range of phenomena.


```{r}
sfm <- stockflow("SIR")
plot(sfm, show_eqn = FALSE)
sfm |>
  simulate() |>
  plot()
```


```{r}
sfm <- stockflow("predator_prey") |>
  sim_settings(stop = 50, time_units = "days")
sfm |>
  simulate() |>
  plot()
```



```{r}
sfm <- stockflow("fitzhugh_nagumo") |>
  meta(name = "FitzHugh-Nagumo (Neuronal Dynamics)")
sfm |>
  simulate() |>
  plot()
```


## Stock-and-Flow Models to Build Psychological Theory

Similarly, we may use stock-and-flow models to formalize psychological theory. For example, we can model the dynamics of knowledge acquisition and forgetting:

```{r}
stockflow() |>
  stock(Stock, label = "Knowledge") |>
  flow(Inflow, label = "Learning", to = Stock) |>
  flow(Outflow, label = "Forgetting", from = Stock) |>
  plot(show_eqn = FALSE)
```


Stock-and-flow models force us to specify how a state increases and decreases over time.

## Building Stock-and-Flow Models in R

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE, fig.height = 5
)
```


Load the stock-and-flow model of people waiting in a queue from the package's model library:

```{r}
sfm <- stockflow("queue")
```

`sfm` stands for **s**tock-and-**f**low **m**odel.

Look at the model's structure:

```{r}
print(sfm)
```

Or inspect it in data frame format:

```{r}
as.data.frame(sfm, properties = "eqn")
```

Plot the stock-and-flow diagram:

```{r}
plot(sfm, show_constants = TRUE)
```

Simulate the model and plot the resulting time series:

```{r}
sfm |>
  simulate() |>
  plot()
```

The pipe operator `|>` passes the result on its left into the function on its right. For example, `x |> f(y)` is interpreted as `f(x, y)`.
 
The timeseries can also be inspected in data frame format:

```{r}
sfm |>
  simulate() |>
  as.data.frame(direction = "wide") |>
  head()
```



## A Simplified Model of Burnout

We now use stock-and-flow modeling to formalize a simplified model of burnout. We first draw the **target phenomenon**:

<!-- On board: draw overshoot and collapse pattern of engagement over six months, time unit day -->



```{r, warning = FALSE, message = FALSE, echo = FALSE, fig.height=5}
fig <- plotly::plot_ly(
  x = numeric(0),
  y = numeric(0),
  type = "scatter",
  mode = "lines"
)
fig <- plotly::layout(fig,
  dragmode = "drawopenpath",
  title = "Burnout"
) |>
  plotly::layout(
    margin = list(t = 100),
    xaxis = list(title = "Time (days)", range = c(0, 180), dtick = 30),
    yaxis = list(
      title = "Engagement", range = c(0, 4),
      dtick = 1
    ),
    font = list(family = font_family, size = 16)
  )

fig <- fig |>
  plotly::config(modeBarButtonsToAdd = list(
    "drawopenpath",
    "eraseshape"
  ))

fig
```



Create an empty stock-and-flow model:

```{r}
sfm <- stockflow()
print(sfm)
```

Change name of the model:

```{r}
sfm <- meta(sfm, name = "Burnout")
```


### Defining the Time Horizon and Time Unit

Change simulation settings:

```{r}
sfm <- sim_settings(sfm,

  # Run simulation for 6 months (~ 180 days)
  stop = round(365 / 2), time_unit = "days",

  # Return all variables in output (not just stocks)
  only_stocks = FALSE
)

print(sfm)
```


### Adding a Stock

```{r}
sfm <- stock(sfm, name = engagement, eqn = .3)
```

Plot stock-and-flow diagram:

```{r, fig.height = 1}
sfm |> plot()
```

Simulate and visualise timeseries:

```{r}
sfm |>
  simulate() |>
  plot(animation = "time")
```


### Adding an Outflow

```{r}
sfm <- constant(sfm, decay_rate, eqn = .05) |>
  flow(decay, eqn = decay_rate, from = engagement)
```


Plot stock-and-flow diagram:

```{r}
sfm |> plot()
```

Simulate and visualise timeseries:

```{r}
sfm |>
  simulate() |>
  plot(animation = "time")
```


### Stock-Dependent Outflow

```{r}
sfm <- update(sfm, decay, eqn = decay_rate * engagement)
```


Plot stock-and-flow diagram:

```{r}
sfm |> plot()
```

Simulate and visualise timeseries:

```{r}
sfm |>
  simulate() |>
  plot(animation = "time")
```



### Adding an Inflow

```{r}
sfm <- constant(sfm, enjoyment, eqn = .3) |>
  flow(motivation, eqn = enjoyment, to = engagement)
```


Plot stock-and-flow diagram:

```{r}
sfm |> plot()
```

Simulate and visualise timeseries:

```{r}
sfm |>
  simulate() |>
  plot(animation = "time")
```

The stock settles at a steady state where inflow == outflow.

### Dynamic Inflow Rate

What if work enjoyment is not static, but erodes over time? Put differently, what if work enjoyment is not a constant, but a stock?

```{r}
sfm <- change_type(sfm, enjoyment, new_type = "stock")
sfm |> plot()
```


### Adding an Outflow from Enjoyment

```{r}
sfm <- sfm |>
  flow(overcommitment,
    eqn = enjoyment * new_projects,
    from = enjoyment
  ) |>
  aux(new_projects, eqn = .1 * engagement)
```


Plot stock-and-flow diagram:

```{r}
sfm |> plot()
```

Simulate and visualise timeseries:

```{r}
sfm |>
  simulate() |>
  plot(animation = "time")
```

### Recap

Iteratively, we have built a stock-and-flow model with:

- A stock for engagement
- A constant outflow, updated to a stock-dependent outflow
- A constant inflow, updated to an inflow with a dynamic rate

In other words, the inflow rate changed from an **exogenous to an endogenous variable**; from a constant to a stock that erodes over time.



The table below provides an overview of each model revision and the behaviour it produces. 


```{r echo = FALSE}
eqn_interpretation_table <- data.frame(
  Panel = LETTERS[1:5],
  `Stocks` = c("engagement", "engagement", "engagement", "engagement", "engagement, enjoyment"),
  `Constants` = c("", "`decay_rate`", "`decay_rate`", "`decay_rate`, `enjoyment`", "`decay_rate`"),
  `Recovery eqn (inflow)` = c("", "", "", "`enjoyment`", "`enjoyment`"),
  `Depletion eqn (outflow)` = c("", "`decay_rate`", "`decay_rate * engagement`", "`decay_rate * engagement`", "`decay_rate * engagement`"),
  Interpretation = c(
    "No process of change",
    "Engagement decreases at a constant rate",
    "Engagement decreases at a rate proportional to its current value",
    "Engagement changes at a rate equal to a constant minus a rate proportional to its current value",
    "Engagement recovers at a rate which itself changes over time"
  ),
  Behaviour = c(
    "Static",
    "Linear decrease",
    "Exponential decrease towards zero",
    "Stability when recovery and depletion are equal",
    "Rise and collapse"
  ),
  check.names = FALSE
)

kableExtra::kbl(eqn_interpretation_table,
  format = "markdown",
  caption = "Connecting Equations to Model Behaviour"
) |>
  kableExtra::kable_styling()
```



## Dependence on Initial Condition

If you did not complete all steps above, load the model from the model library:

```{r}
sfm <- stockflow("burnout")
```

Rather than a fixed value, initialize the engagement stock with a random value between 0 and 4:

```{r}
sfm <- update(sfm, engagement, eqn = runif(1, min = 0, max = 4))
```

Equations are stored as expressions. Note the initial value of engagement:

```{r}
as.data.frame(sfm, properties = "eqn")
```

This means that each simulation will have a different initial value for engagement:

```{r}
sfm |>
  simulate() |>
  plot()
```

```{r}
sfm |>
  simulate() |>
  plot()
```


## Systematically Investigating Dependence on Initial Conditions

Update simulation settings to prepare for ensemble simulations:

```{r}
sfm <- sfm |>
  sim_settings(
    # Save only 50 time points
    save_length = 50,

    # Save only stocks (not flows or auxiliaries)
    only_stocks = TRUE,

    # Save individual simulation runs (not just summary statistics)
    save_sims = TRUE
  )
```

Run 50 simulations with different initial conditions:

```{r ensemble-init, cache = TRUE}
sims <- ensemble(sfm, n = 50)
```

Summary statistics:

```{r}
plot(sims)
```

Individual simulation runs:

```{r}
plot(sims, which = "sims")
```


## Parameter Sensitivity

Ensemble simulations can also be used to explore the sensitivity of the model to changes in parameter values.

```{r ensemble-params, cache = TRUE}
# Reload model from library to reset any changes made above
sfm <- stockflow("burnout") |>
  update(engagement, eqn = runif(1, min = 0, max = 4)) |>
  sim_settings(save_length = 50, only_stocks = TRUE, save_sims = TRUE)

# Specify a range of values of decay_rate to vary
conditions <- list(decay_rate = c(0.005, 0.01, 0.05, 0.1, 0.2))

# Run 50 simulations for each condition
sims <- ensemble(sfm, n = 50, conditions = conditions)
```

```{r}
plot(sims, which = "sims", condition_display = "slider")
```


## Assignment: Expressing Theoretical Ideas with Equations


### Set-up

Let's return to the simpler model with a static recovery rate:

```{r}
sfm <- sfm |>

  # Remove the outflow and auxiliary
  discard(c(overcommitment, new_projects)) |>

  # Change recovery_rate back to a constant
  change_type(enjoyment, new_type = "constant")

sfm |> plot(show_constants = TRUE)
```

This model has a constant inflow. The functional form of the motivation inflow can be changed to explore different theoretical assumptions about how motivation depends on engagement. 


### Functional Forms


```{r, echo = FALSE, message=FALSE, warning=FALSE}
functional_forms_widget()
```

### Questions

Explore the functional forms above:

1. What theoretical ideas does each functional form express? 


2. How do they differ in terms of the dynamics they produce?


3. Is the behaviour dependent on the initial condition of energy?




## Parallel Simulations

If you have a multi-core machine, you can run simulations in parallel to speed up simulations:

```{r, eval = interactive()}
# Set up parallel execution (3 cores less than the total number of available cores)
n_workers <- max(c(1, parallel::detectCores() - 3))
future::plan(future::multisession, workers = n_workers)
```

If configured, `ensemble()` will automatically run simulations in parallel. Otherwise, it will run sequentially.

```{r}
sfm <- sim_settings(sfm, save_sims = TRUE)
```

Run 100 simulations:

```{r ensemble-parallel, cache = TRUE}
sims <- ensemble(sfm, n = 100)
```

Individual simulation runs:

```{r}
plot(sims, which = "sims")
```


If parallel execution was configured above, restore sequential execution:

```{r, eval = interactive()}
future::plan(future::sequential)
```



