---
title: "Unit Testing Stock-and-Flow Models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Unit Testing Stock-and-Flow Models}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  include = TRUE,
  eval = TRUE,
  comment = "#>",
  fig.width = 12, # 1
  fig.height = 4, # 3
  fig.align = "center", # Center all figures
  out.width = "80%" # Display figures at 80% of container width
)
```

```{r, echo=TRUE, message=FALSE}
library(sdbuildR)
```

Models are only as good as their assumptions and equations. Unit tests are an essential tool for verifying that your model behaves as expected under various conditions. As stock-and-flow models can easily become complex, unexpected behaviours can arise as the model is developed. In this vignette, we will demonstrate how to use `sdbuildR` to create and run unit tests for a simple SIR (Susceptible-Infected-Recovered) model.

## Load the model

We start by loading the SIR model and plotting its structure.

```{r}
sfm <- sdbuildR("SIR")
plot(sfm)
```

A quick simulation shows the model's behaviour.

```{r}
sim <- simulate(sfm)
plot(sim)
```

```{r}
diagnose(sfm)
```

`diagnose()` is a useful checkpoint before testing because it catches common
model definition issues early.

## Add simple unit tests

`unit_test()` adds one test at a time. A test expression should evaluate to
`TRUE` when the model behaves as expected.

Below we add three basic checks:

1. Susceptible population never becomes negative.
2. Recovered population should not decrease over time.
3. Total population remains constant in this closed SIR model.

```{r}
sfm <- sfm |>
  unit_test(
    expr = all(Susceptible >= 0),
    label = "Susceptible is non-negative"
  ) |>
  unit_test(
    expr = all(diff(Recovered) >= 0),
    label = "Recovered does not decrease"
  ) |>
  unit_test(
    expr = all(abs(Susceptible + Infected + Recovered - Total_Population) < 1e-8),
    label = "Population is conserved"
  )
```

You can inspect all defined tests with `unit_tests()`.

```{r}
unit_tests(sfm)
```

## Verify model behavior

Run all active tests with `verify()`.

```{r}
verify(sfm)
```

If a test fails, check whether the equation is wrong, the test is too strict,
or your model assumptions need to be revised. 

## Test behaviour under a condition

You can also test behaviour under a specific parameter setting using
`conditions`. Here, we set `Beta = 0`, so no new infections should occur.
Under this condition, the infected stock should not increase.

```{r}
sfm <- sfm |>
  unit_test(
    expr = all(diff(Infected) <= 0),
    label = "Infected does not increase when Beta is zero",
    conditions = list(Beta = 0)
  )

verify(sfm)
```


**explain == is usually fragile

** show how to view simulations

**show sim_index


## Update or remove tests

During model development, you may want to remove outdated tests. Use
`discard_unit_test()` by label or by number.

```{r}
sfm <- discard_unit_test(sfm, nr = 1)
unit_tests(sfm)
```

Unit tests are most useful when kept small, readable, and tied to a specific
model expectation.
