---
title: "Unit Testing Stock-and-Flow Models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Unit Testing Stock-and-Flow Models}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  include = TRUE,
  eval = TRUE,
  comment = "#>",
  fig.width = 12, # 1
  fig.height = 4, # 3
  fig.align = "center", # Center all figures
  out.width = "80%" # Display figures at 80% of container width
)
```

```{r, echo=TRUE, message=FALSE}
library(sdbuildR)
```

Stock-and-flow models can easily become complex, producing unexpected behaviours as the model is developed. To ensure a model behaves as intended, we may make use of unit tests, a concept from software engineering. A unit test compares the behaviour of a small aspect of the model (i.e., a unit) to an explicitly formulated expectation. By implementing an exhaustive test suite, we can systematically and routinely check the model. In this vignette, we will demonstrate how to use `sdbuildR` to create and run unit tests for a simple model.


## Load the model

We will demonstrate unit tests on the SIR (Susceptible-Infected-Recovered) model, a classic model in epidemiology of infectious diseases.

```{r}
sfm <- sdbuildR("SIR")
plot(sfm)
```

A quick simulation shows the model's behaviour.

```{r}
sim <- simulate(sfm)
plot(sim)
```

## Add simple unit tests

`unit_test()` adds one test at a time. A test expression should evaluate to
`TRUE` when the model behaves as expected. For example, we may test that the susceptible population never becomes negative, as people cannot be less than zero:

```{r}
sfm <- sfm |>
  unit_test(
    expr = all(Susceptible >= 0)
  ) 
```

We assess our test with `verify()`, which simulates the model and assess the test on the output data:

```{r}
verify(sfm)
```

The test passed, so we can be confident that the susceptible population does not become negative under the current model specification. As shown above, a test label has been automatically generated. A custom label can also be provided:

```{r}
sfm <- sfm |>
  unit_test(
    expr = all(abs(Susceptible + Infected + Recovered - Total_Population) < 1e-8),
    label = "Population is conserved"
  )

verify(sfm)
```

Unit tests are most useful when kept small, readable, and tied to a specific
model expectation.


## Test behaviour under a condition

You can also test behaviour under a specific parameter setting using
`conditions`. Here, we set `Beta = 0`, so no new infections should occur.
Under this condition, the infected stock should not increase.

```{r}
sfm <- sfm |>
  unit_test(
    expr = all(diff(Infected) <= 0),
    label = "Infected does not increase when Beta is zero",
    conditions = list(Beta = 0)
  )

verify(sfm)
```



## View defined tests

All defined tests can be viewed with `unit_tests()`.

```{r}
unit_tests(sfm)
```

As shown above, the tests are numbered and labeled. The `expr` column shows the test expression, and the `conditions` column shows any parameter conditions for the test. The `active` column indicates whether the test is currently active (i.e., will be run when calling `verify()`).

Unit tests can be filtered by their number:

```{r}
unit_tests(sfm, nr = 1)
```

Or using a regular expression for their label:

```{r}
unit_tests(sfm, label = "population")
```

```{r}
unit_tests(sfm, label = "susceptible")
```

```{r}
unit_tests(sfm, label = "conserved")
```

By default, `ignore_case = TRUE`, so the regular expression is case-insensitive. Set `ignore_case = FALSE` for a case-sensitive search.

## Modify an existing test

Existing tests can be modified by their numeric position (`nr`):

```{r}
# Modify test number 1 to be less strict
sfm <- unit_test(sfm, nr = 1, expr = all(Susceptible >= -1))

unit_tests(sfm, nr = 1)
```

Or their label (exact match):

```{r}
sfm <- unit_test(sfm, 
    expr = all(abs(Susceptible + Infected + Recovered - Total_Population) < 1e-5),
    label = "Population is conserved")

unit_tests(sfm, label = "Population is conserved")
```

## Deactivate / reactivate tests

Use the `active` flag to temporarily disable a test without removing it.

```{r}
sfm <- unit_test(sfm, nr = 1, active = FALSE)
verify(sfm)

# Reactivate
sfm <- unit_test(sfm, nr = 1, active = TRUE)
verify(sfm)
```

## Remove tests

Tests can be removed by number or by their (exact)label.

```{r}
sfm <- discard_unit_test(sfm, nr = 1)
unit_tests(sfm)
```


## Specify simulation settings for tests

By default, `verify()` runs a single simulation with the default settings. Different simulation settings (as in `sim_settings`) can be passed to `verify`:

```{r}
verify(sfm, seed = 123)
```

```{r}
verify(sfm, dt = .1)
```

## Ensemble unit tests

When there are stochastic elements in the model, it is useful to run unit tests multiple times. For example, the SIR model can be given random initial conditions:

```{r}
sfm <- sfm |>
  update(Susceptible, eqn = runif(1, 500, 1000)) |>
  update(Infected, eqn = runif(1, 0, 100)) |>
  update(Recovered, eqn = runif(1, 0, 100))
```

Use `n > 1` to run multiple simulations per test:

```{r}
verify(sfm, n = 100)
```

## Inspecting `verify()` results and returned simulations

`verify()` returns the simulations used when running tests if `return_sims = TRUE` is specified. These can be plotted or analyzed further.

```{r}
sfm <- sim_settings(sfm, return_sims = TRUE)
```

Or specify `return_sims` directly in `verify()`:

```{r}
res <- verify(sfm, return_sims = TRUE)
```

```{r}
plot(res)
```

Plot the simulation used for a specific test:

```{r}
plot(res, nr = 1L)
```

View results as a data frame:

```{r}
as.data.frame(res)
```

Retrieve results for a specific test by number:

```{r}
as.data.frame(res, nr = 1)
```

## Inspect failed tests


```{r}
plot(res, status = "fail")
```

```{r}
as.data.frame(res, status = "fail")
```
