---
title: "Unit Testing Stock-and-Flow Models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Unit Testing Stock-and-Flow Models}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  include = TRUE,
  eval = TRUE,
  comment = "#>",
  fig.width = 12, # 1
  fig.height = 4, # 3
  fig.align = "center", # Center all figures
  out.width = "80%" # Display figures at 80% of container width
)
```

```{r, echo=TRUE, message=FALSE}
library(sdbuildR)
```

Stock-and-flow models can easily become complex, producing unexpected behaviours as the model is developed. To ensure a model behaves as intended, we may make use of unit tests, a concept from software engineering. A unit test compares the behaviour of a small aspect of the model (i.e., a unit) to an explicitly formulated expectation. By implementing an exhaustive test suite, we can systematically and routinely check the model. In this vignette, we will demonstrate how to use `sdbuildR` to create and run unit tests for a simple model.


## Load the model

We will demonstrate unit tests on the SIR (Susceptible-Infected-Recovered) model, a classic model in epidemiology of infectious diseases.

```{r}
sfm <- sdbuildR("SIR")
plot(sfm)
```

A quick simulation shows the model's behaviour.

```{r}
sim <- simulate(sfm)
plot(sim)
```

## Add simple unit tests

`unit_test()` adds one test at a time. A test expression should evaluate to
`TRUE` when the model behaves as expected. For example, we may test that the susceptible population never becomes negative, as people cannot be less than zero:

```{r}
sfm <- sfm |>
  unit_test(
    expr = all(susceptible >= 0)
  ) 
```

Note that a test expression can refer to any variable in the model. It is evaluated on the simulated output data, meaning that `susceptible` refers to the entire timeseries of the susceptible population (not just its initial value). 

We assess our test with `verify()`, which simulates the model and assess the test on the output data:

```{r}
verify(sfm)
```

The test passed, so we can be confident that the susceptible population does not become negative under the current model specification. As shown above, a test label has been automatically generated. A custom label can also be provided, which we demonstrate with a test that checks that the total population is conserved (i.e., the sum of susceptible, infected, and recovered equals the total population at all times):

```{r}
sfm <- sfm |>
  unit_test(
    expr = all(abs(susceptible + infected + recovered - total_population) < 1e-8),
    label = "Population is conserved"
  )

verify(sfm)
```

Unit tests are most useful when kept small, readable, and tied to a specific
model expectation.

## Test behaviour under a condition

Intended model behaviour can also be tested under a specific scenario using
`conditions`. `conditions` should be a named list, and can only specify constants and initial conditions of stock variables.

```{r}
as.data.frame(sfm, type = c("constant", "stock"), 
properties = c("eqn", "label"))
```

Here, we set `recovery_rate = 0`, so the recovered stock should not increase:

```{r}
sfm <- sfm |>
  unit_test(
    expr = all(diff(recovered) == 0),
    label = "Recovered does not increase when recovery_rate is zero",
    conditions = list(recovery_rate = 0)
  )

verify(sfm)
```

Multiple conditions can be specified, which together define one scenario under which the expression is evaluated. In other words, all conditions are applied simultaneously. For example, when all stocks are initialized at zero (except for `recovered`, as this would lead to a `total_population` of zero and thus division by zero), they should remain at zero:

```{r}
sfm <- sfm |>
  unit_test(
    expr = all(susceptible == 0) & all(infected == 0) & all(diff(recovered) == 0),
    conditions = list(susceptible = 0, infected = 0, recovered = 1)
  )

verify(sfm)
```

## View defined tests

All defined tests can be viewed with `unit_tests()`.

```{r}
unit_tests(sfm)
```

Each test shows its number, label, expression, and any conditions. Inactive tests (those with `active = FALSE`) are shown with a dotted icon and are skipped when `verify()` is called.

Unit tests can be filtered by their number:

```{r}
unit_tests(sfm, nr = 1)
```

Or using a regular expression for their label:

```{r}
unit_tests(sfm, label = "population")
```

```{r}
unit_tests(sfm, label = "susceptible")
```

By default, `ignore_case = TRUE`, so the regular expression is case-insensitive. Set `ignore_case = FALSE` for a case-sensitive search.

## Modify an existing test

Existing tests can be modified by their numeric position (`nr`):

```{r}
# Modify label of test number 1
sfm <- unit_test(sfm, nr = 1, label = "Susceptible is non-negative")

unit_tests(sfm, nr = 1)
```

Or their label (exact match):

```{r}
sfm <- unit_test(sfm, 
    expr = all(abs(susceptible + infected + recovered - total_population) < 1e-5),
    label = "Population is conserved")

unit_tests(sfm, label = "Population is conserved")
```

## Deactivate / reactivate tests

Use the `active` flag to temporarily disable a test without removing it.

```{r}
sfm <- unit_test(sfm, nr = 1, active = FALSE)
verify(sfm)
```

```{r}
# Reactivate
sfm <- unit_test(sfm, nr = 1, active = TRUE)
verify(sfm)
```

## Remove tests

Tests can be removed by number or by their (exact) label.

```{r}
sfm <- discard_unit_test(sfm, nr = 3)
unit_tests(sfm)
```


## Specify simulation settings for tests

By default, `verify()` runs a single simulation with the default settings. Different simulation settings (as in `sim_settings`) can be passed to `verify`:

```{r}
verify(sfm, seed = 123)
```

```{r}
verify(sfm, dt = .1)
```

## Ensemble unit tests

When there are stochastic elements in the model, it is useful to run unit tests multiple times. For example, the SIR model can be given random initial conditions:

```{r}
sfm <- sfm |>
  update(c(susceptible, infected, recovered), eqn = runif(1, 500, 1000)) 
```

Reduce the output size for speed:

```{r}
sfm <- sim_settings(sfm, save_n = 100)
```

Use `n > 1` to run multiple simulations per test:

```{r}
verify(sfm, n = 10)
```

To speed up the simulation, we can use Julia:

```{r}
sfm <- sim_settings(sfm, language = "Julia", return_sims = FALSE)
use_julia()
```

```{r}
verify(sfm, n = 100)
```

Close the Julia session after use:

```{r}
use_julia(stop = TRUE)
sfm <- sim_settings(sfm, language = "R")
```

Alternatively, set up multi-session parallel simulation with the `future` package: 

```{r, eval = FALSE}
future::plan(future::multisession, workers = parallelly::availableCores() - 1)
```


```{r}
verify(sfm, n = 100)
```

Close parallel workers after use:

```{r, eval = FALSE}
future::plan(future::sequential)
```

## Inspecting `verify()` results and returned simulations

To inspect the simulations on which tests are evaluated, set `return_sims = TRUE`.

```{r}
sfm <- sim_settings(sfm, return_sims = TRUE)
```

Or specify `return_sims` directly in `verify()`:

```{r}
res <- verify(sfm, return_sims = TRUE)
```

This allows the simulations used for testing to be retrieved and inspected:

```{r}
plot(res)
```

Plot the simulation used for a specific test:

```{r}
plot(res, nr = 1L)
```

View results as a data frame:

```{r}
as.data.frame(res)
```

Retrieve results for a specific test by number:

```{r}
as.data.frame(res, nr = 1)
```

## Inspect failed tests

To demonstrate how to inspect failed tests, we will deliberately break the model by making the susceptible population negative:

```{r}
sfm <- update(sfm, susceptible, eqn = -10)
```

```{r}
res <- verify(sfm, return_sims = TRUE)
print(res)
```

```{r}
plot(res, status = "fail")
```

```{r}
as.data.frame(res, status = "fail")
```
